---
title: "20_CT26_010_Synergy_test.Rmd"
author: "MX"
date: "7/24/2020"
output: html_document
---

Summary

20_CT26_010 is a study with compound F20885 alone and in combination with aVEGFR Antibody treatment. This analysis is to see if there is synergy at any dose
This is not a real synergy analysis, since there is only a single dose point, but we want to know if there is any synergy at each dose combination, 
assuming the F20885 and aVEGFR treatment are completely independent and an additive effect is the product of A x B


```{r}
library(tidyverse)
library(dplyr)
library(magrittr)
library(tibble)
library(ggplot2)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
sum_mx <- function(x) {sum(x,na.rm=TRUE)}
mean_mx <- function(x) {mean(x,na.rm=TRUE)}
```

Load data
Each group has 10 mice
```{r}
library(ggbeeswarm)
data <- read.table("/Volumes/Picard/FLX/GCN2/20_CT26_010_synergy/20_CT26_010_tumor_vol_Day25.txt", sep="\t", header=TRUE) %>% dfilter(!is.na(Tumor_size)) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle","F20855_5mpk","F20855_10mpk","F20855_15mpk","VEGFR","F20855_5mpk_VEGFR","F20855_10mpk_VEGFR","F20855_15mpk_VEGFR")))
unique(data$Treatment)
data %>% ggplot(aes(Treatment, Tumor_size, group=Treatment)) + geom_boxplot() + geom_beeswarm() + theme(axis.text.x = element_text(angle=45,size=12, vjust=1, hjust=1),strip.text=element_text(size=12)) 
```

Set max tumor volume at 1500
```{r}
data_max <- data %>% mutate(Tumor_size= if_else(Tumor_size >= 1500, 1500, Tumor_size))

data_max %>% ggplot(aes(Treatment, Tumor_size, group=Treatment)) + geom_boxplot() + geom_beeswarm() + theme(axis.text.x = element_text(angle=45,size=12, vjust=1, hjust=1),strip.text=element_text(size=12)) 
```

Without a maximum at 1500mm3
Calculate the expected % therapeutic efficacy by random sampling 885/Vehicle, VEGFR/Vehicle and calculating the product of that
```{r fig.height=6, fig.width=4}
sample_mx <- function(x) {
  rand <- sample(1:10,3, replace= TRUE) # pick a vehicle sample
  veh <- data %>% dfilter(Treatment=="Vehicle") %>% pull(Tumor_size) %>% .[rand[1]]
  var1 <- data %>% dfilter(Treatment=="F20855_10mpk") %>% pull(Tumor_size) %>% .[rand[2]]
  var2 <- data %>% dfilter(Treatment=="VEGFR") %>% pull(Tumor_size) %>% .[rand[3]]
  exp <- (var1/veh) * (var2/veh) 
  #mpk20 <- var1/veh
  return(exp)
  }

#exp_vegf <- sapply(seq_along(1:10000), sample_mx)
#exp_mpk10 <- sapply(seq_along(1:10000), sample_mx)
#exp_mpk20 <- sapply(seq_along(1:10000), sample_mx)
exp_5mpk_combo2 <- sapply(seq_along(1:10000), sample_mx)
exp_10mpk_combo2 <- sapply(seq_along(1:10000), sample_mx)
exp_15mpk_combo2 <- sapply(seq_along(1:10000), sample_mx)

median(exp_5mpk_combo)
median(exp_10mpk_combo)
median(exp_20mpk_combo)

median(exp_5mpk_combo2)
median(exp_10mpk_combo2)
median(exp_15mpk_combo2)

getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
getmode(exp_5mpk_combo2)
median(exp_10mpk_combo2)
median(exp_20mpk_combo2)

data.frame(combo_5mpk= exp_5mpk_combo2, combo_10mpk=exp_10mpk_combo2, combo_15mpk=exp_15mpk_combo2) %>% gather(Combo,Effect) %>% mutate(Dose=if_else(grepl("5mpk",Combo),"5mpk",if_else(grepl("10mpk",Combo),"10mpk","15mpk"))) %>% mutate(Dose=factor(Dose,levels=c("5mpk","10mpk","15mpk"))) %>% mutate(Type=if_else(grepl("exp",Combo),"Expected","Observed"))  %>% mutate(Type= factor(Type)) %>% ggplot(aes(x=Effect,alpha=0.8, fill=Type,color=Type)) + geom_histogram(binwidth=0.05) +facet_wrap(~Dose,ncol=1) + xlim(0,2) + nolegend() 
  
  stat_density_ridges(binwidth=0.1,height=0.9,size=0.5, rel_min_height=0.01) + theme(text = element_text(size = 16)) +xlim(0,3)
```
Single agent
Calculate the observed % therapeutic efficacy by random sampling 885/Vehicle, VEGFR/Vehicl
```{r fig.height=6, fig.width=4}

sample2 <- function(x) {
  rand <- sample(1:10,3, replace= TRUE) # pick a vehicle sample
  veh <- data %>% dfilter(Treatment=="Vehicle") %>% pull(Tumor_size) %>% .[rand[1]]
  var1 <- data %>% dfilter(Treatment=="F20858_20mpk") %>% pull(Tumor_size) %>% .[rand[2]]
  #var2 <- data %>% dfilter(Treatment=="VEGFR") %>% pull(Tumor_size) %>% .[rand[3]]
  #exp <- (var1/veh) * (var2/veh) 
  obs <- var1/veh
  return(obs)
  }

#exp_vegf <- sapply(seq_along(1:10000), sample_mx)
#exp_mpk10 <- sapply(seq_along(1:10000), sample_mx)
#exp_mpk20 <- sapply(seq_along(1:10000), sample_mx)
obs_5mpk <- sapply(seq_along(1:10000), sample2)

obs_10mpk <- sapply(seq_along(1:10000), sample2)

obs_20mpk <- sapply(seq_along(1:10000), sample2)

data.frame(obs_5mpk, obs_10mpk, obs_20mpk) %>% gather(Combo,Effect) %>% mutate(Dose=if_else(grepl("5mpk",Combo),"5mpk",if_else(grepl("10mpk",Combo),"10mpk","20mpk"))) %>% mutate(Dose=factor(Dose,levels=c("5mpk","10mpk","20mpk"))) %>% ggplot(aes(x=Effect,alpha=0.8, fill=Type,color=Type)) + geom_histogram(binwidth=0.05) +facet_wrap(~Dose,ncol=1) + xlim(0,2) + nolegend() 
  
median(obs_5mpk_combo2)
median(obs_10mpk_combo2)
median(obs_20mpk_combo2)
```

Combo: Calculate the observed median combo effect by random sampling the n=10 data
```{r fig.height=6, fig.width=5}
library(ggridges)

sample_combo <- function(x) {
  rand <- sample(1:10,2, replace= TRUE) # pick a vehicle sample
  veh <- data %>% dfilter(Treatment=="Vehicle") %>% pull(Tumor_size) %>% .[rand[1]]
  var1 <- data %>% dfilter(Treatment=="F20855_5mpk_VEGFR") %>% pull(Tumor_size) %>% .[rand[2]]
  Obs <- var1/veh
  return(Obs)
}

#Obs_5mpk <- sapply(seq_along(1:10000), sample_combo)
#Obs_10mpk <- sapply(seq_along(1:10000), sample_combo)
#Obs_20mpk <- sapply(seq_along(1:10000), sample_combo)

Obs_5mpk2 <- sapply(seq_along(1:10000), sample_combo)
Obs_5mpk_combo2 <- Obs_5mpk2
Obs_10mpk2 <- sapply(seq_along(1:10000), sample_combo)
Obs_10mpk_combo2 <- Obs_10mpk2
Obs_15mpk2 <- sapply(seq_along(1:10000), sample_combo)
Obs_15mpk_combo2 <- Obs_15mpk2
median(Obs_20mpk)

obs <- cbind(obs_5mpk_combo,obs_10mpk_combo,obs_15mpk_combo) %>% data.frame

#obs_exp <- cbind(exp_5mpk_combo,exp_10mpk_combo,exp_20mpk_combo,obs_5mpk_combo=Obs_5mpk,obs_10mpk_combo=Obs_10mpk,obs_20mpk_combo=Obs_20mpk) %>% data.frame

obs_exp2 <- cbind(exp_5mpk_combo2,exp_10mpk_combo2,exp_15mpk_combo2,Obs_5mpk_combo2,Obs_10mpk_combo2,Obs_15mpk_combo2) %>% data.frame


obs %>% gather(Combo, Effect) %>% mutate(Dose=if_else(grepl("15mpk",Combo),"15mpk",if_else(grepl("10mpk",Combo),"10mpk","5mpk"))) %>% mutate(Type=if_else(grepl("exp",Combo),"Expected","Observed")) %>% mutate(Dose=factor(Dose,levels=c("5mpk","10mpk","15mpk"))) %>% ggplot(aes(Effect, alpha=0.5, color=Type, fill=Type)) + geom_histogram(binwidth=0.05) + facet_wrap(~Dose,ncol=1) + nolegend() +xlim(0,2)

obs_exp2 %>% gather(Combo, Effect) %>% dfilter(Effect < 1) %>% mutate(Dose=if_else(grepl("15mpk",Combo),"15mpk",if_else(grepl("10mpk",Combo),"10mpk","5mpk"))) %>% mutate(Type=if_else(grepl("exp",Combo),"Expected","Observed")) %>% ggplot(aes(Effect,colour=Type,fill=Type, alpha=0.5)) + geom_freqpoly(binwidth=0.1) + facet_wrap(~Dose)

obs_exp2 %>% gather(Combo, Effect)  %>% mutate(Dose=if_else(grepl("15mpk",Combo),"15mpk",if_else(grepl("10mpk",Combo),"10mpk","5mpk"))) %>% mutate(Dose=factor(Dose,levels=c("15mpk","10mpk","5mpk"))) %>% mutate(Type=if_else(grepl("exp",Combo),"Expected","Observed")) %>% mutate(Type= factor(Type)) %>% ggplot(aes(y=Dose,x=Effect,alpha=0.8, fill=Type,color=Type)) + stat_density_ridges(quantile_lines=TRUE,scale=0.9, binwidth=0.5,height=0.9,size=0.5, rel_min_height=0.1, joint_bandwidth=0.3) + theme(text = element_text(size = 16)) + nolegend() +xlim(0,2)
#%>% dfilter(Effect < 1)
obs_exp %>% gather(Combo, Effect) %>% dfilter(Effect < 1) %>% mutate(Dose=if_else(grepl("5mpk",Combo),"5mpk",if_else(grepl("10mpk",Combo),"10mpk","20mpk"))) %>% mutate(Dose=factor(Dose,levels=c("5mpk","10mpk","20mpk"))) %>% mutate(Type=if_else(grepl("exp",Combo),"Expected","Observed")) %>% mutate(Type= factor(Type)) %>% ggplot(aes(y=Dose,x=Effect,alpha=0.8, fill=Type,color=Type)) + stat_density_ridges(quantile_lines=TRUE,scale=0.9, binwidth=0.5,height=0.9,size=0.5, rel_min_height=0.1, joint_bandwidth=0.3) + nolegend() + theme(text = element_text(size = 16))  

median(Obs_5mpk_combo2)
median(Obs_10mpk_combo2)
median(Obs_15mpk_combo2)

geo.mean(exp_5mpk_combo2)
geo.mean(exp_10mpk_combo2)
geo.mean(exp_15mpk_combo2)
geo.mean(Obs_5mpk_combo2)
geo.mean(Obs_10mpk_combo2)
geo.mean(Obs_15mpk_combo2)

x <- c(0.80, 0.83, 1.89, 1.04, 1.45, 1.38, 1.91, 1.64, 0.73, 1.46) # median 1.4
y <- c(1.15, 0.88, 0.90, 0.74, 1.21) #median 0.9
median(x)
median(y)
wilcox.test(x, y, alternative = "g")        # greater
```
Wilcoxon Signed rank test, non-paired data
Greater = refers to the first listed variable being greater than the second one, not the other way around
```{r}
wilcox.test(obs_exp2$exp_5mpk_combo2,obs_exp2$Obs_5mpk_combo2, paired=FALSE,alternative="greater") #W = 59415274, p-value < 2.2e-16
wilcox.test(obs_exp2$exp_10mpk_combo2,obs_exp2$Obs_10mpk_combo2, paired=FALSE,alternative="greater") #W = 62386543, p-value < 2.2e-16
wilcox.test(obs_exp2$exp_15mpk_combo2,obs_exp2$Obs_15mpk_combo2, paired=FALSE,alternative="less") #W = 50577838, p-value = 0.157

wilcox.test(obs_exp2$exp_5mpk_combo2[1:1000],obs_exp2$Obs_5mpk_combo2[1:1000], paired=FALSE, conf.int=TRUE) #W = 599275, p-value = 1.496e-14
wilcox.test(obs_exp2$exp_5mpk_combo2[1:500],obs_exp2$Obs_5mpk_combo2[1:500], paired=FALSE) #W = 143888, p-value = 3.534e-05
wilcox.test(obs_exp2$exp_5mpk_combo2[1:100],obs_exp2$Obs_5mpk_combo2[1:100], paired=FALSE, alternative="less") # W = 4893, p-value = 0.7947
```
Cut off the Effect < 1, and test again
```{r}
obs_exp2 %>% mutate(stupid= seq_along(1:10000)) %>% gather(combo, effect, -stupid) %>% dfilter(effect <= 1) %>% mutate(combo=factor(combo)) %>% pivot_wider(names_from=combo, values_from=effect, id_cols=stupid) #  

obs_fil <- obs_exp2 %>% mutate(across(everything(), function(x) {x[x>1] <- NA; x}))

wilcox.test(obs_fil$exp_5mpk_combo2,obs_fil$Obs_5mpk_combo2, paired=FALSE) #W = 59415274, p-value = 0.08075
wilcox.test(obs_fil$exp_10mpk_combo2,obs_fil$Obs_10mpk_combo2, paired=FALSE) #W = 62386543, p-value = 0.9
wilcox.test(obs_fil$exp_15mpk_combo2,obs_fil$Obs_15mpk_combo2, paired=FALSE,alternative="less") #W = 50577838, p-value = 0.157
```

