---
title: "Azizi_PBMC.rmd"
author: "Mengshu"
date: "2/22/2019"
output: html_document
---

The Azizi et al. PBMC dataset from breast cancer patients is sequenced with the InDrop system. InDrop is the non-commercial equivalent of 10X scRNA seq. 
I will perform analysis with raw read counts without imputation for variable gene selection, defining PCs, clustering, and differential analysis.
When all these are finished, I will use imputed data for data display purposes only. 

A focus for this analysis will be to find all possible DC and monocyte subtypes, with special attention to presence of regulatory markers. Use the classification system defined by Villani et al. from the Satija Regev collaboration scRNA study.  

Data is parse from GSE114725_rna_raw.csv using code "Azizi_parse_raw.Rmd"
PBMCs extracted from normal tissues, blood, and tumor of the same patient are all included
47016 cells by 14875 genes

Load packages
```{r}
#library(SAVER)
library(Seurat)
library(ggplot2)
library(dplyr)
library(readr) # contains write_tsv()
library(magrittr)
library(tidyr)
library(monocle)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
```


Load data into table
47016 cells by 14875 genes, split up into 5000 genes per batch
Cells in each category: 
    BLOOD LYMPHNODE    NORMAL     TUMOR 
    16056      5167      4540     21253 
```{r}
azizi_p <- read.table("Azizi_pbmc_raw_all_tissue.csv", header=TRUE, row.names=1, sep=",")
azizi_pbmc <- CreateSeuratObject(raw.data =azizi_p) # don't filter anything out
saveRDS(azizi_pbmc,file="Azizi_PBMC_Seurat.rds")
azizi_pbmc <- readRDS("Azizi_PBMC_Seurat.rds")

levels(myeloid@ident)
saveRDS(myeloid, "Azizi_Myeloid.rds")
myeloid <- readRDS("Azizi_Myeloid.rds")

saveRDS(DCsubgroup, "Azizi_DCsubgroup.rds")
DCsubgroup <- readRDS("Azizi_DCsubgroup.rds")
 ```

Check Markers

```{r fig.height=20, fig.width=30}
FeaturePlot(object=azizi_pbmc, features.plot = c("IKBKE","TBK1"), cols.use=c("grey","blue"), no.legend=FALSE,reduction.use="umap", use.imputed=TRUE) 
VlnPlot(object=azizi_pbmc, features.plot = c("IKBKE","TBK1"), x.lab.rot = TRUE,use.imputed=TRUE) 
```

Insert the SAVER Imputed data into the Imputed slot
There was an error due to non-matching cell names in colnames and azizi_pbmc@cell.names, just insert one into the other
```{r}
imputed <- readRDS("azizi_orig_saver_Mo.rds")
#str(saver_imp)
saver_imp <- data.frame(imputed)
#fix the cell names 
azizi_pbmc@imputed <- saver_imp
colnames(azizi_pbmc@imputed) <- azizi_pbmc@cell.names
#str(azizi_pbmc@imputed)
```

Add metadata for tissue and donor
```{r}
cell_names <- colnames(azizi_pbmc@raw.data) # save all the cell names, then parse out metadata from them
tissue <- sapply(strsplit(cell_names, "_"),`[`,2) #this splits the string and recovers the second element
donor <- sapply(strsplit(cell_names, "_"),`[`,1)
umi <- azizi_pbmc@meta.data$nUMI
metadata <- data.frame(donor,tissue,umi)
metadata %>% group_by(donor) %>% count(tissue)
summary_pbmc <- metadata %>% group_by(donor) %>% count(tissue)
write.table(spread(summary_pbmc, "donor", "n"),"Tissue sampling Azizi.txt", row.names=FALSE, sep="\t")
table(as.factor(tissue))
table(as.factor(donor))
azizi_pbmc@meta.data$tissue <- as.factor(tissue)
azizi_pbmc@meta.data$donor <- as.factor(donor)
```
Add mitochondrial and ribosomal protein metadata

```{r}
mito.genes <- grep(pattern="^MT.", x=rownames(azizi_pbmc@data),value=TRUE,perl= TRUE)
percent.mito <- Matrix::colSums(azizi_pbmc@raw.data[mito.genes,])/Matrix::colSums(azizi_pbmc@raw.data)
azizi_pbmc <- AddMetaData(object=azizi_pbmc, metadata=percent.mito, col.name= "percent.mito")

ribo.genes <- grep(pattern="^RP[S|L]", x=rownames(x=azizi_pbmc@raw.data),value=TRUE)
percent.ribo <- Matrix::colSums(azizi_pbmc@raw.data[ribo.genes,])/Matrix::colSums(azizi_pbmc@raw.data)
azizi_pbmc <- AddMetaData(object=azizi_pbmc, metadata=percent.ribo, col.name= "percent.ribo")
```

Inspect data
```{r fig.height=15, fig.width=15}
VlnPlot(object=azizi_pbmc, features.plot=c("nGene","nUMI"), group.by="donor",nCol=2)
ggsave("Azizi_PBMC_UMI_donor.jpg", dpi= 600, width=10, height=5, units= "in", plot=last_plot())
VlnPlot(object=azizi_pbmc, features.plot=c("nGene","nUMI","percent.mito","percent.ribo"), group.by="tissue",nCol=2,point.size.use = 0.05) +
ggsave("Azizi_PBMC_UMI_tissue.jpg", dpi= 600, width=15, height=15, units= "in", plot=last_plot())
```

Normalize data
Scale data by regressing out certain factors
Find variable genes #3220

```{r fig.height=12, fig.width=15}
memory.limit(size=56000)
azizi_pbmc <- NormalizeData(object=azizi_pbmc)
azizi_pbmc <- ScaleData(object=azizi_pbmc, vars.to.regress=c("percent.ribo","percent.mito","nUMI"), display.progress = TRUE, assay.type = "RNA", check.for.norm = TRUE)

```

Variable Genes
```{r}
azizi_pbmc <- FindVariableGenes(azizi_pbmc, mean.function = ExpMean,
  dispersion.function = LogVMR, do.plot = TRUE, set.var.genes = TRUE,
  x.low.cutoff = 0.0125,x.high.cutoff = 3.5, y.cutoff = 0.5,
  y.high.cutoff = Inf, num.bin = 20, binning.method = "equal_width",
  selection.method = "mean.var.plot", top.genes = 1000, do.recalc = TRUE,
  sort.results = TRUE, do.cpp = TRUE, display.progress = TRUE)


length(azizi_pbmc@var.genes) #3220
```


PCA
#run on Variable genes
PC1: CD8 vs Monocyte

```{r}
azizi_pbmc <- RunPCA(object=azizi_pbmc, pc.genes=azizi_pbmc@var.genes, do.print=TRUE, pcs.print=1:10,pcs.compute = 25)
PrintPCA(object = azizi_pbmc, pcs.print = 20:25, genes.print = 5, use.full = FALSE)
```

Visualize PCAs
```{r}
VizPCA(object = azizi_pbmc, pcs.use = 1:3)
PCAPlot(object = azizi_pbmc, dim.1 = 1, dim.2 = 2)
PCHeatmap(object = azizi_pbmc, pc.use = 1, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
```

Assess PCS
Jackstraw is taking too long to run
```{r}
azizi_pbmc <- JackStraw(object = azizi_pbmc, num.replicate = 100, display.progress = FALSE)
JackStrawPlot(object = azizi_pbmc, PCs = 1:25)
PCElbowPlot(object = azizi_pbmc,PCs = 1:25) #20 clusters looks fine
```

Cluster
```{r}
azizi_pbmc <- FindClusters(object = azizi_pbmc, reduction.type = "pca", dims.use = 1:25, 
    resolution = 1, print.output = 0, save.SNN = TRUE)
```
##################
Run UMAP for visualization as comparison
###################
Seurat has a UMAP function, install the python version of umap on the local machine first, Seurat will call the python code from R
In Terminal> sudo pip install umap-learn, then restart the R session
n_neighbors = 30L determines preservation of global structures (larger numbers). Typical range 5-50.
Metrics: for this dataset, "manhattan"" works the best
metric: chebyshev = manhattan (better borders) > cosine  > correlation > canberra (chaotic and clumpy) >>>> Jaccard (for binary data) >> Haversine (doens't run)
n_neighbours 10 or 30 didn't make much difference, the clusters are very tight
```{r fig.height=10, fig.width=15}
library(umap)
azizi_pbmc <- RunUMAP(azizi_pbmc, cells.use = NULL, dims.use = 1:25, reduction.use = "pca",
  genes.use = NULL, assay.use = "RNA", max.dim = 2L,
  reduction.name = "umap", reduction.key = "UMAP", n_neighbors = 10L,
  min_dist = 0.2, metric = "manhattan", seed.use = 42)

DimPlot(object = azizi_pbmc, do.label = TRUE, reduction.use = 'umap', pt.size = 0.8, label.size = 4) 
ggsave("Azizi_PBMCs_UMAP.jpg", width= 12, height=8, dpi=600, plot= last_plot(), units = "in")
```
Markers
T cell
```{r fig.height=20, fig.width=26}
FeaturePlot(object=azizi_pbmc, features.plot = c("CD3D","CD8A","CD4","FOXP3","CCR7","GZMB","LYZ","ITGAX","KLRB1","SELL","PRF1","GNLY","NKG7","CD14","ITGAE","CD69","CD79A","CXCL13","ITGAM","PDCD1","TWIK2"), cols.use=c("grey","blue"), no.legend=FALSE,reduction.use="umap") 
ggsave("Azizi_PBMC_Global Markers.jpg", width= 25, height=18, dpi=600, plot= last_plot(), units = "in")

FeaturePlot(object=azizi_pbmc, features.plot = c("CD276","PDCD1","NRG1","LILRA1"), cols.use=c("grey","blue"), no.legend=FALSE,reduction.use="umap") 
```
#########
#####
#####################
######
###############################
Subset into Myeloid lineage
Clusters: 9, 12, 15, 16, 20, 22, and 23 is DC1, (far away from the rest)
#########
#####
#############################################
######
############################################################

```{r}
myeloid <- SubsetData(azizi_pbmc, ident.use=c("9","12","15","16","20","22","23"), do.clean=TRUE, subset.raw=TRUE)
levels(myeloid@ident)
saveRDS(myeloid, "Azizi_Myeloid.rds")

```

Re-process Myeloid cells alone

```{r fig.height=25, fig.width=30}
VlnPlot(myeloid, features.plot = c("nUMI","nGene","CD276","C3AR1","C5AR1"))
myeloid <- NormalizeData(myeloid)
myeloid <- ScaleData(myeloid,display.progress = TRUE, assay.type = "RNA", check.for.norm = TRUE)
myeloid <- FindVariableGenes(myeloid, mean.function = ExpMean,
  dispersion.function = LogVMR, do.plot = TRUE, set.var.genes = TRUE,
  x.low.cutoff = 0.0125,x.high.cutoff = 10, y.cutoff = 0,
  y.high.cutoff = Inf, num.bin = 20, binning.method = "equal_width",
  selection.method = "mean.var.plot", top.genes = 2000, do.recalc = TRUE,
  sort.results = TRUE, do.cpp = TRUE, display.progress = TRUE,cex.text.use=1)
VariableGenePlot(myeloid, cex.text.use=1)
```

Skip both Normalization and ScaleData, then put raw.data into the imputed slot, and run RunPCA on the raw data in "imputed" slot
Variable genes don't include ITGAX or ITGAM
The PCAs don't include ITGAM and ITGAX, so CDs and Macrophages are not resolving
Try running PCA using all genes

```{r fig.height=25, fig.width=30}
length(myeloid@var.genes) #3350
myeloid@imputed <- myeloid@raw.data
myeloid <- RunPCA(object=myeloid, pc.genes=myeloid@var.genes, do.print=TRUE, pcs.print=1:10,pcs.compute = 25, use.imputed=TRUE)
# use all genes
??RunPCA
myeloid <- RunPCA(object=myeloid, pc.genes=rownames(myeloid@data), do.print=TRUE, pcs.print=1:10,pcs.compute = 25)
PCElbowPlot(myeloid, num.pc=25)
PCE
PrintPCA(myeloid, pcs.print= 1:20, genes.print= 50)

grep("ITGAX",myeloid@var.genes)
myeloid@var.genes[782:786]
```
Myeloid Markers to cluster on:

```{r}
myeloid_markers <- c("CD14","CD33","HLA.DRA","CD163","CD14","MRC1","ITGAX","ITGAM","CX3CR1","THBD","CD83","CD86","CD1A","CD1C","CD83","CD86","SLAMF8","CD59","CD1C","CLEC10A","S100A9","LYZ","FCGR3A","LST1","DAB2","IL3RA","GZMB")
```

Regulatory Markers
```{r}
regulatory_markers <- c("PDCD1","TGFB1","IL10","ITGAM","ITGAX","IFNG","IL1B","CD276","FAS", "PDCD1","PDCD1LG2", "IL6", "VTCN1","ITGAE", "ILT2", "ILT3","ILT4")

```


Cluster
```{r}
myeloid <- FindClusters(object = myeloid,
    resolution = 0.7, print.output = 0, save.SNN = TRUE, reduction.type = "pca", dims.use = 1:25, plot.SNN=TRUE,force.recalc=TRUE) # genes.use=c("SLAMF8","CD59","CD1C","CLEC10A","S100A9","FCGR3A","DAB2","NRP1","UGCG")
??FindClusters()
#TSNE
myeloid <- RunTSNE(myeloid, reduction.use = "pca", seed.use=1, tsne.method = "Rtsne",dims.use=1:15, check_duplicates = FALSE, perplexity=100)

TSNEPlot(myeloid, do.label =TRUE)

#UMAP
myeloid <- RunUMAP(myeloid, cells.use = NULL, dims.use = 1:25, reduction.use = "pca",
  genes.use = NULL, assay.use = "RNA", max.dim = 2L,
  reduction.name = "umap", reduction.key = "UMAP", n_neighbors = 50L,
  min_dist = 0.1, metric = "manhattan", seed.use = 13)

DimPlot(object = myeloid, do.label = TRUE, reduction.use = 'umap', pt.size = 0.8, label.size = 4) +
ggsave("Azizi_myeloid_UMAP_allgene_25PC.jpg", width= 9, height=6, dpi=600, plot= last_plot(), units = "in")
```
Using 25 PCs makes too many clusters (16), reduce resolution from 1 to 0.3
CD141 is THBD
```{r fig.height=10, fig.width=15}
FeaturePlot(object=myeloid, features.plot = c("CSF1","SLAMF8","CD59","CD1C","CLEC10A","S100A9","LYZ","FCGR3A","LST1","DAB2","IL3RA","GZMB","CD276"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap", nCol=4) 
 FeaturePlot(object=myeloid, features.plot = c("SLAMF8","CD59","CD1C","CLEC10A","S100A9","LYZ","FCGR3A","LST1","DAB2","IL3RA","GZMB","CD276"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap", use.imputed = TRUE)
ggsave("Myeloid_ClusterGenePlot.jpg", width= 15, height=10, dpi=600, plot= last_plot(), units = "in")

FeaturePlot(object=myeloid, features.plot = c("CD14","CD84","FCGR3A","CD1C","CD1E","IL3RA","LILRA1","ITGAX","ITGAM","CD276","SPP1","CXCR3","MARCO", "CD163","CX3CR1","LILRB2","THBD","CLEC9A","GZMB"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap") +
ggsave("Myeloid_LineageGenePlot.jpg", width= 15, height=10, dpi=600, plot= last_plot(), units = "in")
FeaturePlot(object=myeloid, features.plot = c("ITGAX","ITGAM","CXCR3","MARCO", "CD163","CD59", "HLA.DRA","HLA.DRB1","nUMI"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap")
FeaturePlot(object=myeloid, features.plot = c("LILRB1","LILRB2","TGFB1","IL1B","KYNU","CD276","C3AR1","C5AR1","SIGLEC1"), cols.use=c("grey90","blue"),no.legend=TRUE,reduction.use="umap",)
ggsave("Myeloid_RegGenePlot.jpg", width= 10, height=12, dpi=600, plot= last_plot(), units = "in")
FeaturePlot(object=azizi_pbmc, features.plot = c("AXL","IL3RA","CLEC4C","GZMB"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap") 
```

Make heatmap of markers
```{r fig.height=35, fig.width=35}
PCAPlot(myeloid)
PCASigGenes(myeloid)
myeloid.markers <- FindAllMarkers(object = myeloid, only.pos = TRUE, min.pct = 0.25, 
    thresh.use = 0.25)
myeloid.markers %>% group_by(cluster) %>% top_n(20, avg_logFC)
#This is the cluster 2 that includes SPP1, CD59, CD276 possibly regulatory
Cluster_2_Reg <-  myeloid.markers  %>% dplyr::filter(cluster==2) %>% select(gene)
grep("CXCR3", myeloid.markers)
myeloid.markers[7]
write.table(Cluster_2_Reg,file ="Cluster2_markers.txt",quote=FALSE, row.names=FALSE)
top10 <- myeloid.markers %>% group_by(cluster) %>% top_n(20, avg_logFC)
DoHeatmap(object = myeloid, genes.use=top10$gene,slim.col.label = TRUE) +
ggsave("Azizi_myeloid_UMAP_allgene_10PC.jpg", width= 35, height=30, dpi=600, plot= last_plot(), units = "in")
```
Heatmap of potential regulatory markers
CD123=IL3RA is DC marker
CD59: Marker of DC1 cells
CLEC9A: DC1 marker
SLAMF8
CDC141 BDCA3 THBD: Classic DC1 marker that is also expressed on DC5 DC6",
Cluster 5
```{r fig.height=5, fig.width=10}
regulatory_markers <- c("HLA.DRB1","CD14","ITGAM","ITGAX","CD163","CD169","SIGLEC1","SERPING","MARCO","CD59","SLAMF8","IL3RA","CD1C","THBD","CLEC9A","IL12","KYNU","IDO1","PDCD1","TGFB1","IL10","IFNG","IL1B","CD273","CD276","FAS", "PDCD1","PDCD1LG2", "IL6", "VTCN1","ITGAE","LILRA1","LILRB1", "LILRB2","LAIR1","RGCC","CX3CR1")
DoHeatmap(object = myeloid, genes.use=regulatory_markers,slim.col.label = TRUE) +
ggsave("Azizi_myeloid_heatmap_my_markers.jpg", width= 10, height=5, dpi=600, plot= last_plot(), units = "in")
```









Get marker lists
```{r}
myeloid.markers %>% dplyr::filter(cluster==16) %>% top_n(10, avg_logFC) %>% select(gene)
```

Markers of regulatory DC cells
Not expressed: "ARG1"
```{r fig.height=10, fig.width=20}
FeaturePlot(object=myeloid, features.plot = c("TGFB1","IL10","ITGAM","IFNG","IL1B","CD276","FAS", "PDCD1","PDCD1LG2", "IL6","ITGAX","KYNU","CD163","THBD","IL3RA"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap") 
VlnPlot(object=myeloid, features.plot = c("CSF1","CD34","TGFB1","IL10","ITGAM","ITGAX","IFNG","IL1B","CD276","FAS", "PDCD1","PDCD1LG2", "IL6"), x.lab.rot = TRUE, do.sort=TRUE, nCol=4)
#FeaturePlot(object=myeloid, features.plot = c("TGFB1","IL10","ITGAM","IL1B","CD276","IDO1", "PDCD1","PDCD1LG2", "IL6"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap", use.imputed = TRUE) 
ggsave("Myeloid_regDC_markers_Feature_noNorm.jpg", width= 12, height=6, dpi=600, plot= last_plot(), units = "in")
```
Label cell clusters

```{r}
current.cluster.ids <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)

current.cluster.ids <- c("0. DC1","1. DC2","2. Immature DC","3. DC2", "4. DC1", "5. DC3","6. DC4","7. DC5","8. Unknown","9. DC6","10.DC3","11")
new.cluster.ids <- c("0. DC1","1. DC2","2. Immature DC","3. DC2", "4. DC1", "5. DC3","6. DC4","7. DC5","8. Unknown","9. DC6","10.DC3","11. DC1")
myeloid@ident <- plyr::mapvalues(x = myeloid@ident, from = current.cluster.ids, to = new.cluster.ids)
levels(myeloid@ident)
```
################
############
Isolate clusters that express regulatory markers, and adjacent clusters to sub-cluster for the regDC-like group
Include precursor DC to have an anchor for Pseudotime analysis later
############
#################

```{r}
DCsubgroup <- SubsetData(myeloid, ident.use=c("1","2","5","6","7","9","10","12","14","15"), subset.raw= TRUE, do.clean= TRUE )
levels(DCsubgroup@ident)
saveRDS(DCsubgroup, "Azizi_DCsubgroup.rds")
saveRDS(myeloid_m, "Azizi_myeloid_Monocle.rds")
```
Process Data
#Changing parameters does not increase clustering above 9 
#Skip Normalization -> Doesn't help the signal inflation at all, did change the UMAP clustering
#Can't Skip ScaleData
```{r}
DCsubgroup <- NormalizeData(DCsubgroup)
DCsubgroup <- ScaleData(DCsubgroup,do.scale =FALSE,do.center=FALSE, display.progress = TRUE, assay.type = "RNA", check.for.norm = TRUE)
DCsubgroup <- FindVariableGenes(DCsubgroup, mean.function = ExpMean,
  dispersion.function = LogVMR, do.plot = TRUE, set.var.genes = TRUE,
  x.low.cutoff = 0.0125,x.high.cutoff = 3.5, y.cutoff = 0.5,
  y.high.cutoff = Inf, num.bin = 20, binning.method = "equal_width",
  selection.method = "mean.var.plot", top.genes = 1000, do.recalc = TRUE,
  sort.results = TRUE, do.cpp = TRUE, display.progress = TRUE)
length(DCsubgroup@var.genes) #3820
#trick the program into using raw.data inserted into imputed slot
DCsubgroup@imputed <- DCsubgroup@raw.data
DCsubgroup <- RunPCA(object=DCsubgroup, pc.genes=DCsubgroup@var.genes, do.print=TRUE, pcs.print=1:10,pcs.compute = 25, use.imputed=TRUE)

DCsubgroup <- RunPCA(object=DCsubgroup, pc.genes=rownames(DCsubgroup@data), do.print=TRUE, pcs.print=1:10,pcs.compute = 25, use.imputed=TRUE)

PCElbowPlot(DCsubgroup)
DCsubgroup <- FindClusters(object = DCsubgroup,
    resolution = 0.8, print.output = 0, save.SNN = TRUE, reduction.type = "pca", dims.use = 1:25, plot.SNN=TRUE,force.recalc =TRUE) # genes.use=c("SLAMF8","CD59","CD1C","CLEC10A","S100A9","FCGR3A","DAB2","NRP1","UGCG")
#exclude PC 1 because it has RP genes
DCsubgroup <- RunUMAP(DCsubgroup, cells.use = NULL, dims.use = 2:20, reduction.use = "pca",
  genes.use = NULL, assay.use = "RNA", max.dim = 2L,
  reduction.name = "umap", reduction.key = "UMAP", n_neighbors = 50L,
  min_dist = 0.1, metric = "manhattan", seed.use = 13)

DimPlot(object = DCsubgroup, do.label = TRUE, reduction.use = 'umap', pt.size = 0.8, label.size = 6, cols.use=c("#0000FF","#8B0000","#9400D3","#FF0000","#FFA500","#FF1493", "#00FF00","#FFD700","#32CD32","#00FFFF")) 

ggsave("Azizi_DCsubgroup_UMAP.jpg", width= 6, height=4, dpi=600, plot= last_plot(), units = "in")


table(DCsubgroup@ident)
```


```{r}
DoHeatmap(object = DCsubgroup, genes.use=regulatory_markers,slim.col.label = TRUE) +
ggsave("Azizi_Regsubgroup_heatmap_my_markers.jpg", width= 5, height=5, dpi=600, plot= last_plot(), units = "in")
#Nonregsubgroup <- SubsetData(myeloid, ident.use=c("3","4","8","11","13","16"), subset.raw= TRUE)
DoHeatmap(object = Nonregsubgroup, genes.use=regulatory_markers,slim.col.label = TRUE)
ggsave("Azizi_NonRegsubgroup_heatmap_my_markers.jpg", width= 5, height=5, dpi=600, plot= last_plot(), units = "in")
```

Violin plot showing increased gene content in CD276+ cells
```{r fig.height=8, fig.width=20}
VlnPlot(DCsubgroup, features.plot = c("nUMI","nGene","GAPDH","RPL10","CD276","HLA.DRA"), use.raw=TRUE, nCol=6)
ggsave("Azizi_DCsubgroup_RNAcontent.jpg", width= 20, height=8, dpi=600, plot= last_plot(), units = "in")
```

Macrophages (ITGAM CD11B) vs. DCs (ITGAX CD11C)
CD33: MDSC marker
CD163: Macrophage/M2 marker
SIGLEC1: Macrophage marker
CD169: Macrophage marker
CD206/MRC1: M2 marker
CD141/THBD thrombomodulin
CD1A, CD1C: Marker of some DCs, DC 2, 3
CD83
CD86
```{r fig.height=15, fig.width=15}
FeaturePlot(object=DCsubgroup, features.plot = c("ITGAM","ITGAX"), min.cutoff=c(0.0001, 0.0001),cols.use=c("grey","red","blue","green"), overlay= TRUE, reduction.use = "umap") 
FeaturePlot(object=DCsubgroup, features.plot = c("ITGAM","ITGAX","IL1B", "KYNU","LILRB1","LILRB2", "MRC1","CX3CR1","CD276","NCOR2"), min.cutoff=c(0.0001, 0.0001),cols.use=c("grey90","blue"), overlay= FALSE, reduction.use = "umap") 
FeaturePlot(object=DCsubgroup, features.plot = c("CD33","HLA.DRA","CD163","CD14","MRC1","ITGAX","THBD","CD83","CD86","CD1A","CD1C"), min.cutoff=c(0.1), overlay=FALSE, reduction.use = "umap", no.legend=FALSE, cols.use=c("grey90","blue")) 
VlnPlot(DCsubgroup, features.plot = c("CD33","HLA.DRA","CD163","CD14","MRC1","ITGAX","ITGAM","CX3CR1","THBD","CD83","CD86","CD1A","CD1C"), use.raw=TRUE, nCol=3)
ggsave("CT26_mMDSC.jpg", width= 8, height=5, dpi=600, plot=last_plot(), units = "in")
Differential Analysis
```



```{r}
Cluster_1 <- FindMarkers(object=DCsubgroup, ident.1=1, ident.2 =c(2,9,0,5), min.pct=0.1, logfc.threshold = 0.3) #8,
Cluster_1 %<>% mutate(gene=rownames(Cluster_1)) %>% arrange(avg_logFC)
Cluster_1 %<>% dfilter(gene != "^RP*")
write.table(Cluster_1,"Cluster_1_diff_2905.txt", quote=FALSE, sep="\t")
```


Check groupings

```{r}
FeaturePlot(DCsubgroup, features.plot = c("HLA.DPB1","TGFB1","ITGAM","ITGAX", "PDCD1","CD276","KLF4","CXCL8"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap") 
#ggsave("Azizi_DCsubgroup_Feature.jpg", width= 10, height=7, dpi=600, plot= last_plot(), units = "in")
VlnPlot(DCsubgroup, features.plot = c("HLA.DPB1","TGFB1","ITGAM","ITGAX", "PDCD1","CD276"), x.lab.rot = TRUE, nCol=3) #,use.raw=TRUE

FeaturePlot(object=DCsubgroup, features.plot = c("SLAMF8","CD59","CD1C","CLEC10A","S100A9","LYZ","FCGR3A","LST1","DAB2","IL3RA","GZMB","CD276"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap", nCol=4) 

FeaturePlot(object=DCsubgroup, features.plot = c("SLAMF8","CD59","CD1C","CLEC10A","S100A9","LYZ","FCGR3A","LST1","DAB2","IL3RA","GZMB","CD276"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap", nCol=4) 

FeaturePlot(object=DCsubgroup, features.plot = c("HLA.DPB1","TGFB1","ITGAM","ITGAX", "PDCD1","HCK"), cols.use=c("grey90","blue"),no.legend=FALSE,reduction.use="umap") +
ggsave("Azizi_DCsubgroup_Feature_HCK.jpg", width= 12, height=10, dpi=600, plot= last_plot(), units = "in")


ggsave("Azizi_DCsubgroup_Feature.jpg", width= 10, height=7, dpi=600, plot= last_plot(), units = "in")
VlnPlot(DCsubgroup, features.plot = c("SLAMF8","CD59","CD1C","CLEC10A","S100A9","LYZ","FCGR3A","LST1","DAB2","IL3RA","GZMB","CD276"), x.lab.rot = TRUE, nCol=4, use.raw=TRUE )
ggsave("Azizi_DCsubgroup_Violin.jpg", width= 10, height=7, dpi=600, plot= last_plot(), units = "in")
```
Tissue or Donor specific expression of regDC markers?

```{r fig.height=10, fig.width=10}
VlnPlot(DCsubgroup, features.plot = c("HLA.DPB1","TGFB1","ITGAM","ITGAX", "PDCD1","CD276"), x.lab.rot = TRUE, nCol=3, group.by="donor", use.raw=TRUE)
ggsave("Azizi_DCsubgroup_Vln_donor.jpg", width= 12, height=8, dpi=600, plot= last_plot(), units = "in")
table(DCsubgroup@meta.data$tissue)
table(DCsubgroup@meta.data$donor)
```

Overlap of DC1, 2, 3 markers in the same cells?
```{r}

```


Pseudotime with Monocle 3

```{r}
library(monocle)

```
##########################################
Import Seurat object into Monocle 3
#################################
#######################
##############
########
```{r}
myeloid_m <- importCDS(myeloid, import_all= TRUE)
DCs <- importCDS(DCsubgroup, import_all=TRUE)
regDCm <- importCDS(regDC, import_all=TRUE)

save(myeloid_m, file='azizi_myeloid_Monocle.RData')
load('azizi_myeloid2_Monocle.RData') # don't pass to anything
load("azizi_regDC_Monocle.RData")

save(myeloid_m, file='azizi_myeloid2_Monocle.RData')
save(DCs, file='azizi_Myeloid_May3_Monocle.RData')
```
Export cluster information from Seurat to label cells in Monocle 3
```{r}
pData(myeloid_m)$cluster <- myeloid@ident

c("0. DC1","1. DC2","2. Immature DC","3. DC2", "4. DC1", "5. DC3","6. DC4","7. DC5","8. Unknown","9. DC6","10.DC3","11. DC1")

cell_type_colour <- c("0. DC1" = "#FF0000", #red
                      "1. DC2" ="#32CD32", #lgreen 
                      "2. Immature DC" ="#00FF00", #lime
                      "3. DC2" ="#228B22",  #green
                      "4. DC1" = "#FFA500", # orange
                      "5. DC3" = "#0000FF", #blue
                      "6. DC4" = "#696969", #grey
                      "7. DC5" ="#FFD700", #gold
                      "8. Unknown" ="#00FFFF", #aqua
                      "9. DC6" ="#FF1493", #hotpink
                      "10.DC3" = "#00008B", #navyblue
                      "11. DC1" ="#8B0000") #crimson

pData(DCs)$cluster <- DCsubgroup@ident

cell_type_colour <- c("0" = "#0000FF", #blue
                      "1" ="#8B0000", #crimson
                      "2" ="#9400D3", #purple
                      "3" ="#FF0000", #red
                      "4" = "#FFA500", #orange
                      "5" = "#FF1493", #hot pink
                      "6" = "#00FF00", #lime green
                      "7" ="#FFD700",  #gold
                      "8" ="#32CD32", #lgreen 
                      "9" ="#00FFFF") #aqua

pData(regDCm)$cluster <- regDC@ident

#Mar 25
cell_type_colour <- c("0" = "#9400D3", #purple
                      "1" ="#000000", #grey
                      "2" ="#9400D3", #purple
                      "3" ="#FFD700",  #gold
                      "4" = "#FFA500", #orange
                      "5" = "#FF1493", #hot pink
                      "6" = "#FF0000", #red
                      "7" ="#696969", #grey
                      "8" ="#32CD32", #lgreen 
                      "9" ="#696969", #grey
                      "10" ="#0000FF", #blue
                      "11" ="#0000FF", #blue
                      "12" ="#696969", #grey
                      "13" ="#00FF00", #lime green
                      "14" ="#696969", #grey
                      "15" ="#FF0000", #red
                      "16" = "#00008B") #navyblue

#d8d9d3

cell_type_colour <- c("1" ="#8B0000", #crimson
                      "2" ="#0000FF", #blue
                      "5" = "#9400D3", #purple
                      "6" = "#00FF00", #lime green
                      "7" ="#FF0000", #red
                      "9" ="#FFA500", #orange
                      "10" ="#FFD700",  #gold
                      "12" = "#FF1493", #hot pink
                      "14" = "#8B0000", #crimson
                      "15" ="#FF1493") #hot pink) 
```
Process

```{r}
myeloid_m <- estimateSizeFactors(myeloid_m)
myeloid_m <- estimateDispersions(myeloid_m)

DCs <- estimateSizeFactors(DCs)
DCs <- estimateDispersions(DCs)

regDCm <- estimateSizeFactors(regDCm)
regDCm <- estimateDispersions(regDCm)
```

Preprocess data to find pricipal components
```{r}
myeloid_m <- preprocessCDS(myeloid_m, num_dim=25)
myeloid_m <- reduceDimension(myeloid_m, reduction_method = 'UMAP')

myeloid_m <- clusterCells(myeloid_m,method = 'louvain',
                        res = 1e-6,
                        louvain_iter = 1,
                        verbose = T)

DCs <- preprocessCDS(DCs, num_dim=20)
DCs <- reduceDimension(DCs, reduction_method = 'UMAP')

DCs <- clusterCells(DCs,method = 'louvain',
                        res = 1e-6,
                        louvain_iter = 1,
                        verbose = T)

regDCm <- preprocessCDS(regDCm, num_dim=20)
regDCm <- reduceDimension(regDCm, reduction_method = 'UMAP')

regDCm <- clusterCells(regDCm,method = 'louvain',
                        res = 1e-6,
                        louvain_iter = 1,
                        verbose = T)
```

3. Partition cells into Supergroups
```{r}
myeloid_m <- partitionCells(myeloid_m)

DCs <- partitionCells(DCs)

regDCm <- partitionCells(regDCm)
```

4. Learn the Pricipal graph
SimplePPT and DDRTree are not giving different results
```{r}
myeloid_m <- learnGraph(myeloid_m, RGE_method ='SimplePPT') # partition_component +T
myeloid_m <- learnGraph(myeloid_m, RGE_method ='DDRTree')

DCs <- learnGraph(DCs, RGE_method ='DDRTree')

regDCm <- learnGraph(regDCm, RGE_method ='DDRTree')
```

5. Visualize trajectory
```{r fig.height=6, fig.width=8}
plot_cell_trajectory(myeloid_m , 
                     color_by = "cluster",
                     show_backbone=TRUE, show_branch_points = TRUE, alpha = 0.9)  +
                     scale_color_manual(values = cell_type_colour) + 
                     theme(legend.position = "bottom")  +
                     guides(colour=guide_legend(nrow=3))


ggsave("Azizi_myeloid_Supercluster_path_cluster.jpg", plot = last_plot(), dpi=400, height= 6, width = 10, units = "in")

plot_cell_trajectory(myeloid_m , color_by = "louvain_component",show_backbone = TRUE, backbone_color = "black") 
ggsave("Azizi_myeloid_Supercluster_louvain_component.jpg", plot = last_plot(), dpi=400, height= 6, width = 10, units = "in")

plot_cell_clusters(myeloid_m, color_by = pData(myeloid_m)$cluster) 


plot_cell_trajectory(DCs, 
                     color_by = "cluster",
                     show_backbone=TRUE, show_branch_points = TRUE, alpha = 0.9)  +
                     scale_color_manual(values = cell_type_colour) + 
                     theme(legend.position = "bottom")  +
                     guides(colour=guide_legend(nrow=3))
ggsave("Azizi_DCs_Psedotime_Path_cluster.jpg", plot = last_plot(), dpi=400, height= 8, width = 10, units = "in")


plot_cell_trajectory(regDCm, 
                     #color_by = "cluster",
                     show_backbone=TRUE, show_branch_points = TRUE, alpha = 0.9) # +
                     #scale_color_manual(values = cell_type_colour) + 
                     #theme(legend.position = "bottom")  +
                     #guides(colour=guide_legend(nrow=3))
                      
ggsave("Azizi_regDC_Psedotime_Path_cluster.jpg", plot = last_plot(), dpi=400, height= 8, width = 10, units = "in")

```
Root the tree

```{r}
azizi_node_ids = get_correct_root_state(myeloid_m,
                                      cell_phenotype =
                                        'cluster', "2. Immature DC")
myeloid_m <- orderCells(myeloid_m, root_pr_nodes = azizi_node_ids)
plot_cell_trajectory(myeloid_m)
ggsave("Azizi_myeloid_Supercluster_heatmap.jpg", plot = last_plot(), dpi=400, height= 6, width = 10, units = "in")


azizi_node_ids = get_correct_root_state(DCs,
                                      cell_phenotype =
                                        'cluster', "10")
DCs <- orderCells(DCs, root_pr_nodes = azizi_node_ids)
plot_cell_trajectory(DCs)
ggsave("Azizi_DCs_Supercluster_heatmap.jpg", plot = last_plot(), dpi=400, height= 6, width = 8, units = "in")


azizi_node_ids = get_correct_root_state(regDCm,
                                      cell_phenotype =
                                        'cluster', "6")
regDCm <- orderCells(regDCm, root_pr_nodes = azizi_node_ids)
plot_cell_trajectory(regDCm)
ggsave("Azizi_regDC_Supercluster_heatmap.jpg", plot = last_plot(), dpi=400, height= 6, width = 8, units = "in")
```
PLot Markers

```{r fig.height=15, fig.width=20} 
#"ITGAM","ITGAX",
plot_cell_trajectory(myeloid_m , 
                     markers = c("TGFB1","CD276","SPP1","LILRB2","CXCL10"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") + scale_color_gradient(low="grey90", high="blue") +
ggsave("Azizi_Macrophage_Supercluster_diffMarkers.jpg", plot = last_plot(), dpi=400, height= 10, width = 10, units = "in")
#"TGFB1","ITGAM","ITGAX","KYNU"

plot_cell_trajectory(DCs, 
                     markers = c("HLA.DPB1","TGFB1","ITGAM","ITGAX", "PDCD1","HCK"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") + scale_color_gradient(low="grey90", high="blue")
plot_cell_trajectory(DCs, 
                     markers = c("HCK"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") + scale_color_gradient(low="grey90", high="blue")

plot_cell_trajectory(regDCm, 
                     markers = c("APOE","ACP5","PLD3"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") + scale_color_gradient(low="grey90", high="blue")

plot_cell_trajectory(regDCm, 
                     markers = c("APOE","SPP1","FN1","CD9","IL10","CD59","C1QB","CHIT1","PDCD1","MARCO","CD276","CD84","APOE","NRP2","CD59","C2","CCL18"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") + scale_color_gradient(low="grey90", high="blue")

ggsave("Azizi_Myeloid_HCK.jpg", plot = last_plot(), dpi=400, height= 6, width = 8, units = "in")
ggsave("Azizi_myeloid_Supercluster_MHCII.jpg", plot = last_plot(), dpi=400, height= 12, width = 15, units = "in")
```
Find the HLA genes

```{r}
greprownames(myeloid@data)
grep(pattern="^HLA", x= rownames(x=myeloid@data), value=TRUE)
```

Differential expression along the lineage

```{r}
pr_graph_test <- principalGraphTest(regDCm, k=3, cores=10, interactive=TRUE)

DClineage <- dplyr::add_rownames(pr_graph_test) %>%
    dplyr::arrange(plyr::desc(morans_test_statistic), plyr::desc(-qval)) %>% head(200)
write.table(DClineage, "Pseudotime_lineage_DE.txt",sep="\t", quote=FALSE)

```

Refine the lineage further to just regDCs
This eliminates branching points and makes a cleaner Moral statistic
Go Back to Monocle Pseudotime analysis and come back to this point: do a musical CODA
```{r}
regDC <- SubsetData(DCsubgroup, ident.use=c("1","3","4","7","8","6"), do.clean=TRUE, subset.raw=TRUE)
```

Count cells expressing specific markers

```{r}
cd276 <- DCsubgroup@raw.data["CD276",] > 0 # 
length(cd276[cd276 == TRUE]) #908

cd276 <- SubsetByPredicate(object = DCsubgroup,
                      vars.use = c("CD276"),
                      predicate = "CD276 > 0")
table(cd276@ident)


table(cd276@meta.data$donor)
table(DCsubgroup@meta.data$donor)
"SLAMF8","CD59","S100A9","LYZ","FCGR3A","LST1","DAB2","IL3RA","CD276"
multiDC <- SubsetByPredicate(object = DCsubgroup,
                      vars.use = c("SLAMF8","CD59","S100A9","LYZ","FCGR3A","LST1","DAB2","IL3RA","CD276"),
                      predicate = "SLAMF8 > 0 &
                                    CD59 > 0 &
                                    S100A9 > 0 &
                                    LYZ > 0 &
                                    FCGR3A > 0 &
                                    LST1 > 0 &
                                    DAB2 > 0 &
                                    IL3RA > 0 &
                                    CD276 > 0")
table(DCsubgroup@ident)
table(multiDC@ident)

multiDC2 <- SubsetByPredicate(object = DCsubgroup,
                      vars.use = c("CD59","S100A9","FCGR3A","DAB2","CD276"),
                      predicate = " CD59 > 2 &
                                    S100A9 > 2 &
                                    FCGR3A > 2 &
                                    DAB2 > 2 &
                                    CD276 > 2")

Both <- SubsetByPredicate(object = DCsubgroup,
                      vars.use = c("ITGAX","ITGAM"),
                      predicate = " ITGAM > 0 &
                                    ITGAX > 0")
macrophage <- SubsetByPredicate(object = DCsubgroup,
                      vars.use = c("ITGAX","ITGAM"),
                      predicate = " ITGAM > 0 &
                                    ITGAX == 0")

dcs <- SubsetByPredicate(object = DCsubgroup,
                      vars.use = c("ITGAX","ITGAM"),
                      predicate = " ITGAM == 0 &
                                    ITGAX > 0")
table(DCsubgroup@ident)
table(macrophage@ident)
table(dcs@ident)
table(Both@ident)
```

###########
################
#############
########
Perform correlation analysis with regDC markers
CD276, SPP1, CHIT, CD9, PDCD1
###########
################
#############
########
For rcorr, the variable have to be the columns, not the rows, so the matric has to be transposed
rcorr(x,y,type=c("pearson"))
specify "y" if you want to just run one gene, y will be concatenated
in trying to rcorr with just CD276, the tool actually ran all vs. all correlations, and gave
$r, $n, $P in separate slots
```{r}
library(Hmisc)
install.packages("corrplot")
library(corrplot)
DC_counts <- data.frame(t(as.matrix(DCsubgroup@raw.data)))
DC_counts <- as.matrix(DC_counts)
#cd276_corr <- rcorr(DC_counts,CD276,type="pearson")
cd276_corr <- rcorr(DC_counts,type="pearson")
dim(cd276_corr$r)
str(cd276_corr)
saveRDS(cd276_corr,file ="Myeloid_gene_corr.rds")
write.table(cd276_corr$r, file="DC_correlations.txt", sep="\t", quote=FALSE)
reg_markers <- c("CD276", "SPP1", "CHIT1","CD9","PDCD1","MARCO","IL10","APOE","NRP2","CD59","C2","CCL18")
```

```{r fig.height=30, fig.width=10}
reg_corr <- data.frame(cd276_corr$r) %>% dselect(reg_markers) %>% dfilter(CD276 != "NaN")
reg_corr <- tibble::rownames_to_column(reg_corr)
reg_corr %<>% arrange(-CD276)
rownames_dc <- reg_corr$rowname
colnames_dc <- colnames(reg_corr)
reg_corr %<>% dselect(-rowname)
reg_corr <- matrix(as.numeric(unlist(reg_corr)),nrow=nrow(reg_corr))
rownames(reg_corr) <- rownames_dc 
colnames(reg_corr) <- colnames_dc[2:length(colnames_dc)] 
write.table(reg_corr, file="regDC_correlations.txt", sep="\t", quote=FALSE)
rownames_dc <- reg_corr$rowname
DC_reg_corr <- as.matrix(reg_corr[,2:8])
rownames(DC_reg_corr) <- rownames_dc
corrplot(reg_corr[1:200,], method="color")
dev.copy(jpeg,"Azizi_reg_corr_heatmap.jpg")

heatmap(reg_corr[1:50,])
ggplot(reg_corr, aes()) +
  geom_tile(aes(fill = rescale), colour = "white", scale_fill_gradient(low = "white", high = "steelblue"))
```

