---
title: "Azizi T cell Monocle"
author: "Mengshu"
date: "1/7/2019"
output: html_document
---
I want to determine the various activation states and naive-memory-exhaustion states of T cells
Monocle has a pseudotime analysis tool that I will use for this
Dataset: Azizi T cell dataset made from the 10X 5' Seq TCR dataset: 24K cells

```{r}
library(Seurat)
library(monocle)
library(dplyr)
library(magrittr)
library(ggplot2)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
```
INSTALLATION
Determine which version of Monocle is downloaded: monocle_2.10.0
Update version
```{r}
sessionInfo() #monocle_2.99.2   This must be what they named beta 3
install.packages("devtools")
#Install DDRTree (simple-ppt-like branch) from our GitHub repo:
devtools::install_github("cole-trapnell-lab/DDRTree", ref="simple-ppt-like")

#Install the latest version of L1-graph from our GitHub repo:
devtools::install_github("cole-trapnell-lab/L1-graph")

#Dependencies
install.packages("reticulate") #this seems to be an R version of python, thus the name?
library(reticulate)
py_install('umap-learn', pip = T, pip_ignore_installed = T) # Ensure the latest version of UMAP is installed
py_install("louvain")

#install xquartz from xquartz.org
#Then pull the monocle3_alpha branch of the Monocle GitHub repo:
devtools::install_github("cole-trapnell-lab/monocle-release", ref="monocle3_alpha", force = TRUE)
install.packages("rgl")
library(monocle)
#Install DelayedArray, because estimateDispersions(azizi_t) throws and error
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("DelayedArray", version = "3.8")
```
-1. Load data though Seurat -> Monocle
Monocle is able to import Seurat datasets with the importCDS() function
importCDS(dataset, import_all= TRUE)

Apparently Imputed data won't work for this, because the data type is inferred, and the logic check is data == floor(data), so the numbers have to be integers.
Load raw (unimputed) data into Seurat, because it's easier to use, then import into Monocle. 
```{r}
#BRCA_singlet <- readRDS("BRCA_singlet_Azizi.rds")
expression_matrix <- read.table("GSE114724_rna_raw.tsv", skip=1, header=TRUE, row.names=1, sep="\t",colClasses=c("character",rep("numeric",28013)))

brca <- readRDS("BRCA_singlet_Azizi.rds")

brca_tcell <- CreateSeuratObject(raw.data = expression_matrix)
#Tranasfer over the  cluster identities
brca_tcell@ident <- BRCA_singlet@ident

#Add the cell identities to a meta.data column
brca_tcell@meta.data$cell_ID <- brca_tcell@ident

# Check the labels make sense. They do. it all matches
VlnPlot(object=brca_tcell, features.plot = c("CD8A","FOXP3","GZMK"), use.raw = TRUE, x.lab.rot = TRUE,size.x.use = 10, point.size.use = 0.1 ) 

# Transfer over more metadata information: variable genes
brca_tcell@var.genes <- BRCA_singlet@var.genes

saveRDS(brca_tcell,file="Azizi_t_cell_raw_Seurat.rds")
brca_tcell <- readRDS("Azizi_t_cell_raw_Seurat.rds")
```

The Cell_ID is available only for a subset of all cells, namely 27792 of 28013 cells
I need to export all cell-barcodes, and all cell_ids, and merge together, filling in the missing cols with NA, then put it back into the Monocle object
```{r}
IDS <- as.data.frame(colnames(brca@raw.data))
colnames(IDS) <- "barcode"
IDS %<>% rename(barcode='colnames(brca@raw.data)')

c_type <- as.data.frame(brca@ident)
barcodes <- rownames(c_type)
c_type %<>% mutate(barcode=barcodes)

head(c_type)

merge_ids <- left_join(IDS,c_type, by="barcode")
#dim(merge_ids)
#head(merge_ids)
rownames(merge_ids) <- merge_ids$barcode
#rownames(merge_ids)
merge_ids %<>% dselect(-barcode)
pData(azizi_t)$cluster <- as.vector(merge_ids[,1])
```

Add percent.ribo to cell information
```{r}
rp <- as.data.frame(brca@meta.data$percent.ribo)
rp %<>% mutate(barcode= rownames(brca@meta.data))

RPS10 <- as.data.frame(brca@meta.data$RPS10)
RPS10 %<>% mutate(barcode= rownames(brca@meta.data))

RPL17 <- as.data.frame(brca@meta.data$RPL17)
RPL17 %<>% mutate(barcode= rownames(brca@meta.data))

merge_rp <-left_join(IDS,rp, by="barcode")
merge_rps10 <-left_join(IDS,RPS10, by="barcode") #('colnames(brca@raw.data)','brca@meta.data$percent.ribo')
merge_rpl17 <-left_join(IDS,RPL17, by="barcode")
dim(merge_rp) #28013 2
rownames(merge_rp) <- merge_rp[,1]
merge_rp %<>% select(-barcode)
pData(azizi_t)$percent_ribo <- merge_rp
pData(azizi_t)$RPS10 <- merge_rps10 
pData(azizi_t)$RPL17 <- merge_rpl17
```

```{r}
azizi_t <- importCDS(brca_tcell, import_all=TRUE)
save(azizi_t, file='azizi_t_Monocle.RData')
load("azizi_t_Monocle.RData")
```

0. Label cells with predetermined groups and assign colours
levels(BRCA_singlet@ident)
 "0. Naive CD4 T"           "1. Effector GZMK CD8 T"   "2. Memory CD4 T"         
 "3. Effector GNLY CD8 T"   "4. CD4 T reg"             "5. Helper CXCL13 CD4 T"  
 "6. Effector CD8 T"        "7. Resident memory CD8 T" "8. CD4 T" 
```{r}
options(warn=-1)

#these are two slots for storing cell metadata
length(pData(azizi_t)$cluster)
length(brca_tcell@ident)
pData(azizi_t)$cluster <- merge_ids
pData(azizi_t)$cell_type2 <- as.vector.factor(merge_ids)

cell_type_colour <- c("0. Naive CD4 T" = "#f6ff60", 
                     "1. Effector GZMK CD8 T" ="#00fbdb", 
                     "2. Memory CD4 T" = "#f9d62e", 
                     "3. Effector GNLY CD8 T" ="#198908", 
                     "4. CD4 T reg" ="#ff4e50",
                     "5. Helper CXCL13 CD4 T" ="#fc913a", 
                     "6. Effector CD8 T" = "#231aff",
                     "7. Resident memory CD8 T" = "#131313",
                     "8. CD4 T" ="#a300ff",
                     "NA" = "#808080")
```

1. Estimate Size Factors and Dispersions
Needed when working with negbinomail.size() or negbinomial()expression family, which is for UMI data
```{r}
azizi_t <- estimateSizeFactors(azizi_t)
azizi_t <- estimateDispersions(azizi_t)
#glm.fit: algorithm did not convergeRemoving 16 outliers
#Error in .local(object, ...) : Error: estimateDispersions only works, and is only needed, when you're using a CellDataSet with a negbinomial or negbinomial.size expression family
#This suggests that the data in not in the right expression family format
#Using un-imputed data fixed the above error
```
Preprocess data to find pricipal components
```{r}
azizi_t <- preprocessCDS(azizi_t, num_dim=20)
```
2. UMAP dimension reduction
2.5 Regress out cell cycle markers
#Added percent.ribo as a metadata col and using it as a residualModelFormulaStr to regress out ribosome content
```{r}
azizi_t <- reduceDimension(azizi_t, reduction_method = 'UMAP', residualModelFormulaStr ="~percent_ribo + RPL17 + RPS10 + cellcycle_genes")

cellcycle_genes <- c('RRM2','TYMS','BIRC5',"UHRF1","CDC20","DLGAP5","UBE2C","PKMYT1","CDT1","CCNB2","SPC25","HJURP","CDC45","CDCA5")
# use_pca = TRUE,
```

```{r}
azizi_t <- clusterCells(azizi_t,method = 'louvain',
                        res = 1e-6,
                        louvain_iter = 1,
                        verbose = T)
plot_cell_clusters(azizi_t, color_by = "cluster")
```


3. Partition cells into Supergroups
```{r}
azizi_t <- partitionCells(azizi_t,residualModelFormulaStr ="~percent_ribo + RPL17 + RPS10 + cellcycle_genes")
```
4. Learn the Pricipal graph
SimplePPT and DDRTree are not giving different results
```{r}
azizi_t <- learnGraph(azizi_t, RGE_method ='SimplePPT',residualModelFormulaStr ="~percent_ribo + RPL17 + RPS10 + cellcycle_genes") # partition_component +T
azizi_t <- learnGraph(azizi_t, RGE_method ='DDRTree')
```
5. Visualize trajectory
```{r fig.height=6, fig.width=10}
plot_cell_trajectory(azizi_t, 
                     color_by = "cluster",
                     show_backbone=TRUE, show_branch_points = TRUE, alpha = 0.9)  +
                     scale_color_manual(values = cell_type_colour) + 
                    theme(legend.position = "bottom")  +
                    guides(colour=guide_legend(nrow=3))
ggsave("Azizi_t_Supercluster_path_cluster.jpg", plot = last_plot(), dpi=400, height= 6, width = 10, units = "in")

plot_cell_trajectory(azizi_t, color_by = "louvain_component",show_backbone = TRUE, backbone_color = "black") scale_color_manual(values = cell_type_colour)
ggsave("Azizi_t_Supercluster_louvain_component.jpg", plot = last_plot(), dpi=400, height= 6, width = 10, units = "in")

plot_cell_clusters(azizi_t, color_by = pData(azizi_t)$cell_type2) +
                     scale_color_manual(values = cell_type_colour)
```
Specify the Root of the large cluster
```{r}
# a helper function to identify the root principal points:
get_correct_root_state <- function(cds, cell_phenotype, root_type){
  cell_ids <- which(pData(cds)[, cell_phenotype] == root_type)

  closest_vertex <-
    cds@auxOrderingData[[cds@rge_method]]$pr_graph_cell_proj_closest_vertex
  closest_vertex <- as.matrix(closest_vertex[colnames(cds), ])
  root_pr_nodes <-
    V(cds@minSpanningTree)$name[as.numeric(names
      (which.max(table(closest_vertex[cell_ids,]))))]

  root_pr_nodes
}

azizi_node_ids = get_correct_root_state(azizi_t,
                                      cell_phenotype =
                                        'cluster', "0. Naive CD4 T")
azizi_t<- orderCells(azizi_t, root_pr_nodes = azizi_node_ids)
plot_cell_trajectory(azizi_t)
ggsave("Azizi_t_Supercluster_pseudotime.jpg", plot = last_plot(), dpi=400, height= 6, width = 10, units = "in")
```
DIFFERENTIAL ANALYSIS

```{r}
pr_graph_test <- principalGraphTest(azizi_t, k=3, cores=6)
my_list <- dplyr::add_rownames(pr_graph_test) %>%
    dplyr::arrange(plyr::desc(morans_test_statistic), plyr::desc(-qval)) %>% head(8000)
grep("MTRNR*", my_list)
write.table(trajectory_genes, file="Azizi_t_trajectory_genes_50.txt", sep="\t", quote=FALSE)
genes
nrow(subset(pr_graph_test, qval < 0.01))
# 6517 genes are significant
```
```{r fig.height=10, fig.width=12}
#T RM cycling cluster
plot_cell_trajectory(azizi_t, 
                     markers = c('RRM2','TYMS','BIRC5',"UHRF1","CDC20","DLGAP5","UBE2C","PKMYT1","CDT1","CCNB2","SPC25","HJURP","CDC45","CDCA5"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") + scale_color_gradient(low="grey90", high="blue")
#test
plot_cell_trajectory(azizi_t, 
                     markers = c("MTRNR2L12","MTRNR2L8"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") +scale_color_gradient(low="grey90", high="blue")
#main
plot_cell_trajectory(azizi_t, 
                     markers = c('NKG7','CXCL13','CCL5',"CD8A","TRDV2","FOXP3","GZMK","CCL4","GZMA","CST7","CD8B","GZMB","KRT86","HLA-DRB1","IFIT1"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") +scale_color_gradient(low="grey90", high="blue")
Global separator
plot_cell_trajectory(azizi_t, 
                     markers = c("RPS10","RPL41","EEF1G","MTRNR2L12","RPL21"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") + scale_color_gradient(low="grey90", high="blue")
ggsave("Azizi_t_Monocle_RRM2.jpg", plot = last_plot(), dpi=400, height= 6, width = 10, units = "in")

Humanin
plot_cell_trajectory(azizi_t, 
                     markers = c("MTRNR2L12","RPS10"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") +scale_color_gradient(low="grey90", high="blue")
ggsave("Azizi_t_Monocle_Humanin_RPS10.jpg", plot = last_plot(), dpi=400, height= 6, width = 10, units = "in")

# CCR4
plot_cell_trajectory(azizi_t, 
                     markers = c("CCR4","GATA3","CXCR3","FOXP3","SELL"), use_color_gradient = TRUE,
                     show_backbone=TRUE, alpha = 0.9)  +
                    theme(legend.position = "bottom") + scale_color_gradient(low="grey90", high="blue")

```
Visualizing genes by pseudotemporal expression pattern
```{r}
diff_test_res <- differentialGeneTest(HSMM_myo[marker_genes,],
              fullModelFormulaStr = "~sm.ns(Pseudotime)")
sig_gene_names <- row.names(subset(diff_test_res, qval < 0.1))
plot_pseudotime_heatmap(HSMM_myo[sig_gene_names,],
                num_clusters = 3,
                cores = 1,
                show_rownames = T)
```

