---
title: "Talmadge_Tcell_Trajectory.Rmd"
author: "MX"
date: "8/27/2020"
output: html_document
---


```{r}
library(Seurat)
library(data.table)
library(magrittr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(stringr)
#library(limma) #quantileNormalization
library(tibble)
#library("DESeq2")
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
source("D:/FLX/Files_from_Gene/R_functions.r")
install.packages("hdf5r")
devtools::install_github("hhoeflin/hdf5r")
```

Load Batch 4 Seurat object and subset to just T cells
```{r}
load("/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/Cellranger3.1.0_batch4/batch4.Robj")
batch4_Tcell <- subset(batch4, idents=c(0,1,26,4,18,19))
save(batch4_Tcell, file="batch4_Tcell.Robj") 
```


#######################
#######################

Import into Monocle2
monocle_2.16.0
This was a disaster previously, but
Try again on the Mac trash can, Aug 17, 2020
Monocle 2 has worked out before, in 2018
#######################
#######################
 
```{r}
library(monocle)
sessionInfo() # monocle 2.16.0
#load data from line 101, batch4_myeloid

batch4_Tcell[["ident"]] <- batch4_Tcell@active.ident 
counts.matrix <- as(as.matrix(batch4_Tcell@assays$RNA@counts), 'sparseMatrix') # raw counts data
pheno <- as(batch4_Tcell@meta.data,"AnnotatedDataFrame") # phenoData where rows are cells, and columns are attributes
gene_meta <- as(data.frame(gene_short_name = row.names(counts.matrix), row.names = row.names(counts.matrix)), "AnnotatedDataFrame")
#fd <- new('AnnotatedDataFrame', data = fData)


#Construct monocle cds
batch4_Tcell_monocle2 <- newCellDataSet(counts.matrix,
                         phenoData = pheno,
                         featureData = gene_meta,
                         expressionFamily=negbinomial.size()) 

save(batch4_Tcell_monocle2,file="batch4_Tcell_monocle2.rdata") # current saved version has CITE data as meta data in pData
save(batch4_Tcell_monocle2_CITE,file="batch4_Tcell_monocle2_CITE.rdata")
```
 
```{r}
#Add in CITE data

CITE_Tcell <- FetchData(batch4_Tcell, vars= c("CD4-CITE","CCR4-CITE","CCR6-CITE","CD11b-CITE","CD11c-CITE","CD14-CITE","CD15-CITE","CD33-CITE" ,"CD34-CITE","CD45RA-CITE", "CD45RO-CITE", "CSF-1R-CITE","CXCR3-CITE","LAG-3-CITE","PD-1-CITE","PD-L1-CITE"), cells = T_monocle) # Get CITE markers from the Seurat object, for the cells in the Monocle T cell object

colnames(CITE_Tcell) <- c("CD4_CITE","CCR4_CITE","CCR6_CITE","CD11b_CITE","CD11c_CITE","CD14_CITE","CD15_CITE","CD33_CITE" ,"CD34_CITE","CD45RA_CITE", "CD45RO_CITE", "CSF_1R_CITE","CXCR3_CITE","LAG_3_CITE","PD_1_CITE","PD_L1_CITE")

CITE_t <- as(as.matrix(CITE_Tcell %>% t), 'sparseMatrix') 

counts.matrix.CITE <- rbind(exprs(batch4_Tcell_monocle2),CITE_t)

matrix.CITE <- rbind(exprs(batch4_Tcell_monocle2),CITE_t) # merge the existing raw data with CITE data
matrix.CITE.dense <- as.matrix(matrix.CITE)

batch4_Tcell_monocle2_CITE
str(matrix.CITE.dense)
batch4_Tcell_monocle2@assayData <- matrix.CITE.dense # put the merged data back into the assayData slot

gene_meta <- as(data.frame(gene_short_name = row.names(matrix.CITE), row.names = row.names(counts.matrix)), "AnnotatedDataFrame")
pheno <- as(pData(batch4_Tcell_monocle2), "AnnotatedDataFrame")
#Construct monocle cds with CITE data
batch4_Tcell_monocle2_CITE <- newCellDataSet(matrix.CITE,
                         phenoData = pheno,
                         featureData = gene_meta,
                         expressionFamily=negbinomial.size()) 

save(batch4_Tcell_monocle2,file="batch4_Tcell_monocle2.rdata") # current saved version has CITE data as meta data in pData
save(batch4_Tcell_monocle2_CITE,file="batch4_Tcell_monocle2_CITE.rdata")

grep("CITE",row.names(batch4_Tcell_monocle2_CITE), value=TRUE)
```

Pre-Process
Process
```{r}
batch4_Tcell_monocle2 <- estimateSizeFactors(batch4_Tcell_monocle2)
batch4_Tcell_monocle2 <- estimateDispersions(batch4_Tcell_monocle2)

batch4_Tcell_monocle2_CITE <- estimateSizeFactors(batch4_Tcell_monocle2_CITE)
batch4_Tcell_monocle2_CITE <- estimateDispersions(batch4_Tcell_monocle2_CITE)

disp_table <- dispersionTable(batch4_Tcell_monocle2_CITE)
unsup_clustering_genes <- subset(disp_table, mean_expression >= 0.01)
length(unsup_clustering_genes$gene_id) #7568 genes # 8779 cells #10902 CITE
batch4_Tcell_monocle2_CITE  <- setOrderingFilter(batch4_Tcell_monocle2_CITE , unsup_clustering_genes$gene_id)
plot_ordering_genes(batch4_Tcell_monocle2_CITE)
```
This is the PCA analysis step, can take quite a while
```{r}
# HSMM@auxClusteringData[["tSNE"]]$variance_explained <- NULL
plot_pc_variance_explained(batch4_Tcell_monocle2, return_all = F,max_components = 15) # norm_method='log'

plot_pc_variance_explained(batch4_Tcell_monocle2_CITE, return_all = F,max_components = 9) # norm_method='log'
```
reduceDimension(cds, max_components = 2, reduction_method = c("DDRTree",
  "ICA", "tSNE", "SimplePPT", "L1-graph", "SGL-tree"), norm_method = c("log",
  "vstExprs", "none"), residualModelFormulaStr = NULL, pseudo_expr = 1,
  relative_expr = TRUE, auto_param_selection = TRUE, verbose = FALSE,
  scaling = TRUE, ...)

```{r fig.height=8, fig.width=12}
batch4_Tcell_monocle2  <- reduceDimension(batch4_Tcell_monocle2 , max_components = 2, num_dim = 9,
                reduction_method = 'tSNE', verbose = T,residualModelFormulaStr = "~orig.ident")
batch4_Tcell_monocle2  <- clusterCells(batch4_Tcell_monocle2 , num_clusters = 10) #Distance cutoff calculated to 5.928387 


batch4_Tcell_monocle2_CITE  <- reduceDimension(batch4_Tcell_monocle2_CITE , max_components = 2, num_dim = 9,
                reduction_method = 'tSNE', verbose = T, residualModelFormulaStr = "~orig.ident")
batch4_Tcell_monocle2_CITE  <- clusterCells(batch4_Tcell_monocle2_CITE , num_clusters = 10) #Distance cutoff calculated to 5.928387 

plot_cell_clusters(batch4_Tcell_monocle2_CITE , 1, 2, color = "CellType",
    markers = c("CD4","CD8A","FOXP3","IFNG","CXCR3","IL17","LAG3","TOX","EOMES","CCR7","SELL","TNF","TGFB","HAVCR2","CD274","CD69","GZMA","KLRG1","PRF1","PDCD1","CCL5","TCF7","BAX","CCNA2","CCNB1","MKI67","NR4A1","IL2","IL4","IL6","IL25","CCR4"),cell_size = 0.5)
```


Clusters

```{r fig.height=5, fig.width=6}
plot_cell_clusters(batch4_Tcell_monocle2_CITE, color_by = 'as.factor(Cluster)',cell_size=0.3)
```


Clusters by orig.ident
```{r fig.height=7, fig.width=8}

plot_cell_clusters(batch4_Tcell_monocle2_CITE, color_by = 'as.factor(Cluster)', cell_size=0.3) + facet_wrap(~orig.ident)
```

 I see cells cluster by the original donor: regress out this contribution in the next block of code
```{r fig.height=6, fig.width=8}
plot_cell_clusters(batch4_Tcell_monocle2_CITE, 1, 2, color = "orig.ident",cell_size = 0.5)
plot_cell_clusters(batch4_Tcell_monocle2_CITE, 1, 2, color = "nFeature_RNA",cell_size = 0.5)
plot_cell_clusters(batch4_Tcell_monocle2_CITE, 1, 2, color = "seurat_clusters",cell_size = 0.5) # the seurat clusters are better

```

CITE markers

```{r fig.height=8, fig.width=12}
plot_cell_clusters(batch4_Tcell_monocle2_CITE, 1, 2, color = "CellType",
    markers = c("CD4_CITE","CCR4_CITE","CCR6_CITE","CD11b_CITE","CD11c_CITE","CD14_CITE","CD15_CITE","CD33_CITE" ,"CD34_CITE","CD45RA_CITE", "CD45RO_CITE", "CSF_1R_CITE","CXCR3_CITE","LAG_3_CITE","PD_1_CITE","PD_L1_CITE"),cell_size = 0.5)


grep("PD_L1_CITE",rownames(batch4_Tcell_monocle2_CITE), value=TRUE)
```


Filter out high MT cells, which form a portion of AML3 CD8 cells that cluster with themselves
```{r}
batch4_Tcell_monocle2 <- batch4_Tcell_monocle2[,pData(batch4_Tcell_monocle2)$percent.mt < 15]
```


Monocle 2 Trajectory


1. Try trajectory along clusters
Select genes with min expression of 0.1, and expressed in > 5% of cells
```{r}
batch4_Tcell_monocle2_CITE <- detectGenes(batch4_Tcell_monocle2_CITE, min_expr = 0.1)
fData(batch4_Tcell_monocle2_CITE)$use_for_ordering <-
    fData(batch4_Tcell_monocle2_CITE)$num_cells_expressed > 0.05 * ncol(batch4_Tcell_monocle2_CITE)
```
PCA
```{r}
plot_pc_variance_explained(batch4_Tcell_monocle2_CITE, return_all = F,max_components = 9)
```
tSNE Add the ,residualModelFormulaStr = "~orig.ident" term here too to regress of the second arm of CD8 T cells
```{r}
batch4_Tcell_monocle2 <- reduceDimension(batch4_Tcell_monocle2,
                              max_components = 2,
                              norm_method = 'log',
                              num_dim = 12,
                              reduction_method = 'tSNE',
                              verbose = T,
                              residualModelFormulaStr = "~orig.ident")

batch4_Tcell_monocle2_CITE <- reduceDimension(batch4_Tcell_monocle2_CITE,
                              max_components = 2,
                              norm_method = 'log',
                              num_dim = 9,
                              reduction_method = 'tSNE',
                              verbose = T,
                              residualModelFormulaStr = "~orig.ident")
```

```{r}
batch4_Tcell_monocle2_CITE <- clusterCells(batch4_Tcell_monocle2_CITE, verbose = F, num_clusters=10)
```

```{r fig.height=8, fig.width=10}
plot_cell_clusters(batch4_Tcell_monocle2_CITE, color_by = 'as.factor(Cluster)')
```
Check markers
```{r fig.height=8, fig.width=12}
plot_cell_clusters(batch4_Tcell_monocle2_CITE , 1, 2, color = "CellType",
    markers = c("CD4", "CD8A","FOXP3","IFNG","CXCR3","IL17","LAG3","TOX","EOMES","CCR7","SELL","TNF","TGFB","HAVCR2","CD274","CD69","GZMA","KLRG1","PRF1","PDCD1","CCL5","TCF7","BAX","CCNA2","CCNB1","MKI67","NR4A1","IL2","IL4","IL6","IL25","CCR4"),cell_size = 0.5)
```


Check some parameters
Rho p is the local cell density, while delta is the nearest distance to another cell 
```{r}
plot_rho_delta(batch4_Tcell_monocle2_CITE, rho_threshold = 5, delta_threshold = 5 )

```

Re-run with Delta and Rho_sigma
rho 2
sigma 5
```{r}
batch4_Tcell_monocle2 <- clusterCells(batch4_Tcell_monocle2,
                 rho_threshold = 5,
                 delta_threshold = 5,
                 skip_rho_sigma = T,
                 verbose = F)
```


Check markers
```{r fig.height=8, fig.width=12}
plot_cell_clusters(batch4_Tcell_monocle2_CITE , 1, 2, color = "CellType",
    markers = c("CD4_CITE","CD8A","State"),cell_size = 0.5)
```


DE analysis between clusters
differentialGeneTest takes a while, set cores=1 because cores=6 threw an error
For T cells: taking a long time
This takes a long time!!!
```{r}
Tcell_expressed_genes <-  row.names(subset(fData(batch4_Tcell_monocle2_CITE),
num_cells_expressed >= 10))

clustering_DEG_genes <-
    differentialGeneTest(batch4_Tcell_monocle2_CITE[Tcell_expressed_genes,],
          fullModelFormulaStr = '~Cluster',
          cores = 1)

save(clustering_DEG_genes, file="clustering_Tcell_DEG_genes.rdata")

clustering_DEG_genes %>% arrange(qval) %>% dfilter(qval >1e-45)
```


Set the top 2000 significant genes as the ordering genes and run DDRTree


```{r}
Tcell_ordering_genes <-
    row.names(clustering_DEG_genes)[order(clustering_DEG_genes$qval)][1:2000]

batch4_Tcell_monocle2_CITE <-
    setOrderingFilter(batch4_Tcell_monocle2_CITE,
        ordering_genes = Tcell_ordering_genes)

batch4_Tcell_monocle2_CITE <-
    reduceDimension(batch4_Tcell_monocle2_CITE, method = 'DDRTree')

batch4_Tcell_monocle2_CITE <-
    orderCells(batch4_Tcell_monocle2_CITE)

#batch4_Tcell_monocle2 <-
#    orderCells(batch4_Tcell_monocle2, root_state = GM_state(batch4_Tcell_monocle2))
```


```{r fig.height=6, fig.width=6}
plot_cell_trajectory(batch4_Tcell_monocle2_CITE, color_by = "State",cell_size = 1)
```


Plot scatter
```{r fig.height=8, fig.width=9}
library(viridis)
markers = c("CD4", "CD8A","FOXP3","IFNG","CXCR3","IL17","LAG3","TOX","EOMES","CCR7","SELL","TNF","TGFB","HAVCR2","CD274","CD69","GZMA","KLRG1","PRF1","PDCD1","CCL5","TCF7","BAX","CCNA2","CCNB1","MKI67","NR4A1","IL2","IL4","IL6","IL25")

plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, markers = c("CD8A","CD4","CCL5","CCL6","CCL12","GZMA","CCR7","KLRG1","PRF1","SELL","IFNG","TOX","FOXP3","TCF7"),use_color_gradient = TRUE, cell_size=0.5,markers_linear= FALSE, show_branch_points = FALSE) + scale_color_viridis(option="B",begin=0, end=1, direction= 1, limit = c(0,0.8),oob=squish) + theme(panel.background = element_rect(fill = "grey90"))
```
Plot by State

```{r fig.height=5, fig.width=5}
plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, colour_by= "State", cell_size=0.5)
batch4_Tcell_monocle2_CITE$State %>% table
```


Put CITE markers into the Object as Metadata
```{r}
T_monocle <- colnames(batch4_Tcell_monocle2) # get the cells in the T cell object

CITE_Tcell <- FetchData(batch4_Tcell, vars= c("CD4-CITE","CCR4-CITE","CCR6-CITE","CD11b-CITE","CD11c-CITE","CD14-CITE","CD15-CITE","CD33-CITE" ,"CD34-CITE","CD45RA-CITE", "CD45RO-CITE", "CSF-1R-CITE","CXCR3-CITE","LAG-3-CITE","PD-1-CITE","PD-L1-CITE"), cells = T_monocle) # Get CITE markers from the Seurat object, for the cells in the Monocle T cell object

colnames(CITE_Tcell) <- c("CD4_CITE","CCR4_CITE","CCR6_CITE","CD11b_CITE","CD11c_CITE","CD14_CITE","CD15_CITE","CD33_CITE" ,"CD34_CITE","CD45RA_CITE", "CD45RO_CITE", "CSF_1R_CITE","CXCR3_CITE","LAG_3_CITE","PD_1_CITE","PD_L1_CITE")

CITE_Tcell


pData(batch4_Tcell_monocle2) <- cbind(pData(batch4_Tcell_monocle2)[,1:30] , CITE_Tcell)
pData(batch4_Tcell_monocle2)$CD4_CITE
```

Plot CITE markers
```{r fig.height=8, fig.width=8}
plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, markers = c("CD4_CITE","CCR4_CITE","CCR6_CITE","CD11b_CITE","CD11c_CITE","CD14_CITE","CD15_CITE","CD33_CITE" ,"CD34_CITE","CD45RA_CITE", "CD45RO_CITE", "CSF_1R_CITE","CXCR3_CITE","LAG_3_CITE","PD_1_CITE","PD_L1_CITE"),use_color_gradient = TRUE, cell_size=0.5,markers_linear= FALSE, show_branch_points = FALSE) + scale_color_viridis(option="B",begin=0, end=1, direction= 1, limit = c(0,0.3),oob=squish) + theme(panel.background = element_rect(fill = "grey90"))

plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, markers = c("CD4_CITE","CCR4_CITE","CCR6_CITE","CD15_CITE","CD33_CITE" ,"CD45RA_CITE", "CD45RO_CITE","CXCR3_CITE","PD_1_CITE","PD_L1_CITE"),use_color_gradient = TRUE, cell_size=0.5,markers_linear= FALSE, show_branch_points = FALSE) + scale_color_viridis(option="B",begin=0, end=1, direction= 1, limit = c(0,0.8),oob=squish) + theme(panel.background = element_rect(fill = "grey90"))

plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, markers = c("CD14_CITE","CD34_CITE","LAG_3_CITE","CD11b_CITE","CD11c_CITE"),use_color_gradient = TRUE, cell_size=0.5,markers_linear= FALSE, show_branch_points = FALSE) + scale_color_viridis(option="B",begin=0, end=1, direction= 1, limit = c(0,0.8),oob=squish) + theme(panel.background = element_rect(fill = "grey90"))
```


nFeature_RNA
```{r fig.height=10, fig.width=10}
plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, color = "nFeature_RNA",use_color_gradient =FALSE, cell_size=0.75) + scale_color_continuous(type="viridis", limit = c(0,3000),oob=squish) 

plot_cell_trajectory(batch4_Tcell_monocle2, 1, 2, color = "percent.ribo",use_color_gradient =FALSE, cell_size=0.75) + scale_color_continuous(type="viridis", limit = c(0,50),oob=squish) 

plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, color = "percent.mt",use_color_gradient =FALSE, cell_size=0.75) + scale_color_continuous(type="viridis", limit = c(0,50),oob=squish) 

plot_cell_trajectory(batch4_Tcell_monocle2, 1, 2, color = "RNA.Per.Feature",use_color_gradient =FALSE, cell_size=0.75) + scale_color_continuous(type="viridis", limit = c(0,5),oob=squish) 

plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, color = "nCount_RNA",use_color_gradient =FALSE, cell_size=0.75) + scale_color_continuous(type="viridis", limit = c(0,14000),oob=squish) 

plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, markers = c("nCount_RNA","CD8A","CD4","CCL5","GZMA","CCR7","KLRG1","PRF1","SELL","IFNG","TOX","LAG3","FOXP3","CCR4","PDCD1","CD274","TCF7"),use_color_gradient = TRUE, cell_size=0.2,markers_linear= FALSE, show_branch_points = FALSE) + scale_color_viridis(option="B",begin=0, end=1, direction= 1, limit = c(0,0.8),oob=squish) + theme(panel.background = element_rect(fill = "grey90"))

```

```{r fig.height=8, fig.width=8}
library(scales)
plot_cell_trajectory(batch4_Tcell_monocle2_CITE, 1, 2, color_by = "orig.ident",use_color_gradient =FALSE, cell_size=0.75) + nolegend() +  scale_color_manual(values=c("darkorchid1","darkorchid1","turquoise3","turquoise3","turquoise3","orchid2","orchid2","orchid2")) + facet_wrap(~orig.ident)
```


Semi-supervised Trajectory

```{r}
CD4_id <-
    row.names(subset(fData(batch4_Tcell_monocle2_CITE), gene_short_name == "CD4_CITE")) #  pData(batch4_Tcell_monocle2)$CD4_CITE
CD8_id <-
    row.names(subset(fData(batch4_Tcell_monocle2_CITE), gene_short_name == "CD8A"))

cth <- newCellTypeHierarchy()

cth <- addCellType(cth,
                   "CD4 T cell",
                   classify_func = function(x) { x[CD4_id,] >= 1 })

cth <- addCellType(cth,
                   "CD8 T cell",
                   classify_func = function(x) { x[CD8_id,] >= 1 })

batch4_Tcell_monocle2_CITE <- classifyCells(batch4_Tcell_monocle2_CITE, cth)
```

Select genes that covary with CD4 and CD8

Does not work, there are no markers, everything is not significant
```{r}
marker_diff <- markerDiffTable(batch4_Tcell_monocle2_CITE[Tcell_expressed_genes,],
                       cth,
                       cores = 1)
marker_diff %>% arrange(qval) %>% dfilter(gene_short_name=="CD4_CITE")

semisup_clustering_genes <-
    row.names(marker_diff)[order(marker_diff$qval)][1:1000]
```

Trajectory from CD4 CD8 and covarying genes
```{r}
batch4_Tcell_monocle2_CITE <- setOrderingFilter(batch4_Tcell_monocle2_CITE, semisup_clustering_genes)
#plot_ordering_genes(HSMM_myo)
batch4_Tcell_monocle2_CITE <- reduceDimension(batch4_Tcell_monocle2_CITE, max_components = 2,
    method = 'DDRTree', norm_method = 'log')
batch4_Tcell_monocle2_CITE <- orderCells(batch4_Tcell_monocle2_CITE)
batch4_Tcell_monocle2_CITE <- orderCells(batch4_Tcell_monocle2_CITE, root_state = GM_state(batch4_Tcell_monocle2_CITE))
plot_cell_trajectory(batch4_Tcell_monocle2_CITE, color_by = "CellType") +
    theme(legend.position = "right")
```



Count cells by State by Donor
```{r fig.height=4, fig.width=10}
data.frame(State= batch4_Tcell_monocle2@phenoData$State, source = batch4_Tcell_monocle2@phenoData@data$orig.ident) %>% table %>% data.frame %>% group_by(source) %>% mutate(Group_sum= sum(Freq)) %>% ungroup %>% mutate(T_Cell_percent =Freq/Group_sum) %>% mutate(source=as.character(source)) %>% mutate(Type=substr(source,1,nchar(source)-1)) %>% left_join(key,by="State") %>% left_join(key2,by="source") %>% mutate(Total_cell_percent = Freq/Total_cell) %>% dselect(source, Freq, Identity) %>% dfilter(!is.na(Identity)) %>% ggplot(aes(x=source,y=Freq)) + geom_bar(stat="identity") + facet_wrap(~Identity)

key <- data.frame(State=as.factor(c(1,3)),Identity=c("CD8","CD4"))
key2 <- data.frame(source=c("pbmc1","pbmc2","pbmc3","pdac1","pdac2","pdac3","aml1","aml3"),Total_cell=c(2460,3464,4699,5110,4108,2611,6436,2418))
my_counts <- data.frame(State= batch4_Tcell_monocle2_CITE@phenoData$State, Donor = batch4_Tcell_monocle2_CITE@phenoData@data$orig.ident) %>% mutate(Type=substr(Donor,1,nchar(Donor)-1)) %>% count(Donor)

data.frame(State= batch4_Tcell_monocle2@phenoData$State, Donor = batch4_Tcell_monocle2@phenoData@data$orig.ident) %>% mutate(Type=substr(Donor,1,nchar(Donor)-1)) %>% left_join(my_counts,by="Donor") %>% mutate(T_Cell_percent =Freq/n) %>% left_join(key,by="State") %>% left_join(key2,by="Donor") %>% mutate(Total_cell_percent = Freq/Total_cell) %>% ggplot(aes(x=Type, y=Total_cell_percent, group=Type)) + geom_point() + facet_wrap(~Identity, ncol=5)

data.frame(State= batch4_Tcell_monocle2@phenoData$State, source = batch4_Tcell_monocle2@phenoData@data$orig.ident) %>% table %>% data.frame %>% group_by(source) %>% mutate(Group_sum= sum(Freq)) %>% ungroup %>% mutate(T_Cell_percent =Freq/Group_sum) %>% mutate(source=as.character(source)) %>% mutate(Type=substr(source,1,nchar(source)-1)) %>% left_join(key,by="State") %>% left_join(key2,by="source") %>% mutate(Total_cell_percent = Freq/Total_cell) %>% ggplot(aes(x=Identity, y=Total_cell_percent, group=Type)) + geom_point()+ facet_wrap(~Type,ncol=3)

data.frame(State= batch4_Tcell_monocle2@phenoData$State, source = batch4_Tcell_monocle2@phenoData@data$orig.ident) %>% table %>% data.frame %>% group_by(source) %>% mutate(Group_sum= sum(Freq)) %>% ungroup %>% mutate(T_Cell_percent =Freq/Group_sum) %>% mutate(source=as.character(source)) %>% mutate(Type=substr(source,1,nchar(source)-1)) %>% left_join(key,by="State") %>% left_join(key2,by="source") %>% mutate(Total_cell_percent = Freq/Total_cell) %>% ggplot(aes(x=Type, y=Total_cell_percent, group=Type)) + geom_point()+ facet_wrap(~Identity,ncol=5)

data.frame(State= batch4_Tcell_monocle2@phenoData$State, source = batch4_Tcell_monocle2@phenoData@data$orig.ident) %>% table %>% data.frame %>% group_by(source) %>% mutate(Group_sum= sum(Freq)) %>% ungroup %>% mutate(T_Cell_percent =Freq/Group_sum) %>% mutate(source=as.character(source)) %>% mutate(Type=substr(source,1,nchar(source)-1)) %>% left_join(key,by="State") %>% left_join(key2,by="source") %>% mutate(Total_cell_percent = Freq/Total_cell) %>% ggplot(aes(x=Type, y=T_Cell_percent, group=Type, colour=source)) + geom_point()+ facet_wrap(~Identity,ncol=5)    
```



Plot cells by State

```{r fig.height=6, fig.width=6.5}
plot_cell_trajectory(batch4_Tcell_monocle2, color_by = "State",cell_size = 1)

plot_cell_trajectory(batch4_Tcell_monocle2, color_by = "State",cell_size = 0.5) + facet_wrap(~State, nrow=2)
```
Pseudotime DE
VGAM

DE analysis between States

Took 1 hour
This doesn't record the state of the marker either!!!!
```{r}

Expressed_genes <-  row.names(subset(fData(batch4_Tcell_monocle2_CITE),
num_cells_expressed >= 10))

Tcell_State_DEG_genes <-
    differentialGeneTest(batch4_Tcell_monocle2_CITE[Tcell_expressed_genes,],
          fullModelFormulaStr = '~State',
          cores = 1)

Tcell_State_DEG_genes %<>% tibble::rownames_to_column("Gene")
save(Tcell_State_DEG_genes, file="Tcell_clustering_DEG_genes.rdata")

Tcell_markers <- Tcell_State_DEG_genes %>% dfilter(qval < 1e-100, num_cells_expressed > 500) %>% arrange(qval) %>% pull(Gene) #300+

Tcell_State_DEG_genes %>% dfilter(Gene =="CD8A")


cluster_heatmap <- plot_pseudotime_heatmap(batch4_Tcell_monocle2_CITE[Tcell_markers,],
                num_clusters = 3,
                cores = 1, return_heatmap = TRUE,
                show_rownames = T)
```

BEAM Branched expression amalysis modeling

```{r}

BEAM_res <- BEAM(batch4_Tcell_monocle2_CITE, branch_point = 1, cores = 1)
BEAM_res <- BEAM_res[order(BEAM_res$qval),]
BEAM_res <- BEAM_res[,c("gene_short_name", "pval", "qval")]

plot_genes_branched_heatmap(batch4_Tcell_monocle2_CITE[row.names(subset(BEAM_res,
                                          qval < 1e-16)),],
                                          branch_point = 1,
                                          num_clusters = 4,
                                          cores = 1,
                                          use_gene_short_name = T,
                                          show_rownames = T)
```

Lineage scatter

```{r fig.height=7, fig.width=5}
Tcell_genes <- row.names(subset(fData(batch4_Tcell_monocle2_CITE),
          gene_short_name %in% c("CD8A", "KLRG1", "LAG_3_CITE","TOX","TIGIT","HAVCR2","PD_1_CITE")))

HPK1_genes <- row.names(subset(fData(batch4_Tcell_monocle2_CITE),
          gene_short_name %in% c("MAP4K1", "LCP2", "ZAP70","PTGER2","IL2")))

CD4_genes <- row.names(subset(fData(batch4_Tcell_monocle2_CITE),
          gene_short_name %in% c("CD4_CITE","CCR4_CITE","CCR6_CITE","CXCR3_CITE","CCL22","FOXP3")))

plot_genes_branched_pseudotime(batch4_Tcell_monocle2_CITE[CD4_genes,],
                       branch_point = 1,
                       color_by = "State",
                       ncol = 1)
```


Here is where the cell order and cell state information are stored
```{r}

batch4mye_monocle2@phenoData$Pseudotime
colnames(batch4mye_monocle2)
pseudotime_cell <- data.frame(Cell=colnames(batch4mye_monocle2), Pseudotime=batch4mye_monocle2@phenoData$Pseudotime) %>% arrange(Pseudotime) %>% pull(Cell)
batch4mye_monocle2@phenoData$State

batch4mye_monocle2@featureData@data$gene_short_name
batch4mye_monocle2@assayData$exprs # this is a sparse matrix

batch4mye_monocle2@assayData$exprs[cluster_DEG,pseudotime_cell] %>% data.frame %>% tibble::rownames_to_column("Gene") %>% write.table(file="Pseudotime_myeloid.txt",quote=FALSE,row.names = FALSE, sep="\t")
```