---
title: "RAPT_CD3_Gene_scRNA"
author: "Mengshu Xu"
date: "3/18/2020"
output: html_document
---

Load packages

```{r}
library(Seurat)
library(data.table)
library(magrittr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(tibble)
#source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
source("D:/FLX/Files_from_Gene/R_functions.r")
theme_set(theme_rapt())
theme_set(theme_bw())
```

Load Seurat object
This didn't work: Gene's rdata loaded into a bunch of files but no Seurat object
```{r}
load("D:/FLX/scRNA_data/RAPT_CD3/seurat.rdata")
CD3 <- UpdateSeuratObject()
str(tf)
saveRDS(CD3,file="D:/FLX/scRNA_data/RAPT_CD3/CD3_seurat3.rds")
```


Save and Load files
```{r}
saveRDS(CD3,file="D:/FLX/scRNA_data/RAPT_CD3/CD3_seurat3.rds")
CD3 <- readRDS(file="D:/FLX/scRNA_data/RAPT_CD3/CD3_seurat3.rds") #CD3
```


Load in h5 data files
Create Seurat objects with filter
Merge with tags
```{r}
install.packages("hdf5r")
library(hdf5r)

cntrl_D1_20K <- Read10X_h5("D:/FLX/scRNA_data/RAPT_CD3/control-20k-filtered_gene_bc_matrices_h5.h5") 
cntrl_D1_33K <- Read10X_h5("D:/FLX/scRNA_data/RAPT_CD3/control-33k-filtered_gene_bc_matrices_h5.h5") 
cntrl_D1_46K <- Read10X_h5("D:/FLX/scRNA_data/RAPT_CD3/control-46k-filtered_gene_bc_matrices_h5.h5")
NoTrp_D1 <- Read10X_h5("D:/FLX/scRNA_data/RAPT_CD3/treatment-A-filtered_gene_bc_matrices_h5.h5")
NoArg_D1 <- Read10X_h5("D:/FLX/scRNA_data/RAPT_CD3/treatment-B-filtered_gene_bc_matrices_h5.h5")
cntrl_D2 <- Read10X_h5("D:/FLX/scRNA_data/RAPT_CD3/treatment-C-filtered_gene_bc_matrices_h5.h5")
NoArg_D2 <- Read10X_h5("D:/FLX/scRNA_data/RAPT_CD3/treatment-D-filtered_gene_bc_matrices_h5.h5")
NoTrp_D2 <- Read10X_h5("D:/FLX/scRNA_data/RAPT_CD3/treatment-E-filtered_gene_bc_matrices_h5.h5")

ctrl_20k  <- CreateSeuratObject(counts= cntrl_D1_20K, project= "Ctrl_1", min.cells=1, min.features=200, assay="RNA")
ctrl_33k  <- CreateSeuratObject(counts= cntrl_D1_33K, project= "Ctrl_1", min.cells=1, min.features=200, assay="RNA")
ctrl_46k  <- CreateSeuratObject(counts= cntrl_D1_46K, project= "Ctrl_1", min.cells=1, min.features=200, assay="RNA")
NoTrp_1  <- CreateSeuratObject(counts= NoTrp_D1, project= "NoTrp_1", min.cells=1, min.features=200, assay="RNA")
NoArg_1  <- CreateSeuratObject(counts= NoArg_D1, project= "NoArg_1", min.cells=1, min.features=200, assay="RNA")
Ctrl_2  <- CreateSeuratObject(counts= cntrl_D2, project= "Ctrl_2", min.cells=1, min.features=200, assay="RNA")
NoArg_2  <- CreateSeuratObject(counts= NoArg_D2, project= "NoArg_2", min.cells=1, min.features=200, assay="RNA")
NoTrp_2  <- CreateSeuratObject(counts= NoTrp_D2, project= "NoTrp_2", min.cells=1, min.features=200, assay="RNA")

CD3 <- merge(x=ctrl_20k,y=c(ctrl_33k,ctrl_46k,NoTrp_1,NoArg_1,Ctrl_2,NoArg_2,NoTrp_2),add.cell.ids = c("Ctrl_1","Ctrl_1","Ctrl_1","NoTrp_1","NoArg_1","Ctrl_2","NoArg_2","No_Trp2"), merge.data= TRUE, project = "CD3")
CD3$orig.ident %>% table
CD3@active.ident %>% table
```


######################
QC Biological state of cells
  Mitochondrial gene expression
  Check Apoptosis
    nCount_RNA is # of unique UMIs I think
    Cut off percent.mt at < 30
mt genes are not being found here?? How did this work before??
#######################
```{r fig.height=3, fig.width=10}
CD3[["percent.mt"]] <- PercentageFeatureSet(CD3, pattern = "MT-")
CD3[["percent.HLA"]] <- PercentageFeatureSet(CD3, pattern = "HLA-") #HLA-D in mouse
CD3[["RNA.Per.Feature"]] <- CD3[["nCount_RNA"]] / CD3[["nFeature_RNA"]]
CD3[["percent.ribo"]] <- PercentageFeatureSet(CD3, pattern = c("^RP[LS]"))

CD3@assays$RNA@data@Dimnames[[1]][grepl("^MT-",CD3@assays$RNA@data@Dimnames[[1]])]
CD3@assays$RNA@data@Dimnames[[1]][grepl("^RP[LS]",CD3@assays$RNA@data@Dimnames[[1]])]

VlnPlot(CD3, features=c("nFeature_RNA","nCount_RNA","RNA.Per.Feature","percent.mt","percent.ribo"),ncol =3,pt.size = 0.1,log=TRUE) + 
  ggsave("CD3_QC_violin_cell_clusters.jpeg", width= 10, height=10, dpi=150, plot=last_plot(), units = "in")
VlnPlot(CD3, features=c("nCount_RNA"),ncol =4,pt.size = 0.1, y.max=20000)
VlnPlot(CD3, features=c("nCount_RNA","nFeature_RNA","percent.mt"),ncol =3,pt.size = 0.1, log=TRUE)

VlnPlot(CD3, features=c("BCL2","BAX","MLKL","HMGB1"),ncol=4,pt.size = 0.1, log=TRUE) +
  ggsave("CD3_QC_Apoptosis_Necrosis.jpeg", width= 10, height=5, dpi=150, plot=last_plot(), units = "in")

VlnPlot(CD3, features=c("BCL2_norm","BAX_norm","MLKL_norm","HMGB1_norm"),ncol=4,pt.size = 0.1, log=TRUE) +
  ggsave("CD3_norm_QC_Apoptosis_Necrosis.jpeg", width= 10, height=5, dpi=150, plot=last_plot(), units = "in")

CD3[["BCL2_norm"]] <- PercentageFeatureSet(CD3, pattern = "BCL2")
CD3[["BAX_norm"]] <- PercentageFeatureSet(CD3, pattern = "BAX")
CD3[["MLKL_norm"]] <- PercentageFeatureSet(CD3, pattern = "MLKL")
CD3[["HMGB1_norm"]] <- PercentageFeatureSet(CD3, pattern = "HMGB1")
  
#VlnPlot(CT26_immunome, features=c("Cd63","Fcgr3b","Ltf","S100a9","percent.ribo"),ncol =5,pt.size = 0.1, log=TRUE)
## Cell cycle

VlnPlot(CD3, features=c("MKI67","CCNA2","CCNB1","GAPDH"),ncol =4,pt.size = 0.1, log=TRUE) + ggsave("CD3_QC_Cell_Cycle.jpeg", width= 15, height=5, dpi=150, plot=last_plot(), units = "in")

```

######################
Normalize data
#######################
By default, we employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result. Normalized values are stored in pbmc[["RNA"]]@data
```{r}
CD3 <- NormalizeData(CD3,normalization.method = "LogNormalize",scale.factor=10000) # this goes into the @data slot
```
######################
Feature Selection
#######################
Top variable genes: 
 [1] "Jchain"  "Igkc"    "Hba-a1"  "Car2"    "Col3a1"  "Ighm"    "Hbb-bs"  "Col1a1"  "Car1"    "Hba-a2"  "Rarres2" "Col5a2"  "Hbb-bt"  "Igfbp7" 
[15] "Ccl8"    "Meg3"    "Pf4"     "Fstl1"   "Ccl11"   "Bgn"  
```{r fig.height=5, fig.width=12}
CD3 <- FindVariableFeatures(CD3, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(CD3),30)

plot1 <- VariableFeaturePlot(CD3)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))
```

######################
Data scaling to median = 0
Z-score
Stored in @scale.data
#######################
```{r}
all.genes <- rownames(CD3)
CD3 <- ScaleData(CD3,features = all.genes, assay = "RNA")
```

####################################
PCA analysis
UMAP
FindClusters
####################################

```{r}
CD3 <- RunPCA(CD3,features =c(VariableFeatures(object=CD3)))
```


Heapmap of PCs
Cut off should be at 12 PCs from this plot
```{r fig.height=25, fig.width=25}
DimHeatmap(CD3, dims=1:24, cells=500, balanced = TRUE)
```


####################################
DimHeatMap shows first ~12 are distinguishible as groups
Elbowplot shows first ~12 are significant
####################################
```{r}
ElbowPlot(CD3)
```

###################################
KNN Clustering
###################################

Based on marker clusters, 15 clusters is too many. Reduce resolution 
```{r}
CD3 <- FindNeighbors(CD3, dims=1:12)
CD3 <- FindClusters(CD3, resolution=0.5) #Resolution 1.2 > 17 clusters, Resolution 1.0 > 15 clusters, 0.5 > 8
table(Idents(CD3))

```

UMAP 

Looks better than TSNE
```{r fig.height=5, fig.width=7}
CD3 <- RunUMAP(CD3, dims=1:12, min.dist=0.6, verbose=FALSE)
DimPlot(CD3,reduction="umap", label=TRUE, pt.size=1.5, label.size = 5) + NoLegend() + ggsave("CD3_UMAP_numbers.jpg", width=7, height=5, dpi=150, plot= last_plot(), units = "in")
```

Find Top markers of each cluster
```{r}
cd3.markers <- FindAllMarkers(CD3, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.8)
save(cd3.markers, file="D:/FLX/scRNA_data/RAPT_CD3/CD3_all_vs_all_DE.rdata")
write.table(cd3.markers, file="D:/FLX/scRNA_data/RAPT_CD3/CD3_all_vs_all_DE.txt", sep="\t")

cd3.markers %>% group_by(cluster) %>% top_n(n=15)
#Cluster
# 0: CD79 B Cells Coronin high, motility
# 1: B cells
# 2: LBT, SELL high
# 3: Ribo high
# 4: 
# 5 
# 6: 
# 7: CD8 T exh CCL3 CCL5 PRF1 LAG3 NKG7 GNLY
# 8: CEN TOP2A highly dividing
# 9: S100A4 Myeloid or T
# 10: NK TIGIT T or NK cells
# 11: NK XCL1 positive
# 12: S100A11
# 13: 
# 15:
# 16: 
```


Plot Features
Not expressed: ,"CCR6"
```{r fig.height=9, fig.width=15}
FeaturePlot(CD3,slot = "data", cols=c("grey","red"),features=c("PTPRC","CD3D","CD4","CD8A","NKG7","GZMB","PRF1","FOXP3","CXCR3","LAG3","SELL","TOP2A"))
FeaturePlot(CD3,slot = "data", cols=c("grey","red"),features=c("HPGD"))
```
Non-CD3 call types

```{r fig.height=6, fig.width=13}
FeaturePlot(CD3,slot = "data", cols=c("grey","red"),features=c("PTPRC","CD3D","HLA-DRA","IDO1","CD14","ITGAM","FCGR3A","BDCA1","BDCA3","CD141","ADGRE1","ITGAX","MARCO","S100A9"),ncol=4)
```

Sample of origin

```{r fig.height=9, fig.width=17}
Ctrl <- DimPlot(CD3, cells.highlight = WhichCells(CD3,expression= orig.ident ==c("Ctrl_1")),cols.highlight="#0A95A7", pt.size=0.1) + NoLegend()
noArg <- DimPlot(CD3, cells.highlight = WhichCells(CD3,expression= orig.ident ==c("NoArg_1")),cols.highlight="purple", pt.size=0.1) + NoLegend()
noTrp <- DimPlot(CD3, cells.highlight = WhichCells(CD3,expression= orig.ident ==c("NoTrp_1")),cols.highlight="red", pt.size=0.1) + NoLegend()
Ctrl2 <- DimPlot(CD3, cells.highlight = WhichCells(CD3,expression= orig.ident ==c("Ctrl_2")),cols.highlight="#0A95A7", pt.size=0.1) + NoLegend()
noArg2 <- DimPlot(CD3, cells.highlight = WhichCells(CD3,expression= orig.ident ==c("NoArg_2")),cols.highlight="purple", pt.size=0.1) + NoLegend()
noTrp2 <- DimPlot(CD3, cells.highlight = WhichCells(CD3,expression= orig.ident ==c("NoTrp_2")),cols.highlight="red", pt.size=0.1) + NoLegend()

#This below doesn't work. The groups can't be combined
Ctrl_all <- DimPlot(CD3, cells.highlight = WhichCells(CD3,expression= orig.ident ==c("Ctrl_1","Ctrl_2")),cols.highlight="#0A95A7", pt.size=0.1) + NoLegend()
Trp_all <- DimPlot(CD3, cells.highlight = WhichCells(CD3,expression= orig.ident ==c("NoTrp_1","NoTrp_2")),cols.highlight="red", pt.size=0.1) + NoLegend()
Arg_all <- DimPlot(CD3, cells.highlight = WhichCells(CD3,expression= orig.ident ==c("NoArg_1","NoArg_2")),cols.highlight="purple", pt.size=0.1) + NoLegend()

library(gridExtra)
grid.arrange(Ctrl,noArg,noTrp,Ctrl2,noArg2,noTrp2, nrow = 2)  
grid.arrange(Ctrl_all,Trp_all,Arg_all, nrow = 1)  
CD3$orig.ident %>% table
```


####################
Signatures
#####################

Export and check data

```{r}
rna_matrix <- as.matrix(CD3$RNA@scale.data)
save(rna_matrix,file = "CD3_counts_matrix.rdata")
load("rna_scale.data_matrix.rdata")
rna_matrix[1:5,1:5] #looks normalized
boxplot(rna_matrix[,1:50])
```

To put data back into batch1, the data has to be a data.frame, with cells as rowsnames and a single column with a name 
dataframe changes the cell name from Ctrl_1_AAACCGTGCCTCCA-1_1 to Ctrl_1_AAACCGTGCCTCCA.1_1, so 
```{r}
rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% Trp_AAS) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% summarise_all(median) %>% t %>% data.frame


CD3[["GCN2_pathway"]] <- rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% GCN2_h) %>% summarize_if(is.numeric,sum_mx) %>% t %>% data.frame %>% tibble::rownames_to_column("Cell") %>%  mutate(Cell= sub("\\.","-",Cell)) %>% tibble::column_to_rownames("Cell") 
CD3[["PERK_pathway"]] <- rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% PERK_h) %>% summarize_if(is.numeric,sum_mx) %>% t %>% data.frame %>% tibble::rownames_to_column("Cell") %>%  mutate(Cell= sub("\\.","-",Cell)) %>% tibble::column_to_rownames("Cell") 
CD3[["HRI_pathway"]] <- rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% HRI_h) %>% summarize_if(is.numeric,sum_mx) %>% t %>% data.frame %>% tibble::rownames_to_column("Cell") %>%  mutate(Cell= sub("\\.","-",Cell)) %>% tibble::column_to_rownames("Cell") 
CD3[["ISR_pathway"]] <- rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("DDIT3","TRIB3","PPP1R15A","ASNS","ATF4")) %>% summarize_if(is.numeric,sum) %>% t %>% data.frame  %>% tibble::rownames_to_column("Cell") %>%  mutate(Cell= sub("\\.","-",Cell)) %>% tibble::column_to_rownames("Cell") 
CD3[["AAS"]] <- rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% aas_sig) %>% summarize_if(is.numeric,sum) %>% t %>% data.frame  %>% tibble::rownames_to_column("Cell") %>%  mutate(Cell= sub("\\.","-",Cell)) %>% tibble::column_to_rownames("Cell") 
CD3[["Arg_AAS"]] <- rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% Arg_AAS_h) %>% summarize_if(is.numeric,sum) %>% t %>% data.frame  %>% tibble::rownames_to_column("Cell") %>%  mutate(Cell= sub("\\.","-",Cell)) %>% tibble::column_to_rownames("Cell") 

CD3[["Trp_AAS"]] <- rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% Trp_AAS) %>% summarize_if(is.numeric,sum) %>% t %>% data.frame  %>% tibble::rownames_to_column("Cell") %>%  mutate(Cell= sub("\\.","-",Cell)) %>% tibble::column_to_rownames("Cell") 

CD3[["Hypoxia"]] <- rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% hypoxia) %>% summarize_if(is.numeric,sum_mx) %>% t %>% data.frame  %>% tibble::rownames_to_column("Cell") %>%  mutate(Cell= sub("\\.","-",Cell)) %>% tibble::column_to_rownames("Cell") 


Arg_AAS_h <- Arg_Sig %>% pull(Gene.Symbol)
Arg_AAS_h <- Arg_Sig %>% pull(Gene.Symbol)
Trp_AAS <- read_tsv("D:/FLX/Reference_tables/Trp_sig_h.txt") %>% pull(x)
hypoxia <- read.table("D:/FLX/GCN2/ISR_data/RNA_seq/GSEA_hypoxia.txt",sep="\n", header=TRUE) %>% pull(HALLMARK_HYPOXIA)

```

```{r fig.height=4, fig.width=15}
FeaturePlot(CD3,slot = "data", cols=c("grey","red"),features=c("GCN2_pathway","AAS","ISR_pathway","Hypoxia"), ncol=4, max.cutoff=80, label=TRUE,label.size=4)
```

Signature correlations

```{r}
CD3[["GCN2_pathway"]] %>% 
  cbind(CD3[["AAS"]]) %>% 
  cbind(CD3$orig.ident) %>% dfilter(CD3$orig.ident != "NoTrp_2") %>% data.frame %>%
  ggplot(aes(GCN2_pathway,AAS, colour=CD3.orig.ident, group=NA)) + 
  geom_point(size=1.5, alpha=0.7) +
  scale_color_rapt() +
  geom_smooth(method="lm", size=0.25) +
  #annotate("text", x=25,y=1500, label = "paste(italic(R) ^ 2, \" = 0.86\")", parse = TRUE) # +
  ggsave("GCN2_vs_Arg_AAS.jpg", width=7, height=5, dpi=150, plot= last_plot(), units = "in")
  
CD3[["Arg_AAS"]] %>% 
  cbind(CD3[["AAS"]]) %>% 
  cbind(CD3$orig.ident) %>% dfilter(CD3$orig.ident != "NoTrp_2") %>% data.frame %>%
  ggplot(aes(Arg_AAS,AAS, colour=CD3.orig.ident, group=NA)) + 
  geom_point(size=3, alpha=0.5) +
  scale_color_rapt() +
  geom_smooth(method="lm", size=0.25) +
  #annotate("text", x=25,y=1500, label = "paste(italic(R) ^ 2, \" = 0.86\")", parse = TRUE) # +
  ggsave("GCN2_vs_Arg_AAS.jpg", width=7, height=5, dpi=150, plot= last_plot(), units = "in")

CD3[["ISR_pathway"]] %>% 
  cbind(CD3[["AAS"]]) %>% 
  cbind(CD3$orig.ident) %>% dfilter(CD3$orig.ident != "NoTrp_2") %>% data.frame %>%
  ggplot(aes(ISR_pathway,AAS, colour=CD3.orig.ident, group=NA)) + 
  geom_point(size=3, alpha=0.5) +
  scale_color_rapt() +
  geom_smooth(method="lm", size=0.25) +
  annotate("text", x=10,y=200, label = "paste(italic(R) ^ 2, \" = 0.90\")", parse = TRUE) +
  ggsave("GCN2_vs_Arg_AAS.jpg", width=7, height=5, dpi=150, plot= last_plot(), units = "in")

CD3[["Arg_AAS"]] %>% 
  cbind(CD3[["GCN2_pathway"]]) %>% 
  cbind(CD3$orig.ident) %>% cbind(CD3@active.ident) %>% dfilter(CD3$orig.ident != "NoTrp_2")  %>% data.frame %>%
  ggplot(aes(Arg_AAS,GCN2_pathway, colour=CD3.orig.ident, group=NA)) + 
  geom_point(size=3, alpha=0.5) +
  scale_color_rapt() +
  geom_smooth(method="lm", size=0.25) +
  annotate("text", x=200,y=22, label = "paste(italic(R) ^ 2, \" = 0.39\",\"; Pearson = 0.63\")", parse = TRUE) +
  ggsave("GCN2_vs_Arg_AAS.jpg", width=7, height=5, dpi=150, plot= last_plot(), units = "in")
  
CD3[["Arg_AAS"]] %>% 
  cbind(CD3[["GCN2_pathway"]]) %>% 
  cbind(CD3$orig.ident) %>% cbind(CD3@active.ident) %>% dfilter(CD3@active.ident == "Macrophage")  %>% data.frame %>%
  ggplot(aes(Arg_AAS,GCN2_pathway, colour=CD3.active.ident, group=NA)) + 
  geom_point(size=3, alpha=0.5) +
  scale_color_rapt() +
  geom_smooth(method="lm", size=0.25) +
  annotate("text", x=200,y=22, label = "paste(italic(R) ^ 2, \" = 0.43\",\"; Pearson = 0.66\")", parse = TRUE) +
  ggsave("GCN2_vs_Arg_AAS_macrophage.jpg", width=7, height=5, dpi=150, plot= last_plot(), units = "in")

#AAS correlates less well


cor(CD3[["Arg_AAS"]],CD3[["GCN2_pathway"]]) #0.63

mac <- CD3[["Arg_AAS"]] %>% 
  cbind(CD3[["GCN2_pathway"]]) %>% 
  cbind(CD3$orig.ident) %>% cbind(CD3@active.ident) %>% dfilter(CD3@active.ident == "Macrophage") 

all <- CD3[["Arg_AAS"]] %>% 
  cbind(CD3[["GCN2_pathway"]]) %>% 
  cbind(CD3$orig.ident) %>% cbind(CD3@active.ident) 
  
cor(mac[["Arg_AAS"]],mac[["GCN2_pathway"]]) #0.66

cor(CD3[["AAS"]],CD3[["ISR_pathway"]]) # 0.90

cor(CD3[["AAS"]],CD3[["ISR_pathway"]]) # 0.90

lm <- lm(GCN2_pathway ~ Arg_AAS, data= all )
summary(lm) 
```




Split data by donor
 Ctrl_1  Ctrl_2 NoArg_1 NoArg_2 NoTrp_1 NoTrp_2 
   1449     326     731     406     666    1001 
```{r}
Ctrl1 <- subset(CD3, subset= orig.ident=="Ctrl_1")
Ctrl2 <- subset(CD3, subset= orig.ident=="Ctrl_2")
Arg_Starve_1 <- subset(CD3, subset= orig.ident=="NoArg_1")
Arg_Starve_2 <- subset(CD3, subset= orig.ident=="NoArg_2")
Trp_Starve_1 <- subset(CD3, subset= orig.ident=="NoTrp_1")
Trp_Starve_2 <- subset(CD3, subset= orig.ident=="NoTrp_2")
#CD3_2 <- subset(CD3, idents=c("Ctrl_1","Ctrl_2","NoArg_1","NoArg_2","NoTrp_1"))
CD3_2$orig.ident %>% table
```

```{r fig.height=4, fig.width=16}
FeaturePlot(Ctrl2,slot = "data", cols=c("grey","red"),features=c("GCN2_pathway","PERK_pathway","HRI_pathway","ISR_pathway"), ncol=4, max.cutoff=15)
FeaturePlot(Arg_Starve_2,slot = "data", cols=c("grey","red"),features=c("GCN2_pathway","PERK_pathway","HRI_pathway","ISR_pathway"), ncol=4, max.cutoff=15)
FeaturePlot(Trp_Starve_2,slot = "data", cols=c("grey","red"),features=c("GCN2_pathway","PERK_pathway","HRI_pathway","ISR_pathway"), ncol=4, max.cutoff=15)
```

RidgePlot
```{r fig.height=3, fig.width=6}
RidgePlot(CD3,features = c("GCN2_pathway"),slot="counts", group.by="orig.ident", idents=c("Ctrl_1","NoTrp_1","NoArg_1","Ctrl_2","NoTrp_2","NoArg_2"),ncol=4) + NoLegend() + ggsave("All_GCn2_ridge.jpg", width=6, height=3, dpi=150, plot= last_plot(), units = "in")

RidgePlot(CD3,features = c("AAS"),slot="counts", group.by="orig.ident",sort="increasing", idents=c("Ctrl_1","NoTrp_1","NoArg_1","Ctrl_2","NoTrp_2","NoArg_2"),ncol=4) + NoLegend() + ggsave("All_GCn2_ridge.jpg", width=6, heigh=3, dpi=150, plot= last_plot(), units = "in")
# 
RidgePlot(CD3,features = c("ISR_pathway"),slot="counts", group.by="orig.ident",sort="increasing", idents=c("Ctrl_1","NoTrp_1","NoArg_1","Ctrl_2","NoTrp_2","NoArg_2"),ncol=4) + NoLegend() #+ ggsave("All_GCn2_ridge.jpg", width=6, height=3, dpi=150, plot= last_plot(), units = "in")

#,cols=c("orange","chartreuse3","dodgerblue2","orange","chartreuse3","dodgerblue2")

RidgePlot(CD3,features = c("HRI_pathway"),slot="counts", group.by="orig.ident",sort="increasing", idents=c("Ctrl_1","NoTrp_1","NoArg_1","Ctrl_2","NoTrp_2","NoArg_2"),ncol=4) + NoLegend() # ggsave("All_GCn2_ridge.jpg", width=23, height=3, dpi=150, plot= last_plot(), units = "in")

,cols=c("orange","chartreuse3","dodgerblue2","orange","chartreuse3","dodgerblue2")

CD3$orig.ident %>% table
```
Plot Ridge
```{r fig.height=7, fig.width=15}
install.packages("ggridges")
library(ggridges)
CD3[["GCN2_pathway"]] %>% cbind(CD3$orig.ident) %>% data.frame %>% arrange(desc(CD3.orig.ident)) %>% dfilter(CD3.orig.ident != "NoTrp_2") %>% ggplot(aes(y=CD3.orig.ident,x=GCN2_pathway, fill=CD3.orig.ident)) + geom_density_ridges() + xlim(-2.5,15) + NoLegend() + theme(axis.title.y = element_blank(), axis.text= element_text(size=14))

CD3[["ISR_pathway"]] %>% cbind(CD3$orig.ident) %>% data.frame %>% arrange(desc(CD3.orig.ident)) %>% dfilter(CD3.orig.ident != "NoTrp_2") %>% ggplot(aes(y=CD3.orig.ident,x=ISR_pathway, fill=CD3.orig.ident)) + geom_density_ridges(alpha=0.8) + xlim(-5,30) + NoLegend() + theme(axis.title.y = element_blank(), axis.text= element_text(size=12))

CD3[["AAS"]] %>% cbind(CD3$orig.ident) %>% data.frame %>% arrange(desc(CD3.orig.ident)) %>% dfilter(CD3.orig.ident != "NoTrp_2") %>% ggplot(aes(y=CD3.orig.ident,x=AAS, fill=CD3.orig.ident)) + geom_density_ridges(alpha=0.8) + xlim(-5,150) + NoLegend() + theme(axis.title.y = element_blank(), axis.text= element_text(size=12))

sigs <- CD3[["AAS"]] %>% cbind(CD3$orig.ident,CD3$GCN2_pathway,CD3$Arg_AAS,CD3$Trp_AAS,CD3$ISR_pathway,CD3$PERK_pathway,CD3$HRI_pathway) 
colnames(sigs) <- new_colnames
sigs %>% data.frame %>% dfilter(orig.ident != "NoTrp_2") %>% dselect(-GCN2_pathway,-HRI_pathway,-PERK_pathway) %>% mutate(orig.ident= factor(orig.ident, levels=c("Ctrl_1","NoArg_1","NoTrp_1","Ctrl_2","NoArg_2"))) %>% gather(Signature, Score, -orig.ident) %>% ggplot(aes(y=orig.ident,x=Score, fill=orig.ident))  + geom_density_ridges(alpha=0.7, rel_min_height=0.01) + facet_wrap(~Signature, scale="free_x") + NoLegend() + theme(axis.title.y = element_blank(), axis.text= element_text(size=12)) + scale_fill_manual(values=c("#5B666F","#0A95A7","#E5541B","#5B666F","#0A95A7")) + ggsave("All_Sigs_ridge.jpg", width=15, height=6, dpi=150, plot= last_plot(), units = "in")

sigs %>% data.frame %>% dfilter(orig.ident != "NoTrp_2") %>% dselect(-AAS,-Arg_AAS,-Trp_AAS,-ISR_pathway) %>% mutate(orig.ident= factor(orig.ident, levels=c("Ctrl_1","NoArg_1","NoTrp_1","Ctrl_2","NoArg_2"))) %>% gather(Signature, Score, -orig.ident) %>% ggplot(aes(y=orig.ident,x=Score, fill=orig.ident))  + geom_density_ridges(alpha=0.7, rel_min_height=0.01) + facet_wrap(~Signature, scale="free_x") + xlim(-2,18) + NoLegend() + theme(axis.title.y = element_blank(), axis.text= element_text(size=12)) + scale_fill_manual(values=c("#5B666F","#0A95A7","#E5541B","#5B666F","#0A95A7")) + ggsave("All_Sigs_ridge2.jpg", width=15, height=6, dpi=150, plot= last_plot(), units = "in")

my_colnames <- sigs %>% colnames
new_colnames <- gsub("macrophage.","",my_colnames)

#blue="#0A95A7", orange="#E5541B", gray="#5B666F", yellow="#FE9F33", green="#4DA167", lilac="#947EB0"
```



Ridgeplot for Myeloid clusters 6 and 8
```{r fig.height=4, fig.width=16}
RidgePlot(CD3,features = c("GCN2_pathway"),group.by="orig.ident", idents="Macrophage",ncol=4) + NoLegend() # + ggsave("All_GCn2_ridge.jpg", width=6, height=3, dpi=150, plot= last_plot(), units = "in")
RidgePlot(CD3,features = c("GCN2_pathway","AAS","ISR_pathway"),group.by="orig.ident",slot="counts", idents="Macrophage") + NoLegend()

sigs <- macrophage[["AAS"]] %>% cbind(macrophage$orig.ident,macrophage$GCN2_pathway,macrophage$Arg_AAS,macrophage$ISR_pathway,macrophage$PERK_pathway,macrophage$HRI_pathway) 

new_colnames <- gsub("macrophage.","",my_colnames)
colnames(sigs) <- new_colnames

sigs %>% data.frame %>% dfilter(!grepl("_2",orig.ident)) %>% dselect(-GCN2_pathway,-HRI_pathway,-PERK_pathway) %>% mutate(orig.ident= factor(orig.ident, levels=c("Ctrl_1","NoArg_1","NoTrp_1","Ctrl_2","NoArg_2"))) %>% gather(Signature, Score, -orig.ident) %>% ggplot(aes(y=orig.ident,x=Score, fill=orig.ident))  + geom_density_ridges(alpha=0.7, rel_min_height=0.01) + facet_wrap(~Signature, scale="free_x") + NoLegend() + theme(axis.title.y = element_blank(), axis.text= element_text(size=12)) + theme(strip.text = element_text(size=15)) + scale_fill_manual(values=c("#5B666F","#0A95A7","#E5541B","#5B666F","#0A95A7")) + ggsave("All_Sigs_macrophage_ridge.jpg", width=15, height=3, dpi=150, plot= last_plot(), units = "in")

sigs %>% data.frame %>% dfilter(!grepl("_2",orig.ident)) %>% dselect(-AAS,-Arg_AAS,-ISR_pathway) %>% mutate(orig.ident= factor(orig.ident, levels=c("Ctrl_1","NoArg_1","NoTrp_1","Ctrl_2","NoArg_2"))) %>% gather(Signature, Score, -orig.ident) %>% ggplot(aes(y=orig.ident,x=Score, fill=orig.ident))  + geom_density_ridges(alpha=0.7, rel_min_height=0.01) + facet_wrap(~Signature, scale="free_x") + xlim(-15,200) + NoLegend() + theme(axis.title.y = element_blank(), axis.text= element_text(size=12)) + theme(strip.text = element_text(size=15)) + scale_fill_manual(values=c("#5B666F","#0A95A7","#E5541B","#5B666F","#0A95A7")) + ggsave("All_Sigs_macrophage_ridge2.jpg", width=15, height=3, dpi=150, plot= last_plot(), units = "in")

```
Subset Macrophages

```{r fig.height=3, fig.width=4}
macrophage <- subset(CD3, idents="Macrophage")
DimPlot(macrophage,group.by="orig.ident")

rna_matrix <- as.matrix(macrophage$RNA@counts)

rna_matrix %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% GCN2_h) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% summarise_all(median) %>% t %>% data.frame %>% count(.data$.)
macrophage$orig.ident %>% table
```



#################################
Name the clusters for now

Ctrl1: 0
No Arg 1: 4
No Trp 1 : 3, 7
Ctrl 2: 1, 0 ,5
No Arg 2: 1, 4, 5
No Trp 2: 2
###################################

```{r}

mRNA_ident <- c("Ctrl_1",1,"NoTrp_2","NoTrp_1","NoArg_1",5,"Macrophage","NoTrp_1","Macrophage")
#make the quivalency of the old labels to the new label names

names(mRNA_ident) <- levels(CD3) # 0 to 8
CD3 <- RenameIdents(CD3,mRNA_ident)
CD3@active.ident %>% table
CD3[["Cell_type_call"]] <- CD3@active.ident

```