---
title: "Cellclassifier"
output: html_document
---

Load data

```{r}
my_data <- read.delim("C:/Users/lgutierrez/Desktop/Downloads/10X_pbmc_t_variable.txt")
```

Removing high abundance genes

```{r}
GENES <- names(my_data.caret)
index<-grep("^RP.*|^MT.*", GENES, perl =T) #Ribosomal proteins & Mitochondrial genes
GENES[index]
```

```{r}
my_data.caret <- my_data[, -index]
```

```{r}
library(tidyverse)
library(data.table)
my_data.caret <- data.table(my_data.caret)
```


Joining labels 4 & 6 (They are both Naive Cells) & Switching labels 10 and 11

```{r}
my_data.caret[id_num == "6", id_num :="3",]
my_data.caret[id_num == "10", id_num :="22",]
my_data.caret[id_num == "11", id_num :="10",]
my_data.caret[id_num == "22", id_num :="11",]

my_data.caret[id_num == "3", id:= "3. Naive CD4 T"]
my_data.caret[id_num == "10", id:= "10. NKT cell"]
my_data.caret[id_num == "11", id:= "11. Immature Monocyte"]
my_data.caret[id_num == "17", id:= "17. Mix T & B cells"]


```


```{r}
my_data.caret$id_num <- as.factor(my_data.caret$id_num)
```

```{r}
unique(my_data.caret$id_num)
```


Remove the id from the dataset

```{r}
my_data.caret2 <- my_data.caret[, -2,]
```

Splitting data

```{r}
library(caret)
set.seed(1)
inTrain <- createDataPartition(
  y = my_data.caret2$id_num,
  p = 0.75,  ##The percentage of data in the training set
  list = FALSE
)

```

Defining training and testing datasets

```{r}
training.caret <- my_data.caret2[inTrain,]
testing.caret <- my_data.caret2[-inTrain,]
```

Testing labels

```{r}
labels <- testing.caret$id_num
```



Applying PCA to both training and testing datasets

```{r}
pca_generator <- preProcess(training.caret, method= "pca", thresh=0.75)
training.pcs <- predict(pca_generator, training.caret)

```

```{r}
testing.pcs <- predict(pca_generator, testing.caret)
```

Training

Cross-validation

```{r}
fitControl <- trainControl(
  method = "repeatedcv", 
  repeats = 5,
  n = 5
)

fitControl2 <- trainControl(
  method = "none"
)
```


LDA

```{r}
metric <- "Accuracy"

set.seed(123)
LDAFit.pcs <- train(id_num ~., data = training.pcs, method = "lda", trControl = fitControl, metric = metric )
```

Training accuracy score

```{r}
LDAFit.pcs
```

Testing accuracy score

```{r}
LDAFit.pcs.test <- predict(LDAFit.pcs, testing.pcs[, -1,])
confusionMatrix(LDAFit.pcs.test, labels)
```

Model seems to missclassify many labels 2. Active Monocyte and 5. Monocyte

Random Forest (mtry = 21)

```{r}
mtry <- sqrt(ncol(training.pcs))

randomForestGrid.pcs <- expand.grid(mtry = c(7, 8, 9, 10, 21, 22, 23))

set.seed(123)
randomForestFit.pcs <- train(id_num~., data = training.pcs, method = "rf", trControl = fitControl, tuneGrid = randomForestGrid.pcs, metric = metric)
```


```{r}
randomForestFit.pcs.test <- predict(randomForestFit.pcs, testing.pcs[, -1, ])
confusionMatrix(randomForestFit.pcs.test, labels)
```

```{r}
plot(randomForestFit.pcs)
```

random Forest (mtry = 10) --> Without cross-validation

```{r}
randomForestFit.pcs.10.test <- predict(randomForestFit.pcs.10, testing.pcs)
confusionMatrix(randomForestFit.pcs.10.test, labels)
```

gbm (interaction.depth = 8, n.trees = 250, shrinkage = 0.1, n.minobsinnode = 30)

```{r}
gbmGrid.pcs <- expand.grid(interaction.depth = c(2,4,8), n.trees = c(150, 200, 250), shrinkage = 0.1, n.minobsinnode = 30)

set.seed(123)
gbmFit.pcs <- train(id_num~., data = training.pcs, method = "gbm", trControl = fitControl, tuneGrid = gbmGrid.pcs, verbose = TRUE, metric = metric)
```


```{r}
gbmFit.pcs
```

```{r}
gbmFit.pcs.test <- predict(gbmFit.pcs, testing.pcs[, -1,])
confusionMatrix(gbmFit.pcs.test, labels)
```

```{r}
plot(gbmFit.pcs)
```

**** MOST ACCURATE ****

gbmFit (interaction.depth = 8, n.trees = 150, shrinkage = 0.1, n.minobsinnode = 30)--> Without cross-validation

```{r}
gbmFit.pcs.150.test <- predict(gbmFit.pcs.150, testing.pcs[, -1,])
confusionMatrix(gbmFit.pcs.150.test, labels)
```






