---
title: "Check Azizi Seurat data for Manu Shiny tool"
author: "Mengshu"
date: "1/2/2019"
output: html_document
---


```{r}
library(Seurat)
#update.packages("Seurat")
library(data.table)
library(magrittr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(tibble)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
#source("D:/FLX/Files_from_Gene/R_functions.r")
```

Load data from rds file

This version of the Azizi Seurat object has the t-sne projection, but no UMAP
I preferred the t-sne in this case because the cells are distributed well with less "clumping"
```{r fig.height=6, fig.width=12}
pbmc<- readRDS("Azizi_PBMC_Seurat.rds")
azizi <- UpdateSeuratObject(pbmc)
saveRDS(azizi,file="azizi_v3.rds")
readRDS("azizi_v3.rds")

#azizi@reductions$tsne
DimPlot(azizi, reduction="tsne", label=TRUE) +
ggsave("CITE_Clusters_UMAP.jpg", width=12, height=6, dpi=300, plot= last_plot(), units = "in")

azizi_cluster <- AverageExpression(azizi)
save(azizi_cluster, file="Azizi_PBMC_clusters_ave.rdata")
write.table(azizi_cluster, file="Azizi_PBMC_clusters_ave.txt", sep="\t")
```
Clusters 10. NKT and 11. Premature Monocyte look switched
They *are* switched: 11. is GNLY high, CD14-, while 10. is CD14+ GNLY-, so 10. is clearly monocytic, while 11. is NKT (T because it's CD3+ but CD8-, and NK beccause GNLY+)
```{r fig.height=12, fig.width=24}
FeaturePlot(azizi,slot = "data",reduction="tsne", features = c("CD14", "CD3D","GNLY","CD8A"))
```
Fix cluster labels

```{r fig.height=6, fig.width=12}
#current labels
levels(azizi@active.ident)
levels(azizi)
fix.cluster.ids <- c("0. Naive CD8 T","1. Active CD4 T","2. Active Monocyte","3. Naive CD4 T","4. B cell precursor","5. Monocyte","6. Naive CD4 T","7. T EM","8. B cells","9. NK cell","10.Immature Monocyte","11. NKT cell","12. FCGR3A+ DC4", "13. DC2","14. T effector","15. Apoptotic T","16. T RM","17. CD3+ B","18. DC1","19. Megakaryocyte" ,"20. NK precursor")  
names(fix.cluster.ids) <- levels(azizi)
azizi <- RenameIdents(azizi, fix.cluster.ids)
DimPlot(object = azizi, label = TRUE, reduction= 'tsne') # fixed. 
```







Don't know why UMAP slot is empty here, since I ran it before: rerun with same dims=1:24
```{r fig.height=8, fig.width=12}
library(umap)
azizi <-  RunUMAP(azizi, dims=1:24) 
DimPlot(object = azizi, label = TRUE, reduction.use = 'umap')
DimPlot(object = azizi, do.label = TRUE, reduction.use = 'umap', pt.size = 0.8, label.size = 4) 
ggsave("Azizi_PBMCs_UMAP.jpg", width= 12, height=8, dpi=600, plot= last_plot(), units = "in")
```



