---
title: "HPK1 Correlation PBMCs"
author: "Mengshu"
date: "November 19, 2018"
output: html_document
---

Perform correlation analysis for HPK1 in scRNA datasets. Starting with the healthy PBMC dataset. 
Load the Imputed, normalized and filtered data from Seurat

```{r}
library(tidyverse)
library(dplyr)
library(magrittr)
library(data.table)
library(ggplot2)
source("R_functions.r")

install.packages("Hmisc")
library("Hmisc")
detach("package:Hmisc")
detach("package:data.table")
detach("package:Seurat")
```

Load Seurat dataset: 10X_8K_PBMC_Seurat_Imputed_original

```{r}
pbmc<- readRDS("10X_8K_PBMC_Seurat_Imputed_original.rds")
pbmc_matrix <- as.matrix(pbmc@data)
dim(pbmc_matrix)
#18408  8483 : 18K genes, 8K cells
```

Run rcorr from the Hmisc package
##This takes too long to run. It ran on the 8 RAM laptop for 2 whole days and never finished

```{r}
corr_result <- rcorr(pbmc_matrix, type="spearman")
```

Cluster on cell type subclusters, the 21 classes that were found, so the table would be 18408 by 21
Go back to the Seurat file and export the average expression values of each cluster. 
```{r}
not done yet
```
Gene suggested running HPK1 correlation with all other genes
1. Transpose the matrix so that Genes are cols and Cells are rows. Then save HPK1 into vector y, save matrix as x
2. HPK1_corr <- rcorr(x,y, type="spearman")
3. This returns, r: correlations, NA if one number is missing, n: number of observations used, for self-self diagonal cells, n is number of the single variable, and ,p: P-values
Correlation ran from Mon to Thur morning. 4 Days!
Unexpectedly, the correlation actually ran all genes vs. all genes, not just HPK1 vs. all genes. 
```{r}
pbmc_t <- t(pbmc_matrix)
dim(pbmc_t)
#pbmc_t[1:5,1:5]
HPK1_vector <- pbmc_t[,"MAP4K1"]
head(HPK1_vector)

hpk1_corr <- rcorr(pbmc_t, HPK1_vector, type="spearman")
save(hpk1_corr, file="8K_PBMC_HPK1_corr.rdata")
hpk1_corr <- load("8K_PBMC_HPK1_corr.rdata")
```

Look at a ranked list of HPK1 correlated genes
# Don't use Tibble because it doesn't allow row names. 
```{r}
hpk1_matrix <- data.frame(hpk1_corr$r)
hpk1_matrix %<>% mutate(gene=colnames(hpk1_matrix)) #added as last col

class(hpk1_matrix)
head(hpk1_matrix)
str(hpk1_matrix)
head(remove_RP)

remove_RP <- colnames(hpk1_matrix %>% dselect(starts_with("RP")))
hpk1_matrix %<>% dselect(-remove_RP) %>% dfilter(MAP4K1 >= 0.5)
hpk1_matrix[,c(ncol(hpk1_matrix),1:(ncol(hpk1_matrix)-1))]
n <- hpk1_matrix$gene
p <- colnames(hpk1_matrix)
t_hpk1 <- as.data.frame(t(hpk1_matrix[,-1]))
colnames(t_hpk1) <- n
rownames(t_hpk1) <- head(p,-1) #error
t_hpk1 %<>% mutate(gene=head(p,-1))
t_hpk1[,c(ncol(t_hpk1),1:(ncol(t_hpk1)-1))]
head(t_hpk1)
tail(p)
remove_RP2 <- colnames(t_hpk1 %>% dselect(starts_with("RP")))
t_hpk1 %<>% dfilter(MAP4K1 >= 0.5) %<>% dselect(-remove_RP2)
write.table(t_hpk1, file="Corr_sort_PBMC_filter.txt", quote= FALSE, sep="\t", row.names=FALSE)


hpk1_matrix %<>% dfilter(rownames)
hpk1_corr_sort <- sort(hpk1_corr$r[,"MAP4K1"], decreasing = TRUE) 
write.csv(hpk1_corr_sort, file="HPK1_corr_sort_PBMC.txt", quote= FALSE, sep="\t")
head(hpk1_corr_sort)
tail(hpk1_corr_sort)
```



```{r}
hpk1_matrix <- data.frame(hpk1_corr$r)
keeping <- hpk1_matrix$MAP4K1 >= 0.5
str(keeping)
hpk1_filter <- hpk1_matrix[keeping,keeping]
str(hpk1_filter)
head(hpk1_filter)
hpk1_filter %<>% mutate(gene= rownames(hpk1_filter))
hpk1_filter <-hpk1_filter[,c(ncol(hpk1_filter),1:(ncol(hpk1_filter)-1))]
write.table(hpk1_filter, file="Corr_sort_PBMC_filter.txt", quote= FALSE, sep="\t", row.names=FALSE)
```

