---
title: "Deconvolution_scRNA_evaluation.Rmd"
author: "MX"
date: "4/7/2021"
output: html_document
---

```{r}
library(Seurat)
library(data.table)
library(magrittr)
library(dplyr)
library(tidyverse)
library(ggplot2)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
```

Data objects for this analysis:
```{r}
Actual_cell_percents # cell types by donor
pseudobulk_m # pseudo-bulk data for 8 donors, in matrix format for immunedeconv 
as.matrix(pseudobulk)
```




Install Deconvolution software

https://icbi-lab.github.io/immunedeconv/
install.packages("remotes")
remotes::install_github("icbi-lab/immunedeconv")

Load scRNA datasets, calculate cell types per sample
```{r}
load("/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/Cellranger3.1.0_batch4/batch4B.Robj")
```


Make a table of immune cell content by donor

```{r}
cluster_ident2 <- data.frame(Ident.ident=as.factor(c(0:27)), Cell_type=c("CD4 T", "CD4 T", "Neutrophil","Neutrophil","CD8 T","Neutrophil Apoptotic","NK","Monocyte","Neutrophil","Apoptotic Lymphoid","Apoptotic monocyte","Monocyte","mMDSC","Neutrophil Apoptotic","Platelets","Monocyte","Apoptotic NK","B cell","CD8 T","CD8 T","gMDSC","Neutrophil","DC","B cell","NK","Monocyte","CD4 T","DC"))

cbind(Ident=FetchData(batch4B, vars= c("ident", "orig.ident"))) %>% data.frame %>% left_join(cluster_ident2, by="Ident.ident") %>% dselect(-Ident.ident) %>% table
```

Count T regs by Foxp3 expression alone
There are only 105 CD4+ FOXP3+ cells in the entire dataset. Leave T regs out of the analysis. We know the decovonvolution software are terrible at predicting them anyway
Gene defined CTLA4, IKZF2, CCR8 and FOXP3, but CCR8 is not detected, IKZF2 is non-specific, and CTLA4 looks similar to FOXP3
```{r fig.height=10, fig.width=10}
FeaturePlot(batch4B,reduction="umap",slot ="data",cols=c("grey90","red"),features = c( "CTLA4","IKZF2","CCR8","FOXP3","LAG3"), ncol=2, max.cutoff = 1) 
# Fetch cells with FOXP3 expression that are not in the CD8 cluster
Treg <- FetchData(batch4B, vars= c("FOXP3","ident", "orig.ident"), slot = "data") %>% dfilter(FOXP3>0, ident != c(1,0,26)) %>% rownames#%>% count(orig.ident) 
```

Merge Treg counts into the rest of the cells by replacing by cell IDS

```{r fig.height=3, fig.width=10}
cell_type_counts <- cbind(Ident=FetchData(batch4B, vars= c("ident", "orig.ident"))) %>% data.frame %>% rownames_to_column("Cell") %>% left_join(cluster_ident2, by="Ident.ident") %>% mutate(Cell_type= if_else(Cell %in% Treg,"Treg",Cell_type)) %>% dfilter(!is.na(Cell_type)) %>% dselect(-Ident.ident) %>% count(Ident.orig.ident,Cell_type, .drop=FALSE) # group_by(Ident.orig.ident,Cell_type) %>% summarise(n=n()) 
#this skips B cell in aml1 because there are none

cell_type_counts$Cell_type %>% unique
cell_order <- rev(c("Neutrophil","Neutrophil Apoptotic","CD4 T","Treg","Monocyte","CD8 T","mMDSC","Platelets","NK","B cell","gMDSC","DC"))
cell_type_counts %>% mutate(Cell_type= factor(Cell_type, levels=cell_order)) %>% group_by(Ident.orig.ident) %>% mutate(Total_cells=sum(n)) %>% ungroup %>% mutate(Immune_cell_percent= round((n/Total_cells)*100,digits=1)) %>% ggplot(aes(x=Ident.orig.ident, fill=Cell_type,y=Immune_cell_percent)) + 
  geom_bar(position="stack",stat="identity") + coord_flip() + 
  theme(text = element_text(size=18)) +
scale_fill_manual(values=rev(c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920","#083D77"))) #,"#356E46","#69597D","#052344"

#Cell proportions per donor data
Actual_cell_percents <- cell_type_counts %>% group_by(Ident.orig.ident) %>% mutate(Total_cells=sum(n)) %>% ungroup %>% mutate(Immune_cell_percent= round((n/Total_cells)*100,digits=1))
```

Pool mRNA counts per donor into pseudo-bulk data

pull RNA counts out and transpose, pull orig.idents out
Merge into a data.frame of cells in rows and genes in columns
Use data.table to group sum counts by donor
transpose back into genes in rows and samples in columns
```{r}
raw_counts <- as.matrix(batch4B@assays$RNA@counts) %>% t
raw_counts_dt[1:5,1:5]
#This is too much for data.frame
pseudo_bulk <- raw_counts %>% t %>% data.frame %>% tibble::rownames_to_column("donor") %>% group_by(donor) %>% summarise_all(sum)
library(data.table)
# cbind changes the numbers in raw_counts into characters, apparently this is default of cbind, so join using data.frame, then convert to data.table
raw_counts_dt <- data.frame(donor=batch4B$orig.ident,raw_counts) %>% data.table

pseudo_bulk <- raw_counts_dt[,lapply(.SD,sum),by=donor] #.SD is like the summarise_all function, it contains all columns except the column specified in by=
pseudobulk <- pseudo_bulk[1:8,] %>% tibble::column_to_rownames("donor") %>% t %>% data.frame %>% tibble::rownames_to_column("Gene")
save(pseudobulk, file="batch4B_pseudobulk_RNAseq.rdata")
save(pseudobulk_m, file="batch4B_pseudobulk_RNAseq_m.rdata")
pseudobulk_m <- pseudobulk %>% tibble::column_to_rownames("Gene") %>% as.matrix
```


```{r}
UMAPPlot(batch4B, label=TRUE)
c("CD4 T", "CD4 T", "Neutrophil","Neutrophil","CD8 T","Neutrophil Apoptotic","NK","Monocyte","Neutrophil","Apoptotic Lymphoid","Apoptotic monocyte","Monocyte","mMDSC","Neutrophil Apoptotic","Platelets","Monocyte","Apoptotic NK","B cell","CD8 T","CD8 T","gMDSC","Neutrophil","DC","B cell","NK","Monocyte","CD4 T","DC")
```

Deconvolute with  library(immunedeconv)

"quantiseq" : percent content that is comparable intra-sample and inter-sample
"mcp_counter" : arbitrary number unit estimate that is comparable inter-sample
"xcell" : produces enrichment scores that are comparable inter-sample, not intra-sample
 "epic" 
 "timer"

"cibersort"  "cibersort_abs"         
!Cibersort requires a commercial license so skip that one

 pbmc1  pbmc2    pdac1    pdac2     aml1  pbmc3 aml3

```{r}
library(immunedeconv)
res_Quantiseq <- deconvolute(pseudobulk_m,"quantiseq",tumor=FALSE) # Signature genes found in data set: 148/153 (96.73%)
res_mcp <- deconvolute(pseudobulk_m,"mcp_counter") 
res_epic <- deconvolute_epic(pseudobulk_m, scale_mrna = TRUE,tumor= FALSE)
res_xcell <- deconvolute_xcell(pseudobulk_m,array=FALSE) # cell.types.use can't find options for this anywhere on the original xcell code
res_timer <- deconvolute_timer(pseudobulk_m, indications=c("UVM","UVM","PAAD","PAAD","ACC","UVM","PAAD","ACC"))  #needs the cancer type for each case, which is stupid. say UVM uveal melanoma for healthy since it's a cold tumor type?

```
QuantiSeq evaluate
Predictions for Macrophage M1 = 0 for all, which is correct!
```{r fig.height=3, fig.width=11}

cell_order_qt <- rev(c("Neutrophil","T cell CD4+ (non-regulatory)","T cell regulatory (Tregs)","Monocyte","T cell CD8+","Macrophage M2","Macrophage M1","NK cell","B cell","uncharacterized cell","Myeloid dendritic cell"))
res_Quantiseq %>% gather(Donor,Prediction_percent, -cell_type) %>% mutate(cell_type= factor(cell_type, levels=cell_order_qt)) %>% 
  ggplot(aes(x=Donor, fill=cell_type,y=Prediction_percent)) + 
  geom_bar(position="stack",stat="identity") + 
  coord_flip() + 
  theme(text = element_text(size=18)) +
  scale_fill_manual(values=rev(c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920","#083D77")))


```
Compare relative estimates
```{r}
res_Quantiseq_groom <- res_Quantiseq %>% gather(Donor,Prediction_percent, -cell_type) %>% mutate(cell_type =case_when(
  grepl(cell_type  "T cell CD4+ (non-regulatory)") ~ "CD4 T",
  cell_type == "T cell regulatory (Tregs)" ~ "Treg",
  cell_type == "T cell CD8+" ~ "CD8 T",
  cell_type == "Macrophage M2" ~ "mMDSC",
  cell_type == "NK cell" ~ "NK",
  cell_type == "Myeloid dendritic cell" ~ "DC",
  TRUE ~ cell_type
  ))

res_mcp_groom <- res_mcp %>% gather(Donor,Enrichment_score, -cell_type) %>% mutate(cell_type =case_when(
  cell_type == "T cell CD8+" ~ "CD8 T",
  cell_type == "NK cell" ~ "NK",
  cell_type == "Myeloid dendritic cell" ~ "DC",
  TRUE ~ cell_type
  ))

res_epic_groom <- res_epic %>% data.frame %>% tibble::rownames_to_column("cell_type") %>% gather(Donor,Enrichment_score, -cell_type) %>% mutate(cell_type =case_when(
  cell_type == "CD4_Tcells" ~ "CD4 T",
  cell_type == "CD8_Tcells" ~ "CD8 T",
  cell_type == "Monocytes" ~ "Monocyte",
  cell_type == "NKcells" ~ "NK",
  cell_type == "Neutrophils" ~ "Neutrophil",
  cell_type == "Bcells" ~ "B cell",
  TRUE ~ cell_type
  ))

res_xcell_groom <- res_xcell %>% data.frame %>% tibble::rownames_to_column("cell_type") %>% gather(Donor,Enrichment_score, -cell_type) %>% mutate(cell_type =case_when(
  grepl("CD4+",cell_type) ~ "CD4 T",
  grepl("CD8+",cell_type) ~ "CD8 T",
  cell_type == "Tregs" ~ "Treg",
  cell_type == "Monocytes" ~ "Monocyte",
  cell_type == "Macrophages M2" ~ "mMDSC",
  cell_type == "NK cells" ~ "NK",
  grepl("DC", cell_type) ~ "DC",
  cell_type == "Neutrophils" ~ "Neutrophil",
  grepl("B-cells",cell_type) ~ "B cell",
  TRUE ~ cell_type
  )) %>% group_by(cell_type, Donor) %>% summarise_if(is.numeric,sum) %>% ungroup %>% unique

res_timer_groom <- res_timer %>% data.frame %>% tibble::rownames_to_column("cell_type") %>% gather(Donor,Enrichment_score, -cell_type) %>% mutate(cell_type =case_when(
  grepl("T_cell.CD4",cell_type) ~ "CD4 T",
  grepl("T_cell.CD8",cell_type) ~ "CD8 T",
  cell_type == "Macrophage" ~ "Monocyte",
  grepl("B_cell",cell_type) ~ "B cell",
  TRUE ~ cell_type
  ))
```
```{r}
Actual_cell_percents %<>% mutate(Donor = Ident.orig.ident, cell_type=Cell_type, Cell_percent= Immune_cell_percent/100) %>% dselect(Donor,cell_type,Cell_percent)
```

Correlate QunatiSeq results

```{r fig.height=6, fig.width=6}
library(ggpubr)
#log
res_Quantiseq_groom %>% left_join(Actual_cell_percents, by=c("Donor","cell_type")) %>% dfilter(cell_type %nin% c("Macrophage M1","uncharacterized cell")) %>%
  ggplot(aes(Cell_percent,Prediction_percent)) + 
  geom_point()  +
  geom_abline(slope=1) + 
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~cell_type, scale="fixed", ncol=3) +
  stat_cor(label.y = -0.3) # +
  stat_regline_equation(label.y = 3.5)
#linear
res_Quantiseq_groom %>% left_join(Actual_cell_percents, by=c("Donor","cell_type")) %>% dfilter(cell_type %nin% c("Macrophage M1","uncharacterized cell")) %>%
  ggplot(aes(Cell_percent,Prediction_percent)) + 
  geom_point()  +
  geom_abline(slope=1) + 
  facet_wrap(~cell_type, scale="fixed", ncol=3) +
  stat_cor(label.y = 0.5) # +
  stat_regline_equation(label.y = 3.5)

```

Correlate MCP

```{r fig.height=6, fig.width=9}
library(ggpubr)
#log
res_mcp_groom %>% left_join(Actual_cell_percents, by=c("Donor","cell_type")) %>% dfilter(!is.na(Cell_percent)) %>%
  ggplot(aes(Cell_percent,Enrichment_score)) + 
  geom_point()  +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~cell_type, scale="fixed", ncol=3) +
  stat_cor(label.y = 1) # +
  stat_regline_equation(label.y = 1.5)
#linear
res_mcp_groom %>% left_join(Actual_cell_percents, by=c("Donor","cell_type")) %>% dfilter(cell_type %nin% c("Macrophage M1","uncharacterized cell")) %>%
  ggplot(aes(Cell_percent,Enrichment_score)) + 
  geom_point()  +
  geom_abline(slope=1) + 
  facet_wrap(~cell_type, scale="fixed", ncol=3) +
  stat_cor(label.y = 0.5) # +
  stat_regline_equation(label.y = 3.5)

```

Correlate EPIC

```{r fig.height=6, fig.width=9}
res_epic_groom %>% left_join(Actual_cell_percents, by=c("Donor","cell_type")) %>% dfilter(!is.na(Cell_percent)) %>%
  ggplot(aes(Cell_percent,Enrichment_score)) + 
  geom_point()  +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope=1) + 
  facet_wrap(~cell_type, scale="fixed", ncol=3) +
  stat_cor(label.y = 1) # +
  stat_regline_equation(label.y = 1.5)

#linear
res_epic_groom %>% left_join(Actual_cell_percents, by=c("Donor","cell_type")) %>% dfilter(!is.na(Cell_percent)) %>%
  ggplot(aes(Cell_percent,Enrichment_score)) + 
  geom_point()  +
  geom_abline(slope=1) + 
  facet_wrap(~cell_type, scale="fixed", ncol=3) +
  stat_cor(label.y = 0.7) # +
  stat_regline_equation(label.y = 3.5)
```

xcell correlate 

```{r fig.height=8, fig.width=12}
res_xcell_groom %>% left_join(Actual_cell_percents, by=c("Donor","cell_type")) %>% dfilter(!is.na(Cell_percent)) %>% mutate(if_else(Enrichment_score==0, 0.0001,Enrichment_score)) %>%
  ggplot(aes(Cell_percent,Enrichment_score)) + 
  geom_point()  +
  scale_x_log10() +
  scale_y_log10() +
  #geom_abline(slope=1) + 
  facet_wrap(~cell_type, scale="free", ncol=4) +
  stat_cor(label.y = 1, method="pearson") # +
  stat_regline_equation(label.y = 1.5)

```
TIMER

```{r fig.height=6, fig.width=9}
res_timer_groom %>% left_join(Actual_cell_percents, by=c("Donor","cell_type")) %>% dfilter(!is.na(Cell_percent)) %>% mutate(if_else(Enrichment_score==0, 0.0001,Enrichment_score)) %>%
  ggplot(aes(Cell_percent,Enrichment_score)) + 
  geom_point()  +
  scale_x_log10() +
  scale_y_log10() +
  #geom_abline(slope=1) + 
  facet_wrap(~cell_type, scale="fixed", ncol=3) +
  stat_cor(label.y = 1, method="pearson") # +
  stat_regline_equation(label.y = 1.5)
```

Signatures
Treg
```{r fig.height=4, fig.width=12}
library(limma)

pseudobulk_m_norm <- normalizeBetweenArrays(pseudobulk_m)
boxplot(pseudobulk_m_norm)
pseudobulk_z <- t(apply(pseudobulk_m_norm,1,mx_zscore2)) #the middle term 1 means rows, and 2 means columns, the apply() will transpose the matrix because it has to do that before applying calculations by column, the second t is to flip it back
boxplot(pseudobulk_z)

Treg_sig_score <- pseudobulk_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("FOXP3","CTLA4")) %>% summarise_if(is.numeric, mean) %>% t %>% data.frame(Treg_score_FOXP3_CTLA4 = .) %>% tibble::rownames_to_column("Donor")
Treg_sig_scor2 <- pseudobulk_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("CTLA4","IKZF2","FOXP3")) %>% summarise_if(is.numeric, mean) %>% t %>% data.frame(Treg_score2_FOXP_CTLA_IKZ = .) %>% tibble::rownames_to_column("Donor")

FOXP3_z_score <- pseudobulk_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("FOXP3")) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame(FOXP3_score = .) %>% tibble::rownames_to_column("Donor")

t_reg <- Actual_cell_percents$cell_type == "Treg"

Actual_cell_percents[t_reg,] %>% left_join(Treg_sig_score, by="Donor") %>% left_join(FOXP3_z_score, by="Donor") %>% left_join(Treg_sig_scor2, by="Donor")
  
Actual_cell_percents[t_reg,] %>% left_join(Treg_sig_score, by="Donor") %>% left_join(FOXP3_z_score, by="Donor") %>% left_join(Treg_sig_scor2, by="Donor") %>% dselect(-Donor,-cell_type) %>% gather(Signature,Scores, -Cell_percent) %>%
  ggplot(aes(Cell_percent,Scores)) + 
  geom_point()  +
  scale_x_log10() +
  #geom_abline(slope=100) +
  facet_wrap(~Signature, ncol=3, scale="fixed") +
  stat_cor(label.y = 2) # +
  stat_regline_equation(label.y = 1.5)

```

Merge All tools results

```{r}
library(brewer)
res_Quantiseq_groom %<>% mutate(Tool ="QuantiSeq")
res_Quantiseq_groom %<>% rename("Enrichment_score"=Prediction_percent)
res_mcp_groom %<>% mutate(Tool= "MCP")
res_epic_groom %<>% mutate(Tool= "Epic")
res_xcell_groom %<>% mutate(Tool= "xCell")
res_timer_groom %<>% mutate(Tool= "TIMER")
colnames(res_Quantiseq_groom) == colnames(res_mcp_groom)
all_dc <- rbind(res_Quantiseq_groom,res_mcp_groom,res_epic_groom,res_xcell_groom,res_timer_groom,match.names=F)

all_dc %>% dfilter(cell_type %in% c("Neutrophil","CD4 T","Treg","Monocyte","CD8 T","mMDSC","Platelets","NK","B cell","gMDSC","DC")) %>% left_join(Actual_cell_percents, by=c("cell_type","Donor")) %>% mutate(Cell_percent= if_else(is.na(Cell_percent),0,Cell_percent)) %>% dselect(-Donor) %>% group_by(cell_type,Tool) %>% summarise(Cor = cor(Enrichment_score,Cell_percent, method="spearman")) %>% ungroup %>% mutate(Tool=factor(Tool, levels=rev(c("QuantiSeq","xCell","Epic","MCP","TIMER")))) %>% mutate(cell_type=factor(cell_type, levels=c("Neutrophil","B cell","CD4 T","CD8 T","Treg","NK","Monocyte","DC","mMDSC","Platelets"))) %>%
  ggplot(aes(x=cell_type, y=Tool, fill=Cor)) + geom_tile() + scale_fill_distiller(limits=c(0,1), palette="RdBu")
```

