---
title: "Dotmatics2Vortex"
author: "Mengshu"
date: "12/10/2019"
output: html_document
---

```{r}
library(dplyr)
library(tidyverse)
library(stringr)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
```

###########
Parsing Plan
###########
Common columns:
ID, Reg ID, SMILES

0. General parsing, fill in rows, groom numbers, filter out assays that we don't need in the output
1. Percent Protein Bound Assay -> PPB Species PCT_Free, 2 columns
2. DMPK -> AUC, %F, Half Life, by Route+Dose+Strain, 76 columns
3. Biochemical assays: 15 columns
4. Biochemical assay "Maximum": only for assays listed as "IL-2", 2 columns
5. Liver Microsome Half life -> Average Half life by species, 3 columns
Total: 98 data columns

read.table does not interpret excel encoding correctly
readxl::read_excel works
```{r}
dot <- readxl::read_excel("/Volumes/Enterprise/FLX/HPK1/Dotmatics2Vortex/19_08_23_HPK1_download_raw.xlsx")
dot2 <- readxl::read_excel("/Volumes/Enterprise/FLX/HPK1/Dotmatics2Vortex/MyTabularView-60HSW7W7.xlsx")
dot3 <- readxl::read_excel("/Volumes/Enterprise/FLX/HPK1/Dotmatics2Vortex/excelexport-CRIVA2OI.xlsx")
dot4 <- readxl::read_excel("/Volumes/Enterprise/FLX/HPK1/Dotmatics2Vortex/excelexport-011020.xlsx")
```

```{r}

dot_cleanup <- dot4 %>% rename("Reg_ID" = `Reg ID`) %>% fill(Reg_ID, SMILES) %>% mutate(Result_num = ifelse(startsWith(Result,'>'), 10*as.numeric(substr(Result,2,nchar(Result))), ifelse(startsWith(Result,'<'), 0.1*as.numeric(substr(Result,2,nchar(Result))), as.numeric(Result)))) %>% dplyr::rename(Species_LM=`Species...7`, Species_PPB=SPECIES, Strain_dmpk=Strain) %>% dplyr::select(-starts_with("Result "), -Result) %>% dfilter(
Validated == 'VALIDATED',
`Protocol Name` != "Library Screening",
!str_detect(`Protocol Name`,"RBC$"),
!str_detect(`Protocol Name`,"^HRI"),
!str_detect(`Protocol Name`,"^PERK"),
!str_detect(`Protocol Name`,"^PKR"),
!str_detect(`Protocol Name`,"^Cytotoxicity"),
!str_detect(`Protocol Name`,"SKOV"),
!str_detect(`Protocol Name`,"^GSK"),
!str_detect(`Protocol Name`,"^HPK1 KO_"),
!str_detect(`Protocol Name`,"^HPK1 WT Clone"), 
!str_detect(`Protocol Name`, "LanthaScreen$"))


### Meta data
dot_meta <- dot_cleanup %>% dselect(ID,Reg_ID,SMILES) %>% separate(ID, into=c("ID",NA),sep= "-") %>% unique

##### PPB section
dot_PPB <- dot_cleanup %>% dselect(Reg_ID,SMILES,Species_PPB,AVG_PCT_FREE) %>% group_by(Reg_ID,Species_PPB) %>% mutate(GeoMean_PCT_Free = geo.mean(AVG_PCT_FREE)) %>% ungroup() %>% dselect(-AVG_PCT_FREE) %>% unique %>% pivot_wider(names_from=Species_PPB, values_from=GeoMean_PCT_Free, names_prefix="PCT_Free_") %>% dselect(Reg_ID,PCT_Free_Rat,PCT_Free_Human)

#### DMPK results : select coluns by name instead of index in case the export is slightly different order in future
#### Creates 68 columns
dot_dmpk <- dot_cleanup %>% dfilter(!is.na(Strain_dmpk)) %>% dselect(Reg_ID,`Strain_dmpk`,`CL (L/hr/kg)`,`%F`,`T1/2 (hr)`,`AUC (0-infinity) (hr*ng/mL)`,`Route`,`Dose (mg/kg)`) %>% group_by(Reg_ID,`Strain_dmpk`,`Route`,`Dose (mg/kg)`) %>% mutate_if(is.numeric, geo.mean) %>% unique  %>% pivot_wider(names_from=c(Strain_dmpk,Route,`Dose (mg/kg)`), values_from=c(`AUC (0-infinity) (hr*ng/mL)`,`%F`,`CL (L/hr/kg)`,`T1/2 (hr)`))
 
 ## In vitro Assays
## 13 columns
dot_invitro <- dot_cleanup %>% dfilter(Validated == "VALIDATED",`Protocol Name` != "Microsomal") %>% dselect(Reg_ID,`Protocol Name`,Result_num) %>% group_by(Reg_ID,`Protocol Name`) %>% mutate_if(is.numeric, geo.mean) %>% ungroup %>% unique %>% pivot_wider(names_from=`Protocol Name`, values_from= Result_num)

#In Vitro: Get Maximum values from the IL2 assays
dot_maxima <- dot_cleanup %>% dplyr::filter(grepl("IL-2",`Protocol Name`),Validated == 'VALIDATED') %>% dselect(Reg_ID,`Protocol Name`,Maximum) %>% mutate(Maximum = as.numeric(Maximum)) %>% group_by(Reg_ID,`Protocol Name`) %>% mutate_if(is.numeric, geo.mean) %>% ungroup %>% unique %>% pivot_wider(names_from=`Protocol Name`, values_from= Maximum, names_prefix="Maximum_")
  
#Liver Microsome
dot_LM <- dot_cleanup %>% dfilter(!is.na(Species_LM)) %>% dselect(Reg_ID,Species_LM,`Avg Half Life (min)`) %>% group_by(Reg_ID,Species_LM) %>% mutate_if(is.numeric,geo.mean) %>% ungroup %>% unique %>% pivot_wider(names_from = c(`Species_LM`), values_from = `Avg Half Life (min)`, names_prefix = "LM_HalfLife_")

#Join everything
rm(vortex)
vortex <- left_join(dot_meta,dot_PPB,by="Reg_ID")
vortex <- left_join(vortex,dot_dmpk,by="Reg_ID")
vortex <- left_join(vortex,dot_invitro,by="Reg_ID")
vortex <- left_join(vortex,dot_maxima,by="Reg_ID")
vortex <- left_join(vortex,dot_LM,by="Reg_ID")
# 101 columns total
#1101 compound as of Jan 10 2020
vortex %>% write.table(file="/Volumes/Enterprise/FLX/HPK1/Dotmatics2Vortex/Dot2Vortex_01102020.txt", sep="\t",col.names = TRUE, row.names = FALSE, quote=FALSE)
vortex %>% write.table(file="/Volumes/Enterprise/FLX/HPK1/Dotmatics2Vortex/Dot2Vortex_01102020_Happy_Rubber_Ducky_Day.txt", sep="\t",col.names = TRUE, row.names = FALSE, quote=FALSE,na = "")
```

