---
title: "GCN2_survival_analysis_Gene"
author: "Gene_MX"
date: "1/30/2020"
output: html_document
---

```{r}
library(tidyverse)
# survival analysis package
library(icenReg)
#package for regression models for interval censored data
install.packages("icenReg")
```

##################################################
#######
Datasets

```{r}
19
CT26_004 <- read.table(file="/Volumes/Enterprise/FLX/GCN2/In_vivo_data/19-CT26-004.txt", header=TRUE, sep="\t")
CT26_021 <- read.table(file="/Volumes/Enterprise/FLX/GCN2/In_vivo_data/19-CT26-021.txt", header=TRUE, sep="\t")
CT26_028 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/19-CT26-028_28d.txt", header=TRUE, sep="\t")
Renca_05 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/19-Renca-005.txt", header=TRUE, sep="\t")
Renca_08 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/19-Renca-008_29d.txt", header=TRUE, sep="\t")
Crown_LL2 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/Crown_LL2.txt", header=TRUE, sep="\t")
Crown_Pan02 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/Crown_Pan02_Comp.txt", header=TRUE, sep="\t")
RENCA_001_20 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/20_RENCA_001.txt", header=TRUE, sep="\t")
Crown_01_20 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/20_Crown_LL2.txt", header=TRUE, sep="\t")
RENCA_004_21 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/21-Renca-004.txt", header=TRUE, sep="\t")
CT26_005 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/19-CT26-005.txt", header=TRUE, sep="\t")

CT26_004 <- CT26_004 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) %>% 
	filter(complete.cases(.))

CT26_021 <- CT26_021 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) %>% 
	filter(complete.cases(.))

CT26_028 <- CT26_028 %>% gather("day","size", -Group, -animal) %>% 
	rename(treatment=Group) %>%
	mutate(day=as.numeric(sub("X","",day))) #%>% 
	filter(complete.cases(.))

Renca_05 <- Renca_05 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) %>% 
	filter(complete.cases(.))

Renca_08 <- Renca_08 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) #%>% 
	filter(complete.cases(.))

Crown_LL2 <- Crown_LL2 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) %>% 
	filter(complete.cases(.))

Crown_Pan02 <- Crown_Pan02 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) 

RENCA_001_20 <- RENCA_001_20 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) #
	
Crown_01_20 <- Crown_01_20 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) %>% dfilter(!is.na(animal)) %>% dfilter(treatment %in% c("Anti_VEGFR","F019282_50_mpk_Anti_VEGFR","Vehicle","F019282_50_mpk"))

Crown_01_20 %<>% dfilter(treatment %in% c("Anti_VEGFR","F019282_50_mpk_Anti_VEGFR","Vehicle","F019282_50_mpk"))

RENCA_004_21 <- RENCA_004_21 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day)))

CT26_005 <- CT26_005 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) 

ggplot(CT26_004, aes(x=day, y=size, color=treatment, group=interaction(treatment, animal))) + geom_line()
ggplot(CT26_021, aes(x=day, y=size, color=treatment, group=interaction(treatment, animal))) + geom_line()
ggplot(CT26_028, aes(x=day, y=size, color=treatment, group=interaction(treatment, animal))) + geom_line()
ggplot(Renca_05, aes(x=day, y=size, color=treatment, group=interaction(treatment, animal))) + geom_line()
ggplot(Renca_08, aes(x=day, y=size, color=treatment, group=interaction(treatment, animal))) + geom_line()
#ggplot(MM1S.vol, aes(x=day, y=size, color=treatment, group=animal)) + geom_line() + facet_wrap(~treatment)
```



Load Prism file: Apparently the Excel file for 21-RENCA-004 is wrong
```{r}
#install.packages("pzfx")
library(pzfx)
RENCA_004_21 <- read_pzfx("/Volumes/Picard/FLX/GCN2/In_vivo_data/21-Renca-004.pzfx") %>% data.frame %>% dfilter(!is.na(Days.Post.Inoculation)) %>% tibble::column_to_rownames("Days.Post.Inoculation") %>% t %>% data.frame %>% mutate(treatment=c(rep("Vehicle",10),rep("Anti-PD1",10),rep("F019282_15mpk",10),rep("F019282_15mpk_PDL1",10))) %>% dselect(treatment,everything()) %>% gather(day, size,-treatment) %>% mutate(day=as.numeric(sub("X","",day)))
```


Find the tumor size that half the tumors reach, to set as threshold
CT26-004 threshold 1000, reached by 31/48 samples: 65%
CT26-021 threshold 1000, reached by 15/15 samples: 100%
CT26-028 threshold 1000, reached by 25/40 samples: 62.5%
RENCA-08 threshold 1000, reached by 26/69 samples: 38%
```{r}
CT26_004 %>% group_by(treatment,animal) %>% dplyr::filter(day==max(day)) %>% dplyr::filter(.data$size>=1000)

CT26_021 %>% group_by(treatment,animal) %>% dplyr::filter(day==max(day)) %>% dplyr::filter(.data$size>=1000)

CT26_028 %>% group_by(treatment,animal) %>% dplyr::filter(day==max(day)) %>% dplyr::filter(.data$size>=2000)

Renca_08 %>% group_by(treatment,animal) %>% dplyr::filter(day==max(day)) %>% dplyr::filter(.data$size>=1000)

RENCA_001_20 %>% group_by(treatment,animal) %>% dplyr::filter(day==max(day)) %>% dplyr::filter(.data$size>=1000)
```





```{r, width=7, height=4}

ggplot(CT26_004, aes(x=day, y=size, group=interaction(treatment,animal))) + geom_line(show.legend = FALSE, aes(color=treatment)) +
	facet_wrap(~treatment) + geom_hline(aes(yintercept=1500, color="grey80")) + theme(legend.position="none") + ggsave("CT26_04_facet.jpg", width=5, height=3, dpi=150, plot= last_plot(), units = "in")

ggplot(CT26_021, aes(x=day, y=size, group=interaction(treatment,animal))) + geom_line(show.legend = FALSE, aes(color=treatment)) +
	facet_wrap(~treatment) + geom_hline(aes(yintercept=1000, color="grey80")) + theme(legend.position="none") + ggsave("CT26_21_facet.jpg", width=5, height=3, dpi=150, plot= last_plot(), units = "in")

ggplot(CT26_028, aes(x=day, y=size, group=interaction(treatment,animal))) + geom_line(show.legend = FALSE, aes(color=treatment)) +
	facet_wrap(~treatment) + geom_vline(aes(xintercept=24, color="grey80")) + theme(legend.position="none") # + ggsave("CT26_28_facet.jpg", width=5, height=3, dpi=150, plot= last_plot(), units = "in")


ggplot(Renca_05, aes(x=day, y=size, color=treatment, group=interaction(treatment,animal))) + geom_line(show.legend = FALSE) +
	facet_wrap(~treatment) + geom_hline(aes(yintercept=1000, color="grey80")) + theme(legend.position="none") + ggsave("Renca_05_facet.jpg", width=5, height=3, dpi=150, plot= last_plot(), units = "in")

ggplot(Renca_08, aes(x=day, y=size, color=treatment, group=interaction(treatment,animal))) + geom_line(show.legend = FALSE) +
	facet_wrap(~treatment) + geom_hline(aes(yintercept=500, color="grey80")) + theme(legend.position="none") + ggsave("Renca_08_facet.jpg", width=5, height=3, dpi=150, plot= last_plot(), units = "in")

ggplot(Crown_LL2, aes(x=day, y=size, color=treatment, group=interaction(treatment,animal))) + geom_line(show.legend = FALSE) +
	facet_wrap(~treatment) + geom_hline(aes(yintercept=2000, color="grey80")) + theme(legend.position="none") + ggsave("Crown_LL2_facet.jpg", width=5, height=3, dpi=150, plot= last_plot(), units = "in")

ggplot(Crown_Pan02, aes(x=day, y=size, color=treatment, group=interaction(treatment,animal))) + geom_line(show.legend = FALSE) +
	facet_wrap(~treatment) + geom_vline(aes(xintercept=32, color="grey80")) + theme(legend.position="none") + ggsave("Crown_Pan02_facet.jpg", width=5, height=3, dpi=150, plot= last_plot(), units = "in")

ggplot(RENCA_001_20, aes(x=day, y=size, color=treatment, group=interaction(treatment,animal))) + geom_line(show.legend = FALSE) +
	facet_wrap(~treatment) + geom_hline(aes(yintercept=500, color="grey80")) + theme(legend.position="none") 

ggplot(Crown_01_20, aes(x=day, y=size, color=treatment, group=interaction(treatment,animal))) + geom_line(show.legend = FALSE) +
	facet_wrap(~treatment) + geom_hline(aes(yintercept=1000, color="grey80")) + theme(legend.position="none") 
```

# make survival data functions
```{r}
# create a data.frame that is ready to go into survival analysis
# cutoff = size cutoff for suvival
# this returns the "death" days as left and right censored
make_survival_data <- function(x, cutoff=2000) {
	x %>% group_by(treatment, animal) %>% arrange(day) %>% summarize(
		which.cutoff = as.integer(min(which(size >= cutoff))), # the actual cutoff day
		reaches.cutoff = is.finite(which.cutoff),   # boolean on whehther it ever reaches 1000
		left = if_else(reaches.cutoff, day[which.cutoff-1], max(day)), # left bound: -1 day when it reaches
		right = if_else(reaches.cutoff, day[which.cutoff], Inf)   # right bound, the day it reaches, or infinity if never reaches
	) %>% ungroup %>% select(-which.cutoff, -reaches.cutoff)
}

# left = interval before survival event
# right = interval after survival event
# make a data.frame that is ready to plot as a survival curve
make_surv_curve_data <- function(x, cutoff=2000) {
	x %>% group_by(treatment) %>% mutate(n.day1=sum(day==min(day))) %>%
		group_by(treatment, day) %>% summarize(frac_alive = sum(size < cutoff)/n.day1[1])
}

```
Make survival data input

```{r}
current_data <- CT26_004 %>% make_survival_data(cutoff=1500)
current_data <- CT26_021 %>% make_survival_data(cutoff=1000)
current_data <- CT26_028 %>% make_survival_data(cutoff=2000)
current_data <- Renca_05 %>% make_survival_data(cutoff=1000)
current_data <- Renca_08 %>% make_survival_data(cutoff=500)
current_data <- Crown_Pan02 %>% make_survival_data(cutoff=1500)
current_data <- Crown_LL2 %>% make_survival_data(cutoff=1000)
current_data <- RENCA_001_20 %>% make_survival_data(cutoff=500)
current_data <- Crown_01_20 %>% make_survival_data(cutoff=2000) #1000 not sig

current_data <- current_data %>% mutate(treatment=fct_relevel(factor(treatment), "19282_15_mpk"))
```

```{r fig.height=4, fig.width=7}
plot.data <- Crown_Pan02  %>% make_surv_curve_data
plot.data <- CT26_004  %>% make_surv_curve_data
plot.data <- CT26_021  %>% make_surv_curve_data
plot.data <- CT26_028  %>% make_surv_curve_data
plot.data <- Renca_05  %>% make_surv_curve_data
plot.data <- Renca_08  %>% make_surv_curve_data(cutoff=500)
plot.data <- Crown_Pan02  %>% make_surv_curve_data(cutoff=1500)
plot.data <- Crown_LL2  %>% make_surv_curve_data
plot.data <- RENCA_001_20 %>% make_surv_curve_data
plot.data <- Crown_01_20 %>% make_surv_curve_data
ggplot(plot.data, aes(x=day, y=frac_alive, color=treatment)) + geom_step(size=2) # ggsave("Crown_Pan02_survival.jpg", width=7, height=4, dpi=150, plot= last_plot(), units = "in")

```

# test different fit models for survival
```{r}
#tmp <- MM1S.surv_data %>% filter(treatment %in% c("18797.50mpkBID","Vehicle"))
#s <- Surv(tmp$left, tmp$right, type = "interval2")

#non_para <- ic_np(cbind(left,right) ~ treatment, data=MM1S.surv_data)
ph <- ic_sp(cbind(left,right) ~ treatment, model='ph', bs_samples=100, data=current_data) 
#po <- ic_sp(cbind(left,right) ~ treatment, model='po', bs_samples=100, data=MM1S.surv_data)
#par <- ic_par(cbind(left,right) ~ treatment, model='ph', dist="gamma", data=MM1S.surv_data)
#par.ph_weib <- ic_par(cbind(left,right) ~ treatment, model='ph', dist="weibull", data=MM1S.surv_data)

#non_para %>% plot(.,  xlab="Time", ylab="Surv", main="non parametric")
ph %>% plot(.,  xlab="Time", ylab="Surv", main="cox ph")
#po  %>% plot(., xlab="Time", ylab="Surv", main="prop odds")
#par  %>% plot(., xlab="Time", ylab="Surv", main="parametric cox ph")
#par.ph_weib  %>% plot(., xlab="Time", ylab="Surv", main="parametric cox ph - Weibull")

summary(ph)

```

ANOVA for PAN02 and CT26
Check conditions for ANOVA with Levene test: Not met for Pan02 and CT26_028

```{r}
#check normality assumption of ANOVA
plot(res.aov,2)

# check homogeneity of variance assumptions
plot(res.aov, 1)

# there are some outliers, Levene's test for homogeneity of variance
#results show there is indeed significance differences in variance, so assumptions for ANOVA are not met
#In this case, run the Kruskal-Wallis rank sum test

library(car)
leveneTest(size ~ treatment, data=Crown_Pan02)
leveneTest(size ~ treatment, data=CT26_028)

``` 


```{r}
Crown_Pan02 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/Crown_Pan02_Comp.txt", header=TRUE, sep="\t")
Crown_Pan02 <- Crown_Pan02 %>% gather("day","size", -Treatment, -animal) %>% 
	rename(treatment=Treatment) %>%
	mutate(day=as.numeric(sub("X","",day))) %>% mutate(treatment=factor(treatment, levels=c("Group_01", "Group_08")))

levels(Crown_Pan02$treatment)

#Crown_Pan02 on Day 32
#compute variance
res.aov <- aov(size~ treatment, data=Crown_Pan02 %>% dplyr::filter(day==32))
summary(res.aov)

#ANOVA CT26 Day 24
CT26_028 <- read.table(file="/Volumes/Picard/FLX/GCN2/In_vivo_data/19-CT26-028_28d.txt", header=TRUE, sep="\t")

CT26_028 <- CT26_028 %>% gather("day","size", -Group, -animal) %>% 
	rename(treatment=Group) %>%
	mutate(day=as.numeric(sub("X","",day))) %>%
  mutate(treatment=factor(treatment, levels=c("Vehicle","F019282_15mpk ","Anti-VEGFR", "Combo")))

res.aov <- aov(size~ treatment, data=CT26_028 %>% dplyr::filter(day==24))
summary(res.aov)
levels(CT26_028$treatment)

# Tukey's HSD honest significant differences
TukeyHSD(res.aov)


#two-sample T test
control <- Crown_Pan02 %>% dplyr::filter(day==32) %>% dplyr::filter(treatment=="Group_01") %>% pull(size)
treated <- Crown_Pan02 %>% dplyr::filter(day==32) %>% dplyr::filter(treatment=="Group_08") %>% pull(size)
t.test(control,treated,alternaitive="two.sided", var.equal=FALSE)
```


Kruskal-Wallis rank sum test
Wilcoxon rank sum for 2 groups

```{r}
unique(Crown_Pan02$day) # 4  7 11 14 18 21 25 28 32
attach(Crown_Pan02)
my_day <- day==28
my_veh <- treatment=="Group_01"
my_treat <- treatment=="Group_08"
wilcox.test(Crown_Pan02[my_day&my_veh,]$size, Crown_Pan02[my_day&my_treat,]$size, alternative = "two.sided")

Day 4: not sig
Day 7: 0.0089
11: 0.0007
14: 0.0004
18: 0.0007
21: 0.002
25  0.005
32: 0.005
kruskal.test(size ~ treatment, data = Crown_Pan02) # 10-5

```

CROWN LL2
Crown_LL2 is the one in the paper
```{r}
#unique(Crown_01_20$day) # 0  2  5  9 12 16 19 23 26 28
unique(Crown_LL2$day)  # 12 14 18 21 25 28
unique(Crown_LL2$treatment) # "Group_01" "Group_08"
attach(Crown_LL2)
my_day <- day==28
my_veh <- treatment=="Group_01"
my_treat <- treatment=="Group_08"
wilcox.test(Crown_LL2[my_day&my_veh,]$size, Crown_LL2[my_day&my_treat,]$size, alternative = "two.sided")
12: NS
14: NS
18: NS
21: 0.0001
25: 0.0001
28: 0.002

```

19-RENCA_005
```{r}
unique(Renca_05$day) # 6 10 13 17 20 24 27
attach(Renca_05)
my_day <- day==13
my_veh <- treatment=="Vehicle"
my_treat <- treatment=="F019282_15mpk"
wilcox.test(Renca_05[my_day&my_veh,]$size, Renca_05[my_day&my_treat,]$size, alternative = "two.sided")
6: 0.002
10: 0.012
13: 0.003
17: 0.0007
20: 0.0004
24: 1-5
27: 0.003

kruskal.test(size ~ treatment, data = Renca_05 %>% dplyr::filter(day==6)) #0.0015
kruskal.test(size ~ treatment, data = Renca_05 %>% dplyr::filter(day==24)) #0.0001
kruskal.test(size ~ treatment, data = Renca_05 %>% dplyr::filter(day==27)) #0.005
pairwise.wilcox.test(Renca_05 %>% dplyr::filter(day==6) %>% dselect(day), Crown_Pan02$treatment %>% dplyr::filter(day==6),
                 p.adjust.method = "BH")
```
19-RENCA_008
```{r}
unique(Renca_08$day) #  6  9 12 16 19 23 25 27 29
unique(Renca_08$treatment) "19282_15mpk_CD8"  "19282_15mpk_PDL1"
attach(Renca_08)

kruskal.test(size ~ treatment, data = Renca_08 %>% dplyr::filter(day==12)) #NS
6: NS
9: NS
12: 0.22
16: 0.01
19: 0.003
23: 0.0004
25: 0.0003
27: 0.0001

my_day <- day==29
my_veh <- treatment=="19282_15mpk_CD8"
my_treat <- treatment=="19282_15mpk_PDL1"
wilcox.test(Renca_08[my_day&my_veh,]$size, Renca_08[my_day&my_treat,]$size, alternative = "two.sided")

19: NS
23: NS
25: NS
27: 0.04

282 vs. Combo
27: 0.06
29: 0.05
```

19-CT26-028

```{r}
unique(CT26_028$day) #  6 10 13 17 19 24 26 28
unique(CT26_028$treatment) # "Vehicle"        "F019282_15mpk " "Anti-VEGFR"     "Combo"      
attach(CT26_028)

kruskal.test(size ~ treatment, data = CT26_028 %>% dplyr::filter(day==28)) #NS
6: NS
9: NS
13: 0.199
17: 0.13
19: 0.01
24: 0.001
26: 0.025
27: 0.0235

my_day <- day==24
my_veh <- treatment=="Vehicle"
my_treat <- treatment== "Combo"
wilcox.test(CT26_028[my_day&my_veh,]$size, CT26_028[my_day&my_treat,]$size, alternative = "two.sided")
```

20-RENCA-001
```{r}
RENCA_001_20 

unique(RENCA_001_20$day) #  6 10 13 17 20 24
unique(RENCA_001_20$treatment) # ""Vehicle" "F019282_30_mpk" "Anti_VEGFR" "F019282_30_mpk_VEGFR" "Anti-PDL1" "F019282_30_mpk_PDL1" 
attach(RENCA_001_20)
kruskal.test(size ~ treatment, data = RENCA_001_20 %>% dplyr::filter(day==13)) #NS
6: NS
9: NS
13: 0.0006
17: 4 -12
20:  5 -15
24: 0.001
26: 0.025
27: 0.0235

my_day <- day==17
my_veh <- treatment=="F019282_30_mpk"
my_treat <- treatment== "F019282_30_mpk_VEGFR"
wilcox.test(RENCA_001_20[my_day&my_veh,]$size, RENCA_001_20[my_day&my_treat,]$size, alternative = "two.sided")
```

LL2
```{r}
Crown_01_20
unique(Crown_01_20$day) #   0  2  5  9 12 16 19 23 26 28
unique(Crown_01_20$treatment) # "Vehicle"    "F019282_50_mpk"  "Anti_VEGFR"    "F019282_50_mpk_Anti_VEGFR"
attach(Crown_01_20)
kruskal.test(size ~ treatment, data = Crown_01_20 %>% dplyr::filter(day==5)) #NS
2: NS
5: 0.02
9: 3 -5
12: 2 -7
16: 3 -8
19: 4 -8
23: 7 -7
26: 3 -7

my_day <- day==19
my_veh <- treatment=="F019282_50_mpk"
my_treat <- treatment== "F019282_50_mpk_Anti_VEGFR"
wilcox.test(Crown_01_20[my_day&my_veh,]$size, Crown_01_20[my_day&my_treat,]$size, alternative = "two.sided")
```

21_RENCA_004

```{r}
RENCA_004_21
unique(RENCA_004_21$day) # 0  9 13 15 19 22 26
unique(RENCA_004_21$treatment) # "Vehicle"            "Anti-PD1"           "F019282_15mpk"      "F019282_15mpk_PDL1"
attach(RENCA_004_21)
kruskal.test(size ~ treatment, data = RENCA_004_21 %>% dplyr::filter(day==0)) #NS
0, 4, 5, 10, 13: NS
17: 0.009898 # **
20: 0.02966 # *

my_day <- day== 22
my_veh <- treatment=="Vehicle"
my_treat <- treatment== "F019282_15mpk_PDL1"
wilcox.test(RENCA_004_21[my_day&my_veh,]$size, RENCA_004_21[my_day&my_treat,]$size, alternative = "two.sided")

21_RENCA_004
DAy 22 19282 vs vehicle : NS
DAy 22 combo vs vehicle : 0.02
DAy 26 19282 vs vehicle : NS
DAy 26 combo vs vehicle : 0.002
```

19-CT26-005
```{r}
CT26_005
unique(CT26_005$day) # 7 11 14 17 19 21 24
unique(CT26_005$treatment) # "Vehicle"       "F019282_15mpk"
attach(CT26_005)

14: NS
17: 0.04
19: NS
21: NS
24: 0.1 NS
my_day <- day==14
my_veh <- treatment=="Vehicle"
my_treat <- treatment== "F019282_15mpk"
wilcox.test(CT26_005[my_day&my_veh,]$size, CT26_005[my_day&my_treat,]$size, alternative = "two.sided")
```

```{r}
library(pzfx)
pzfx_tables("/Volumes/Picard/FLX/GCN2/In_vivo_data/19-CT26-004.pzfx") # look at what tabs are available
ct26_004 <- read_pzfx("/Volumes/Picard/FLX/GCN2/In_vivo_data/19-CT26-004.pzfx", table="Tumor Volumes")

ct26_004 %<>% dselect("Days Post Inoculation",starts_with("Vehicle "), starts_with("F019282 30mpk BID")) %>% dfilter(!is.na(`Days Post Inoculation`)) %>% gather(Sample,size,-`Days Post Inoculation`) %>% mutate(treatment=case_when(
  grepl("Vehicle",Sample) ~ "Vehicle",
  grepl("F019282",Sample) ~ "F019282_30mpk"
)) %>% dselect(-Sample) %>% dfilter(!is.na(size)) %>% rename(day = `Days Post Inoculation`) 

unique(ct26_004$day) # 0  6  9 12 14 16 19 21
unique(ct26_004$treatment) # "Vehicle"       "F019282_30mpk"
attach(ct26_004)
9: 0.2 NS
12: 0.01
14: 0.01
16: 0.01
19: 0.03
21: 0.01
my_day <- day==9
my_veh <- treatment=="Vehicle"
my_treat <- treatment== "F019282_30mpk"
wilcox.test(ct26_004[my_day&my_veh,]$size, ct26_004[my_day&my_treat,]$size, alternative = "two.sided")
```

