---
title: "GCN2_Signature_TCGA"
author: "Mengshu Xu"
date: "2/14/2020"
output: html_document
---

With the GCN2 signature, find 
1) Immune cell correlates in the TCGA
  A) xCell 64 cell types
  B) Cibersort 22 immune cell types
  C) QuantiSeq 10 immune cell types
  D) Timer 6 immune cell types
  
QuantiSeq and CiberSort are the only signatures that are comparable across immune cell types and across sapmles, because they are absolute fractions of the sample

2) Correalates to onco-signatures

3) Correlates to cBioPOrtal Meta signatures

4) Correlates to specific genes of interest

Load packages
```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(magrittr)
library(tibble)
library(ggplot2)
library(scales)
library(Hmisc)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
source("d:/FLX/Files_from_Gene/R_functions.r")
```


Load Immune_estimate datasets

Cibersort has 8243 samples 
QuantiSeq has 8488 samples : M1, M2, Mono, Neu, CD8, T Reg
All of the samples overlap: however the cell types do not all overlap because the signature sets are different

TIMER only has 6 categories: B, CD8 , CD4, Neu, Macro, DC.
```{r}
list.files("D:/FLX/Reference_tables/TIL_estimates/")
Quanti_Ciber <- read.table("D:/FLX/Reference_tables/TIL_estimates/QuantiSeq_CiberSort_CellTypeFractionsData.tsv", sep="\t") 
Quanti_Ciber <- read.table("/Volumes/Enterprise/FLX/Reference_tables/TIL_estimates/QuantiSeq_CiberSort_CellTypeFractionsData.tsv", sep="\t") 
colnames(Quanti_Ciber) <- c("Sample","TT","Celltype","QuantiSeq","CiberSort")

# Quanti_Ciber %<>% unique                             #365414
# Quanti_Ciber %>% dfilter(!is.na(QuantiSeq)) #101638
# Quanti_Ciber %>% dfilter(!is.na(CiberSort))  %>% dselect(Celltype) %>% unique # 8243 samples
# Quanti_Ciber %>% dfilter(!is.na(QuantiSeq))  %>% dselect(Celltype) %>% unique # 8488 samples
# inner_join(Quanti_Ciber %>% dfilter(!is.na(CiberSort))  %>% dselect(Sample) %>% unique,Quanti_Ciber %>% dfilter(!is.na(QuantiSeq))  %>% dselect(Sample) %>% unique, by="Sample" ) #8243
# Quanti_Ciber %>% dfilter(!is.na(QuantiSeq)) %>% dfilter(!is.na(CiberSort)) %>% dselect(TT) %>% table #98916


TIMER <- read.table("D:/FLX/Reference_tables/TIL_estimates/TCGA_TIMER_estimates.csv", sep=",", header=TRUE) %>% mutate(Sample=substr(X,1,12)) %>% dselect(Sample,B_cell,T_cell.CD4,T_cell.CD8,Neutrophil,Macrophage,DC)

library(Hmisc) # for sedit
xCell <- read.table("D:/FLX/Reference_tables/TIL_estimates/xCell_TCGA_RSEM.txt", sep="\t", header=TRUE) %>% data.frame %>% dfilter(X %in% c("aDC","B-cells","CD4+ memory T-cells","CD4+ naive T-cells","CD8+ naive T-cells","CD8+ T-cells","CD8+ Tcm","CD8+ Tem","DC","Macrophages M1","Macrophages M2","Monocytes","Neutrophils","NK cells","Tregs")) %>% tibble::column_to_rownames("X") %>% t %>% data.frame %>% tibble::rownames_to_column("TCGA") %>% mutate(Sample=substr(TCGA,1,12)) %>% mutate(Sample = sedit(Sample,'.','-')) %>% dselect(-TCGA) %>% dselect(Sample,everything())
```

Load TCGA Z_Score data

To generate scores for hypoxia and pyruvate metabolism
```{r}
load("D:/FLX/Gabrilovich_data/allMergedData.quantileNormZscore.rdata") # Cancers_zscore
hypoxia <- read.table("D:/FLX/GCN2/ISR_data/RNA_seq/GSEA_hypoxia.txt",sep="\n", header=TRUE) %>% pull(HALLMARK_HYPOXIA)
hypoxia <- hypoxia[2:201]
pyruvate_meta <- read.table("D:/FLX/GCN2/ISR_data/RNA_seq/GSEA_pyruvate_metabolism.txt",sep="\n", header=TRUE) %>% pull(KEGG_PYRUVATE_METABOLISM)
pyruvate_meta <- pyruvate_meta[2:41]

hypoxia_sig_score <- Cancers_zscore %>% dfilter(Gene %in% hypoxia) %>% summarize_if(is.numeric,sum) %>% t %>% data.frame %>% rename(Hypoxia_score=.data$.) %>% tibble::rownames_to_column("Sample")
pyruvate_sig_score <- Cancers_zscore %>% dfilter(Gene %in% pyruvate_meta) %>% summarize_if(is.numeric,sum) %>% t %>% data.frame %>% rename(Pyruvate_metabolism_score=.data$.) %>% tibble::rownames_to_column("Sample")

```


```{r}

```

Load TCGA RNA expression data and Meta data

```{r}
load("/Volumes/Picard/FLX/TCGA/allMergedData.quantileNorm.rdata")
load("D:/FLX/TCGA/allMergedData.quantileNorm.rdata")
load("/Volumes/Picard/FLX/TCGA/allMergedData.SampleInfo.rdata")
load("D:/FLX/TCGA/allMergedData.SampleInfo.rdata")
str(allMergedData.quantileNorm)
cancer_t <- allMergedData.SampleInfo %>% dselect(Sample,TumorAbr,DzNormal) #,PrimaryDisease,Dataset,
# Z-scored data signatures of interest
# this dataset was generated in MDSC_Signature.Rmd
load("D:/FLX/Gabrilovich_data/Signature_zscores_Jan2020.rdata")
load("/Volumes/Picard/FLX/Gabrilovich_data/Signature_zscores_Jan2020.rdata")
Sig_z <- Signature_zscores_Jan2020 %>% t %>% data.frame %>% tibble::rownames_to_column("Sample")
Sig_z %<>% dfilter(grepl("TCGA",Sample))
GCN2_z <- Sig_z %>% dselect(Sample,GCN2,gMDSC,mMDSC)
PERK_z <- Sig_z %>% dselect(Sample,PERK,gMDSC,mMDSC)
HRI_z <- Sig_z %>% dselect(Sample,HRI,gMDSC,mMDSC)
```

Merge and Cor
QuantiSeq sample names need -01 added to the end to match the expression Sig Sample names
pivot_wider only works when there is a col with all unique numbers for each pivoted group: don't understand why
```{r}
GCN2_Ciber <- Quanti_Ciber %>% dfilter(!is.na(CiberSort)) %>% dselect(-QuantiSeq,-TT) %>% group_by(Celltype) %>% mutate(row = row_number()) %>% pivot_wider(names_from=Celltype,values_from=CiberSort) %>% dselect(-row) %>% mutate(Sample= sedit(Sample,"$","-01",Sample)) %>% inner_join(Sig_z, by="Sample") %>% unique

GCN2_Ciber_Cor <- round(cor(GCN2_Ciber[2:34],use = "complete.obs"),2) %>% data.frame %>% tibble::rownames_to_column("Cell_type") 

Ciber_order <- GCN2_Ciber_Cor %>% arrange(-GCN2) %>% dselect(Cell_type) %>% mutate(Cell_type= sedit(Cell_type," ",".")) %>% .[,1]


GCN2_Quanti <- Quanti_Ciber %>% dfilter(!is.na(QuantiSeq)) %>% dselect(-CiberSort,-TT) %>% group_by(Celltype) %>% mutate(row = row_number()) %>% pivot_wider(names_from=Celltype,values_from=QuantiSeq) %>% dselect(-row) %>% mutate(Sample= sedit(Sample,"$","-01",Sample)) %>% dselect(-`Regulatory T cells`) %>% mutate_if(is.numeric,mx_zscore2) %>% inner_join(Sig_z, by="Sample") %>% unique

GCN2_Quanti_Cor <- round(cor(GCN2_Quanti[2:17],use = "complete.obs"),2) %>% data.frame %>% tibble::rownames_to_column("Cell_type") 

Quanti_order <- GCN2_Quanti_Cor %>% arrange(-GCN2) %>% dselect(Cell_type) %>% mutate(Cell_type= sedit(Cell_type," ",".")) %>% .[,1]
```

Plot heatmap
```{r fig.height=10, fig.width=10}
GCN2_Ciber_Cor %>% mutate(Cell_type= sedit(Cell_type," ",".")) %>% gather(Signature,Correlation,-Cell_type) %>% mutate(Cell_type=factor(Cell_type,levels=rev(Ciber_order)), Signature=factor(Signature,levels=Ciber_order)) %>% ggplot(aes(x=Signature,y=Cell_type, fill=Correlation)) + geom_tile() + scale_fill_gradient2(low = "dodgerblue4", high = "red", na.value="white", limit = c(-1,1), name="Score",oob=squish) + theme(axis.text.x = element_text(angle=90,size=15,hjust=1,vjust=0.5), axis.text.y=element_text(size=15))  + ggsave("GCN2_Ciber_Cor.jpg", width=10, height=10, dpi=150, plot= last_plot(), units = "in")
#+ ggsave("PERK_xcell_cor.jpg", width=15, height=15, dpi=150, plot= last_plot(), units = "in")

GCN2_Quanti_Cor %>% mutate(Cell_type= sedit(Cell_type," ",".")) %>% gather(Signature,Correlation,-Cell_type) %>% mutate(Cell_type=factor(Cell_type,levels=rev(Quanti_order)), Signature=factor(Signature,levels=Quanti_order)) %>% ggplot(aes(x=Signature,y=Cell_type, fill=Correlation)) + geom_tile() + scale_fill_gradient2(low = "dodgerblue4", high = "red", na.value="white", limit = c(-1,1), name="Score",oob=squish) + theme(axis.text.x = element_text(angle=90,size=15,hjust=1,vjust=0.5), axis.text.y=element_text(size=15)) + ggsave("GCN2_QuantiSeq_Cor.jpg", width=10, height=10, dpi=150, plot= last_plot(), units = "in")
```

Correlate by Tumor type

pivot_wider: 1. Group_by the pivot column 2. mutate(row= row_number()) to give a unique value to each row, no idea why 3. pivot 4. dselect(-row)

CIBER
```{r fig.height=6, fig.width=8}

Ciber <- Quanti_Ciber %>% dselect(-QuantiSeq) %>% dfilter(!is.na(CiberSort)) %>% mutate(Sample= sedit(Sample,"$","-01",Sample)) %>% unique

Cor_MX <- function(i) { i %>% group_by(Celltype) %>% mutate(row = row_number()) %>% pivot_wider(names_from=Celltype,values_from=CiberSort) %>% dselect(-row) %>% inner_join(pyruvate_sig_score,by="Sample") %>% inner_join(GCN2_z, by="Sample") %>% unique %>% dselect(-Sample) %>% cor %>% data.frame %>% tibble::rownames_to_column("Cell_type") %>% dselect(Cell_type,Pyruvate_metabolism_score) }

Ciber_by_cancer <- Ciber %>% group_by(TT) %>% group_modify(~Cor_MX(.x)) 

Ciber_cancer_order <- Ciber %>% group_by(TT) %>% group_modify(~Cor_MX(.x)) %>% group_by(TT) %>% mutate(row = row_number()) %>% pivot_wider(names_from=TT, values_from = Pyruvate_metabolism_score) %>% dselect(-row) %>% mutate(RowSums=rowSums(.[2:20], na.rm=TRUE)) %>% arrange(RowSums) %>% pull(Cell_type)

Ciber_by_cancer %>% mutate(Cell_type= factor(Cell_type, levels=Ciber_cancer_order)) %>% ggplot(aes(y=Cell_type, x=TT, fill=Pyruvate_metabolism_score)) + geom_tile() + scale_fill_gradient2(low = "dodgerblue4", high = "red", na.value="white", limit = c(-1,1), name="Score",oob=squish) + theme(axis.text.x = element_text(angle=90,size=10,hjust=1,vjust=0.5), axis.text.y=element_text(size=10),axis.title = element_blank()) + ggsave("Pyruvate_Ciber_Cor.jpg", width=8, height=6, dpi=150, plot= last_plot(), units = "in")
```

QUANTI
```{r fig.height=3, fig.width=6}

Quanti <- Quanti_Ciber %>% dselect(-CiberSort) %>% dfilter(!is.na(QuantiSeq)) %>% mutate(Sample= sedit(Sample,"$","-01",Sample)) %>% unique %>% dfilter(grepl("TCGA",Sample))

Cor_MX_QS <- function(i) { i %>% group_by(Celltype) %>% mutate(row = row_number()) %>% pivot_wider(names_from=Celltype,values_from=QuantiSeq) %>% dselect(-row) %>% inner_join(GCN2_z,by="Sample")%>% inner_join(pyruvate_sig_score,by="Sample") %>% unique %>% dselect(-Sample) %>% cor %>% data.frame %>% tibble::rownames_to_column("Cell_type") %>% dselect(Cell_type,Pyruvate_metabolism_score) }

Quanti_by_cancer <- Quanti %>% group_by(TT) %>% group_modify(~Cor_MX_QS(.x)) 

Quanti_cancer_order <- Quanti %>% group_by(TT) %>% group_modify(~Cor_MX_QS(.x)) %>% group_by(TT) %>% mutate(row = row_number())  %>% pivot_wider(names_from=TT, values_from = Pyruvate_metabolism_score) %>% dselect(-row) %>% mutate(RowSums=rowSums(.[2:20], na.rm=TRUE)) %>% arrange(RowSums) %>% pull(Cell_type)

Quanti_by_cancer %>% mutate(Cell_type= factor(Cell_type, levels=Quanti_cancer_order)) %>% ggplot(aes(y=Cell_type, x=TT, fill=Pyruvate_metabolism_score)) + geom_tile() + scale_fill_gradient2(low = "dodgerblue4", high = "red", na.value="white", limit = c(-1,1), name="Score",oob=squish) + theme(axis.text.x = element_text(angle=90,size=10,hjust=1,vjust=0.5), axis.text.y=element_text(size=10),axis.title = element_blank()) + ggsave("Pyruvate_Quanti_Cor.jpg", width=6, height=3, dpi=150, plot= last_plot(), units = "in")
```

xCELL
```{r fig.height=6, fig.width=9}
xcell_TCGA <- read.table("D:/FLX/Reference_tables/TIL_estimates/xCell_TCGA_RSEM.txt",sep="\t", header=TRUE) # 64 cell types, 9948 TCGA cases
xcell_TCGA <- read.table("/Volumes/Enterprise/FLX/Reference_tables/TIL_estimates/xCell_TCGA_RSEM.txt",sep="\t", header=TRUE) # 64 cell types, 9948 TCGA cases

xcell_TCGA_t_z <- xcell_TCGA %>% tibble::column_to_rownames("X") %>% t %>% data.frame %>% tibble::rownames_to_column("Sample") %>% mutate_if(is.numeric,mx_zscore2) %>% dselect(Sample,aDC,B.cells,starts_with("CD4"), starts_with("CD8"),cDC,iDC,starts_with("Macrophage"), Monocytes,Neutrophils,NK.cells,NKT, pDC,Th1.cells, Th2.cells,Tregs) %>% mutate(Sample = sedit(Sample,".","-"))

Xcell <- xcell_TCGA_t_z %>% left_join(cancer_t, by="Sample") %>% inner_join(pyruvate_sig_score,by="Sample") %>% inner_join(GCN2_z,by="Sample") %>% unique 

Cor_xcell <- function(i) { i %>% dselect(-Sample,-DzNormal,-PrimaryDisease,-Dataset) %>% cor %>% data.frame %>% tibble::rownames_to_column("Cell_type") %>% dselect(Cell_type,Pyruvate_metabolism_score) }

Xcell_by_cancer <- Xcell %>% group_by(TumorAbr) %>% group_modify(~Cor_xcell(.x))

xcell_cancer_order <- Xcell %>% group_by(TumorAbr) %>% group_modify(~Cor_xcell(.x)) %>% group_by(TumorAbr) %>% mutate(row = row_number())  %>% pivot_wider(names_from=TumorAbr, values_from = Pyruvate_metabolism_score) %>% dselect(-row) %>% mutate(RowSums=rowSums(.[2:29], na.rm=TRUE)) %>% arrange(RowSums) %>% pull(Cell_type)

library(umap)
hri_xcell_umap <- umap(Xcell %>% group_by(TumorAbr) %>% group_modify(~Cor_xcell(.x)) %>% group_by(TumorAbr) %>% mutate(row = row_number())  %>% pivot_wider(names_from=TumorAbr, values_from = HRI) %>% dselect(-row, -Cell_type) )

cell_rows <- Xcell %>% group_by(TumorAbr) %>% group_modify(~Cor_xcell(.x)) %>% group_by(TumorAbr) %>% mutate(row = row_number())  %>% pivot_wider(names_from=TumorAbr, values_from = HRI) %>% pull(Cell_type) 

hri_xcell_umap$layout %>% data.frame %>% mutate(Celltype= cell_rows) %>% ggplot(aes(x=X1,y=X2,label=Celltype)) + geom_point() + geom_label()

Xcell_by_cancer %>% mutate(Cell_type= factor(Cell_type, levels=xcell_cancer_order)) %>% ggplot(aes(y=Cell_type, x=TumorAbr, fill=Pyruvate_metabolism_score)) + geom_tile() + scale_fill_gradient2(low = "dodgerblue4", high = "red", na.value="white", limit = c(-1,1), name="Score",oob=squish) + theme(axis.text.x = element_text(angle=90,size=10,hjust=1,vjust=0.5), axis.text.y=element_text(size=10),axis.title = element_blank()) + ggsave("Pyruvate_Xcell_Cor.jpg", width=9, height=6, dpi=150, plot= last_plot(), units = "in")

```

Correlate GCN2 signature with all genes, by Tumor type
```{r}
allMergedData.quantileNorm

cancer_t

```

```{r fig.height=4, fig.width=8}
data.frame(Hotness = Hotness_score %>% t) %>% tibble::rownames_to_column("Sample")%>% left_join(cancer_t, by="Sample") %>% dfilter(DzNormal==c("Cancer","Metastatic")) %>% dfilter(TumorAbr %in% choice) %>% mutate(TumorAbr= fct_reorder(TumorAbr,Hotness, .fun = median)) %>% ggplot(aes(x=TumorAbr,y=Hotness, color=TumorAbr)) + geom_jitter(pch=20, size=0.8, width=0.3) + coord_flip() + nolegend()

data.frame(Hotness = Hotness_score %>% t) %>% tibble::rownames_to_column("Sample")%>% left_join(cancer_t, by="Sample") %>% dfilter(DzNormal==c("Cancer","Metastatic"))  %>% dfilter(TumorAbr %in% choice) %>% mutate(TumorAbr= fct_reorder(TumorAbr,Hotness, .fun = median)) %>% ggplot(aes(x=TumorAbr,y=Hotness, color=TumorAbr)) + geom_jitter(pch=20, size=0.5, width=0.3) + nolegend() + theme(axis.title = element_text(size=12), axis.text.y = element_text(size=10), axis.text.x = element_text(angle=90,size=11,hjust=1))

choice <- data.frame(Hotness = Hotness_score %>% t) %>% tibble::rownames_to_column("Sample")%>% left_join(cancer_t, by="Sample")  %>% dplyr::count(TumorAbr, sort=TRUE) %>% dfilter(n > 80) %>% pull(TumorAbr)

```

Plot medians
```{r}
median_cancer_TNF_hotness <- Cancers_zscore %>% dfilter(Gene =="TNF") %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("Sample") %>% left_join(data.frame(Hotness = Hotness_score %>% t) %>% tibble::rownames_to_column("Sample") , by="Sample") %>% left_join(cancer_t, by="Sample") %>% dfilter(TumorAbr %in% choice, DzNormal %in% c("Cancer","Metastatic"), TumorAbr %nin% c("ALL","AML","GL")) %>% group_by(TumorAbr) %>% mutate(Median_TNF= median(TNF), Median_Hotness=median(Hotness)) %>% ungroup %>% dselect(-Sample,-TNF,-Hotness,-DzNormal) %>% unique
ggplot(median_cancer_TNF_hotness,aes(x=Median_TNF, y=Median_Hotness, colour=TumorAbr)) + geom_label(aes(label=TumorAbr, alpha=0.5)) +nolegend()
```

