---
title: "HPK1_OTI_batch2"
author: "MX"
date: "9/29/2020"
output: html_document
---

```{r}
library(limma)
library(tidyverse)
library(dplyr)
library(magrittr)
library(tibble)
library(ggplot2)
library(DESeq2)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
#source("D:/FLX/Files_from_Gene/R_functions.r")
sum_mx <- function(x) {sum(x,na.rm=TRUE)}
mean_mx <- function(x) {mean(x,na.rm=TRUE)}
```

Save and Load data
```{r}
save(tpm_gene2,file="/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/tpm_gene2.rdata") # normalized and summed by gene
load("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/tpm_gene2.rdata")
save(tpm_z2,file="/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/tpm_z2.rdata")
load("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/tpm_z2.rdata")
load("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/tpm_z.rdata")
load("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/OTI1_2_DESeq2_object.rdata")
```




Read in the data
```{r}
file_names <- list.files("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/", pattern=".tsv") %>% lapply(function(i) {strsplit(i,"_abundance")[[1]][1]})
files <- base::list.files("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/", pattern=".tsv", full.names=TRUE)
#data_files[1]
data_files <- lapply(files, function(i){read.table(i,header=TRUE)})
tpms2 <- cbind.data.frame(lapply(data_files, function(j) {j %>% dselect(tpm)}))
rownames(tpms2) <- data_files[[1]]$target_id
colnames(tpms2) <- file_names
```

Normalize data for plotting
Sum by genes
and take gene-wise z_score
```{r}
boxplot(tpms2 %>% as.matrix)
tpm_norm2 <- normalizeBetweenArrays(tpms2 %>% as.matrix)
boxplot(tpm_norm2)

gencode_symbol <- read.table("/Volumes/Picard/FLX/Reference_tables/GRCm38_GENCODE2SYMBOL.txt",sep="\t",header=TRUE) %>% data.frame

tpm_gene2 <- tpm_norm2 %>% data.frame %>% tibble::rownames_to_column("Gencode") %>% left_join(gencode_symbol, by="Gencode") %>% dselect(Symbol,everything()) %>% dselect(-Gencode) %>% dfilter(!is.na(Symbol)) %>% group_by(Symbol) %>% mutate_if(is.numeric,sum_mx) %>% unique %>% ungroup

tpm_m2 <- tpm_gene2 %>% tibble::column_to_rownames("Symbol") %>% as.matrix
tpm_z2 <- t(apply(tpm_m2,1,mx_zscore)) #the middle term 1 means rows, and 2 means columns
tpm_z2
boxplot(tpm_gene2)
boxplot(tpm_z2)
save(tpm_gene2,file="tpm_gene_batch2.rdata") 
#load("tpm_gene.rdata")

save(tpm_z2,file="tpm_z_batch2.rdata")
load(file="/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/tpm_z_batch2.rdata")
```

DESeq2 PCA plot
Batch comparison for samples

https://support.bioconductor.org/p/77960/

```{r}
library(DESeq2)
library(tximport)
files <- base::list.files("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/", pattern=".tsv", full.names=TRUE)
gencode_symbol <- read.table("/Volumes/Picard/FLX/Reference_tables/GRCm38_GENCODE2SYMBOL.txt",sep="\t",header=TRUE) %>% data.frame
names <- base::list.files("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/RNA_seq_batch2/", pattern=".tsv") %>% lapply(function(i) {strsplit(i,"_abundance")[[1]][1]}) %>% unlist
ot1 <- tximport(files, type="kallisto", tx2gene=gencode_symbol)
colnames(ot1$counts) <- names #<- colnames(tpm_gene)[2:length(colnames(tpm_gene))]
meta <- data.frame(Sample= names) %>% mutate(Name= case_when(
  grepl("_1-",Sample) ~ "Vehicle_Day3",
  grepl("_2-",Sample) ~ "F019324_100mpk_Day3",
  grepl("_3-",Sample) ~ "Vehicle_Day7",
  grepl("_4-",Sample) ~ "F019324_100mpk_Day7"
)) %>% mutate(Day= case_when(
  grepl("_1-",Sample) ~ "Day3",
  grepl("_2-",Sample) ~ "Day3",
  grepl("_3-",Sample) ~ "Day7",
  grepl("_4-",Sample) ~ "Day7"
)) %>% mutate(Treatment= case_when(
  grepl("_1-",Sample) ~ "Vehicle",
  grepl("_2-",Sample) ~ "Treated",
  grepl("_3-",Sample) ~ "Vehicle",
  grepl("_4-",Sample) ~ "Treated"
)) 
```


```{r fig.height=4, fig.width=7}
OT1_2 <- DESeqDataSetFromTximport(ot1,
                                   colData = meta,
                                   design= ~ Name)
OT1_2@assays@data$counts
OT1_2 <- DESeq(OT1_2)
rld <- rlog(OT1_2) 
plotPCA(rld,intgroup=c("Name")) + scale_color_manual(values= c("orange","red","grey60","grey30"))
plotPCA(rld,intgroup=c("Treatment")) + scale_color_manual(values= c("orange","grey60"))
plotPCA(rld,intgroup=c("Day"))
data <- plotPCA(rld,intgroup=c("Sample"), returnData=TRUE)
ggplot(data, aes(PC1,PC2, colour=group, label=Sample, alpha=0.1)) + geom_label(vjust = "inward", hjust = "inward",size=3) 
```

Explore genes in PCs
https://www.biostars.org/p/339919/

PC1: Ighv, Itgax, Cx3cr1
PC2: Retng, Batf3, S100A9
```{r}
rv <- rowVars(assay(rld)) 
selected <- order(rv, decreasing=TRUE)[seq_len(min(500, length(rv)))]
pc <- prcomp(t(assay(rld)[selected,]))
loadings <- as.data.frame(pc$rotation)
aload = abs(loadings)
sweep(aload, 2, colSums(aload), "/")
aload %>% arrange(-PC1) %>% tibble::rownames_to_column("Gene") %>% dfilter(!grepl("Gm|Rik",Gene))
```


DE
Levels: Vehicle_Day3 F019324_100mpk_Day3 Vehicle_Day7 F019324_100mpk_Day7 
There is a lot of noise from B cells, but since B cell carry-over is noisy, I should rank DEGs on padj instead of log2FoldChange, so look for consistent changes instead of dramatic ones
```{r}
OT1_2$Name <- relevel(OT1_2$Name, ref = "Vehicle_Day7")
OT1_2$Name <- relevel(OT1_2$Name, ref = "Vehicle_Day3")
OT1_2$Day <- relevel(OT1_2$Day, ref = "Day3")

OT1_2 <- DESeq(OT1_2)
resultsNames(OT1_2)
save(OT1_2,file="OTI1_2_DESeq2_object.rdata")

res_324_UP <- results(OT1_2, name="Name_F019324_100mpk_Day7_vs_Vehicle_Day7") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.1,log2FoldChange >= 0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
res_324_UP %>% write.table(file="Day7_324_vs_Veh_UP_padj.txt", sep="\t",row.names = FALSE,quote=F)

res_324_DN <- results(OT1_2, name="Name_F019324_100mpk_Day7_vs_Vehicle_Day7") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.05,log2FoldChange <= -0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
res_324_DN %>% write.table(file="Day7_324_vs_Veh_DN.txt", sep="\t",row.names = FALSE,quote=F)

Vehicle_Day7_vs_Day3_UP <- results(OT1_2, name="Name_Vehicle_Day7_vs_Vehicle_Day3") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.2,log2FoldChange >= 0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
Vehicle_Day7_vs_Day3_UP %>% write.table(file="Vehicle_Day7_vs_Day3_UP.txt", sep="\t",row.names = FALSE,quote=F)

Vehicle_Day7_vs_Day3_DN <- results(OT1_2, name="Name_Vehicle_Day7_vs_Vehicle_Day3") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.05,log2FoldChange <= -0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
Vehicle_Day7_vs_Day3_DN %>% write.table(file="Vehicle_Day7_vs_Day3_DN.txt", sep="\t",row.names = FALSE,quote=F)
```

Day 3 324 vs. Vehicle
```{r}
Day3_324_vs_Vehicle_UP <- results(OT1_2, name="Name_F019324_100mpk_Day3_vs_Vehicle_Day3") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.1,log2FoldChange >= 0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
Day3_324_vs_Vehicle_UP %>% write.table(file="Day3_324_vs_Vehicle_UP.txt", sep="\t",row.names = FALSE,quote=F)

Day3_324_vs_Vehicle_DN <- results(OT1_2, name="Name_F019324_100mpk_Day3_vs_Vehicle_Day3") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.1,log2FoldChange <= -0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
Day3_324_vs_Vehicle_DN %>% write.table(file="Day3_324_vs_Vehicle_DN.txt", sep="\t",row.names = FALSE,quote=F)
```

Genes activated by LMP1_OVA1 that are further activated by 324 treatment

```{r}
intersect(res_324_UP$Gene,Vehicle_Day7_vs_Day3_UP$Gene)
intersect_m <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% intersect(res_324_UP$Gene,Vehicle_Day7_vs_Day3_UP$Gene)) %>% dselect(Gene,everything()) %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(intersect_m,cluster_cols=F)
```


Volcano Plot

```{r fig.height=8, fig.width=6}

library(EnhancedVolcano)

EnhancedVolcano(results(OT1_2,name="Name_F019324_100mpk_Day3_vs_Vehicle_Day3"), #"Name_F019324_100mpk_Day7_vs_Vehicle_Day7"
                 lab = rownames(results(OT1_2)),
                 x = 'log2FoldChange',
                 y = 'padj',
                 ylim = c(0, 15),
                 xlim = c(-7, 8),
                 ylab = bquote(~-Log[10]~adjusted~italic(P)),
                 title = 'OTI Day 3 19324 vs Vehicle',
                 pCutoff = 0.05,
                 FCcutoff = 0.5,col=c('grey75', 'grey50', 'grey25', 'red'),
                 colAlpha = 1)

```

Heatmap of DE genes between Day 3 Vehicle and Day 7 100 mpk

```{r fig.height=6, fig.width=6}
library(RColorBrewer)
UP <- Vehicle_Day7_vs_Day3_UP %>% top_n(n=500) %>% pull(Gene)
DN <- Vehicle_Day7_vs_Day3_DN %>% top_n(n=500) %>% pull(Gene)
Vehicle_m <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c(UP,DN)) %>% dselect(Gene,everything()) %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(Vehicle_m,cluster_cols=F)

UP_324 <- res_324_UP %>% top_n(n=500)%>% pull(Gene) # 
DN_324 <- res_324_DN %>% top_n(n=500) %>% pull(Gene)
Vehicle_m <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c(UP_324,DN_324)) %>% dselect(Gene,everything()) %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(Vehicle_m,cluster_cols=F)
```

https://davetang.org/muse/2018/05/15/making-a-heatmap-in-r-with-the-pheatmap-package/

Heatmap of 1000 most variable genes across all samples (as measured by variace)
Variance is Sum of (value - mean)^2/(n-1)
Get a list of most variable genes
```{r fig.height=110, fig.width=6}
rv <- rowVars(assay(rld))# save row variance to rv
top1000 <- order(rv, decreasing=TRUE)[seq_len(min(1000, length(rv)))] # take the top 500 most variable genes, or the length of rv, whichever is less
top1000_varGenes <- row.names(assay(rld)[top1000,])
top1000_genes <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% top1000_varGenes) %>% dselect(Gene,everything()) %>% dfilter(!grepl("Gm|Rik",Gene))
top1000_genes_m <- top1000_genes %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(top1000_genes_m,cluster_cols=F)
```

Effector 
c("Prf1","Gzma")
Lineage
Il7r is Cd127 is Central Memory
Late Effector/Exhaustion
Suppressor
Entpd1/CD39
TCR pathway
Proliferation

Heatmap of interesting markers
```{r fig.height=2.6, fig.width=6}
Cd3 markers
c("Cd8a","Cd8b1","Cd3e","Cd3g","Cd3d","Cd247","Cd4","Tcf7")
Cytokine
c("Ifng","Tnf") # 1.3
Effector 
c("Prf1","Gzma","Gzmb","Gzmk","Gnly","Klrg1","Nkg7") # 2.1
Lineage
c("Ccr7","Sell","Cd69","Il7r","Tcf1") # 1.8
Late Effector/Exhaustion
c("Tigit", "Havcr2","Tox") # 1.6
Exhaustion
c("Tcf7","Lef1","Lag3","Ctla4","Pdcd1","Nr4a1","Eomes") # removed Cd247/Pdcd1 because not expressed in T cells
Suppressor
c("Entpd1","Foxp3","Tgfb","Il35","Il4")
TCR pathway
c("Cd28","Cd3a","Lck","Zap70","Slp76","Prdm1")
Proliferation
c("Mki67","Slamf7")
MAPK pathway: Map2k1 is MEK1
c("Gab2","Nras","Kras","Kit","Braf","Map2k1")
Lineage transcription factors: Lef1 is Tcf1 is a positive regulator of exhaustion: Tcf1 and Tcf7 are the same thing
c("Lef1")

Interesting_markers_m <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Mki67","Slamf7")) %>% tibble::column_to_rownames("Gene") %>% as.matrix
Interesting_markers_m <- Interesting_markers_m[c("Mki67","Slamf7"),]
pheatmap(Interesting_markers_m,cluster_cols=F, cluster_rows = F)

Interesting_markers_m <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Tcf7","Lef1","Tbx21","Lag3","Ctla4","Pdcd1","Nr4a1","Eomes","Cd69")) %>% tibble::column_to_rownames("Gene") %>% as.matrix
Interesting_markers_m <- Interesting_markers_m[c("Tcf7","Lef1","Tbx21","Lag3","Ctla4","Pdcd1","Nr4a1","Eomes","Cd69"),]
pheatmap(Interesting_markers_m,cluster_cols=F, cluster_rows = F)

Interesting_markers_m <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Kit","Ptpn11","Gab2","Nras","Kras","Braf","Map4k1","Map3k1","Map2k1","Mapk3","Nf1","Spred1","Pik3ca","Akt1")   ) %>% tibble::column_to_rownames("Gene") %>% as.matrix
Interesting_markers_m <- Interesting_markers_m[c("Tcf7","Lef1","Kit","Ptpn11","Gab2","Nras","Kras","Braf","Map4k1","Map3k1","Map2k1","Mapk3","Nf1","Spred1","Pik3ca","Akt1"),] 
pheatmap(Interesting_markers_m,cluster_cols=F, cluster_rows = F)

# batch1 data
Interesting_markers_m <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Tcf7","Lef1","Ptpn11","Gab2","Nras","Kras","Braf","Map4k1","Map3k1","Map2k1","Mapk3","Nf1","Spred1","Pik3ca","Akt1")   ) %>% dselect(Gene,sample_order_dot) %>% tibble::column_to_rownames("Gene") %>% as.matrix
Interesting_markers_m <- Interesting_markers_m[c("Tcf7","Lef1","Ptpn11","Gab2","Nras","Kras","Braf","Map4k1","Map3k1","Map2k1","Mapk3","Nf1","Spred1","Pik3ca","Akt1"),] 
pheatmap(Interesting_markers_m,cluster_cols=F, cluster_rows = F)
```


CD161 = Klrb1 is a marker of MAIT T cells
Tim3 = Havcr2
CD39 = Entpd exhaustion marker
PGE2= PTGER2
CD127 naive marker
```{r fig.height=7, fig.width=10}
library(ggbeeswarm)
tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Ccr7","Sell","Klrg1","Mki67","Ccr7","Il7r","Cd161","Cd27","Lck","Zap70","Tox","Lag3","Havcr2","Entpd1","Ifng","Tnf","Ptger2","Cd69","Cd127","Tigit","Tcf7","Gzmk","Nkg7")  ) %>% gather(Sample, RNA_z_score,-Gene) %>% mutate(Sample = str_replace(Sample,"\\.","-")) %>% left_join(meta, by="Sample") %>% ggplot(aes(Treatment,RNA_z_score, colour=Sample)) + geom_beeswarm() + facet_wrap(~Gene) + theme(axis.text.x = element_text(angle=90,size=12, vjust=0.5, hjust=1),strip.text=element_text(size=12))

tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Prdm1","Lag3","Havcr2","Tigit")  ) %>% gather(Sample, RNA_z_score,-Gene) %>% mutate(Sample = str_replace(Sample,"\\.","-")) %>% left_join(meta, by="Sample") %>% ggplot(aes(Treatment,RNA_z_score, colour=Sample)) + geom_beeswarm() + facet_wrap(~Gene) + theme(axis.text.x = element_text(angle=90,size=12, vjust=0.5, hjust=1),strip.text=element_text(size=12))



meta %<>% mutate(Treatment=factor(Treatment, levels=c("Vehicle_Day3","F019324_30mpk_Day3","F019324_100mpk_Day3","Vehicle_Day7","F019324_30mpk_Day7","F019324_100mpk_Day7")))

```

```{r}
boxplot(tpm_z)
tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Kit","Cd19")  ) %>% gather(Sample, RNA_z_score,-Gene) %>% mutate(Sample = str_replace(Sample,"\\.","-")) %>% left_join(meta, by="Sample") %>% ggplot(aes(Treatment,RNA_z_score, colour=Sample)) + geom_beeswarm() + facet_wrap(~Gene) + theme(axis.text.x = element_text(angle=90,size=12, vjust=0.5, hjust=1),strip.text=element_text(size=12))
```

Heatmap of Listeria response signature in T cells
```{r fig.height=2.5, fig.width=6}

T_cell_listeria_genes <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Cx3cr1","Klrg1","Tbx21","Ifng","Gzmb","Cd8b1")) 
T_cell_listeria_genes_m <- T_cell_listeria_genes %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(T_cell_listeria_genes_m,cluster_cols=F)
```
Lavanya's markers
```{r fig.height=2, fig.width=6}
library(ggbeeswarm)
Flow_markers <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Tnf","Ifng","Gzmb")) 
Flow_markers_m <- Flow_markers %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(Flow_markers_m,cluster_cols=F, cluster_rows=F)

Flow_markers %>% dselect(!ends_with("_E")) %>% gather(Sample, mRNA_zscore,-Gene) %>% mutate(Treatment=c(rep("Vehicle",15),rep("19324","15") ) ) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle","19324"))) %>% mutate(Gene=factor(Gene, levels=c("Tnf","Ifng","Gzmb"))) %>%
  ggplot(aes(x=Treatment, y=mRNA_zscore, group=Treatment)) + facet_wrap(~Gene) + geom_quasirandom() + theme_bw()
```

Plot proliferation markers vs. Ifng and Tnf
```{r}
Flow_markers <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Tnf","Ifng","Gzmb")) 
Flow_markers_m <- Flow_markers %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(Flow_markers_m,cluster_cols=F, cluster_rows=F)
```

Proliferation markers
```{r fig.height=3.5, fig.width=6}
Cell_cycle <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Cdk4","Top2a","Cenpe","Polr1c","Rpl4","Rpl30","Rpl3","Cox7b","Eif4a1","Cox17","Ndufb9","Atp5j")) 
Cell_cycle_m <- Cell_cycle %>% tibble::column_to_rownames("Gene") %>% as.matrix
Cell_cycle_m <- Cell_cycle_m[c("Cdk4","Top2a","Cenpe","Polr1c","Rpl4","Rpl30","Rpl3","Eif4a1","Cox7b","Cox17","Ndufb9","Atp5j"),]
pheatmap(Cell_cycle_m,cluster_cols=F, cluster_rows=F)

Cell_cycle <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Cdk4","Top2a","Cenpe","Polr1c","Rpl4","Rpl30","Rpl3","Cox7b","Eif4a1","Cox17","Ndufb9","Atp5j")) %>% dselect(Gene, sample_order_dot)
Cell_cycle_m <- Cell_cycle %>% tibble::column_to_rownames("Gene") %>% as.matrix
Cell_cycle_m <- Cell_cycle_m[c("Cdk4","Top2a","Cenpe","Polr1c","Rpl4","Rpl30","Rpl3","Eif4a1","Cox7b","Cox17","Ndufb9","Atp5j"),]
pheatmap(Cell_cycle_m,cluster_cols=F, cluster_rows=F)
```
```{r}
tpm_gene2 %>% dfilter(Symbol %in% c("Cdk4","Top2a","Cenpe","Polr1c","Rpl4","Rpl30","Rpl3","Cox7b","Eif4a1","Cox17","Ndufb9","Atp5j")) %>% dselect(!ends_with("_E")) %>% gather(Sample, mRNA_TPM,-Symbol) %>% mutate(Treatment=c(rep("Vehicle",60),rep("19324",60) ) ) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle","19324"))) %>% mutate(Gene=factor(Symbol, levels=c("Cdk4","Top2a","Cenpe","Polr1c","Rpl4","Rpl30","Rpl3","Cox7b","Eif4a1","Cox17","Ndufb9","Atp5j"))) %>%
  ggplot(aes(x=Treatment, y=mRNA_TPM)) + facet_wrap(~Gene, scales="free_y") + geom_boxplot(aes(fill=Treatment)) + geom_quasirandom() + theme_bw() + scale_fill_manual(values=c("#0A95A7","#E5541B"))
```
Cell cycle markers across Day 3 and Day 7

```{r fig.height=2.5, fig.width=8}
tpm_gene2 %>% dfilter(Symbol %in% c("Cdk4","Top2a","Cenpe")) %>% gather(Sample, mRNA_TPM,-Symbol) %>% mutate(Treatment=c(rep("Vehicle_Day3",12),rep("19324_Day3",15),rep("Vehicle_Day7",15),rep("19324_Day7",15) ) ) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle_Day3","19324_Day3","Vehicle_Day7","19324_Day7"))) %>%
  ggplot(aes(x=Treatment, y=mRNA_TPM)) + facet_wrap(~Symbol, scales="free_y", ncol=3) + geom_boxplot(aes(fill=Treatment)) + geom_quasirandom() + theme_bw() + scale_fill_manual(values=c("#FE9F33","#0A95A7","#FE9F33","#0A95A7"))
```


Naive cell markers
```{r fig.height=1.4, fig.width=6}
naive <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Ccr7","Sell","Ptprc")) 
naive_m <- naive %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(naive_m,cluster_cols=F, cluster_rows=F)
```
Apoptosis markers 
Necrosis markers: "MLKL","HMGB1" Hmgb1 might not be a necrosis marker, inflammaotry cells can secrete Hmgb1 without dying so illicit damage repair 

AICD markers: FasL transcription is regulated by a lot TFs: cMyc cJun, Irf1, Irf2, 

```{r fig.height=2.1, fig.width=6}
apop <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Casp3","Bid")) 
apop_m <- apop %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(apop_m,cluster_cols=F, cluster_rows=F)

necro <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Mlkl","Hmgb1","Bcl2l11","E2f1")) 
necro_m <- necro %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(necro_m,cluster_cols=F, cluster_rows=F)

aicd <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Fasl","Irf1","Irf2","Myc","Jun","Nfatc1")) 
aicd_m <- aicd %>% tibble::column_to_rownames("Gene") %>% as.matrix
aicd_m <- aicd_m[c("Fasl","Irf1","Irf2","Myc","Jun","Nfatc1"),]
pheatmap(aicd_m,cluster_cols=F, cluster_rows=F)
```
AICD markers plot scatter
```{r fig.height=1.5, fig.width=4}
aicd %>% dselect(!ends_with("_E")) %>% gather(Sample, mRNA_zscore,-Gene) %>% mutate(Treatment=c(rep("Vehicle",15),rep("19324","15") ) ) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle","19324"))) %>% mutate(Gene=factor(Gene, levels=c("Fasl","Irf1","Irf2"))) %>%
  ggplot(aes(x=Treatment, y=mRNA_zscore, group=Treatment)) + facet_wrap(~Gene) + geom_quasirandom() + theme_bw()

tpm_gene2 %>% dfilter(Symbol %in% c("Fasl","Irf1","Irf2")) %>% dselect(!ends_with("_E")) %>% gather(Sample, mRNA_TPM,-Symbol) %>% mutate(Treatment=c(rep("Vehicle",15),rep("19324",15) ) ) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle","19324"))) %>% mutate(Gene=factor(Symbol, levels=c("Fasl","Irf1","Irf2") )) %>%
  ggplot(aes(x=Treatment, y=mRNA_TPM)) + facet_wrap(~Gene, scales="free_y") + geom_boxplot() + geom_quasirandom(aes(color=Treatment)) + theme_bw() + scale_color_manual(values=c("#0A95A7","#E5541B")) + nolegend()
```

ISR/UPR markers
```{r}
tpm_gene2 %>% dfilter(Symbol %in% c("Ddit3","Trib3","Asns","Atf4","Atf6","Eif2ak1","Eif2ak2","Eif2ak3","Eif2ak4")) %>% dselect(!ends_with("_E")) %>% gather(Sample, mRNA_TPM,-Symbol) %>% mutate(Treatment=c(rep("Vehicle",45),rep("19324",45) ) ) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle","19324"))) %>% mutate(Gene=factor(Symbol, levels=c("Ddit3","Trib3","Asns","Atf4","Atf6","Eif2ak1","Eif2ak2","Eif2ak3","Eif2ak4"))) %>%
  ggplot(aes(x=Treatment, y=mRNA_TPM)) + facet_wrap(~Gene, scales="free_y") + geom_boxplot(aes(fill=Treatment)) + geom_quasirandom() + theme_bw() + scale_fill_manual(values=c("#0A95A7","#E5541B"))
```
```


Si et al. 2020 B6 model HPK1 k/o data
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE156204

Heatmap
```{r fig.height=6.5, fig.width=6}
pheatmap(FPKM_z,cluster_cols=F, annotation_names_row=FALSE)
pheatmap(FPKM_z[c("Gata3","Tbpl1","Nfat5","Pdcd1","Ctla4","Havcr2","Lag3","Cd274","Cdk2","Ccnb1","Casp3","Tbx21"),],cluster_cols=F)

Si <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Gata3","Tbpl1","Nfat5","Sema6d","Stat5a","Jun","Foxp3","Prdm1","Tbx21","Pdcd1","Ctla4","Havcr2","Lag3","Klrd1","Cd274","Cd200","Cd244","Cd25","Cd137","Cd28","Cd69","Cd27","Gadd45b","Cdkn2a","Cdk2","Ccnb1","Mad1l1","Tfdp1","Cdc25b","Cdc25c","Anxa1","Fas","Bcl2l2","Casp3","Casp1","Casp4","Bid","Bcl2","Bax"))
Si_m <- Si %>% tibble::column_to_rownames("Gene") %>% as.matrix
rownames(Si_m)
Si_m <- Si_m[c("Gata3","Tbpl1","Nfat5","Sema6d","Stat5a","Jun","Foxp3","Prdm1","Tbx21","Pdcd1","Ctla4","Havcr2","Lag3","Klrd1","Cd274","Cd200","Cd28","Cd69","Cd27","Gadd45b","Cdkn2a","Cdk2","Ccnb1","Mad1l1","Tfdp1","Cdc25b","Anxa1","Bcl2l2","Casp3","Casp1","Casp4","Bid","Bcl2","Bax"),]

pheatmap(Si_m,cluster_cols=F, cluster_rows=F,color= colorRampPalette(rev(brewer.pal(n = 7, name =
  "RdYlBu")))(length(mat_breaks) - 1),breaks= mat_breaks,)
mat_breaks <- seq(min(-2), max(2.5), length.out = 10)

#compare to batch 1 data: doesn't look similar at all
Cell_cycle <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Gata3","Tbpl1","Nfat5","Sema6d","Stat5a","Jun","Foxp3","Prdm1","Tbx21","Pdcd1","Ctla4","Havcr2","Lag3","Klrd1","Cd274","Cd200","Cd244","Cd25","Cd137","Cd28","Cd69","Cd27","Gadd45b","Cdkn2a","Cdk2","Ccnb1","Mad1l1","Tfdp1","Cdc25b","Cdc25c","Anxa1","Fas","Bcl2l2","Casp3","Casp1","Casp4","Bid","Bcl2","Bax")) %>% dselect(Gene, sample_order_dot)
Cell_cycle_m <- Cell_cycle %>% tibble::column_to_rownames("Gene") %>% as.matrix
Cell_cycle_m <- Cell_cycle_m[c("Gata3","Tbpl1","Nfat5","Sema6d","Stat5a","Jun","Foxp3","Prdm1","Tbx21","Pdcd1","Ctla4","Havcr2","Lag3","Klrd1","Cd274","Cd200","Cd28","Cd69","Cd27","Gadd45b","Cdkn2a","Cdk2","Ccnb1","Mad1l1","Tfdp1","Cdc25b","Anxa1","Bcl2l2","Casp3","Casp1","Casp4","Bid","Bcl2","Bax"),]
pheatmap(Cell_cycle_m,cluster_cols=F, cluster_rows=F)
Normali
```

Si et al  markers plot scatter

si_markers <- c("Cdc25c","Anxa1","Fas","Bcl2l2","Casp3","Casp1","Casp4","Bid")
TF : "Gata3","Tbpl1","Nfat5","Sema6d","Stat5a","Jun","Foxp3","Prdm1","Tbx21"
Exhaustion "Pdcd1","Ctla4","Havcr2","Lag3","Klrd1","Cd274","Cd200"
TCR "Cd244","Cd25","Cd137","Cd28","Cd69","Cd27"
Cell cycle "Gadd45b","Cdkn2a","Cdk2","Ccnb1","Mad1l1","Tfdp1","Cdc25b"
Apoptosis "Cdc25c","Anxa1","Fas","Bcl2l2","Casp3","Casp1","Casp4"
length(si_markers) # 39
```{r fig.height=1.6, fig.width=6}
si_markers <- c("Tcf7","Lef1","Tbx21") # "Kit"
tpm_gene2 %>% dfilter(Symbol %in% si_markers) %>% dselect(!ends_with("_E")) %>% gather(Sample, mRNA_TPM,-Symbol) %>% mutate(Treatment=c(rep("Vehicle",5*length(si_markers)),rep("19324",5*length(si_markers)) ) ) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle","19324"))) %>% mutate(Gene=factor(Symbol, levels=si_markers)) %>% ggplot(aes(x=Treatment, y=mRNA_TPM)) + facet_wrap(~Gene, scales="free_y", ncol=9) + theme_bw() + stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median, geom = "crossbar", width = 0.5) + geom_quasirandom(aes(color=Treatment)) + scale_color_manual(values=c("#0A95A7","#E5541B")) + nolegend()

ter <- tpm_gene2 %>% dfilter(Symbol=="Tbx21") %>% dselect(!ends_with("_E")) %>% dselect(-Symbol) %>% gather(Sample, mRNA_TPM) %>% mutate(Treatment=c(rep("Vehicle",5),rep("19324",5) ))
t.test(unlist(x=ter$mRNA_TPM[1:5]), y=unlist(ter$mRNA_TPM[6:10]), paired=F, alternative="two.sided")
```
INF-G downstream markers
```{r fig.height=2, fig.width=8}
tpm_gene2 %>% dfilter(Symbol %in% c("Cxcl9","Cxcl10")) %>% gather(Sample, mRNA_TPM,-Symbol) %>% mutate(Treatment=c(rep("Vehicle_Day3",8),rep("19324_Day3",10),rep("Vehicle_Day7",10),rep("19324_Day7",10)) ) %>% mutate(Treatment=factor(Treatment,levels=c("Vehicle_Day3","19324_Day3","Vehicle_Day7","19324_Day7"))) %>% ggplot(aes(x=Treatment, y=mRNA_TPM)) + facet_wrap(~Symbol, scales="free_y", ncol=9) + theme_bw() + stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median, geom = "crossbar", width = 0.5) + geom_quasirandom(aes(color=Treatment)) + scale_color_manual(values=c("#0A95A7","#E5541B","#0A95A7","#E5541B")) + nolegend()
```


PGE2 pathway
```{r fig.height=2, fig.width=8}
tpm_gene2 %>% dfilter(Symbol %in% c("Ptger2","Ptger4")) %>% dselect(!ends_with("_E")) %>% gather(Sample, mRNA_TPM,-Symbol) %>% mutate(Treatment=c(rep("Vehicle",5*2),rep("19324",5*2) ) ) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle","19324"))) %>% mutate(Gene=factor(Symbol, levels=si_markers)) %>% ggplot(aes(x=Treatment, y=mRNA_TPM)) + facet_wrap(~Symbol, scales="free_y", ncol=9) + theme_bw() + stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median, geom = "crossbar", width = 0.5) + geom_quasirandom(aes(color=Treatment)) + scale_color_manual(values=c("#0A95A7","#E5541B")) + nolegend()

tpm_gene2 %>% dfilter(Symbol %in% c("Ptger2","Ptger4")) %>% gather(Sample, mRNA_TPM,-Symbol) %>% mutate(Treatment=c(rep("Vehicle_Day3",8),rep("19324_Day3",10),rep("Vehicle_Day7",10),rep("19324_Day7",10)) ) %>% mutate(Treatment=factor(Treatment,levels=c("Vehicle_Day3","19324_Day3","Vehicle_Day7","19324_Day7"))) %>% ggplot(aes(x=Treatment, y=mRNA_TPM)) + facet_wrap(~Symbol, scales="free_y", ncol=9) + theme_bw() + stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median, geom = "crossbar", width = 0.5) + geom_quasirandom(aes(color=Treatment)) + scale_color_manual(values=c("#0A95A7","#E5541B","#0A95A7","#E5541B")) + nolegend()
```


Plot for batch 1 data
```{r fig.height=2, fig.width=8}
si_markers <- c("Map4k1","Map3k1","Map2k1","Mapk3") # "Kit"
tpm_gene %>% dfilter(Symbol %in% si_markers) %>% dselect(!ends_with("d3")) %>% gather(Sample, mRNA_TPM,-Symbol) %>% arrange(Sample) %>% mutate(Treatment=c(rep("Vehicle",3*length(si_markers)),rep("19324_30",3*length(si_markers)),rep("19324_100",2*length(si_markers)) ) ) %>% mutate(Treatment=factor(Treatment, levels=c("Vehicle","19324_30","19324_100"))) %>% mutate(Gene=factor(Symbol, levels=si_markers)) %>% ggplot(aes(x=Treatment, y=mRNA_TPM)) + facet_wrap(~Gene, scales="free_y", ncol=9) + theme_bw() + stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median, geom = "crossbar", width = 0.5) + geom_quasirandom(aes(color=Treatment)) + scale_color_manual(values=c("#0A95A7","#E5541B","#E5541B")) + nolegend()

ter <- tpm_gene %>% dfilter(Symbol=="Map2k1") %>% dselect(!ends_with("_d3")) %>% dselect(-Symbol) %>% gather(Sample, mRNA_TPM) %>% mutate(Treatment=c(rep("Vehicle",3),rep("19324_30",3),rep("19324_100",2) ) ) 
t.test(unlist(x=ter$mRNA_TPM[1:6]), y=unlist(ter$mRNA_TPM[7:8]), paired=F, alternative="two.sided")
```

Gene in HPK1 supposed pathway
Prdm1/Blimp Gata3, Tbx21
```{r fig.height=2, fig.width=6}
library(pheatmap)
Alleged_genes <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Prdm1","Gata3","Tbx21","Tox")) %>% dselect(Gene,sample_order_dot)
Alleged_genes_m <- Alleged_genes %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(Alleged_genes_m,cluster_cols=F)
```
NFkB pathway
```{r}
nfkb_genes <- tpm_z2 %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Rela","Nfkbil1","Bcl3","Traf1","Nfkbia","Icam1","Nfkb2","Relb","Il15","Traf3","Traf4","Nfkb1")) 
nfkb_genes_m <- nfkb_genes %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(nfkb_genes_m,cluster_cols=F)
```




