---
title: "HPK1_survival_analysis_Gene"
author: "Gene_MX"
date: "5/24/2021"
output: html_document
---
Useful tutorial:
https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Landmark_method



```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(magrittr)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
# survival analysis package
library(icenReg)
#package for regression models for interval censored data
install.packages("icenReg")
library(readxl)

# Other survival analysis package that allows right censor-only analysis
library("survival")
library("survminer")

```

##################################################
all 21-CT26-*, where * = 1,2,3,5,6,7,8,9,10
All these studies have CTLA4 control
21-CT26-001 F022162
21-CT26-002 F022019
21-CT26-003 F022054
21-CT26-004 is CCR4
21-CT26-005 F022005
21-CT26-006 F022692
21-CT26-007 F022234
21-CT26-008 F022054 NSG
21-CT26-009 F022835 
21-CT26-010 F022054 QD BID
21-CT26-011 F022054 with CTLA4, PD1, PDL1
21-CT26-012 F022019 low dose 

```{r}
# Group 02: Vehicle + anti-CTLA4
CT26_001 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-001_(3_10_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
# Group 02: Vehicle + anti-CTLA4
CT26_002 <-  readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-002_(3_11_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
# Group 02: Vehicle + anti-CTLA4
CT26_003 <-  readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-003_(5_4_2021)MX.xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
# Group 2: Vehicle + ⍺CTLA-4
CT26_005 <-  readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-005_(3_25_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
# Group 2: Vehicle + ⍺CTLA-4
CT26_006 <-  readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-006_(4_26_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
# Group 2: Vehicle + ⍺CTLA-4
CT26_007 <-  readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-007_(4_23_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
# Group 01: Vehicle (1% HPMC) BID
CT26_008 <-  readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-008_(4_12_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
# Group 2: Vehicle + ⍺CTLA-4
CT26_009 <-  readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-009_(5_24_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
#Group 06: Vehicle (1% HPMC) BID + anti-CTLA-4
CT26_010 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-010_(5_10_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
# Group 2: Vehicle + ⍺CTLA-4
CT26_012 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-012_(5_24_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) # %>% fill(size)
# Group 1: Vehicle (1% HPMC) BID + Isotype antibody
# Group 2: Vehicle (1% HPMC) + ⍺CTLA-4
# Group 3: Vehicle (1% HPMC) + ⍺PD-1
# Group 4: Vehicle (1% HPMC) + ⍺PD-L1
CT26_011 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/21-CT26-011_(6_3_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% mutate(Type= rep(c("Vehicle","CTLA4","PD1","PDL1","Vehicle","CTLA4","PD1","PDL1"),each=10)) %>% gather("day","size", -Treatment, -Animal,-Type) %>% mutate(day= as.numeric(day))
```

Load all file together
Hazard of reaching 1500mm3
```{r fig.height=3, fig.width=6}

files <- base::list.files("/Volumes/Picard/FLX/HPK1/In_vivo_data/", pattern=".xls", full.names=TRUE)

parse_mouse <- function(i) {
  readxl::read_xls(i,sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day)) %>% arrange(Animal,day) %>% fill(size)
}
data_files <- lapply(files, parse_mouse)
merge_CT26 <- do.call("rbind",data_files)

merge_CT26 %>% dselect(day,size) %>% dfilter(!is.na(day)) %>% group_by(day) %>% summarise(Threshold= length(day[size>=1500]), All= n(), Hazard= Threshold/All) %>% dfilter(All > 30) %>%
  ggplot(aes(x=day,y=Hazard)) + geom_line(colour="#0A95A7") + geom_point(colour="#0A95A7") + theme_bw() + scale_x_continuous(breaks = seq(0,80,by=5)) + ggtitle("Hazard of reaching 1500mm3 in all CT26 mice")
```

Average Graphs
```{r fig.height=4, fig.width=8}
CT26_001 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=exp(mean(log(size), na.rm=T))) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1500), color="grey20") + ylim(0,2000) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54"))

CT26_003 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=exp(mean(log(size), na.rm=T))) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1500), color="grey20") + ylim(0,2000) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))
geo.mean()
CT26_008 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=exp(mean(log(size), na.rm=T))) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2000) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))

CT26_011 %>% dselect(-Animal) %>% group_by(Type, Treatment,day) %>% summarise(Geom_mean_size=exp(mean(log(size), na.rm=T))) %>% ggplot(aes(x=day, y=Geom_mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1000), color="grey20") + facet_wrap(~Type) + ylim(0,2000) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))

CT26_011 %>% dselect(-Animal) %>% group_by(Type, Treatment,day) %>% summarise(Mean_size=mean(size,na.rm=T)) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2)  + geom_hline(aes(yintercept=1000), color="grey20") + facet_wrap(~Type) + ylim(0,NA) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920")) 
```

Faceted Graphs
! For these, use data that does not have the last measured tumor volume filled in

```{r fig.height=5, fig.width=11}
CT26_001 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20")+ ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
CT26_002 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
CT26_003 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20")  + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
CT26_005 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
CT26_006 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
CT26_007 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
CT26_008 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
CT26_009 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
CT26_010 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920")) + theme_bw()  + nolegend() 
CT26_012 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 

CT26_011 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment, ncol=4) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 

CT26_011 %>% dfilter(day > 25, Type == "Vehicle")
```


# make survival data functions

This function makes interval censored data, which is not right for mouse data. Since all mice are started on the same day, there is no need to left censor, only right censor. I can't find a way to only set right censoring with this Icenreg tool. When using coxph with right censoring, the results make a lot more sense
```{r}
# create a data.frame that is ready to go into survival analysis
# cutoff = size cutoff for suvival

# this returns left and right interval censored data for icenReg
make_survival_data <- function(x, cutoff=1000) {
	x %>% group_by(Treatment, Animal) %>% arrange(day) %>% summarize(
		which.cutoff = as.integer(min(which(size >= cutoff))), # the actual cutoff day
		reaches.cutoff = is.finite(which.cutoff),   # boolean on whether it ever reaches cutoff size
		left = if_else(reaches.cutoff, day[which.cutoff-1], max(day)), # left bound: -1 day when it reaches
		right = if_else(reaches.cutoff, day[which.cutoff], Inf)   # right bound, the day it reaches, or infinity if never reaches
	) %>% ungroup %>% select(-which.cutoff, -reaches.cutoff)
}

#returns right-censored data for Survival::coxph analysis

make_right_censor_survival_data <- function(x, cutoff=1000) {
	x %>% group_by(Treatment, Animal) %>% arrange(day) %>% summarize(
		which.cutoff = as.integer(min(which(size >= cutoff))), # the actual cutoff day
		reaches.cutoff = is.finite(which.cutoff),   # boolean on whether it ever reaches cutoff size
		right = if_else(reaches.cutoff, day[which.cutoff], max(day)), # day threshold is reached: or the last day of the study
		status = if_else(reaches.cutoff,1,0)
	) %>% ungroup %>% select(-which.cutoff, -reaches.cutoff)
}

# make a data.frame that is ready to plot as a survival curve
make_surv_curve_data <- function(x, cutoff=1000) {
	x %>% group_by(Treatment) %>% mutate(n.day1=sum(day==min(day))) %>%
		group_by(Treatment, day) %>% summarize(frac_alive = sum(size < cutoff)/n.day1[1])
}

```

Make survival data input that is interval censored for icenReg analysis

```{r}
current_data <- CT26_001 %>% make_survival_data(cutoff=1000) #Group 02: Vehicle + anti-CTLA4
current_data <- CT26_002 %>% make_survival_data(cutoff=1000) #Group 02: Vehicle + anti-CTLA4 1000 is non-sig even for 324
current_data <- CT26_003 %>% make_survival_data(cutoff=1000) # Group 02: Vehicle + anti-CTLA4
current_data <- CT26_005 %>% make_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_data <- CT26_006 %>% make_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_data <- CT26_007 %>% make_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_data <- CT26_008 %>% make_survival_data(cutoff=1000) # Group 01: Vehicle (1% HPMC) BID
current_data <- CT26_009 %>% make_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_data <- CT26_010 %>% make_survival_data(cutoff=1000) # Group 06: Vehicle (1% HPMC) BID + anti-CTLA-4
current_data <- CT26_012 %>% make_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_data <- CT26_011 %>% make_survival_data(cutoff=1000) # multiple grouops

current_data <- current_data %>% mutate(Treatment=fct_relevel(factor(Treatment), "Group 02: Vehicle + anti-CTLA4")) # Group 2: Vehicle + ⍺CTLA-4 # Group 02: Vehicle + anti-CTLA4
# Group 2: Vehicle (1% HPMC) + ⍺CTLA-4
# Group 3: Vehicle (1% HPMC) + ⍺PD-1
# Group 4: Vehicle (1% HPMC) + ⍺PD-L1
```
Make survival data input that is right censored for library(survival) coxph analysis

```{r}
current_right_data <- CT26_001 %>% make_right_censor_survival_data(cutoff=1000) # "Group 02: Vehicle + anti-CTLA4"
current_right_data <- CT26_002 %>% make_right_censor_survival_data(cutoff=1000) #Group 02: Vehicle + anti-CTLA4 1000 is non-sig even for 324
current_right_data <- CT26_003 %>% make_right_censor_survival_data(cutoff=1000) # Group 02: Vehicle + anti-CTLA4
current_right_data <- CT26_005 %>% make_right_censor_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_right_data <- CT26_006 %>% make_right_censor_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_right_data <- CT26_007 %>% make_right_censor_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_right_data <- CT26_008 %>% make_right_censor_survival_data(cutoff=1000) # Group 01: Vehicle (1% HPMC) BID
current_right_data <- CT26_009 %>% make_right_censor_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_right_data <- CT26_010 %>% make_right_censor_survival_data(cutoff=1000) # Group 06: Vehicle (1% HPMC) BID + anti-CTLA-4
current_right_data <- CT26_012 %>% make_right_censor_survival_data(cutoff=1000) # Group 2: Vehicle + ⍺CTLA-4
current_right_data <- CT26_011 %>% make_right_censor_survival_data(cutoff=1000) # multiple grouops
current_right_data %<>% mutate(Treatment=fct_relevel(factor(Treatment),"Group 02: Vehicle + anti-CTLA4")) # set baseline condition
```


Test assumptions of Cox PH model

In order to check these model assumptions, Residuals method are used. The common residuals for the Cox model include:

Schoenfeld residuals to check the proportional hazards assumption:
Martingale residual to assess nonlinearity
Deviance residual (symmetric transformation of the Martinguale residuals), to examine influential observations

```{r}
library("survival")
library("survminer")
#groom current_Data for Surv input
res.cox <- coxph(Surv(right,status) ~ Treatment, data=current_right_data)
res.cox

# test for proportional hazards assumption:
test.ph <- cox.zph(res.cox)
test.ph # When the covariates are NOT significant, then we can assume proportional hazards
ggcoxzph(test.ph)

```

# test different fit models for survival
```{r}
#tmp <- MM1S.surv_data %>% filter(treatment %in% c("18797.50mpkBID","Vehicle"))
#s <- Surv(tmp$left, tmp$right, type = "interval2")

#non_para <- ic_np(cbind(left,right) ~ treatment, data=MM1S.surv_data)
ph <- ic_sp(cbind(left,right) ~ Treatment, model='ph', bs_samples=1000, data=current_data) 
po <- ic_sp(cbind(left,right) ~ Treatment, model='po', bs_samples=1000, data=current_data)
par <- ic_par(cbind(left,right) ~ Treatment, model='ph', dist="weibull", data=current_data)
#par.ph_weib <- ic_par(cbind(left,right) ~ treatment, model='ph', dist="weibull", data=MM1S.surv_data)

#non_para %>% plot(.,  xlab="Time", ylab="Surv", main="non parametric")
ph %>% plot(.,  xlab="Time", ylab="Surv", main="cox ph")
#po  %>% plot(., xlab="Time", ylab="Surv", main="prop odds")
#par  %>% plot(., xlab="Time", ylab="Surv", main="parametric cox ph")
#par.ph_weib  %>% plot(., xlab="Time", ylab="Surv", main="parametric cox ph - Weibull")

summary(ph)
summary(po)
summary(par)
```


GL261

21-GL261-002 F022162
21-GL261-005 F022019
21-GL261-006 F022054
21-GL261-007 F022005
21-GL261-010 F022234
21-GL261-012 F022692 excel needs fixing
21-GL261-014 F022835 still ongoing on Day 41

```{r}
# Group 02: Vehicle + anti-CTLA4
GL_002 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/GL261/21-GL261-002_(4_19_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day))

GL_005 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/GL261/21-GL261-005_(4_28_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day))

GL_006 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/GL261/21-GL261-006_(4_26_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day))

GL_007 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/GL261/21-GL261-007_(4_28_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day))

GL_010 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/GL261/21-GL261-010_(5_10_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day))

GL_012 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/GL261/21-GL261-012_(5_21_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day))

GL_014 <- readxl::read_xls("/Volumes/Picard/FLX/HPK1/In_vivo_data/GL261/21-GL261-014_(6_1_2021).xls", sheet="Ind VA", skip=8,trim_ws = TRUE) %>% rename(Treatment="...1") %>% dselect(-starts_with("Length"),-starts_with("Width"),-starts_with("..."), -Code,-Day) %>% dfilter(Treatment != "Unassigned") %>% mutate(Animal= c(1:length(Treatment))) %>% dselect(Treatment, Animal, everything()) %>% gather("day","size", -Treatment, -Animal) %>% mutate(day= as.numeric(day))
```
Average Graphs
```{r fig.height=3, fig.width=7}
#GL_002 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=exp(mean(log(size), na.rm=T))) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,1500) + xlim(0,36) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))

GL_002 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=mean(size, na.rm=T)) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=800), color="grey20") + ylim(0,1000) + xlim(0,36) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))

GL_005 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=mean(size, na.rm=T)) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,1000) + xlim(0,36) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))

GL_006 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=mean(size, na.rm=T)) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,1000) + xlim(0,36) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))

GL_007 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=mean(size, na.rm=T)) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,1000) + xlim(0,36) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))

GL_012 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=mean(size, na.rm=T)) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,1000) + xlim(0,45) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))

GL_010 %>% dselect(-Animal) %>% group_by(Treatment,day) %>% summarise(Mean_size=mean(size, na.rm=T)) %>% ggplot(aes(x=day, y=Mean_size, group=Treatment, color=Treatment)) + geom_line(size=1.5) + theme_bw() + geom_point(pch=19, size=2) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,1500) + xlim(0,45) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54","#06707E","#AC3F12","#D07920"))
```
Graphs

```{r fig.height=4, fig.width=7}
GL_002 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=800), color="grey20")+ ylim(0,2000) + xlim(0,36) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
GL_005 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2000) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
GL_006 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=800), color="grey20")  + ylim(0,2000) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
GL_007 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=1000), color="grey20") + ylim(0,2000) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
GL_012 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=800), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 
GL_010 %>% ggplot(aes(x=day, y=size, group= interaction(Treatment, Animal), color=Treatment)) + geom_line() + facet_wrap(~Treatment) + geom_hline(aes(yintercept=800), color="grey20") + ylim(0,2500) + scale_color_manual(values=c("#0A95A7","#E5541B","#5B666F","#FE9F33","#4DA167","#947EB0","#083D77","#444D54")) + theme_bw()  + nolegend() 

```

Make survival data input

```{r}
current_data <- GL_002 %>% make_survival_data(cutoff=800) #Group 01: Vehicle QD
current_data <- GL_005 %>% make_survival_data(cutoff=800) #Group 01: Vehicle QD
current_data <- GL_006 %>% make_survival_data(cutoff=800) # Group 01: Vehicle QD
current_data <- GL_007 %>% make_survival_data(cutoff=1000) # Group 1: Vehicle (1% HPMC) BID
current_data <- GL_012 %>% make_survival_data(cutoff=800) # Group 1: Vehicle (1% HPMC) BID
current_data <- GL_010 %>% make_survival_data(cutoff=800) # Group 1: Vehicle (1% HPMC) BID


current_data <- current_data %>% mutate(Treatment=fct_relevel(factor(Treatment), "Group 1: Vehicle (1% HPMC) BID")) # Group 2: Vehicle + ⍺CTLA-4 # Group 02: Vehicle + anti-CTLA4
#Group 02: F022019 60 mpk BID
```


Test assumptions of Cox PH model

https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Landmark_method

In order to check these model assumptions, Residuals method are used. The common residuals for the Cox model include:
Schoenfeld residuals to check the proportional hazards assumption: suspect this test will fail on the mouse data
Martingale residual to assess nonlinearity
Deviance residual (symmetric transformation of the Martinguale residuals), to examine influential observations

```{r}
library("survival")
library("survminer")
#groom current_Data for Surv input
surv_data <- current_data %>% mutate(status = if_else(right == Inf, 0,1))
surv_data <- surv_data %>% mutate(Treatment=fct_relevel(factor(Treatment), "Group 1: Vehicle (1% HPMC) BID")) 

res.cox <- coxph(Surv(left,status, type="right") ~ Treatment, data=surv_data)
res.cox

# test for proprotional hazards assumption:
test.ph <- cox.zph(res.cox)
test.ph # When the covariates are NOT significant, then we can assume proportional hazards
ggcoxzph(test.ph)

```


# test different fit models for survival

I think in some cases the data is too "clean" for a Cox PH or Proportional Odds model, because they never "die", the tumors that are 0 are censored and don't get consideration
In addition, since Cox PH assumes constant risk of "dying" over time, it doesn't help the model when I try to lie to the model and say all Inf mice survived until the last measured day, or some late date. 
```{r}

#current_data %<>% mutate(left= if_else(right==Inf, Inf, left)) # if a tumor never reaches threshold, remove the left censor and make it also Inf: this is not allowed

np <- ic_np(cbind(left,right) ~ Treatment, maxIter=100, B=c(0,1), data=current_data)  
ph <- ic_sp(cbind(left,right) ~ Treatment, model='ph', bs_samples=100, data=current_data) 
po <- ic_sp(cbind(left,right) ~ Treatment, model='po', bs_samples=100, data=current_data)
par <- ic_par(cbind(left,right) ~ Treatment, model='ph', dist="weibull", data=current_data)

par %>% plot(.,  xlab="Time", ylab="Surv", main="cox ph")

summary(np)
summary(ph)
summary(po)
summary(par)
```
