---
title: "HPK1_OTI"
author: "MX"
date: "9/29/2020"
output: html_document
---

```{r}
library(limma)
library(tidyverse)
library(dplyr)
library(magrittr)
library(tibble)
library(ggplot2)
library(DESeq2)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
source("d:/FLX/Files_from_Gene/R_functions.r")
sum_mx <- function(x) {sum(x,na.rm=TRUE)}
mean_mx <- function(x) {mean(x,na.rm=TRUE)}
```

Save and Load data
```{r}
save(tpm_gene,file="/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/tpm_gene.rdata") # normalized and summed by gene

load("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/tpm_gene.rdata")
tpms
```




Read in the data
```{r}
file_names <- list.files("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/kallisto/", pattern=".tsv") %>% lapply(function(i) {strsplit(i,"\\.")[[1]][1]})
files <- base::list.files("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/kallisto/", pattern=".tsv", full.names=TRUE)
#strsplit("G_AAS_1.kallisto_quant.abundance.tsv","\\.")[[1]][1]
data_files <- lapply(files, function(i){read.table(i,header=TRUE)})
tpms <- cbind.data.frame(lapply(data_files, function(j) {j %>% dselect(tpm)}))
rownames(tpms) <- data_files[[1]]$target_id
colnames(tpms) <- file_names
```

Normalize data for plotting
Sum by genes
and take gene-wise z_score
```{r}
boxplot(tpms %>% as.matrix)
tpm_norm <- normalizeBetweenArrays(tpms %>% as.matrix)
boxplot(tpm_norm)

gencode_symbol <- read.table("/Volumes/Picard/FLX/Reference_tables/GRCm38_GENCODE2SYMBOL.txt",sep="\t",header=TRUE) %>% data.frame

tpm_gene <- tpm_norm %>% data.frame %>% tibble::rownames_to_column("Gencode") %>% left_join(gencode_symbol, by="Gencode") %>% dselect(Symbol,everything()) %>% dselect(-Gencode) %>% group_by(Symbol) %>% mutate_if(is.numeric,sum_mx) %>% unique %>% ungroup

tpm_m <- tpm_gene %>% tibble::column_to_rownames("Symbol") %>% as.matrix
tpm_z <- t(apply(tpm_m,1,mx_zscore)) #the middle term 1 means rows, and 2 means columns
tpm_t[1:5,1:5]
tpm_z[1:5,1:5]
boxplot(tpm_z)
save(tpm_gene,file="tpm_gene.rdata") 
#load("tpm_gene.rdata")

save(tpm_z,file="tpm_z.rdata")
load(file="tpm_z.rdata")
```

DESeq2 PCA plot
Batch comparison for samples

https://support.bioconductor.org/p/77960/

```{r}
library(DESeq2)
library(tximport)
files <- base::list.files("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/kallisto/", pattern=".tsv", full.names=TRUE)
gencode_symbol <- read.table("/Volumes/Picard/FLX/Reference_tables/GRCm38_GENCODE2SYMBOL.txt",sep="\t",header=TRUE) %>% data.frame
names <- base::list.files("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/kallisto/", pattern=".tsv") %>% lapply(function(i) {strsplit(i,"\\.")[[1]][1]}) %>% unlist
ot1 <- tximport(files, type="kallisto", tx2gene=gencode_symbol)
colnames(ot1$counts) #<- colnames(tpm_gene)[2:length(colnames(tpm_gene))]
meta <- data.frame(Sample= names) %>% mutate(Treatment= case_when(
  grepl("RAPT1.*d3",Sample) ~ "Vehicle_Day3",
  grepl("RAPT2.*d3",Sample) ~ "F019324_30mpk_Day3",
  grepl("RAPT3.*d3",Sample) ~ "F019324_100mpk_Day3",
  grepl("RAPT1.*d7",Sample) ~ "Vehicle_Day7",
  grepl("RAPT2.*d7",Sample) ~ "F019324_30mpk_Day7",
  grepl("RAPT3.*d7",Sample) ~ "F019324_100mpk_Day7",
)) 
meta <- data.frame(Sample=names) %>% mutate(Day = if_else(grepl("d3",Sample),"Day_3","Day_7"))
meta <- data.frame(Sample=names) 
sample_order<- data.frame(Sample= names) %>% mutate(Treatment= case_when(
  grepl("1-",Sample) ~ "Vehicle",
  grepl("2-",Sample) ~ "F019324_30mpk",
  grepl("3-",Sample) ~ "F019324_100mpk",
)) %>% mutate(Treatment=factor(Treatment,levels=c("Vehicle","F019324_30mpk","F019324_100mpk"))) %>% mutate(Day=if_else(grepl("_d3",Sample),"Day3","Day7")) %>% arrange(Day,Treatment) %>% pull(Sample)

OT1 <- DESeqDataSetFromTximport(ot1,
                                   colData = meta,
                                   design= ~ Treatment)
OT1@assays@data$counts
#OT1 <- DESeq(OT1)
rld <- rlog(OT1)
plotPCA(rld,intgroup=c("Treatment")) + scale_color_manual(values= c("red","red","orange","orange","grey60","grey60"))
plotPCA(rld,intgroup=c("Day"))
data <- plotPCA(rld,intgroup=c("Sample"), returnData=TRUE)
ggplot(data, aes(PC1,PC2, colour=group, label=Sample, alpha=0.1)) + geom_label(vjust = "inward", hjust = "inward") 
```

Explore genes in PCs
https://www.biostars.org/p/339919/

PC1: CAMP, LTF macrophage gMDSC markers, hemoglobin
PC2: most non genes
```{r}
rv = rowVars(assay(rld)) 
select = order(rv, decreasing=TRUE)[seq_len(min(500, length(rv)))]
pc = prcomp(t(assay(rld)[select,]))
loadings = as.data.frame(pc$rotation)
aload = abs(loadings)
sweep(aload, 2, colSums(aload), "/")
aload %>% arrange(-PC2)
```


DE
Levels: Vehicle_Day3 F019324_100mpk_Day3 F019324_100mpk_Day7 F019324_30mpk_Day3 F019324_30mpk_Day7 Vehicle_Day7
```{r}
OT1$Treatment <- relevel(OT1$Treatment, ref = "Vehicle_Day7")
OT1$Day <- relevel(OT1$Day, ref = "Day_3")
OT1$Sample <- relevel(OT1$Sample, ref = "RAPT1-1_d3") 
OT1 <- DESeq(OT1)
resultsNames(OT1)
save(OT1,file="OTI1_DESeq2_object.rdata")

res_Vehicle <- results(OT1, name="Treatment_Vehicle_Day7_vs_Vehicle_Day3") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.1,log2FoldChange >= 0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
res_Vehicle%>% write.table(file="Day7_vs_Day3_Veh_Gene_2.txt", sep="\t",row.names = FALSE,quote=F)


res_100Day7_vsVehDay3 <- results(OT1, name="Treatment_F019324_100mpk_Day7_vs_Vehicle_Day3") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.1,log2FoldChange >= 0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
res_100Day7_vsVehDay3 %>% write.table(file="Day7100_vs_Day3_Veh_Gene.txt", sep="\t",row.names = FALSE,quote=F)
Day7_100mpk_UP_genes <- res_100Day7_vsVehDay3 %>% pull(Gene)
Day7_100mpk_UP_genes2 <- res_100Day7_vsVehDay3 %>% pull(Gene)
intersect(Day7_100mpk_UP_genes,Day7_100mpk_UP_genes2)

res_100Day7_vsVehDay3_DN <- results(OT1, name="Treatment_F019324_100mpk_Day7_vs_Vehicle_Day3") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.1,log2FoldChange <= -0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
res_100Day7_vsVehDay3_DN %>% write.table(file="Day7100_vs_Day3_Veh_Gene_DN.txt", sep="\t",row.names = FALSE,quote=F)
res_100Day7_vsVehDay3_DN_genes <- res_100Day7_vsVehDay3_DN %>% pull(Gene)

res_30Day7_vsVehDay3 <- results(OT1, name="Treatment_F019324_30mpk_Day7_vs_Vehicle_Day3") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.1,log2FoldChange >= 0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
res_30Day7_vsVehDay3 %>% write.table(file="Day730_vs_Day3_Veh_Gene.txt", sep="\t",row.names = FALSE,quote=F)

"Day_Day_7_vs_Day_3"
res_Day7_vs_Day3_DN <- results(OT1, name="Day_Day_7_vs_Day_3") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.05,log2FoldChange < -0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
res_Day7_vs_Day3_DN

"Treatment_F019324_100mpk_Day7_vs_Vehicle_Day7"
res_100Day7_vsVehDay7 <- results(OT1, name="Treatment_F019324_100mpk_Day7_vs_Vehicle_Day7") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(padj <= 0.5,log2FoldChange <= -0.5) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% arrange(padj)
res_100Day7_vsVehDay7 %>% write.table(file="Day7100_vs_Day3_Veh_Gene_DN.txt", sep="\t",row.names = FALSE,quote=F)

#Check Prdm1/Blimp expression
res_100Day7_vsVehDay7 <- results(OT1, name="Treatment_F019324_100mpk_Day7_vs_Vehicle_Day7") %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene =="Prdm1") 

res %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(pvalue <= 0.05,log2FoldChange >= 0.5) %>% arrange(pvalue) %>% dfilter(!grepl("Gm|Rik",Gene)) #%>% write.table(file="RENCA_19_10_UP.txt",sep="\t", quote=FALSE,row.names = FALSE)
res %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(pvalue <= 0.003,log2FoldChange <= -0.5) %>% arrange(pvalue) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% write.table(file="RENCA_19_10_DN.txt",sep="\t", quote=FALSE,row.names = FALSE)
sig_UP <- res %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(pvalue <= 0.003,log2FoldChange >= 0.5) %>% arrange(-log2FoldChange) %>% dfilter(!grepl("Gm|Rik",Gene)) %>% pull(Gene)
sig_DN <- res %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(pvalue <= 0.003,log2FoldChange <= -0.5) %>% arrange(-log2FoldChange) %>% dfilter(!grepl("Gm|Rik",Gene))  %>% pull(Gene)
sig_UP %>% data.frame
```

Volcano Plot

```{r fig.height=8, fig.width=6}

library(EnhancedVolcano)

pdf("Vehicle_Day7_vs_Day3.pdf",width = 6, height = 8)
EnhancedVolcano(results(OT1,name="Treatment_Vehicle_Day7_vs_Vehicle_Day3"),
                 lab = rownames(results(OT1)),
                 x = 'log2FoldChange',
                 y = 'padj',
                 ylim = c(0, 8),
                 xlim = c(-6, 6),
                 ylab = bquote(~-Log[10]~adjusted~italic(P)),
                 title = 'OTI Vehicle Day 7 vs Day 3',
                 pCutoff = 0.05,
                 FCcutoff = 0.5,col=c('grey75', 'grey50', 'grey25', 'red'),
                 colAlpha = 1)
dev.off()

EnhancedVolcano(results(OT1,name="Treatment_F019324_100mpk_Day7_vs_Vehicle_Day3"),
                 lab = rownames(results(OT1)),
                 x = 'log2FoldChange',
                 y = 'padj',
                 ylim = c(0, 22),
                 xlim = c(-8, 8),
                 ylab = bquote(~-Log[10]~adjusted~italic(P)),
                 title = 'Treatment_F019324_100mpk_Day7_vs_Vehicle_Day3',
                 pCutoff = 0.05,
                 FCcutoff = 0.5,col=c('grey75', 'grey50', 'grey25', 'red'),
                 colAlpha = 1)

EnhancedVolcano(results(OT1,name="Treatment_F019324_30mpk_Day7_vs_Vehicle_Day3"),
                 lab = rownames(results(OT1)),
                 x = 'log2FoldChange',
                 y = 'padj',
                 ylim = c(0, 22),
                 xlim = c(-8, 8),
                 ylab = bquote(~-Log[10]~adjusted~italic(P)),
                 title = 'Treatment_F019324_30mpk_Day7_vs_Vehicle_Day3',
                 pCutoff = 0.05,
                 FCcutoff = 0.5,col=c('grey75', 'grey50', 'grey25', 'red'),
                 colAlpha = 1)

EnhancedVolcano(results(OT1,name="Day_Day_7_vs_Day_3"),
                 lab = rownames(results(OT1)),
                 x = 'log2FoldChange',
                 y = 'padj',
                 ylim = c(0, 50),
                 xlim = c(-7, 7),
                 ylab = bquote(~-Log[10]~adjusted~italic(P)),
                 title = 'Day7_vs_Day3',
                 pCutoff = 0.05,
                 FCcutoff = 0.5,col=c('grey75', 'grey50', 'grey25', 'red'),
                 colAlpha = 1)

EnhancedVolcano(results(OT1,name="Treatment_F019324_100mpk_Day7_vs_Vehicle_Day7"),
                 lab = rownames(results(OT1)),
                 x = 'log2FoldChange',
                 y = 'padj',
                 ylim = c(0, 20),
                 xlim = c(-7, 7),
                 ylab = bquote(~-Log[10]~adjusted~italic(P)),
                 title = '100mpk_Day7_vs_Vehicle_Day7',
                 pCutoff = 0.05,
                 FCcutoff = 0.5,col=c('grey75', 'grey50', 'grey25', 'red'),
                 colAlpha = 1)


```

Heatmap of DE genes between Day 3 Vehicle and Day 7 100 mpk

```{r fig.height=10, fig.width=7}
pheatmap(D7_100mpk_m,cluster_cols=F)
```

https://davetang.org/muse/2018/05/15/making-a-heatmap-in-r-with-the-pheatmap-package/
```{r fig.height=120, fig.width=6}
library(pheatmap)
sample_order_dot <- str_replace(sample_order,"-",".")
D7_100mpk <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% Day7_100mpk_UP_genes2) %>% dselect(Gene,sample_order_dot)

D7_100mpk %>% write.table(file="Gene_zscores_all_samples.txt",sep="\t",quote=F,row.names = F)
#test <- data.frame("one"=c(1,3,5,6,NA,8,4,6),"two"=c(5,NA,5,8,4,NA,3,2))
#test[is.na(test)] <- 0
D7_100mpk_m <- D7_100mpk %>% tibble::column_to_rownames("Gene") %>% as.matrix
D7_100mpk_m[1:10,1:10]
D7_100mpk_m["Hbb-bs",]

pheatmap(D7_100mpk_m,cluster_cols=F)

pheatmap(tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(grepl("^Vmn",Gene)) %>% dfilter(!is.nan(RAPT1.1_d3)) %>% tibble::column_to_rownames("Gene") %>% as.matrix )

pheatmap(tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(grepl("^Traj",Gene)) %>% dfilter(!is.nan(RAPT1.1_d3)) %>% tibble::column_to_rownames("Gene") %>% as.matrix )
```

Heatmap of DN genes
```{r fig.height=180, fig.width=6}

sample_order_dot <- str_replace(sample_order,"-",".")
length(res_100Day7_vsVehDay3_DN_genes) # 1565 genes
D7_100mpk_DN <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% res_100Day7_vsVehDay3_DN_genes) %>% dselect(Gene,sample_order_dot)

D7_100mpk_DN_m <- D7_100mpk_DN %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(D7_100mpk_DN_m,cluster_cols=F)
```
Heatmap of interesting markers
```{r fig.height=6, fig.width=6}
Interesting_markers_m <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Ccr7","Sell","Klrg1","Mki67","Ccr7","Il7r","Cd161","Cd27","Lck","Zap70","Tox","Lag3","Havcr2","Entpd1","Ifng","Tnf","Ptger2","Cd69","Cd127","Tigit","Tcf7","Gzmk","Nkg7","Prdm1")) %>% dselect(Gene,sample_order_dot) %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(Interesting_markers_m,cluster_cols=F)
```


Heatmap of 1000 most variable genes across all samples (as measured by variace)
Variance is Sum of (value - mean)^2/(n-1)
Get a list of most variable genes
```{r fig.height=120, fig.width=6}
rv = rowVars(assay(rld))# save row variance to rv
top1000 <- order(rv, decreasing=TRUE)[seq_len(min(1000, length(rv)))] # take the top 500 most variable genes, or the length of rv, whichever is less
top1000_varGenes <- row.names(assay(rld)[top1000,])
top1000_genes <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% top1000_varGenes) %>% dselect(Gene,sample_order_dot)
top1000_genes_m <- top1000_genes %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(top1000_genes_m,cluster_cols=F)
```
Take top z-score genes
```{r}
RAPT2.3_d3_UP <-tpm_z %>% data.frame %>% arrange(-RAPT2.3_d3) %>% dfilter(RAPT2.3_d3 >= 2.5) %>% tibble::rownames_to_column("Gene") %>% dfilter(!grepl("^Gm",Gene)) %>% pull(Gene)
length(Day7_100mpk_UP_genes2)
intersect(Day7_100mpk_UP_genes2,RAPT2.3_d3_UP) %>% write.table(file="RAPT2.3_d3_UP.txt",sep="\n",row.names = FALSE,quote=F)
RAPT2.3_d3_UP %>% write.table(file="RAPT2.3_d3_UP.txt",sep="\n",row.names = FALSE,quote=F)
```
CD161 = Klrb1 is a marker of MAIT T cells
Tim3 = Havcr2
CD39 = Entpd exhaustion marker
PGE2= PTGER2
CD127 naive marker
```{r fig.height=7, fig.width=10}
library(ggbeeswarm)
tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Ccr7","Sell","Klrg1","Mki67","Ccr7","Il7r","Cd161","Cd27","Lck","Zap70","Tox","Lag3","Havcr2","Entpd1","Ifng","Tnf","Ptger2","Cd69","Cd127","Tigit","Tcf7","Gzmk","Nkg7","Il10")  ) %>% gather(Sample, RNA_z_score,-Gene) %>% mutate(Sample = str_replace(Sample,"\\.","-")) %>% left_join(meta, by="Sample") %>% ggplot(aes(Treatment,RNA_z_score, colour=Sample)) + geom_beeswarm() + facet_wrap(~Gene) + theme(axis.text.x = element_text(angle=90,size=12, vjust=0.5, hjust=1),strip.text=element_text(size=12))

tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Prdm1","Lag3","Havcr2","Tigit")  ) %>% gather(Sample, RNA_z_score,-Gene) %>% mutate(Sample = str_replace(Sample,"\\.","-")) %>% left_join(meta, by="Sample") %>% ggplot(aes(Treatment,RNA_z_score, colour=Sample)) + geom_beeswarm() + facet_wrap(~Gene) + theme(axis.text.x = element_text(angle=90,size=12, vjust=0.5, hjust=1),strip.text=element_text(size=12))



meta %<>% mutate(Treatment=factor(Treatment, levels=c("Vehicle_Day3","F019324_30mpk_Day3","F019324_100mpk_Day3","Vehicle_Day7","F019324_30mpk_Day7","F019324_100mpk_Day7")))

```

```{r}
tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Kit","Cd19")  ) %>% gather(Sample, RNA_z_score,-Gene) %>% mutate(Sample = str_replace(Sample,"\\.","-")) %>% left_join(meta, by="Sample") %>% ggplot(aes(Treatment,RNA_z_score, colour=Sample)) + geom_beeswarm() + facet_wrap(~Gene) + theme(axis.text.x = element_text(angle=90,size=12, vjust=0.5, hjust=1),strip.text=element_text(size=12))
```

Heatmap of Listeria response signature in T cells
```{r}

T_cell_listeria_genes <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Cx3cr1","Klrg1","Tbx21","Ifng","Gzmb","Cd8b1")) %>% dselect(Gene,sample_order_dot)
T_cell_listeria_genes_m <- T_cell_listeria_genes %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(T_cell_listeria_genes_m,cluster_cols=F)
```

Gene in HPK1 supposed pathway
Prdm1/Blimp Gata3, Tbx21
```{r fig.height=2, fig.width=6}
library(pheatmap)
Alleged_genes <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Prdm1","Gata3","Tbx21","Tox")) %>% dselect(Gene,sample_order_dot)
Alleged_genes_m <- Alleged_genes %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(Alleged_genes_m,cluster_cols=F)
```
NFkB pathway
```{r}
nfkb_genes <- tpm_z %>% data.frame %>% tibble::rownames_to_column("Gene") %>% dfilter(Gene %in% c("Rela","Nfkbil1","Bcl3","Traf1","Nfkbia","Icam1","Nfkb2","Relb","Il15","Traf3","Traf4","Nfkb1")) %>% dselect(Gene,sample_order_dot)
nfkb_genes_m <- nfkb_genes %>% tibble::column_to_rownames("Gene") %>% as.matrix
pheatmap(nfkb_genes_m,cluster_cols=F)
```

Si et al. 2020 B6 model HPK1 k/o data
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE156204
```{r}
si_data <- read.table("/Volumes/Picard/Papers/GSE156204_gene.expression_RPKM.txt", header=T)
ENSMUG <- read.table("/Volumes/Picard/FLX/HPK1/Listeria_OVA_HPK1/ENSMUSG.txt",sep="\t",header=F) %>% data.frame
colnames(ENSMUG) <- c("tracking_id", "Gene")
FPKM_gene <- si_data %>% left_join(ENSMUG, by=
"tracking_id") %>% dselect(Gene,everything()) %>% dselect(-tracking_id) %>% group_by(Gene) %>% mutate_if(is.numeric,sum_mx) %>% unique %>% ungroup
FPKM_gene %>% write.table(file="FPKM_gene_sum.txt",sep="\t",quote=F,row.names = F)
FPKM_m <- FPKM_gene %>% dfilter(!grepl("Gm|Rik",Gene)) %>% tibble::column_to_rownames("Gene") %>% as.matrix
FPKM_z <- t(apply(FPKM_m,1,mx_zscore2)) 
```
Heatmap
```{r fig.height=4, fig.width=6}
pheatmap(FPKM_z,cluster_cols=F, annotation_names_row=FALSE)
pheatmap(FPKM_z[c("Gata3","Tbpl1","Nfat5","Pdcd1","Ctla4","Havcr2","Lag3","Cd274","Cdk2","Ccnb1","Casp3","Tbx21"),],cluster_cols=F)

```

```{r}
library(limma)
library(edgeR)
edgeR_hpk <- DGEList(counts=FPKM_m, group=group)
group <- c(1,1,2,2)
design <- model.matrix(~group)
keep <- filterByExpr(edgeR_hpk)
edgeR_hpk <- edgeR_hpk[keep,,keep.lib.sizes=FALSE]

edgeR_hpk <- calcNormFactors(edgeR_hpk)
edgeR_hpk$samples
edgeR_hpk <- estimateDisp(edgeR_hpk)
de <- exactTest(edgeR_hpk)
topTags(de)
de %>% data.frame %>% tibble::rownames_to_column("Gene") %>% arrange(-logFC) #%>% write.table(file="FPKM_EdgeR_DE.txt",sep="\t",quote=F,row.names = F)
Gene_order
```

