---
title: "HPK1_Immuno_TCGA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r }
library(ggplot2)
```


```{r}
hpk1_data <- read.csv("IMMUNO_HPK1_TCGA_clean.txt", sep= "\t", header= TRUE)
hpk1_df <- data.frame(hpk1_data)
head(hpk1_df)
min(hpk1_df$Immune.score)
library(dplyr)
hpk1_df <- hpk1_df %>%
  filter(HPK1_FPKM != "NA" & CD8A_FPKM != "NA" & CD45_FPKM != "NA") %>%
  mutate(Immune.score.pos = Immune.score+2500)
head(hpk1_df)
#unique(hpk1_data$Long_name)
str(hpk1_df)  
```
```{r}
hpk1_omit_OU <- hpk1_df %>%
  filter(Code != "UCEC" & Code != "OV")
head(hpk1_omit_OU)

```

```{r}
##the following sets the order of data plotted as same as input
hpk1_df$Code <- factor(hpk1_df$Code, levels=unique(hpk1_df$Code))
hpk1_omit_OU$Code <- factor(hpk1_omit_OU$Code, levels=unique(hpk1_omit_OU$Code))
```

```{r}

ggplot(hpk1_df, aes(Code, y=HPK1_FPKM, show.legend = FALSE)) +
  geom_boxplot(scale="width", trim = FALSE, aes(fill=Code), show.legend = FALSE) + 
  geom_jitter(height=0, width=0.1, size=0.3, alpha=0.4) + 
  #geom_point(size = 0.1,stroke=FALSE,colour="blue") +
  xlab("TCGA cancer type") +
  scale_y_log10() +
  theme(axis.text.x=element_text(size=16, angle = 90), 
        axis.text.y=element_text(size=16),
        axis.title =element_text(size=16))

```

```{r}
ggsave("TCGA_HPK1_log10.jpg", dpi=300, plot= last_plot(), width = 15, height = 15, units = "cm" )
```

```{r}
ggplot(hpk1_df, aes(Code, y=Immune.score.pos, show.legend = FALSE, aes= 0.4)) +
  geom_boxplot(scale="width", trim = FALSE, aes(fill=Code), show.legend = FALSE) + 
  geom_jitter(height=0, width=0.1, size=0.5, show.legend = FALSE) + 
  ylim(0,5500) +
  xlab("TCGA cancer type") +
  theme(axis.text.x=element_text(size=16, angle = 90), 
        axis.text.y=element_text(size=16),
        axis.title =element_text(size=16))
```
```{r}
ggsave("TCGA_immune.score_log10.jpg", dpi=300, plot= last_plot(), width = 15, height = 15, units = "cm" )
```
```{r}

```

```{r}
ggplot(hpk1_df, aes(x=Immune.score, y=HPK1_FPKM, colour=Code, show.legend = FALSE)) +
  geom_point(height=0, width=0.1, size=1, alpha=0.4, show.legend = FALSE) + 
  geom_abline(intercept = 0, slope = 0.0005,linetype="dashed", show.legend = FALSE) +
  #facet_grid(cols = vars(Code)) +
  facet_wrap( ~ Code, ncol=4) +
  xlab("Immune Score") +
  scale_y_log10() +
  theme(axis.text.x=element_text(size=16, angle = 90), 
        axis.text.y=element_text(size=16),
        axis.title =element_text(size=16))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
ggsave("TCGA_immune.score_cancer_facet_4.jpg", dpi=300, plot= last_plot(), units = "cm", width = 18, height = 15)
```
```{r}

HPK1_mean <-hpk1_df %>%
            group_by(Code) %>%
            summarize(meanImmune.score= median(Immune.score.pos), meanHPK1_FPKM=median(HPK1_FPKM), meanCD45_FPKM=median(CD45_FPKM), mean.ratio=1000*(meanHPK1_FPKM/meanImmune.score)) 
head(HPK1_mean)
```

```{r}
ggplot(HPK1_mean, aes(meanImmune.score, meanHPK1_FPKM, colour= mean.ratio, label = Code, show.legend=FALSE)) +
  geom_point(show.legend=FALSE) +
  geom_abline(intercept = -2.5, slope = 1/700,linetype="dashed", show.legend = FALSE) +
  xlab("Median Immune score") +
  ylab("Median HPK1 FPKM") +
  #scale_y_log10(0.00001,3) +
  xlim(2000,3700) +
  ylim(0.5,3.25) +
  geom_label(show.legend=FALSE)
```

```{r}
ggsave("HPK1_vs_TCGA_immune.jpg", dpi=300, width= 14, height=10, plot= last_plot(), units = "cm" )
```
```{r}
ggplot(HPK1_mean, aes(meanImmune.score, meanHPK1_FPKM, colour= mean.ratio, label = Code, show.legend=FALSE)) +
  geom_point(show.legend=FALSE) +
  geom_abline(intercept = -2.5, slope = 1/700,linetype="dashed", show.legend = FALSE) +
  xlab("Median Immune score") +
  ylab("Median HPK1 FPKM") +
  #scale_y_log10(0.00001,3) +
  xlim(2000,3700) +
  ylim(0.5,3.25) +
  geom_label(show.legend=FALSE)
```

```{r}
ggplot(hpk1_df, aes(Immune.score,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10()
```

# The noise in the correlation comes form OV and UCEC samples. They constitute most of the outliers in this graph
#Test Pearson correlation with and without OV and UCEC
```{r}
ggsave("HPK1_vs_Immune.score.jpg", dpi=300, width= 16, height=14, plot= last_plot(), units = "cm" )
```
```{r}

```


```{r}
cor.test(hpk1_df$HPK1_FPKM,hpk1_df$Immune.score)
```

```{r}
cor.test(hpk1_omit_OU$HPK1_FPKM,hpk1_omit_OU$Immune.score)
```
```{r}
ggplot(hpk1_omit_OU, aes(Immune.score,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10()
```
```{r}
ggsave("HPK1_vs_Immune.score_omit_OU.jpg", dpi=300, width= 16, height=14, plot= last_plot(), units = "cm" )
```

```{r}
ggplot(hpk1_df, aes(Stromal.score,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10()
```
```{r}
cor.test(hpk1_df$Stromal.score,hpk1_df$HPK1_FPKM)
```

```{r}
ggsave("HPK1_vs_stromal.score.jpg", dpi=300, width= 16, height=14, plot= last_plot(), units = "cm" )
```
```{r}
ggplot(hpk1_omit_OU, aes(Stromal.score,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10()
```

```{r}
ggsave("HPK1_vs_stromal.score_omit_OU.jpg", dpi=300, width= 16, height=14, plot= last_plot(), units = "cm" )
```
```{r}

```{r}
cor.test(hpk1_omit_OU$Stromal.score,hpk1_omit_OU$HPK1_FPKM)
```

```{r}
```{r}
ggplot(hpk1_df, aes(tumor.purity,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10()
```
```{r}
cor.test(hpk1_df$HPK1_FPKM,hpk1_df$tumor.purity)
```

```{r}
ggsave("HPK1_vs_tumor.purity.jpg", dpi=300, width= 16, height=14, plot= last_plot(), units = "cm" )
```


```{r}
ggplot(hpk1_omit_OU, aes(tumor.purity,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10()
```
```{r}
ggsave("HPK1_vs_tumor_purity_omit_OU.jpg", dpi=300, width= 16, height=14, plot= last_plot(), units = "cm" )
```

```{r}
cor.test(hpk1_omit_OU$HPK1_FPKM,hpk1_omit_OU$tumor.purity)
```
```{r}
ggplot(hpk1_df, aes(Immune.score,CD45_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10()
```
```{r}
cor.test(hpk1_df$CD45_FPKM,hpk1_df$Immune.score)
```

```{r}
ggsave("CD45_vs_immune_score.jpg", dpi=300, width= 16, height=14, plot= last_plot(), units = "cm" )
```

```{r}
ggplot(hpk1_df, aes(Stromal.score,CD45_FPKM,colour= Code)) +
  geom_point() +
  #facet_wrap( ~ Code, ncol=4) +
  scale_y_log10()
```

```{r}
cor.test(hpk1_df$CD45_FPKM,hpk1_df$Stromal.score)
```

```{r}
ggsave("CD45_Vs_stromal_score.jpg", dpi=300, width= 20, height=15, plot= last_plot(), units = "cm" )
```

```{r}
ggplot(hpk1_df, aes(tumor.purity,CD45_FPKM,colour= Code)) +
  geom_point() +
  #facet_wrap( ~ Code, ncol=4) +
  scale_y_log10()
```

```{r}
cor.test(hpk1_df$CD45_FPKM,hpk1_df$tumor.purity)
```
```{r}
ggsave("CD45_Vs_tumor.purity.jpg", dpi=300, width= 20, height=15, plot= last_plot(), units = "cm" )
```
```{r}
ggplot(hpk1_df, aes(CD45_FPKM,HPK1_FPKM,colour= Code)) +
  geom_point() +
  #facet_wrap( ~ Code, ncol=4) +
  scale_y_log10() +
  scale_x_log10()
```
```{r}
ggsave("HPK1_Vs_CD3D_omit_OU.jpg", dpi=300, width= 20, height=15, plot= last_plot(), units = "cm" )

```
```{r}
## DETERMINE IF HPK1 expression correlates with tumor or stromal content in Ovarian or Uterine cancers
hpk1_OU <- hpk1_df %>%
  filter(Code = OV, Code = UCEC)
```


```{r}
ggplot(hpk1_df, aes(CD8A_FPKM,HPK1_FPKM,colour= Code)) +
  geom_point() +
  #facet_wrap( ~ Code, ncol=4) +
  scale_y_log10() +
  scale_x_log10()

```
```{r}
hpk1_cd3D <- hpk1_df %>%
  filter(CD3D_FPKM != "NA")
hpk1_cd3D$Code <- factor(hpk1_cd3D$Code, levels=unique(hpk1_cd3D$Code))
ggplot(hpk1_cd3D, aes(CD3D_FPKM,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10()
```


```{r}
ggplot(hpk1_omit_OU, aes(CD3D_FPKM,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10()
```

```{r}
```{r}
ggplot(hpk1_omit_OU, aes(CD8A_FPKM,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10()
```
```{r}
cor.test(hpk1_omit_OU$CD45_FPKM,hpk1_omit_OU$HPK1_FPKM)
cor.test(hpk1_omit_OU$CD8A_FPKM,hpk1_omit_OU$HPK1_FPKM)
cor.test(hpk1_omit_OU$CD3D_FPKM,hpk1_omit_OU$HPK1_FPKM)
```

```{r}
hpk1_UO <-hpk1_df %>%
  filter(Code %in% c("OV"))
```
```{r}
ggplot(hpk1_UO, aes(Immune.score,HPK1_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10()
```
```{r}
ggsave("O_HPK1_Vs_Immune.jpg", dpi=300, width= 15, height=15, plot= last_plot(), units = "cm" )
```

```{r}
cor.test(hpk1_UO$HPK1_FPKM,hpk1_UO$Immune.score)
```

```{r}
hpk1_U <-hpk1_df %>%
  filter(Code %in% c("UCEC")) %>%
  mutate(ExpImmRatio = HPK1_FPKM/Immune.score.pos) %>%
  arrange(desc(ExpImmRatio))

head(hpk1_U)

```

```{r}
ggplot(hpk1_U, aes(ExpImmRatio,CD8A_FPKM,colour= Code)) +
  geom_point() +
  scale_y_log10()
```
```{r}
filter_1 <-hpk1_U %>%
  filter(ExpImmRatio >= 0.00125 )
summary(filter_1)
summary(hpk1_U)
```

