---
title: "HRI_regulation_Seq"
author: "Mengshu"
date: "8/9/2019"
output: html_document
---
```{r}
library(data.table)
library(magrittr)
library(dplyr)
library(tidyverse)
detach_package("tidyverse")
library(tibble)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
```


Mouse HRI-/- and WT
GSE119365 RNA seq dataset
Illumina HiSeq 2000 (Mus musculus)
This experiment is done in purified erythroblasts from fetal liver
```{r}
files <- list.files(path="/Volumes/Enterprise/FLX/GCN2/ISR_data/GSE119365_RAW/", pattern="*.txt$", full.names = TRUE)
file_names <-  list.files(path="/Volumes/Enterprise/FLX/GCN2/ISR_data/GSE119365_RAW/", pattern="*.txt$", full.names = FALSE)
data <- lapply(files,read.table)
str(data)
names(data) <- file_names
matrix <- do.call("cbind",data)
#matrix1 <- bind_cols(data)
HRI_meta <- matrix %>% tibble::column_to_rownames("V1") %>% select_if(is.numeric) %>% t %>% data.frame %>% tibble::rownames_to_column("Sample") %>% separate(Sample, sep="_|.txt.",into=c("GSM","rna","seq","genotype","Fe","replicate","trash2","trash3")) %>% dselect(GSM,genotype,Fe,replicate)
save(HRI_meta, file="GSE119365_meta.rdata")
load("GSE119365_meta.rdata")

HRI_matrix <- matrix %>% tibble::column_to_rownames("GSM3371777_rna_seq_wt_plus_r1.txt.V1") %>% select_if(is.numeric)
colnames(HRI_matrix) <- HRI_meta$GSM
save(HRI_matrix,file="GSE119365_matrix.rdata")
load("/Volumes/Enterprise/FLX/GCN2/ISR_data/GSE119365_matrix.rdata")
```

Metadata
```{r}
HRI_meta
```
DESeq2
Make sure the data matrix sample order in the columns matches the order of the metadata rows for the samples
Create deseq2 object
Set the reference conditions
in the design, put the more important factor on the end. In this case, it could be either one. 
DESeq2 can't do multifactor DE analysis, so I have to split up the data into groups and do pair-wise analyses; so stupid
```{r}
library(DESeq2)
wt <- as.matrix(HRI_matrix %>% tibble::column_to_rownames("Rik") %>% dselect(GSM3371777:GSM3371782))
wt_meta <- as.matrix(HRI_meta %>% tibble::column_to_rownames("GSM") %>% dfilter(genotype=="wt"))

dds_wt <- DESeqDataSetFromMatrix(countData = wt,
                              colData = wt_meta,
                              design= ~ Fe)

ko <- as.matrix(HRI_matrix %>% tibble::column_to_rownames("Rik") %>% dselect(GSM3371783:GSM3371788))
ko_meta <- as.matrix(HRI_meta %>% tibble::column_to_rownames("GSM") %>% dfilter(genotype=="ko"))

dds_ko <- DESeqDataSetFromMatrix(countData = ko,
                              colData = ko_meta,
                              design= ~ Fe)

fe_minus <- as.matrix(HRI_matrix %>% tibble::column_to_rownames("Rik") %>% dselect(GSM3371780:GSM3371782,GSM3371786:GSM3371788))
fe_minus_meta <- as.matrix(HRI_meta %>% tibble::column_to_rownames("GSM") %>% dfilter(Fe=="minus"))

dds_fe_minus <- DESeqDataSetFromMatrix(countData = fe_minus,
                              colData = fe_minus_meta,
                              design= ~ genotype)

dds_wt$Fe <- relevel(dds_wt$Fe, ref="plus")
dds_ko$Fe <- relevel(dds_ko$Fe, ref="plus")
dds_fe_minus$genotype <- relevel(dds_fe_minus$genotype, ref="ko")
```
Prefilter data
```{r}
keep <- rowSums(counts(dds_wt)) >= 10
dds_wt <- dds_wt[keep,]

keep <- rowSums(counts(dds_ko)) >= 10
dds_ko <- dds_ko[keep,]

keep <- rowSums(counts(dds_fe_minus)) >= 10
dds_fe_minus <- dds_fe_minus[keep,]
```

DE analysis
```{r}
dds_wt <- DESeq(dds_wt)
result_wt <- results(dds_wt)
resultsNames(dds_wt)

dds_ko <- DESeq(dds_ko)
result_ko <- results(dds_ko)
resultsNames(dds_ko)

dds_fe_minus <- DESeq(dds_fe_minus)
result_fe_minus <- results(dds_fe_minus)
resultsNames(dds_fe_minus)
```

Shrink log fold change for visualization (so the number range isn't so huge)
```{r}
source("https://bioconductor.org/biocLite.R")
biocLite("apeglm")
library(apeglm)
resultLFC_wt <- lfcShrink(dds_wt, coef="Fe_minus_vs_plus", type="apeglm")
resultLFC_ko <- lfcShrink(dds_ko, coef="Fe_minus_vs_plus", type="apeglm")
resultLFC_fe_minus <- lfcShrink(dds_fe_minus, coef="genotype_wt_vs_ko", type="apeglm")
```



```{r}
summary(result_wt) #43 UP 
summary(result_ko) #18 UP
summary(result_fe_minus) #34 UP
res_wt_Ordered <- result_wt[order(result_wt$pvalue),]
res_ko_Ordered <- result_ko[order(result_ko$pvalue),]
res_fe_minus_Ordered <- result_fe_minus[order(result_fe_minus$pvalue),]

res_wt_sig <- subset(res_wt_Ordered, padj < 0.1 & log2FoldChange >1) # 43
res_ko_sig <- subset(res_ko_Ordered, padj < 0.5 & log2FoldChange >1) # 16
res_fe_minus_sig <- subset(res_fe_minus_Ordered, padj < 0.1 & log2FoldChange >1)

sum(res_fe_minus_sig$padj < 0.05, na.rm=TRUE) # 41
```
Differential on the DE
setdiff(A,B) present in A, but not present in B
intersect(A,B)
```{r}
rownames(res_wt_sig)
rownames(res_ko_sig)
rownames(res_fe_minus_sig)
WTonly_Fe_minus_vs_plus <- setdiff(rownames(res_wt_sig),rownames(res_ko_sig))
intersect(rownames(res_wt_sig),rownames(res_ko_sig))

intersect(WTonly_Fe_minus_vs_plus,rownames(res_fe_minus_sig))
```


Plot results
```{r}
plotMA(result_wt)
plotMA(resultLFC_wt, ylim=c(-2,2))
plotMA(result_ko)
plotMA(resultLFC_ko, ylim=c(-2,2))
plotMA(result_fe_minus)
plotMA(resultLFC_fe_minus, ylim=c(-2,2))
```

Export

```{r}
write.csv(as.data.frame(res_wt_Ordered), 
          file="DeSeq2_WT_Fe_minus_vs_plus.csv")
write.csv(as.data.frame(res_ko_Ordered), 
          file="DeSeq2_ko_Fe_minus_vs_plus.csv")
write.csv(as.data.frame(res_fe_minus_Ordered), 
          file="DeSeq2_Fe_minus_wt_vs_ko.csv")
```


