---
title: "Greenberg_ID8_vegf_mouse_line"
author: "Mengshu"
date: "10/10/2019"
output: html_document
---

Load useful packages
```{r}
library(tidyverse)
library(dplyr)
library(magrittr)
library(tibble)
library(ggplot2)
library(data.table)
library(reshape2) #needed for the heatmap
library(limma)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
#source("D:/FLX/Files_from_Gene/R_functions.r")
#load("/Volumes/Enterprise/FLX/Reference tables/mouse_human_gene.rdata")
```
Load the data
```{r}
ID8 <- read.table("/Volumes/Enterprise/FLX/GCN2/Greenberg/GSE120263_series_matrix_data.txt", header=TRUE)
```


This array is the MTA, which I've parsed before
translate to gene names and remove probe names
```{r}
Probe_Gene <- read.table("/Volumes/Enterprise/FLX/GCN2/VEGF/MTA_Affy_Mouse_Probe2Gene.txt",sep="\t", header=FALSE)
colnames(Probe_Gene) <- c("Probe","Gene","Other") 
Probe_Gene %<>% dselect(-Other)

ID8_gene <- left_join(Probe_Gene,ID8, by=c("Probe"="ID_REF")) %>% dfilter(!is.na(GSM3397258),!is.na(Gene)) %>% dselect(-Probe)
#ID8_gene_unique <- ID8_gene[unique(ID8_gene$Gene),] %>% remove_rownames()
ID8_gene_unique <- ID8_gene[!duplicated(ID8_gene$Gene),] %>% remove_rownames()
length(unique(ID8_gene_unique$Gene))
save(ID8_gene,file="Greenberg_gene.rdata")

```

Merge this data with Mosely cell lines
make a matrix of cell lines in rows, genes in columns for Mosely
```{r}
load("/Volumes/Enterprise/FLX/Silpa_data/mouseData_tum.rdata")
load("/Volumes/Enterprise/FLX/Silpa_data/samples.info_tum.rdata")
samples.info_tum
mouseData_tum[1:5,1:93]
mouse <- data.frame(bind_cols(samples.info_tum %>% dselect('SampleID'),mouseData_tum)) %>% tibble::column_to_rownames("SampleID") %>% t %>% data.frame %>% tibble::rownames_to_column("Gene")
```
Merge ID8 with Mosely
transpose the ID8 first to genes in columns
```{r}
Mosely_ID8 <- left_join(ID8_gene_unique,mouse,by="Gene")
my_samples <- c("ID8_T","ID8_T","ID8_T",samples.info_tum$`experimental group`)
Mosely_ID8 %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% mutate(Sample=my_samples) %>% dselect(Sample, everything())
```
Normalize
Take Z-score
```{r}
library(limma)
Mosely_ID8_norm <- normalizeBetweenArrays(Mosely_ID8[,2:97])
boxplot(Mosely_ID8[,2:97])
boxplot(Mosely_ID8_norm)
Mosely_ID8_norm %<>% data.frame %>% mutate(Gene=Mosely_ID8$Gene) %>% dselect(Gene,everything())

Mosely_ID8_zscore <- Mosely_ID8_norm %>% mutate_if(is.numeric,mx_zscore2)


Mosely_ID8 %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% mutate(Sample=my_samples) %>% dselect(Sample, everything())
```

Filter for ISR markers and take double z-score
```{r}
isr_z <- Mosely_ID8_zscore %>% dfilter(Gene %in% ISR_markers)
colnames(isr_z) <- c("Gene",paste0(my_samples,'_',1:96)) 

isr_z_z <- isr_z %>% arrange(match(Gene,ISR_markers)) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("experimental.group") %>% mutate_if(is.numeric,mx_zscore2) %>% tibble::column_to_rownames("experimental.group") %>% t %>% as.data.frame %>% tibble::rownames_to_column("Gene")
```

Calculate GCN2 signature from other ISR signatures
```{r}

Common_s <- isr_z_z %>% dfilter(Gene %in% Common) %>% summarise_if(is.numeric,sum)
PRK_s <- isr_z_z %>% dfilter(Gene %in% PRK) %>% summarise_if(is.numeric,sum)
PERK_s <- isr_z_z %>% dfilter(Gene %in% PERK_try) %>% summarise_if(is.numeric,sum)
HRI_s <- isr_z_z %>% dfilter(Gene %in% HRI_sub) %>% summarise_if(is.numeric,sum)
AAS_s <- isr_z_z %>% dfilter(Gene %in% aas_sig_m) %>% summarise_if(is.numeric,sum)
gMDSC_s <- Mosely_ID8_zscore %>% dfilter(Gene %in% gMDSC_sig_m) %>% dfilter(Gene != "Defa1") %>% summarise_if(is.numeric,sum) # %>% mutate(Gene="gMDSC_score") %>% dselect(Gene, everything())
colnames(gMDSC_s) <- colnames(isr_z_z)[2:97]
mMDSC_s <- Mosely_ID8_zscore %>% dfilter(Gene %in% mMDSC_sig_m) %>% summarise_if(is.numeric,sum)
colnames(mMDSC_s) <- colnames(isr_z_z)[2:97] #c(paste0(c("ID8_T","ID8_T","ID8_T",mouse$`experimental group`),'_',1:96))

ISR_s <- bind_rows(Common_s,PRK_s,PERK_s,HRI_s,AAS_s,gMDSC_s,mMDSC_s)
rownames(ISR_s) <- c("Common","PRK","PERK","HRI","AAS","gMDSC","mMDSC")

exp_grp <- c("ID8_T","ID8_T","ID8_T",samples.info_tum$`experimental group`) # merge the cell line info for the Mosely data with the 3 ID8 lines
ISR_s_t <- ISR_s %>% t %>% data.frame %>% mutate(Group = exp_grp)
#ISR_s_t <- bind_cols(samples.info_tum,ISR_s_t)

ISR_median <- ISR_s_t %>% mutate(PRKi= ifelse(PRK<0,0,PRK), PERKi=ifelse(PERK<0,0,PERK), HRIi=ifelse(HRI<0,0,HRI)) %>% 
  mutate(GCN2_specific = Common - PRKi - PERKi - HRIi) %>% group_by(Group) %>% summarise_if(is.double,median) %>% dfilter(Group %like% "_T$") %>% arrange(desc(Common)) %>% dselect(Group,Group,Common,PRK,PERK,HRI,GCN2_specific,AAS,gMDSC,mMDSC) 

ISR_median %>% write.table(file="ISR_scores_grouped_ZZscore_Oct11.cdt",sep="\t",row.names=FALSE,quote=TRUE)
```
