---
title: "SAVER imputation with Seurat data"
author: "Mengshu"
date: "11/15/2018"
output: html_document
---

#####################
#####################
  
    IMPUTE data using SAVER: This rescues signals in medium to expressing genes, like HPK1 and GCN2

#####################
#####################
```{r}
install.packages("SAVER")
library(SAVER)
packageVersion("SAVER")
```
Use the data.table export of raw data from a Seurat object, or any matrix file can be used.

#Run time for SAVER for this dataset was 2h
$estimate stores the predicted UMIs
$se stores the standard error of the estimates
$info stores the run information
to only generate $estimate, run with command estimates.only=TRUE
```{r}
#export raw seq reads table from Seurat object pbmc
pbmc_matrix <- as.matrix(pbmc@raw.data)
# dim(pbmc_matrix)
#[1] 18408  8492
pbmc.saver <- saver(pbmc_matrix,ncores=12) #the mac desktop has 12 (virtual) cores
save(pbmc.saver,file="8kPBMC_saver.rdata")
#pbmc_matrix[1:10,1:10]
#pbmc.saver$estimate[1:10,1:10]
```

Put SAVER estimates back into the Seurat pbmc object: This has to be put into the @raw.data slot, and then run NormalizeData(), and the imputed, normalized data will be saved in pbmc@data. FeaturePlot plots from the @data slot
```{r}
#pbmc_mito@data <- pbmc.saver$estimate
pbmc@raw.data <- pbmc.saver$estimate
# Run NormalizeData
#Load imputed data
rm(pbmc)
pbmc<- readRDS("10X_8K_PBMC_Seurat_Imputed_original.rds")
FeaturePlot(object=pbmc, features.plot = c("MAP4K1"), cols.use=c("grey90","blue"), no.legend=FALSE) 
VlnPlot(object=pbmc, features.plot = c("EIF2AK4"),x.lab.rot=TRUE, point.size.use = 0.3) 
VlnPlot(object=pbmc, features.plot = c("MAP4K1"),x.lab.rot=TRUE, point.size.use = 0.3) 
ggsave("HPK1_Imputed.jpg", width= 10, height=8, dpi=600, plot= last_plot(), units = "in")
```
