---
title: "Linear_Model_M2_TCGA"
author: "Mengshu"
date: "5/1/2019"
output: html_document
---

Gene's idea: Linear model TCGA M2 data for each gene:
geneX ~ Tumor.Type + Tumor.Purity + M2.Content vs. geneX ~ Tumor.Type + Tumor.Purity

Then you would select the genes for which the M2.Content component was significant.

This basically says, for which genes is M2 content predictive of gene expression after accounting for tumor type and purity?

Load Packages
```{r}
library(tidyverse)
library(data.table)
library(magrittr)
library(dplyr)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
library(lme4)
library(lmerTest)
library(merTools)
install.packages("doMC")
library(doMC)
```

Load the data matrix
```{r}
M2 <- readRDS(file="metadata_limma_input.rds")
TCGA <- readRDS(file="tcgadata_limma_input.rds")
TCGA_t <- data.table(t(TCGA))
genes <- colnames(TCGA_t)
genes3 <- as.list(genes)
M2_TCGA <- data.table(M2,TCGA_t)
M2[1:5,1:19]
TCGA_t[1:5,1:5]
M2_TCGA[1:10,1:25]
M2_TCGA$disease <- as.factor(M2_TCGA$disease) 
M2_TCGA$purity_quant <- as.factor(M2_TCGA$purity_quant) 

```

Genes tested: 
ARF5 Pvalue: 2.11x10-7, Estimate: -1.74
CD8A Pvalue: 2.00x10-16, Estimate: -7
CSF1 Pvalue: 2.00x10-16, Estimate: 12

Genes correlated with M2_fraction when its Estimate is a positive number, with a significant P-value

This crazy syntax is required for lm() to work with lapply, because it doesn't take the variable as pure text correctly:
lm(emal(paste0(variable,'~ factor1 + factor2 + factor3'), data))
linearmodel <- function(gene_input) (lm_results <- lm(eval(paste0(gene_input, '~ disease + Purity + M2_ave')), M2_TCGA))
```{r}
linearmodel <- function(gene_input) { (lm_results <- lm(eval(paste0(gene_input, '~ disease + Purity + M2_ave')), M2_TCGA))
                                      mdl_summary <- summary(lm_results)
                                      parse_lm <- mdl_summary %>% coefficients %>% as.data.frame() %>% tail(1)
                                     return(data.frame(gene_input,parse_lm)) 
                                      }
#Leave filtering until later
%>% dfilter(Estimate > 2,`Pr(>|t|)` <= 0.05)

one_gene <- ("CSF1")
linearmodel(one_gene)
lapply(one_gene,linearmodel)

M2_lm <- lapply(genes3,linearmodel)
#multicore run
M2_lm <- mclapply(genes3,linearmodel, mc.cores=10)
M2_lm_stats <-rbind_all(M2_lm)

M2_lm_sig <- M2_lm_stats %>% dfilter(Pr...t.. < 0.05, gene_input %nlike% "RP", Estimate > 5) %>% arrange(-Estimate)
write.table(M2_lm_sig, file="M2_TCGA_linearModel_sig.txt", sep="\t", row.names=FALSE, quote=FALSE)
saveRDS(M2_lm_stats, file="M2_lm_stats.rds")

#write_out <- function(stuff) write.table(stuff,"M2_lm_results")
#test with one gene

test_lm <- lm(CD8A ~ disease + Purity + M2_ave,  data=M2_TCGA) 
test_lm <- lm(CSF1 ~ disease + Purity + M2_ave,  data=M2_TCGA) 
sum_test <- summary(test_lm)
summary(test_lm) %>% coefficients %>% as.data.frame() %>% tail(1) %>% dfilter(abs(Estimate) > 2,`Pr(>|t|)` <= 0.05)  
sum_test$coefficients

#Gene's code that could be useful for capturing data.tables from lapply
lm(...) %>% map(tidy) %>% unnest()

```

