---
title: "TCGA_Myeloid_DA"
author: "Mengshu"
date: "4/2/2019"
output: html_document
---

This script merges Immune cell infiltration data from QuanTIseq and Xcell, and tumor purity data from Aran et al. 2015 (purity calculated from ESTIMATE, SNV, CNA, and Me-DNA signatures)
In addition, TCGA patient codes, disease code will also be included.

Features cell types to cluster by:
CD8 T cell
CD4 T cell
T reg
Tumor purity
Cancer type (in some)

Immune cells being studied:
M2
DC

```{r}
library(limma)
library(dplyr)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
library(magrittr)
library(tidyr)
library(data.table) # %like% function
```

Load Saved data
```{r}
TCGA_meta_filter <- readRDS("TCGA_meta.rds")
```

Save the cleaned up TCGA expression data
```{r}
#saveRDS(meta_intersect, file="metadata_limma_input.rds")
meta <- readRDS("metadata_limma_input.rds")

#saveRDS(tcga_unique, file="tcgadata_limma_input.rds")
tcga <- readRDS("tcgadata_limma_input.rds")
```

Load Expression data
```{r}
load("/Volumes/Picard/FLX/New Targets/mergedGeneData.TCGA_TARGET_NPTumors.quantileNormalized.rdata")
```

Tumor purity data
purity_s
8665 Tumor samples
# This contains normals samples and cell lines
# Filter for just "-0[1:9][A:Z]" samples
```{r}
purity <- read.table("/Volumes/Picard/FLX/New Targets/TCGA_purity_estimates/TCGA_purity_Aran2015.txt", sep ="\t", header = TRUE)
purity_s <- purity %>% dselect(Sample.ID, Cancer.type, CPE) %>% dfilter(Sample.ID %like% "-0") %>% rename(Purity=CPE) # select just the tumor samples
head(purity_s)
grep("-11", purity_s$Sample.ID)
purity_s %<>% mutate(patientBarcode=substring(Sample.ID,1,12)) %>% dselect(patientBarcode, disease=Cancer.type, Purity, -Sample.ID)
```
Sanity check for samples: only tumor samples included
```{r}
#table(substring(purity_s$Sample.ID,13,16))
```

QuanTISeq data (Includes Cibersort-Relative data)
quanti_ciber_myeloid

Each line is duplicated
```{r}
quantiseq <- read.table("TCGA_quantiSeq.cellTypeFractionsAll.tsv", sep ="\t", header = TRUE)
qts <- quantiseq  %>% distinct(.keep_all=TRUE) 
qts_spread <- qts %>% spread(cell_type,quanTIseq_lsei_TIL10, sep=NULL)
qts_spread_ciber <- qts %>% dselect(-quanTIseq_lsei_TIL10) %>% spread(cell_type,cibersort_LM22, sep=NULL)
qts_spread_ciber %<>% dselect(patientBarcode,M2_ciber="Macrophage M2")
qts_myeloid <- qts_spread %>% select(patientBarcode, disease, CD4="CD4 T cells" ,CD8="CD8 T cells",DC="Dendritic cells",M2="Macrophage M2",Treg="Regulatory T cells")
quanti_ciber_myeloid <- left_join(qts_myeloid,qts_spread_ciber, by="patientBarcode")

hist(qts_myeloid$M2)
hist(qts_myeloid$DC)
#head(qts)
#head(quanti_ciber_myeloid)
#qts_myeloid %>% dfilter(patientBarcode=="TCGA-OR-A5J6")
```

Xcell data
```{r}
xcell <- read.table("xCell_TCGA_RSEM.txt", sep="\t", header=TRUE)
xcell_t <- data.frame(t(xcell %>% tibble::column_to_rownames("X")))
dim(xcell_t)
xcell_t[1:3,1:64]
TCGA_id <- colnames(xcell)
TCGA_id <- TCGA_id[2:length(TCGA_id)]
celltypes <- xcell$X
xcell %<>% dselect(-X)
colnames(xcell) <- ""

colnames(xcell_t) <- celltypes
rownames(xcell_t) <- TCGA_id

xcell_mye <- xcell_t %>% mutate(PatientBarcode=TCGA_id) %>% dselect(PatientBarcode,M2_xcell="Macrophages M2",DC_xcell=DC)
TCGA <- gsub('\\.',"-",TCGA_id) #gsub replaces all instances, and the escape \ needs to used twice as \\
TCGA_final <- sub("-01$","",TCGA)
head(TCGA_final)
head(xcell_mye, n=50)
#grep("-02$",TCGA)
#TCGA[6217]
xcell_mye %<>% dselect(-PatientBarcode) %>% mutate(patientBarcode=TCGA_final) %>% dselect(patientBarcode,M2_xcell,DC_xcell)
hist(xcell_mye$M2_xcell)
hist(xcell_mye$DC_xcell)
```


Merge Metadata columns

purity_s
quanti_ciber_myeloid
xcell_mye
```{r}
TCGA_meta <- left_join(purity_s,quanti_ciber_myeloid,by="patientBarcode")
TCGA_meta <- left_join(TCGA_meta,xcell_mye,by="patientBarcode")
tail(TCGA_meta, n=50)
dim(TCGA_meta)
```

Filter away missing data
7750 after filtering away quanTIseq missing data
Tumor purity missing 66 cases
7684 samples (not counting missing data in Xcell, keeping those because Xcell is only used for DC)
xcell missing 1005 cases

```{r}
TCGA_filter <- TCGA_meta %>% dfilter(!is.na(CD4))
TCGA_filter  %<>% dfilter(!is.na(Purity))

#not run yet
#TCGA_filter %>% dfilter(is.na(M2_xcell))

dim(TCGA_filter)
head(TCGA_filter)
grepl("NA",TCGA_filter) # There are NAs in the xcell data
dim(TCGA_filter %>% dfilter(is.na(M2_xcell)))

TCGA_filter %<>% mutate(M2_cbs=(1-Purity)*M2_ciber) %>% dselect(-M2_ciber) 
TCGA_filter %<>% dselect(-disease.y) 
TCGA_filter %<>% dselect(patientBarcode,disease=disease.x,Purity,CD4,CD8,Treg,M2,M2_cbs,DC_xcell,DC)
#hist(TCGA_filter$M2)
#hist(TCGA_filter$M2_xcell)
#hist(TCGA_filter$M2_cbs)
#hist(TCGA_filter$DC_xcell)
#hist(TCGA_filter$Purity)

```

Save Data

```{r}
#saveRDS(TCGA_meta_filter,"TCGA_meta.rds")
#write.table(TCGA_meta_filter,file="TCGA_meta_myeloid.txt",sep="\t",quote=FALSE,row.names = FALSE)
```

Investigate correlation between M2 and DC estimates between quantiSeq and Xcell
```{r}
library(ggplot2)

ggplot(TCGA_meta_filter, aes(M2,M2_cbs)) +
  geom_point() #+
#ggsave("TCGA_M2_scatter.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in") 
ggplot(TCGA_meta_filter, aes(DC,DC_xcell)) +
  geom_point() #+
#ggsave("TCGA_DC_scatter.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in") 

ggplot(TCGA_meta_filter, aes(M2,Purity)) #+
  geom_point() 
  
cor.test(TCGA_meta_filter$M2, TCGA_meta_filter$M2_cbs,method ="pearson")
cor.test(TCGA_meta_filter$DC, TCGA_meta_filter$DC_xcell,method ="pearson")
```


Find the good bins for Purity
```{r}
library(ggbeeswarm)
head(TCGA_meta_filter)

quartiles <- as.numeric(quantile(TCGA_meta_filter$Purity, 
                                   probs = c(0.25, 0.5, 0.75)))

ggplot(TCGA_meta_filter, aes(x=Purity, y=M2,group = factor(purity_quant))) +
  geom_point(size =0.5, colour="gray") +
  geom_violin(fill="skyblue",alpha=0.5,aes(),draw_quantiles = c(0.25, 0.5, 0.75),scale = "width") #+
#ggsave("TCGA_purity_M2_quartiles.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")  

ggplot(TCGA_meta_filter, aes(x=Purity, y=M2_cbs,group = factor(purity_quant))) +
  geom_point(size =0.5, colour="gray") +
  geom_violin(fill="lightgreen",alpha=0.5,aes(),draw_quantiles = c(0.25, 0.5, 0.75),scale = "width") #+
#ggsave("TCGA_purity_M2_cbs_quartiles.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")  

ggplot(TCGA_meta_filter, aes(x=Purity, y=M2)) +
  geom_violin(fill="skyblue", aes(group = cut_width(Purity, 0.16)))

ggplot(TCGA_meta_filter, aes(x=Purity, y=M2)) +
  geom_beeswarm(size=0.5) +
  geom_violin(fill="skyblue", alpha=0.5, aes(group = cut_width(Purity, 0.3))) 
#ggsave("TCGA_purity_quant.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")

ggplot(TCGA_meta_filter, aes(Purity)) +
  geom_histogram(aes(alpha=0.5), show.legend = FALSE, fill="blue") +
  geom_vline(xintercept = quartiles) 
#ggsave("TCGA_purity_quartiles.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")


```

Split by Purity, then find quantiles of M2 infiltration
Take the average M2 score from CIBERSORT and quanTIseq
At first I split by quartiles, this means each group is the same size, but the purity boundaries fall at 0.5, 0.7, 0.8, 0.9, which is concentrated on the upper range
the 0 to 0.5 bin will give tumor-related DE genes
```{r}
# Divide M2s into 4 and 2 groups
TCGA_meta_filter %<>% mutate(purity_quant= ntile(Purity, 4)) %>% group_by(purity_quant) %>% mutate(M2_bin2= ntile(M2,2))
TCGA_meta_filter %<>% group_by(purity_quant) %>% mutate(M2_quant= ntile(M2,4))

TCGA_meta_filter %<>% group_by(purity_quant) %>% mutate(M2_bin2= ntile(M2,2))
head(TCGA_meta_filter)


# Make groupings with cibersort M2 estimates too
TCGA_meta_filter %<>% group_by(purity_quant) %>% mutate(M2_cbs_quant= ntile(M2_cbs,4))
TCGA_meta_filter %<>% group_by(purity_quant) %>% mutate(M2_cbs_bin2= ntile(M2_cbs,2))

TCGA_meta_filter %>% group_by(purity_quant,M2_quant) %>% summarize(mean_purity=mean(Purity),meanM2=mean(M2))
#plot(quants$meanM2,quants$mean_purity) 
#ggsave("TCGA_purity_M2_quantile.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")

#Average M2 score
TCGA_meta_filter %<>% mutate(M2_ave= (M2+M2_cbs)/2) %>% group_by(purity_quant) %>% mutate(M2_ave_quant= ntile(M2_ave,4))
TCGA_meta_filter %<>% mutate(M2_ave= (M2+M2_cbs)/2) %>% group_by(purity_quant) %>% mutate(M2_ave_bin2= ntile(M2_ave,2))
```

Parse the TCGA expression data
44201 11348
TCGA starts at 814 11348
1) select TCGA samples
2) The meta data has sample types: the TCGA data has -03 samples which are blood cancers, remove those
-01A -01B -01C -01D -01R -02A -02B -05A -06A -06B -07A 
8665  221   26    1    1   51    6    2  388    2    1 
data.table converts - into .

Remaining samples:
 -01  -02  -05  -06  -07  -20 
9186   44   11  392    1    1 
```{r}
dim(mergedGeneData.TCGA_TARGET_NPTumors.quantileNormalized)
mergedGeneData.TCGA_TARGET_NPTumors.quantileNormalized[1:10,11290:11300]
tcga <- mergedGeneData.TCGA_TARGET_NPTumors.quantileNormalized[,814:11348]
sample_name_format <- colnames(mergedGeneData.TCGA_TARGET_NPTumors.quantileNormalized[,814:11348])
tcga <- data.frame(tcga)
colnames(tcga) <- sample_name_format
tcga %<>% dselect(everything(), -ends_with("-03"),-ends_with("-04"),-ends_with("-09")) #10362 samples
tcga %<>% dselect(everything(), -ends_with("-10"),-ends_with("-11"),-ends_with("-12"),-ends_with("-13"),-ends_with("-14")) # get rid of normals 9635 samples
tcga %<>% dselect(everything(), -ends_with("-20")) #9634 samples

table(substring(colnames(tcga),13,16))
tcga[1:5,9000:9010]
dim(tcga)
```


Reorder the data to match the metadata order
There are duplicates in both the metadata and TCGA data after the sample-type extensions are removed
[1] 7684
[1] 7647
[1] 9634
[1] 9561
final tcga: 9634 44202
Final intersect of meta data and tcga data: 7012 samples
```{r}
sample_order <- TCGA_meta_filter$patientBarcode
tcga_presort_samples <- colnames(tcga)
tcga_presort_samples_short <- substring(tcga_presort_samples,1,12)
colnames(tcga) <- tcga_presort_samples_short #shorten TCGA names in tcga

length(sample_order)
length(unique(sample_order))
length(tcga_presort_samples)
length(unique(tcga_presort_samples_short))

TCGA_meta_filter %<>% distinct(patientBarcode,.keep_all=TRUE) 
length(TCGA_meta_filter$patientBarcode) #7667
#Find the indexes where there are duplicates
dups <- grep("TRUE",duplicated(colnames(tcga))) #this gives the duplicated occurences, leaving the first unique one unlisted
#tcga %>% select(dups) #error
#tcga %<>% dselect(sample_order) #error
tcga_t <- t(tcga)
tcga_t <- tibble::rownames_to_column(data.frame(tcga_t)) 
tcga_t %<>% distinct(rowname,.keep_all=TRUE) 
rownames(tcga_t) <- tcga_t$rowname
tcga_t %<>% dselect(-rowname)
dim(tcga_t) #9634 44202
tcga_unique <- t(tcga_t) 
#now the dashes have changed to . again, change back and replace
dots <- colnames(tcga_unique)
dashes <- gsub('\\.',"-",dots) 
colnames(tcga_unique) <- dashes
tcga_unique <- data.frame(tcga_unique)
#tcga_unique %<>% dselect(sample_order) # error due to missing samples
intersect_tcga <- dplyr::intersect(colnames(tcga_unique),sample_order)
length(intersect_tcga) #7012
```
Select samples for final meta-data and expression-data matrices
tcga_unique
meta_intersect
```{r}
#Expression data
tcga_unique %<>% dselect(intersect_tcga)

#Meta data
TCGA_meta_filter %<>% mutate(Group=LETTERS[2*purity_quant+M2_ave_bin2-2]) #THIS WILL GIVE ABCDEFGH
# Add column Group_QTS to group by QuanTIseq and Group_CBS to group by CIBERSORT
#meta %<>% mutate(Group_QTS=LETTERS[2*purity_quant+M2_bin2-2],Group_CBS=LETTERS[2*purity_quant+M2_cbs_bin2-2]) 
#meta %>% dselect(Group_QTS) %>% table
#meta %>% dselect(Group_CBS) %>% table

head(TCGA_meta_filter, n=50)
intersect_tcga_factor <- factor(intersect_tcga, levels=intersect_tcga)
#select the intersect samples
intersect_tcga <- unique(intersect_tcga)
TCGA_meta_filter <- data.frame(TCGA_meta_filter) #######THIS IS CRITICAL, DON'T KEEP THE DATA AS A TIBBLE, IT DOES MESSED UP SHIT!!!!
meta_intersect <- TCGA_meta_filter %>% distinct(patientBarcode,.keep_all=TRUE) %>% filter(patientBarcode %in% intersect_tcga) %>% arrange(factor(patientBarcode, levels=intersect_tcga))
dim(meta_intersect)

meta_intersect$patientBarcode == intersect_tcga
meta_intersect$patientBarcode == colnames(tcga_unique)



```

