---
title: "Mosaic_graphing.Rmd"
author: "MX"
date: "01/04/2021"
output: html_document
---

Parse and graph FLX475 clinical biomarker data on CD8 and FOXP3 IHC stains from patient biopsy cores

```{r}
library(data.table)
library(magrittr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(stringr)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
library(readxl)
```

Load Mosaic data Excel files
```{r}
#/Volumes/Picard/FLX/Mosaic/FLX004 FLX475-02 Raw Data_30Sep2020 RELEASE[1].xlsx  had 191 lines of data, current version has 253 lines
CD8_FOXP3 <- readxl::read_xlsx("/Volumes/Picard/FLX/Mosaic/FLX004 FLX475-02 Raw Data_16Dec2020 RELEASE.xlsx", sheet="MOS487-RD FoxP3+CD8", skip=1,trim_ws = TRUE) %>% dfilter(!is.na(`Subject ID`)) %>% mutate_if(grepl("mm|Percent|Total|%",names(.)),as.numeric) # |Total|Percent|% 
meta_data <- readxl::read_xlsx("/Volumes/Picard/FLX/Mosaic/FLX004 FLX475-02 Raw Data_16Dec2020 RELEASE.xlsx", sheet="Inventory") %>% dfilter(!is.na(`Subject ID`))  #Mosaic_ID
HandE <- readxl::read_xlsx("/Volumes/Picard/FLX/Mosaic/FLX004 FLX475-02 Raw Data_16Dec2020 RELEASE.xlsx", sheet="H&E", skip=1) %>% dfilter(!is.na(`Subject ID`)) %>% mutate_if(grepl("mm|Percent|Total|%",names(.)),as.numeric)  #Mosaic_ID

```


Response score (CONFIDENTIAL)
```{r}
best_response <- readxl::read_xlsx("/Volumes/Picard/FLX/Mosaic/Response_Assessments_04JAN2021.xlsx") 

best_response %<>% mutate(Crossover= if_else(grepl("cross",`FLX475-02 Subject ID`),"Crossover","Original" ) ) %>% separate(`FLX475-02 Subject ID`, into=c("Subject_ID",NA),sep=" ") %>% separate(`Dose Level`, into=c("Dose",NA),sep="mg") %>% separate(`Best Overall Response (Target lesion)`,into=c("Best_Response",NA), sep=" ") %>% separate(`BOR Target (best % change from baseline)`,into=c("BOR",NA),sep=" ") %>% mutate(Best_Response_score= as.numeric(BOR) ) %>% dselect(Subject_ID,"Primary Tumor Type","Stage Phase",Dose, `Mono/Combo`,Crossover,Best_Response,Best_Response_score) %>% dfilter(!is.na(Subject_ID))

best_response_precross <- best_response %>% dfilter(`Mono/Combo` !="Crossover combo")

```


Merge data from Sheets into one
```{r}
library(readr)
#CD8_FOXP3 %>% left_join(meta_data %>% dselect("Mosaic ID","Collection Date","Tissue Site"), by=c("Mosaic ID")) %>% left_join(HandE %>% dselect("Mosaic ID","H&E Quality (P/F/L)","% Cancer Nuclei","% Cancer by Area","% Necrosis"), by="Mosaic ID") %>% write.table("FLX004 FLX475-02 Raw Data_28Oct2020 Merged.txt",sep="\t",quote=F,row.names = F)
ihc <- CD8_FOXP3 %>% left_join(meta_data %>% dselect("Mosaic ID","Collection Date","Tissue Site"), by=c("Mosaic ID")) %>% left_join(HandE %>% dselect("Mosaic ID","H&E Quality (P/F/L)","% Cancer Nuclei","% Cancer by Area","% Necrosis"), by="Mosaic ID")
#fix column header names
param <- gsub(" / "," per ",colnames(ihc))
param <- gsub(" |\\/","_",param)
param <- gsub("\\^|\\+|\\(|\\)","",param)
colnames(ihc) <- param
ihc %>% write.table("FLX004_FLX475-02_Data_16Dec2020_Parsed.txt",sep="\t",quote=F,row.names = F)
save(ihc, file="FLX004_FLX475-02_Data_16Dec2020_Parsed.rdata")
load("FLX004_FLX475-02_Data_16Dec2020_Parsed.rdata")
```


Calculate Total CD8 and FOXP3, and ratios
Add 1 to everything to get real number ratios
Parse the Visit into pre- post- and make into factor with levels
```{r}
ihc_1 <- ihc %>% mutate(Total_CD8_cells = Total_Cells*CD8_Percent_Positive*0.01+1, 
               Total_FOXP3_cells = Total_Cells*FoxP3_Percent_Positive*0.01+1, 
               CD8_FOXP3_ratio = Total_CD8_cells/Total_FOXP3_cells,
               FOXP3_CD8_ratio = Total_FOXP3_cells/Total_CD8_cells,
               CD8_FOXP3_log2 = log2(CD8_FOXP3_ratio),
               FOXP3_CD8_log2 = log2(FOXP3_CD8_ratio)
               ) %>% 
                mutate(Visit_type = case_when(
                Subject_ID == "2102-001" & grepl("Unscheduled",Visit) ~ "Pre_treat",
                grepl("Unscheduled",Visit) ~ "Follow_up",
                grepl("Screen",Visit) ~ "Pre_treat",
                grepl("C3D1",Visit) ~ "Follow_up"
                )) %>% 
                mutate(Visit_type = factor(Visit_type, levels=c("Pre_treat","Follow_up"))) %>%
                mutate(Combo = case_when(
                grepl("1a ",Visit) ~ "single_agent",
                grepl("1b ",Visit) ~ "combo_agent",
                grepl("2a ",Visit) ~ "single_agent",
                grepl("2b ",Visit) ~ "combo_agent"
                )) %>%
                mutate(Phase = case_when(
                grepl("1a ",Visit) ~ "1a",
                grepl("1b ",Visit) ~ "1b",
                grepl("2a ",Visit) ~ "2a",
                grepl("2b ",Visit) ~ "2b"
                )) %>%
                arrange(Subject_ID) %>%
                fill(Combo,.direction="down") %>%
                fill(Phase,.direction="down") %>%
                mutate(Combo=factor(Combo, levels=c("single_agent","combo_agent"))) %>%
                dfilter(Anatomical_Location != "-") %>% arrange(Subject_ID)

```

Add in extra columns of mean log2 #, density, and percent cell data, and take their means for lines

```{r}
ihc_1 %<>% dfilter(!is.na(CD8_Percent_Positive)) %>%
          mutate(CD8_FoxP3_percent_log2=log2((CD8_Percent_Positive+1)/(FoxP3_Percent_Positive+1)),
                 CD8_FoxP3_density_log2=log2((CD8_Cells_per_mm2+1)/(FoxP3_Cells_per_mm2+1)) ) %>%
          group_by(Subject_ID,Visit_type,Anatomical_Location) %>% 
          mutate(CD8_FOXP3_log2_mean = mean(CD8_FOXP3_log2),
                 CD8_FoxP3_percent_log2_mean = mean(CD8_FoxP3_percent_log2, na.rm=T),
                 CD8_FoxP3_density_log2_mean = mean(CD8_FoxP3_density_log2, na.rm=T) ) %>%
          mutate(CD8_density_mean = mean(CD8_Cells_per_mm2,na.rm=T),
               FoxP3_density_mean = mean(FoxP3_Cells_per_mm2,na.rm=T),
               CD8_percent_mean = mean(CD8_Percent_Positive,na.rm=T),
               FoxP3_percent_mean = mean(FoxP3_Percent_Positive,na.rm=T) ) %>%
          ungroup()
save(ihc_1, file="FLX004_FLX475-02_Data_16Dec2020_Parsed_1.rdata")
load("FLX004_FLX475-02_Data_16Dec2020_Parsed_1.rdata")
```



Rearrange columns
```{r}
ihc_1 %<>% dselect(Source_ID,Subject_ID,Tissue_Type,Visit,Visit_type,Collection_Date,Stain_Date,Phase,Combo,Tissue_Site,everything())
```



Examine data for matched samples

Does sample have paired data?
36 patients
18 patients have matched pre-screen and post-treat
 patients have matched pre-screen and post-treat samples at the same tissue site

0102-006 Phase 1b has post-treat data that failed, remove

Matched tissues that don't match on word string
Count these as the same site
0104-002	Abdomen			
0104-002	Abdominal

0101-004	Abdomen			
0101-004	Peritoneal Abdominal Wall
```{r}
#table of all patients by phase
ihc_1 %>% dselect(Subject_ID,Phase) %>% unique %>% arrange(Subject_ID) %>% dselect(Phase) %>% table
#unmatched and matched patients
unmatch <- ihc_1 %>% dselect(Subject_ID,Visit_type) %>% unique %>% group_by(Subject_ID) %>% summarise(Num_sample=n()) %>% dfilter(Num_sample == 1) %>% pull(Subject_ID)
matched <- ihc_1 %>% dselect(Subject_ID,Visit_type) %>% unique %>% group_by(Subject_ID) %>% summarise(Num_sample=n()) %>% dfilter(Num_sample > 1) %>% pull(Subject_ID)
ihc_1 %>% dselect(Subject_ID,Visit_type) %>% unique %>% group_by(Subject_ID) %>% summarise(Num_sample=n()) %>% ungroup %>% mutate(Num_sample=factor(Num_sample))
# table of total unmatched and matched patients
ihc_1 %>% dselect(Subject_ID,Visit_type) %>% unique %>% group_by(Subject_ID) %>% summarise(Num_sample=n()) %>% dselect(Num_sample) %>% table
#table of unmatched patients by phase and which Visit is missing
ihc_1 %>% dselect(Subject_ID,Visit_type,Phase) %>% dfilter(Subject_ID %in% unmatch) %>% unique %>% dselect(Visit_type,Phase) %>% table
ihc_1 %>% dselect(Subject_ID,Visit_type,Phase) %>% dfilter(Subject_ID %in% matched) %>% unique %>% dselect(Visit_type,Phase) %>% table

tissue_matched <- ihc_1 %>% dselect(Subject_ID,Visit_type, Tissue_Site) %>% dfilter(Subject_ID %in% matched) %>% unique %>% group_by(Subject_ID,Tissue_Site) %>% summarise(Num_sample=n()) %>% dfilter(Num_sample == 2) %>% pull(Subject_ID)

ihc %>% dselect(Subject_ID,Visit_type, Tissue_Site,Phase) %>% dfilter(Subject_ID %in% matched) %>% unique %>% group_by(Subject_ID,Tissue_Site,Phase) %>% summarise(Num_sample=n()) %>% ungroup %>% unique %>% dselect(-Tissue_Site,-Subject_ID) %>% table
# 10 mathed exactly, 2 matched by meaning of words
tissue_matched <- c(tissue_matched,"0104-002","0101-004")
length(tissue_matched)
```

Add in column of Screening values to normalize to
44 rows of tissue-matched screening data

ihc_1 %<>% dfilter(!is.na(CD8_Percent_Positive)) %>%
          mutate(CD8_FoxP3_percent_log2=log2((CD8_Percent_Positive+1)/(FoxP3_Percent_Positive+1)),
                 CD8_FoxP3_density_log2=log2((CD8_Cells_per_mm2+1)/(FoxP3_Cells_per_mm2+1)) ) %>%
          group_by(Subject_ID,Visit_type,Anatomical_Location) %>% 
          mutate(CD8_FOXP3_log2_mean = mean(CD8_FOXP3_log2),
                 CD8_FoxP3_percent_log2_mean = mean(CD8_FoxP3_percent_log2, na.rm=T),
                 CD8_FoxP3_density_log2_mean = mean(CD8_FoxP3_density_log2, na.rm=T) ) %>%
          mutate(CD8_density_mean = mean(CD8_Cells_per_mm2,na.rm=T),
               FoxP3_density_mean = mean(FoxP3_Cells_per_mm2,na.rm=T),
               CD8_percent_mean = mean(CD8_Percent_Positive,na.rm=T),
               FoxP3_percent_mean = mean(FoxP3_Percent_Positive,na.rm=T) ) %>%
          ungroup()
          
          
```{r}
count_screen <- ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% select(Subject_ID,Visit_type,Anatomical_Location,CD8_FOXP3_log2_mean,CD8_FoxP3_percent_log2_mean,CD8_FoxP3_density_log2_mean,CD8_density_mean,FoxP3_density_mean,CD8_percent_mean,FoxP3_percent_mean) %>% dfilter(Visit_type=="Pre_treat") %>% dselect(-Visit_type) %>% group_by(Subject_ID,Anatomical_Location) %>% 
  mutate(CD8_FOXP3_log2_screen= CD8_FOXP3_log2_mean,
         CD8_FoxP3_percent_screen= CD8_FoxP3_percent_log2_mean,
         CD8_FoxP3_density_screen=CD8_FoxP3_density_log2_mean) %>% dselect(-CD8_FOXP3_log2_mean,-CD8_FoxP3_percent_log2_mean,-CD8_FoxP3_density_log2_mean) %>% mutate(CD8_density_screen=CD8_density_mean,
         FoxP3_density_screen=FoxP3_density_mean,
         CD8_percent_screen= CD8_percent_mean,
         FoxP3_percent_screen=FoxP3_percent_mean
         ) %>% 
  dselect(-CD8_density_mean,-FoxP3_density_mean,-CD8_percent_mean,-FoxP3_percent_mean) %>% ungroup

count_screen_all <- ihc_1 %>% select(Subject_ID,Visit_type,Anatomical_Location,CD8_FOXP3_log2_mean,CD8_FoxP3_percent_log2_mean,CD8_FoxP3_density_log2_mean,CD8_density_mean,FoxP3_density_mean,CD8_percent_mean,FoxP3_percent_mean) %>% dfilter(Visit_type=="Pre_treat") %>% dselect(-Visit_type) %>% group_by(Subject_ID,Anatomical_Location) %>% 
  mutate(CD8_FOXP3_log2_screen= CD8_FOXP3_log2_mean,
         CD8_FoxP3_percent_screen= CD8_FoxP3_percent_log2_mean,
         CD8_FoxP3_density_screen=CD8_FoxP3_density_log2_mean) %>% dselect(-CD8_FOXP3_log2_mean,-CD8_FoxP3_percent_log2_mean,-CD8_FoxP3_density_log2_mean) %>% mutate(CD8_density_screen=CD8_density_mean,
         FoxP3_density_screen=FoxP3_density_mean,
         CD8_percent_screen= CD8_percent_mean,
         FoxP3_percent_screen=FoxP3_percent_mean
         ) %>% 
  dselect(-CD8_density_mean,-FoxP3_density_mean,-CD8_percent_mean,-FoxP3_percent_mean) %>% ungroup

ihc_1_ratio <- ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% left_join(count_screen, by=c("Subject_ID","Anatomical_Location")) %>% unique %>% 
  mutate(CD8_FOXP3_log2_norm=CD8_FOXP3_log2_mean-CD8_FOXP3_log2_screen,
CD8_FOXP3_percent_log2_norm=CD8_FoxP3_percent_log2_mean-CD8_FoxP3_percent_screen,
CD8_FOXP3_density_log2_norm=CD8_FoxP3_density_log2_mean-CD8_FoxP3_density_screen) # %>% dfilter(Visit_type=="Follow_up")     # Only put in last term for Paul, to filter out the pre-screen 0s 

save(ihc_1_ratio, file="ihc_1_ratio_Dec162020.rdata")
load("ihc_1_ratio_Dec142020.rdata")
ihc_1_ratio %>% write.table(file="ihc_ratio_Dec162020.txt", row.names = F, sep="\t", quote=F)
```

```{r}
colnames(ihc_1_ratio)
```

Prism output
```{r}
Paul <- ihc_1_ratio %>% dfilter(Visit_type=="Follow_up") %>% mutate(CD8_FOXP3_ratio=2^(CD8_FOXP3_log2_norm),CD8_FOXP3_density_ratio=2^(CD8_FOXP3_density_log2_norm),CD8_FOXP3_percent_ratio=2^(CD8_FOXP3_percent_log2_norm) ) %>% dselect(Subject_ID, Tissue_Type, Phase,Combo, Anatomical_Location,CD8_FOXP3_log2_norm,CD8_FOXP3_density_log2_norm,CD8_FOXP3_percent_log2_norm,CD8_FOXP3_ratio,CD8_FOXP3_density_ratio,CD8_FOXP3_percent_ratio) %>% unique
Paul %>% write.table(file="Mosaic_121620_Prism_all_TME.txt",sep="\t", quote=F, row.names = F)

```

LOLLIPOP PLOTS
Grouping parameters for graphing

Phase: Part 1a, Part 1b, Part 2a, Part 2b
Time: Screen, C3D1
Site: Tumor, Stroma, All

Data variables:
Cell stain+ absolute number
Cell stain+ percent
Cell stain+ density

Log2 Ratios for each of these
Normalized Log2 Ratios for each (norm to Screen sample mean)

Ways to deal with 0 values in Ratios:
+1
+0.1

Plot FoxP3 per area,  paired by time, faceted by patients
Plot FoxP3 percent,  paired by time, faceted by patients
```{r fig.height=5, fig.width=12}
#DensityN
ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% 
  ggplot(aes(x=Anatomical_Location,y=FoxP3_Cells_per_mm2, ymin=0, ymax=FoxP3_Cells_per_mm2, color=Visit_type, group=interaction(Visit_type,Anatomical_Location))) + 
  geom_point(position=position_dodge2(width = 0.5)) +  # geom_jitter(width=0.3, height=0) + 
  geom_linerange(position=position_dodge2(width = 0.5)) +
  facet_wrap(~Subject_ID, scales="free_y", ncol=4) + 
  scale_color_manual(values=c("#5B666F","#0A95A7")) + 
  theme(axis.text.x = element_text(angle=45,size=12, vjust=1, hjust=1),strip.text=element_text(size=12))
#Percent
ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% 
  ggplot(aes(x=Anatomical_Location,y=FoxP3_Percent_Positive, ymin=0, ymax=FoxP3_Percent_Positive, color=Visit_type, group=interaction(Visit_type,Anatomical_Location))) + 
  geom_point(position=position_dodge2(width = 0.5)) +  # geom_jitter(width=0.3, height=0) + 
  geom_linerange(position=position_dodge2(width = 0.5)) +
  facet_wrap(~Subject_ID, scales="free_y", ncol=4) + 
  scale_color_manual(values=c("#5B666F","#0A95A7")) + 
  theme(axis.text.x = element_text(angle=45,size=12, vjust=1, hjust=1),strip.text=element_text(size=12))

```

Plot CD8 per area, paired by time, grouped by Site, faceted by patient
Plot CD8 percent positive, paired by time, grouped by Site, faceted by patient

```{r fig.height=5, fig.width=12}
ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% 
  ggplot(aes(x=Anatomical_Location,y=CD8_Cells_per_mm2, ymin=0, ymax=CD8_Cells_per_mm2, color=Visit_type, group=interaction(Visit_type,Anatomical_Location))) + 
  geom_point(position=position_dodge2(width = 0.5)) +  # geom_jitter(width=0.3, height=0) + 
  geom_linerange(position=position_dodge2(width = 0.5)) +
  facet_wrap(~Subject_ID, scales="free_y", ncol=4) + 
  scale_color_manual(values=c("#5B666F","#0A95A7")) + 
  theme(axis.text.x = element_text(angle=45,size=12, vjust=1, hjust=1),strip.text=element_text(size=12))

ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% 
  ggplot(aes(x=Anatomical_Location,y=CD8_Percent_Positive, ymin=0, ymax=CD8_Percent_Positive, color=Visit_type, group=interaction(Visit_type,Anatomical_Location))) + 
  geom_point(position=position_dodge2(width = 0.5)) +  # geom_jitter(width=0.3, height=0) + 
  geom_linerange(position=position_dodge2(width = 0.5)) +
  facet_wrap(~Subject_ID, scales="free_y", ncol=4) + 
  scale_color_manual(values=c("#5B666F","#0A95A7")) + 
  theme(axis.text.x = element_text(angle=45,size=12, vjust=1, hjust=1),strip.text=element_text(size=12))
```
############
RATIOS
#############

LOLLIPOP PLOTS: NORMALIZED

Plot CD8 to FoxP3 cell number ratio, paired by time, grouped by Site, faceted by patient
ihc has ratios from + 0.1
ihc_1 has ratios from + 1
```{r fig.height=5, fig.width=10}
#CD8/FoxP3
library(scales)
ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% 
  ggplot(aes(x=Anatomical_Location,y=CD8_FOXP3_log2, ymin=0,ymax=CD8_FOXP3_log2,color=Visit_type, group=Visit_type)) + 
  geom_point(aes(size=Total_CD8_cells), position=position_dodge2(width = 0.75)) + 
  scale_size_continuous(breaks=c(0,10,100,500,1000,2000,4000)) +
  geom_linerange(position=position_dodge2(width = 0.75)) +
  geom_hline(yintercept=0) +
  facet_wrap(~Subject_ID, ncol=4) + 
  scale_color_manual(values=c("#5B666F","#0A95A7")) + 
  theme_bw()

ihc_1 %>% dfilter(Subject_ID %in% MatchedTissue_Subject_ID) %>% 
  ggplot(aes(x=Anatomical_Location,y=CD8_FOXP3_log2, ymin=0,ymax=CD8_FOXP3_log2,color=Visit_type, group=Visit_type)) + 
  geom_point(aes(size=Total_CD8_cells), position=position_dodge2(width = 0.75)) + 
  geom_linerange(position=position_dodge2(width = 0.75)) +
  geom_hline(yintercept=0) +
  #facet_wrap(~Subject_ID, ncol=4) + 
  scale_color_manual(values=c("#5B666F","#0A95A7")) + 
  theme_bw()

#FoxP3/CD8
ihc %>% dfilter(Subject_ID %in% MatchedTissue_Subject_ID) %>% 
  ggplot(aes(x=Anatomical_Location,y=FOXP3_CD8_log2, ymin=0,ymax=FOXP3_CD8_log2,color=Visit_type, group=Visit_type)) + 
  geom_point(aes(size=Total_FOXP3_cells),position=position_dodge2(width = 0.75)) + 
  geom_linerange(position=position_dodge2(width = 0.75)) +
  geom_hline(yintercept=0) +
  #facet_wrap(~Subject_ID, ncol=4) + 
  scale_color_manual(values=c("#5B666F","#0A95A7")) + 
  theme_bw() +
  theme(axis.text.x = element_text(size=12, vjust=1, hjust=1,strip.text=element_text(size=12)), 
        axis.text.y =element_text(size=12))

ihc_1 %>% dfilter(Subject_ID %in% MatchedTissue_Subject_ID) %>% 
  ggplot(aes(x=Anatomical_Location,y=FOXP3_CD8_log2, ymin=0,ymax=FOXP3_CD8_log2,color=Visit_type, group=Visit_type)) + 
  geom_point(aes(size=Total_FOXP3_cells),position=position_dodge2(width = 0.75)) + 
  geom_linerange(position=position_dodge2(width = 0.75)) +
  geom_hline(yintercept=0) +
  #facet_wrap(~Subject_ID, ncol=4) + 
  scale_color_manual(values=c("#5B666F","#0A95A7")) + 
  theme_bw() +
  theme(axis.text.x = element_text(size=12, vjust=1, hjust=1,strip.text=element_text(size=12) ), 
        axis.text.y = element_text(size=12))
```

CHAOS CRIS-CROSS PLOTS

Plot by total population CD8/FoxP3

CD8_FOXP3_log2_mean 
CD8_FoxP3_percent_log2_mean 
CD8_FoxP3_density_log2_mean 
```{r fig.height=4, fig.width=6}
#counts
ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% 
      ggplot(aes(x=Visit_type,y=CD8_FOXP3_log2)) + 
      stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,geom = "crossbar", width = 0.5) +
      geom_boxplot(aes()) + 
      geom_point(aes(size=Total_CD8_cells)) + #aes(size=Total_CD8_cells, colour=Subject_ID)
      geom_line(aes(x=Visit_type,y=CD8_FOXP3_log2_mean, group=Subject_ID, color=Phase,size=2)) +
      #geom_line(aes(group=interaction(Subject_ID,Visit_type,Anatomical_Location), color=Subject_ID),linetype="dashed",size=1.5) +
      facet_wrap(~Anatomical_Location) + 
      scale_size_continuous(breaks = c(1,10,100, 1000)) + #this is the size of the points
      scale_fill_manual(values=c("#0A95A7","#E5541B","#0A95A7","#E5541B")) +
      scale_color_manual(values=c("#0A95A7","#E5541B","#0A95A7","#E5541B",blue="#0A95A7", orange="#E5541B", gray="#5B666F", yellow="#FE9F33", green="#4DA167", blue="#0A95A7",orange="#E5541B",blue="#0A95A7")) +
      theme_bw() 
#percent
ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% 
      ggplot(aes(x=Visit_type,y=CD8_FoxP3_percent_log2)) + 
      stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,geom = "crossbar", width = 0.5) +
      geom_boxplot(aes()) + 
      geom_point(aes(size=Total_CD8_cells)) + #aes(size=Total_CD8_cells, colour=Subject_ID)
      geom_line(aes(x=Visit_type,y=CD8_FoxP3_percent_log2_mean, group=Subject_ID, color=Phase,size=2)) +
      #geom_line(aes(group=interaction(Subject_ID,Visit_type,Anatomical_Location), color=Subject_ID),linetype="dashed",size=1.5) +
      facet_wrap(~Anatomical_Location) + 
      scale_size_continuous(breaks = c(1,10,100, 1000)) + #this is the size of the points
      scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      scale_color_manual(values=c("#0A95A7","#E5541B","#0A95A7","#E5541B",blue="#0A95A7", orange="#E5541B", gray="#5B666F", yellow="#FE9F33", green="#4DA167", blue="#0A95A7",orange="#E5541B",blue="#0A95A7")) +
      theme_bw() 
#density
ihc_1 %>% dfilter(Subject_ID %in% tissue_matched) %>% 
      ggplot(aes(x=Visit_type,y=CD8_FoxP3_density_log2)) + 
      stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,geom = "crossbar", width = 0.5) +
      geom_boxplot(aes()) + 
      geom_point(aes(size=Total_CD8_cells)) + #aes(size=Total_CD8_cells, colour=Subject_ID)
      geom_line(aes(x=Visit_type,y=CD8_FoxP3_density_log2_mean, group=Subject_ID, color=Phase,size=2)) +
      #geom_line(aes(group=interaction(Subject_ID,Visit_type,Anatomical_Location), color=Subject_ID),linetype="dashed",size=1.5) +
      facet_wrap(~Anatomical_Location) + 
      scale_size_continuous(breaks = c(1,10,100, 1000)) + #this is the size of the points
      scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      scale_color_manual(values=c("#0A95A7","#E5541B","#0A95A7","#E5541B",blue="#0A95A7", orange="#E5541B", gray="#5B666F", yellow="#FE9F33", green="#4DA167", blue="#0A95A7",orange="#E5541B",blue="#0A95A7")) +
      theme_bw() 

# Only plot the two samples that decrease
ihc_1 %>% dfilter(Subject_ID %in% c("0111-002","1802-006","0104-002"), Anatomical_Location =="ALL" ) %>% 
      ggplot(aes(x=Visit_type,y=CD8_FoxP3_percent_log2)) + 
      #stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,geom = "crossbar", width = 0.5) +
      #geom_boxplot(aes()) + 
      geom_point(aes(size=Total_CD8_cells,color=Phase,alpha=0.2)) + #aes(size=Total_CD8_cells, colour=Subject_ID)
      geom_line(aes(x=Visit_type,y=CD8_FoxP3_percent_log2_mean, group=Subject_ID, color=Phase),size=1.5) +
      #geom_line(aes(group=interaction(Subject_ID,Visit_type,Anatomical_Location), color=Subject_ID),linetype="dashed",size=1.5) +
      facet_wrap(~Anatomical_Location) + 
      scale_size_continuous(breaks = c(1,10,100, 1000)) + #this is the size of the points
      scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      scale_color_manual(values=c("#0A95A7","#E5541B","#0A95A7","#E5541B",blue="#0A95A7", orange="#E5541B", gray="#5B666F", yellow="#FE9F33", green="#4DA167", blue="#0A95A7",orange="#E5541B",blue="#0A95A7")) +
      theme_bw() 

ihc_1 %>% dfilter(Subject_ID %in% c("0111-002","1802-006"),Anatomical_Location=="ALL" ) %>% dselect(Subject_ID,Visit,Anatomical_Location,Total_CD8_cells, Total_FOXP3_cells) %>% mutate(Total_CD8_cells=as.integer(Total_CD8_cells),Total_FOXP3_cells=as.integer(Total_FOXP3_cells))

ihc_1 %>% dfilter(Subject_ID %in% c("0104-002"),Anatomical_Location=="ALL" ) %>% dselect(Subject_ID,Visit,Anatomical_Location,Total_CD8_cells, Total_FOXP3_cells) %>% mutate(Total_CD8_cells=as.integer(Total_CD8_cells),Total_FOXP3_cells=as.integer(Total_FOXP3_cells))
```

HALF STARBURST PLOTS
Plot Absolute number Log ratios normalized to Screening Time

```{r}
ihc_1_ratio %>% dfilter(Subject_ID %in% tissue_matched) %>% left_join(best_response, by="Subject_ID") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR"))) %>% 
      ggplot(aes(x=Visit_type,y=CD8_FOXP3_log2_norm)) + 
      geom_point() +
      #geom_label(aes(label=Subject_ID, alpha=0.9)) + #aes(size=Total_CD8_cells, colour=Subject_ID)
      geom_line(aes(x=Visit_type,y=CD8_FOXP3_log2_norm, group=Subject_ID, color=Best_Response),size=1) +
      facet_wrap(~Anatomical_Location) + #, scale="free_y"
      scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7", "#FE9F33")) +
      theme_bw() #+
      #expand_limits(y=c(-40,30))
```


HALF STARBURST PLOTS
Percent
Fig scale breaks in the legend for point size
```{r}
ihc_1_ratio %>% dfilter(Subject_ID %in% tissue_matched) %>% left_join(best_response, by="Subject_ID") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR"))) %>% 
      ggplot(aes(x=Visit_type,y=CD8_FOXP3_percent_log2_norm)) + 
      geom_point() +
      #geom_label(aes(label=Subject_ID, alpha=0.2)) + 
      geom_line(aes(x=Visit_type,y=CD8_FOXP3_percent_log2_norm, group=Subject_ID,color=Best_Response),size=1) +
      #geom_text(aes(label=Subject_ID)) +
      facet_wrap(~Anatomical_Location) + 
      #scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      scale_size_continuous(breaks = c(1,10,100, 1000)) + #this is the size of the points
      scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7", "#FE9F33","#4DA167", "#947EB0","#083D77","#356E46", "#69597D", "#052344")) +
      theme_bw() #+
      #expand_limits(y=c(-40,30))
```
HALF STARBURST PLOTS
Density

```{r}
ihc_1_ratio %>% dfilter(Subject_ID %in% tissue_matched) %>% left_join(best_response, by="Subject_ID") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR"))) %>%
      ggplot(aes(x=Visit_type,y=CD8_FOXP3_density_log2_norm)) + 
      #stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,geom = "crossbar", width = 0.5) +
      #geom_boxplot(aes(alpha=0.9)) + 
      geom_point() + #aes(size=Total_CD8_cells, colour=Subject_ID)
      #geom_label(aes(label=Subject_ID, alpha=0.2)) + 
      geom_line(aes(x=Visit_type,y=CD8_FOXP3_density_log2_norm, group=Subject_ID,color=Best_Response),size=1) +
      #geom_text(aes(label=Subject_ID)) +
      #geom_line(aes(group=interaction(Subject_ID,Visit_type,Anatomical_Location), color=Subject_ID),linetype="dashed",size=1.5) +
      facet_wrap(~Anatomical_Location) + #, scale="free_y"
      scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7", "#FE9F33", "#4DA167", "#947EB0","#083D77","#356E46", "#69597D", "#052344")) +
      theme_bw() #+
```

SUMMARY HALF STARBURST PLOTS

Combined all measures at ALL site location
```{r fig.height=4, fig.width=8}
ihc_1_ratio %>% dfilter(Subject_ID %in% tissue_matched) %>% dselect(CD8_FOXP3_percent_log2_norm,CD8_FOXP3_density_log2_norm,CD8_FOXP3_log2_norm,Visit_type,Subject_ID,Combo,Anatomical_Location,Total_CD8_cells) %>% dfilter(Anatomical_Location=="ALL") %>% gather(measure_type,log2_norm,-Visit_type,-Subject_ID,-Combo,-Anatomical_Location,-Total_CD8_cells) %>% left_join(best_response, by="Subject_ID") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR"))) %>%
      ggplot(aes(x=Visit_type,y=log2_norm)) + 
      geom_point(aes(color=Best_Response)) +
      #geom_label(aes(label=Subject_ID, alpha=0.2)) + 
      geom_line(aes(x=Visit_type,y=log2_norm, group=Subject_ID,color=Best_Response),size=1) +
      #geom_text(aes(label=Subject_ID)) +
      facet_wrap(~measure_type) + 
      scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      #scale_size_continuous(breaks = c(1,10,100, 1000)) + #this is the size of the points
      scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33", "#4DA167", "#947EB0","#083D77","#356E46", "#69597D", "#052344")) +
      theme_bw() #+
      #expand_limits(y=c(-40,30))

ihc_1_ratio %>% dfilter(Subject_ID %in% tissue_matched) %>% dselect(CD8_FOXP3_percent_log2_norm,CD8_FOXP3_density_log2_norm,CD8_FOXP3_log2_norm,Visit_type,Subject_ID,Combo,Anatomical_Location,Total_CD8_cells) %>% dfilter(Anatomical_Location=="ALL") %>% gather(measure_type,log2_norm,-Visit_type,-Subject_ID,-Combo,-Anatomical_Location,-Total_CD8_cells) %>% left_join(best_response, by="Subject_ID") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR"))) %>%
      ggplot(aes(x=Visit_type,y=log2_norm)) + 
      geom_point(aes(color=Combo)) +
      #geom_label(aes(label=Subject_ID, alpha=0.2)) + 
      geom_line(aes(x=Visit_type,y=log2_norm, group=Subject_ID,color=Combo),size=1) +
      #geom_text(aes(label=Subject_ID)) +
      facet_wrap(~measure_type) + 
      scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      #scale_size_continuous(breaks = c(1,10,100, 1000)) + #this is the size of the points
      scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33", "#4DA167", "#947EB0","#083D77","#356E46", "#69597D", "#052344")) +
      theme_bw()




```


Waterfall plots
```{r fig.height=3, fig.width=4.5}
ihc_1_ratio %>% dfilter(Subject_ID %in% tissue_matched, Visit_type=="Follow_up") %>% dfilter(Anatomical_Location =="ALL") %>% left_join(best_response, by="Subject_ID") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR")))  %>% mutate(Subject_ID= fct_reorder(Subject_ID,-CD8_FOXP3_density_log2_norm)) %>%
      ggplot(aes(x=Subject_ID,y=CD8_FOXP3_density_log2_norm, ymin=0, ymax=CD8_FOXP3_density_log2_norm, color=Best_Response)) + 
      geom_point() + #aes(size=Total_CD8_cells, colour=Subject_ID)
      geom_linerange() +
      #facet_wrap(~Anatomical_Location) + #, scale="free_y"
      #scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33","#4DA167", "#947EB0","#083D77","#356E46", "#69597D", "#052344")) +
      theme_bw() +
      geom_hline(yintercept=0) +
      theme(axis.text.x = element_text(angle=45,size=12, vjust=1, hjust=1),strip.text=element_text(size=12)) 
  
ihc_1_ratio %>% dfilter(Subject_ID %in% tissue_matched, Visit_type=="Follow_up") %>% dfilter(Anatomical_Location =="ALL") %>% left_join(best_response, by="Subject_ID") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR")))  %>% mutate(Subject_ID= fct_reorder(Subject_ID,-CD8_FOXP3_percent_log2_norm)) %>%
      ggplot(aes(x=Subject_ID,y=CD8_FOXP3_percent_log2_norm, ymin=0, ymax=CD8_FOXP3_percent_log2_norm, color=Best_Response)) + 
      geom_point() + #aes(size=Total_CD8_cells, colour=Subject_ID)
      geom_linerange() +
      #facet_wrap(~Anatomical_Location) + #, scale="free_y"
      #scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33","#4DA167", "#947EB0","#083D77","#356E46", "#69597D", "#052344")) +
      theme_bw() +
      geom_hline(yintercept=0) +
      theme(axis.text.x = element_text(angle=45,size=12, vjust=1, hjust=1),strip.text=element_text(size=12)) 


ihc_1_ratio %>% dfilter(Subject_ID %in% tissue_matched, Visit_type=="Follow_up") %>% dfilter(Anatomical_Location =="ALL") %>% left_join(best_response, by="Subject_ID") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR")))  %>% mutate(Subject_ID= fct_reorder(Subject_ID,-CD8_FOXP3_log2_norm)) %>%
      ggplot(aes(x=Subject_ID,y=CD8_FOXP3_log2_norm, ymin=0, ymax=CD8_FOXP3_log2_norm, color=Best_Response)) + 
      geom_point() + #aes(size=Total_CD8_cells, colour=Subject_ID)
      geom_linerange() +
      #facet_wrap(~Anatomical_Location) + #, scale="free_y"
      #scale_fill_manual(values=c("#0A95A7","#E5541B")) +
      scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33","#4DA167", "#947EB0","#083D77","#356E46", "#69597D", "#052344")) +
      theme_bw() +
      geom_hline(yintercept=0) +
      theme(axis.text.x = element_text(angle=45,size=12, vjust=1, hjust=1),strip.text=element_text(size=12)) 
```



Correlation between measurement types
```{r fig.height=4, fig.width=4.2}
library(plotly)
plot_ly(x=ihc_1$CD8_FOXP3_log2_norm, y=ihc_1_test$CD8_FOXP3_density_log2_norm, z=ihc_1_test$CD8_FOXP3_percent_log2_norm, type="scatter3d", mode="markers",height=300)

ggplot(ihc_1_ratio %>% dselect(ends_with("_norm"),Subject_ID,Combo,Total_CD8_cells) %>% unique,aes(x=CD8_FOXP3_log2_norm, y=CD8_FOXP3_density_log2_norm, color=Combo)) + geom_point(aes(size=Total_CD8_cells)) + theme_bw()  + nolegend() # geom_label(aes(label=Subject_ID,alpha=0.4),size=3) 

ggplot(ihc_1_ratio %>% dselect(ends_with("_norm"),Subject_ID,Combo,Total_CD8_cells) %>% unique,aes(x=CD8_FOXP3_log2_norm, y=CD8_FOXP3_percent_log2_norm, color=Combo)) + geom_point(aes(size=Total_CD8_cells)) + theme_bw() + theme_bw() + nolegend()# geom_label(aes(label=Subject_ID,alpha=0.4),size=3) + theme_bw()

ggplot(ihc_1_ratio %>% dselect(ends_with("_norm"),Subject_ID,Combo,Total_CD8_cells) %>% unique,aes(x=CD8_FOXP3_density_log2_norm, y=CD8_FOXP3_percent_log2_norm, color=Combo)) + geom_point(aes(size=Total_CD8_cells)) + theme_bw() + nolegend()# geom_label(aes(label=Subject_ID,alpha=0.4),size=3) + theme_bw()

```
Calculate R squared

```{r}
model <- lm(CD8_FOXP3_density_log2_norm ~ CD8_FOXP3_percent_log2_norm, ihc_1_ratio)
model <- lm(CD8_FOXP3_log2_norm ~ CD8_FOXP3_percent_log2_norm, ihc_1_ratio)
model <- lm(CD8_FOXP3_log2_norm ~ CD8_FOXP3_density_log2_norm, ihc_1_ratio)
summary(model)
```



Correlate to Best_Response (CONFIDENTIAL)


Merge and plot for samples with tissue matched data: looking at CD8/FOXP3 ratio normalized to Screening ratio

```{r fig.height=3.5, fig.width=11}
ihc_1_ratio %>% left_join(best_response_precross, by="Subject_ID") %>% dfilter(Visit_type== "Follow_up") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR"))) %>%
  ggplot(aes(y=CD8_FOXP3_log2_norm, x=Best_Response_score, colour=Best_Response)) + 
  geom_point(size=2.5) + #aes(size=Total_CD8_cells)
  #geom_label(aes(label=Subject_ID),alpha=0.2) +
  geom_hline(yintercept=0) +
  facet_wrap(~Anatomical_Location) +
  scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33", "#4DA167", "#947EB0","#083D77","#356E46", "#69597D", "#052344")) +
  scale_x_reverse() +
  theme_bw()
```
Categorize

SKIP THESE, since the ALL doesn't show any correlation
```{r fig.height=3.5, fig.width=11}
library(ggbeeswarm)
ihc_1_ratio %>% left_join(best_response_precross, by="Subject_ID") %>% dfilter(Visit_type== "Follow_up") %>% dfilter(!is.na(Best_Response)) %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR")))  %>%
  ggplot(aes(y=CD8_FOXP3_log2_norm, x=Best_Response)) + 
  geom_boxplot(alpha=0.3, width=0.5) +
  geom_beeswarm(aes(color=Best_Response),size=2.5) +
  facet_wrap(~Anatomical_Location) +
  scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33", "#4DA167", "#947EB0","#083D77","#356E46", "#69597D", "#052344")) +
  theme_bw()
```

SCREENING DATA and BOR

Merge and plot for all samples with screening data: looking at Screening CD8/FOXP3 ratio 


```{r fig.height=3.5, fig.width=11}
ihc_1 %>% left_join(best_response_precross, by="Subject_ID") %>% dfilter(Visit_type== "Pre_treat", !is.na(Best_Response)) %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR")))   %>% dselect(Subject_ID,Anatomical_Location,Best_Response_score,CD8_FOXP3_log2_mean,Best_Response) %>% unique %>%
  ggplot(aes(y=CD8_FOXP3_log2_mean, x=Best_Response_score, colour=Best_Response)) + 
  geom_point(size=2.5) + #aes(size=Total_CD8_cells)
  #geom_label(aes(label=Subject_ID),alpha=0.2) +
  facet_wrap(~Anatomical_Location) +
  scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33")) +
  scale_x_reverse() +
  theme_bw()

ihc_1 %>% left_join(best_response_precross, by="Subject_ID") %>% dfilter(Visit_type== "Pre_treat", !is.na(Best_Response)) %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR")))   %>% dselect(Subject_ID,Anatomical_Location,Best_Response_score,CD8_FoxP3_percent_log2_mean,Best_Response) %>% unique %>%
  ggplot(aes(y=CD8_FoxP3_percent_log2_mean, x=Best_Response_score, colour=Best_Response)) + 
  geom_point(size=2.5) + #aes(size=Total_CD8_cells)
  #geom_label(aes(label=Subject_ID),alpha=0.2) +
  facet_wrap(~Anatomical_Location) +
  scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33")) +
  scale_x_reverse() +
  theme_bw()

ihc_1 %>% left_join(best_response_precross, by="Subject_ID") %>% dfilter(Visit_type== "Pre_treat", !is.na(Best_Response)) %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR")))   %>% dselect(Subject_ID,Anatomical_Location,Best_Response_score,CD8_FoxP3_density_log2_mean,Best_Response) %>% unique %>%
  ggplot(aes(y=CD8_FoxP3_density_log2_mean, x=Best_Response_score, colour=Best_Response)) + 
  geom_point(size=2.5) + #aes(size=Total_CD8_cells)
  #geom_label(aes(label=Subject_ID),alpha=0.2) +
  facet_wrap(~Anatomical_Location) +
  scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33")) +
  scale_x_reverse() +
  theme_bw()


```


Export to Table for Paul
```{r}
ihc_1 %>% left_join(best_response, by="Subject_ID") %>% dfilter(Visit_type== "Pre_treat", !is.na(Best_Response)) %>% mutate(Resp_type= case_when(
  Best_Response > 0.2 ~ "PD",
  Best_Response < -0.3 ~ "R",
  TRUE ~ "SD"
)) %>% mutate(Resp_type= factor(Resp_type, levels=c("PD","SD","R"))) %>% dselect(Subject_ID,Anatomical_Location,Tissue_Type,Visit,Best_Response,Resp_type,CD8_FOXP3_log2_mean,CD8_FoxP3_percent_log2_mean,CD8_FoxP3_density_log2_mean) %>% unique %>% write.table(file="Mosaic_102820_Response_Prism_all_TME.txt",sep="\t", quote=F, row.names = F)
```


SCREENING 
Categorize
```{r fig.height=3.5, fig.width=11}
library(ggbeeswarm)

ihc_1 %>% left_join(best_response, by="Subject_ID") %>% dfilter(Visit_type== "Pre_treat") %>% dfilter(!is.na(Best_Response)) %>% dfilter(Best_Response !="NE") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR","NE"))) %>% dselect(Subject_ID,Anatomical_Location,Best_Response,CD8_FOXP3_log2_mean) %>% unique %>%
  ggplot(aes(y=CD8_FOXP3_log2_mean, x=Best_Response)) + 
  geom_boxplot(alpha=0.3, width=0.5) +
  geom_beeswarm(aes(color=Best_Response),size=2,priority='density',cex=2.5) +
  facet_wrap(~Anatomical_Location) +
  scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33")) +
  theme_bw()
```

Categorize ALL

```{r}
colnames(ihc_1)
```

Plot ALL SCREENING measures together
```{r fig.height=3, fig.width=10}
ihc_1 %>% left_join(best_response, by="Subject_ID") %>% dfilter(Visit_type== "Pre_treat") %>% dfilter(!is.na(Best_Response)) %>% dfilter(Best_Response !="NE") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR","NE"))) %>% dselect(Subject_ID,Anatomical_Location,CD8_FOXP3_log2_mean,CD8_FoxP3_percent_log2_mean,CD8_FoxP3_density_log2_mean,Best_Response) %>% unique %>% dfilter(Anatomical_Location=="ALL") %>% gather(measure_type,log2_norm,-Subject_ID,-Anatomical_Location,-Best_Response) %>%
  ggplot(aes(y=log2_norm, x=Best_Response)) + 
  geom_boxplot(alpha=0.3, width=0.5) +
  geom_beeswarm(aes(color=Best_Response),size=2,priority='density',cex=2.5) +
  facet_wrap(~measure_type) +
  scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7","#FE9F33")) +
  theme_bw()

ihc_1 %>% left_join(best_response, by="Subject_ID") %>% dfilter(Visit_type== "Pre_treat") %>% dfilter(!is.na(Best_Response)) %>% dfilter(Best_Response !="NE") %>% mutate(Best_Response= factor(Best_Response, levels=c("PD","SD","PR"))) %>% 
dselect(CD8_FoxP3_percent_log2_mean,CD8_FoxP3_density_log2_mean,CD8_FOXP3_log2_mean,Visit_type,Subject_ID,Combo,Anatomical_Location,Total_CD8_cells,Best_Response,Best_Response_score) %>% dfilter(Anatomical_Location=="ALL") %>% gather(measure_type,log2_norm,-Visit_type,-Subject_ID,-Combo,-Anatomical_Location,-Total_CD8_cells,-Best_Response,-Best_Response_score) %>% unique %>%
  ggplot(aes(y=log2_norm, x=Best_Response_score)) + 
  geom_point(aes(color=Best_Response) ) +
  facet_wrap(~measure_type,scale="free_y") +
  scale_size_continuous(breaks=c(1,10,100,500,1000)) +
  scale_x_reverse() +
  scale_color_manual(values=c("#E5541B","#5B666F","#0A95A7")) +
  theme_bw()
```


Correlate Screening values to ratio of ratios:
Cell #, sensity, %, CD8/FoxP3 ratio by #, %, density
X values:
CD8_FOXP3_log2_screen
CD8_FoxP3_percent_screen
CD8_FoxP3_density_screen

Y values:
CD8_FOXP3_log2_norm
CD8_FOXP3_percent_log2_norm
CD8_FOXP3_density_log2_norm
```{r fig.height=4, fig.width=4}
ihc_1_ratio %>% dfilter(Visit_type=="Follow_up", Anatomical_Location=="ALL") %>% 
  ggplot(aes(x=CD8_FOXP3_log2_screen,y=CD8_FOXP3_log2_norm)) +
  geom_point(size=2, colour="#0A95A7") +
  theme_bw()

ihc_1_ratio %>% dfilter(Visit_type=="Follow_up", Anatomical_Location=="ALL") %>% 
  ggplot(aes(x=CD8_FoxP3_percent_screen,y=CD8_FOXP3_percent_log2_norm)) +
  geom_point(size=2, colour="#0A95A7") +
  theme_bw()

ihc_1_ratio %>% dfilter(Visit_type=="Follow_up", Anatomical_Location=="ALL") %>% 
  ggplot(aes(x=CD8_FoxP3_density_screen,y=CD8_FOXP3_density_log2_norm)) +
  geom_point(size=2, colour="#0A95A7") +
  theme_bw()
  #geom_smooth(method="glm", se=FALSE, aes(color="grey50"))
  #scale_size_continuous(breaks=c(1,10,100,500,1000,2000))
```


Matrix of corr for paired samples
```{r fig.height=8, fig.width=8}
install.packages("GGally")
library(GGally)
install.packages("ggcorrplot")
library(ggcorrplot)
install.packages("rstatix")
library(rstatix)

#ihc_1_ratio %>% dfilter(Anatomical_Location=="ALL", Visit_type=="Follow_up") %>% dselect(ends_with("_screen"),ends_with("_norm")) %>% unique %>% ggpairs(progress=NULL, switch="both")

#ihc_1_ratio %>% dfilter(Anatomical_Location=="ALL", Visit_type=="Follow_up") %>% dselect(ends_with("_screen"),ends_with("_norm")) %>% unique %>% ggcorr(hjust = 0.9, size = 4, color = "grey50",progress=NULL)

#with Response to correlate Screening values with Response
for_corr <- ihc_1_ratio %>% dfilter(Anatomical_Location=="ALL", Visit_type=="Follow_up") %>% dselect(Subject_ID, ends_with("_screen"),ends_with("_norm")) %>% unique %>% left_join(best_response %>% dselect(Subject_ID,Best_Response_score), by="Subject_ID") %>% dfilter(!is.na(Best_Response_score)) %>% mutate(Best_Response_score = -1*Best_Response_score) %>% dfilter(!is.na(Best_Response_score)) %>% dfilter(!is.na(CD8_FOXP3_log2_screen)) %>% dselect(-Subject_ID)

#withour Response

for_corr <- ihc_1_ratio %>% dfilter(Anatomical_Location=="ALL", Visit_type=="Follow_up") %>% dselect(Subject_ID, ends_with("_screen"),ends_with("_norm")) %>% dfilter(!is.na(CD8_FOXP3_log2_screen)) %>% unique %>% tibble::column_to_rownames("Subject_ID")
#ihc_1_ratio %>% dfilter(Anatomical_Location=="ALL", Visit_type=="Follow_up") %>% dselect(Subject_ID, ends_with("_screen"),ends_with("_norm")) %>% unique %>% dselect(-Subject_ID)

#export for Paul
ihc_1_ratio %>% dfilter(Anatomical_Location=="ALL", Visit_type=="Follow_up") %>% dselect(Subject_ID, ends_with("_screen"),ends_with("_norm")) %>% unique %>% write.table(file="CD8_FoxP3_Ratio_vs_Screening_values.txt", sep="\t",row.names = FALSE, quote=FALSE)

corr <- round(cor(for_corr),1)
p.mat <- cor_pmat(for_corr) %>% as.matrix
dim(corr)
dim(p.mat)

ggcorrplot(corr, type="upper",show.diag=TRUE,p.mat = p.mat)
ggcorrplot(corr)
```

Correlation metric for all samples and Response

```{r}
screen_corr <- ihc_1 %>% dfilter(Anatomical_Location=="ALL", Visit_type=="Pre_treat") %>% 
   left_join(count_screen_all, by=c("Subject_ID","Anatomical_Location")) %>% unique %>% dselect(Subject_ID,ends_with("_screen")) %>%
   left_join(best_response, by="Subject_ID") %>% unique %>% mutate(Best_Response = -1*Best_Response) %>% dfilter(!is.na(Best_Response)) %>% dselect(-Subject_ID)

corr <- round(cor(screen_corr),1)
p.mat <- cor_pmat(screen_corr) %>% tibble::column_to_rownames("rowname") %>% as.matrix
dim(corr)
dim(p.mat)

ggcorrplot(corr, type="upper",show.diag=TRUE,p.mat = p.mat)
ggcorrplot(corr)
```

###########################
###########################

TILE PLOTS

###########################
###########################

Examine patients for those who have and don't have followup

Set the patient order as the current BOR arranged by scores
```{r}
Subject_order <- ihc_1 %>% dselect(Subject_ID,Phase) %>% unique %>% left_join(best_response %>% dselect(Subject_ID,Best_Response,Best_Response_score), by="Subject_ID") %>% mutate(Best_Response=factor(Best_Response, levels=c("PD","SD","PR","CR","NE"))) %>% arrange(Best_Response_score) %>% dselect(Subject_ID) %>% unique %>% pull(Subject_ID)

Subject_order <- c("1602-002","1804-002","1804-006","1601-006","2102-001","0104-010","0102-006","0109-002","0113-002","0111-005","0111-002","1801-005","0102-003","1802-006","0104-008","0102-007","0106-005","0101-004","1806-007","0110-002","0101-001","0106-001","0103-002",
"1806-006","0111-006","1804-010","1806-013","1601-003","0104-002","0104-003","0113-003","0101-005","0101-006","0106-002","0107-002",
"1601-004")

# Cancer type text
ihc_1 %>% dselect(Subject_ID,Tissue_Type) %>% mutate(Subject_ID = factor(Subject_ID,levels=(Subject_order))) %>% unique %>% arrange(Subject_ID) %>% write.table(file="patient_cancer.txt",sep="\t", quote=F, row.names = F)

```


```{r fig.height=8, fig.width=2.5}
ihc_1 %>% dselect(Subject_ID,Visit_type)  %>% unique %>% table 
ihc_1 %>% dselect(Subject_ID,Visit_type)  %>% unique %>% table %>% as.data.frame %>% group_by(Subject_ID) %>% summarise(paired=sum(Freq)) %>% dfilter(paired == 2)
# Pre post
is_pre_post <- ihc_1 %>% dselect(Subject_ID,Visit_type)  %>% unique %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(y=Subject_ID,x=Visit_type, fill=Visit_type)) + geom_tile(colour="white") + scale_x_discrete(position = "top") + scale_fill_manual(values=c("#0A95A7","#0A95A7")) + scale_x_discrete(position = "top") + theme(axis.text.x=element_text(angle=45, size=12,hjust=0)) # + nolegend()
# Phase
ihc_1 %>% dselect(Subject_ID,Phase) %>% unique %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(y=Subject_ID,x=Phase, fill=Phase)) + geom_tile(colour="white") + nolegend() + scale_x_discrete(position = "top") 
  write.table(file="patient_phase.txt",sep="\t", quote=F, row.names = F)
#plot those that are Paired  
is_pair <- ihc_1 %>% dselect(Subject_ID,Visit_type,Phase) %>% unique %>% group_by(Subject_ID,Phase) %>% summarise(Num_sample=n()) %>% ungroup %>% mutate(Num_sample=factor(Num_sample)) %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(y=Subject_ID,x=Num_sample)) + geom_tile(colour="white", fill="#0A95A7") + nolegend() 
# tissue matched
is_match <- ihc_1 %>% dselect(Subject_ID, Phase) %>% unique %>% mutate(Tissue_match=if_else(Subject_ID %in% tissue_matched,1,0)) %>% mutate(Tissue_match=factor(Tissue_match)) %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(y=Subject_ID,x=Tissue_match)) + geom_tile(colour="white", fill="#0A95A7") + nolegend() #+ scale_fill_manual(values=c("grey90","#0A95A7"))
#Best Response outcome plot
is_pre_post %>% left_join(is_pair, by="Subject_ID") %>% left_join(is_match,by="Subject_ID")

best_response
ihc_1 %>% dselect(Subject_ID,Phase)  %>% unique %>% left_join(best_response %>% dselect(Subject_ID,Best_Response), by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% mutate(Best_Response=factor(Best_Response, levels=c("PD","SD","PR","CR","NE"))) %>% ggplot(aes(y=Subject_ID,x=1, fill=Best_Response)) + geom_tile() + scale_x_discrete(position = "top") + scale_fill_manual(values=c("#E5541B","#0A95A7","#052344","grey80")) 
#Best Response score plot
best_response
ihc_1 %>% dselect(Subject_ID,Phase)  %>% unique %>% left_join(best_response %>% dselect(Subject_ID,Best_Response_score), by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(y=Subject_ID,x=1, fill=Best_Response_score)) + geom_tile() + scale_x_discrete(position = "top",labels=c("CD8+ % Screening","FOXP3+ % Screening")) + theme(axis.text.x=element_text(angle=45, size=12,hjust=0)) + scale_fill_distiller(palette = "RdBu")  #scale_fill_brewer(name="RdBu",na.value="grey80",limits=c(-2,2),oob=squish)

#Molecular marker plots
library(viridis)
library(scales)
library(gridExtra)

#Screening PERCENT
# There is a lot of missing data here. I need to take all patients, and left join the available data onto the full list to make a full plot with missing data as placeholders
Subjects <- ihc_1 %>% dselect(Subject_ID) %>% unique 

#PERCENT
percent_cd8 <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>%  dselect(Subject_ID,CD8_Percent_Positive)  %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(x=1,y=Subject_ID,fill=CD8_Percent_Positive)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#0A95A7", limits=c(0,40),na.value="grey80") + theme(axis.ticks =element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend()  #,CD8_Percent_Positive, FoxP3_Percent_Positive

#Density
density_foxp3 <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>%  dselect(Subject_ID,FoxP3_Cells_per_mm2) %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(x=1,y=Subject_ID,fill=FoxP3_Cells_per_mm2)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#0A95A7",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend()

density_cd8 <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>%  dselect(Subject_ID,CD8_Cells_per_mm2) %>% ggplot(aes(x=1,y=Subject_ID,fill=CD8_Cells_per_mm2)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#0A95A7",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend()

#Screening RATIO
screen_num_ratio <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,CD8_FOXP3_log2) %>% ggplot(aes(x=1,y=Subject_ID,fill=CD8_FOXP3_log2)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#E5541B",na.value="grey80",limits=c(0,3),oob=squish) + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend()

# Change in RATIO
delta_ratio <- Subjects %>% left_join(ihc_1_ratio %>% dfilter(Visit_type=="Follow_up",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,CD8_FOXP3_log2_norm) %>% ggplot(aes(x=1,y=Subject_ID,fill=CD8_FOXP3_log2_norm)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient2(low="#0A95A7", mid="white",high="#E5541B",na.value="grey80",limits=c(-1.6,2),oob=squish) + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend() # ,plot.margin=unit(c(0,-1,0,-1), unit="cm")

# PERCENT FOXP3 CD8 double positive cells @ SCREEN
#Percent_CD8_FoxP3_Cells

foxp2_cd8_screen <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,Percent_CD8_FoxP3_Cells) %>% ggplot(aes(x=1,y=Subject_ID,fill=Percent_CD8_FoxP3_Cells)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#E5541B",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend() # ,limits=c(0,2),oob=squish

# PERCENT CD8+ that are also FOXP3+ @ SCREEN
cd8_also_foxp3_screen <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,Percent_of_CD8_Cells_that_are_FoxP3) %>% ggplot(aes(x=1,y=Subject_ID,fill=Percent_of_CD8_Cells_that_are_FoxP3)) + geom_tile(aes(width=1,height=1))  + scale_fill_gradient(low="white",high="#E5541B",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend() #

# PERCENT CD8+ that are also FOXP3+ @ FOLLOW UP
cd8_also_foxp3_followup <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Follow_up",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,Percent_of_CD8_Cells_that_are_FoxP3) %>% ggplot(aes(x=1,y=Subject_ID,fill=Percent_of_CD8_Cells_that_are_FoxP3)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#E5541B",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend() #

grid.arrange(percent_cd8,density_foxp3,density_cd8,screen_num_ratio,delta_ratio,foxp2_cd8_screen,cd8_also_foxp3_screen,cd8_also_foxp3_followup, ncol=8,padding=0)
```

######
LEGENDs

#######
```{r}
library(cowplot) #grab legend
#PERCENT
cd8_l <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>%  dselect(Subject_ID,CD8_Percent_Positive)  %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(x=1,y=Subject_ID,fill=CD8_Percent_Positive)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#0A95A7", limits=c(0,40),na.value="grey80")   #,CD8_Percent_Positive, FoxP3_Percent_Positive
as_ggplot(cowplot::get_legend(cd8_l))
#Density
density_foxp3 <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>%  dselect(Subject_ID,FoxP3_Cells_per_mm2) %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(x=1,y=Subject_ID,fill=FoxP3_Cells_per_mm2)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#0A95A7",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend()

density_cd8 <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>%  dselect(Subject_ID,CD8_Cells_per_mm2) %>% ggplot(aes(x=1,y=Subject_ID,fill=CD8_Cells_per_mm2)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#0A95A7",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend()

#Screening RATIO
screen_num_ratio <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,CD8_FOXP3_log2) %>% ggplot(aes(x=1,y=Subject_ID,fill=CD8_FOXP3_log2)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#E5541B",na.value="grey80",limits=c(0,3),oob=squish) + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend()

# Change in RATIO
delta_ratio <- Subjects %>% left_join(ihc_1_ratio %>% dfilter(Visit_type=="Follow_up",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,CD8_FOXP3_log2_norm) %>% ggplot(aes(x=1,y=Subject_ID,fill=CD8_FOXP3_log2_norm)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient2(low="#0A95A7", mid="white",high="#E5541B",na.value="grey80",limits=c(-1.6,2),oob=squish) + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend() # ,plot.margin=unit(c(0,-1,0,-1), unit="cm")

# PERCENT FOXP3 CD8 double positive cells @ SCREEN
#Percent_CD8_FoxP3_Cells

foxp2_cd8_screen <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,Percent_CD8_FoxP3_Cells) %>% ggplot(aes(x=1,y=Subject_ID,fill=Percent_CD8_FoxP3_Cells)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#E5541B",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend() # ,limits=c(0,2),oob=squish

# PERCENT CD8+ that are also FOXP3+ @ SCREEN
cd8_also_foxp3_screen <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Pre_treat",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,Percent_of_CD8_Cells_that_are_FoxP3) %>% ggplot(aes(x=1,y=Subject_ID,fill=Percent_of_CD8_Cells_that_are_FoxP3)) + geom_tile(aes(width=1,height=1))  + scale_fill_gradient(low="white",high="#E5541B",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend() #

# PERCENT CD8+ that are also FOXP3+ @ FOLLOW UP
cd8_also_foxp3_followup <- Subjects %>% left_join(ihc_1 %>% dfilter(Visit_type=="Follow_up",Anatomical_Location=="ALL"),by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% dselect(Subject_ID,Percent_of_CD8_Cells_that_are_FoxP3) %>% ggplot(aes(x=1,y=Subject_ID,fill=Percent_of_CD8_Cells_that_are_FoxP3)) + geom_tile(aes(width=1,height=1)) + scale_fill_gradient(low="white",high="#E5541B",na.value="grey80") + theme(axis.ticks = element_blank(),axis.text=element_blank(),axis.title=element_blank(),panel.background = element_rect(fill="white"),plot.margin=unit(c(0,0,0,0), unit="lines")) + nolegend() #

grid.arrange(percent_cd8,density_foxp3,density_cd8,screen_num_ratio,delta_ratio,foxp2_cd8_screen,cd8_also_foxp3_screen,cd8_also_foxp3_followup, ncol=8,padding=0)
```

##################
##############
Genomic data from Gene, Personalis

#############
##################
```{r}
load("/Volumes/Picard/FLX/Mosaic/qc_stats.20210121.rdata") # nested
```

```{r}
Ploidy <- nested %>% filter(group=="dna_statistics") %>% pull(data) %>% .[[1]] %>% dfilter(feature=="Tumor ploidy") %>% dselect(Subject.No,value) %>% mutate(value=as.numeric(value))

Total_neoantigen <- nested %>% filter(group=="neoantigen_statistics") %>% pull(data) %>% .[[1]] %>% dfilter(feature =="Total Neoantigens: Count") %>% dselect(Subject.No,value) %>% mutate(value=as.numeric(value))

Viral_DNA_count <- nested %>% filter(group=="oncovirus_statistics") %>% pull(data) %>% .[[1]] %>% dfilter(grepl("Screen",`Visit Name`),grepl("DNA",feature)) %>% group_by(Subject.No) %>% dselect(Subject.No,value) %>% mutate(value=as.numeric(value)) %>% summarise_if(is.numeric,sum) 
#grepl("Screen",`Visit Name`),  #%>% dfilter(grepl("DNA",feature) )  
# Here Viral counts = 0 is just not included, instead of being reported as 0, so I have to add the 0 back in with an if_else statement
# except can't add 0 because the viral counts has to be on a log10 scale, add 0.001 
all_viral <-  samples_with_data %>% left_join (Viral_DNA_count, by="Subject.No" ) %>% mutate(value= if_else(is.na(value),0.001,value) )


nested %>% filter(group=="dna_statistics") %>% pull(data) %>% .[[1]] %>% dselect(Subject.No) %>% unique #22
samples_with_data <- nested %>% filter(group=="rna_statistics") %>% pull(data) %>% .[[1]] %>% dselect(Subject.No) %>% unique #%>% pull(Subject.No)

```

```{r fig.height=9, fig.width=4.5}
# Neoantigen count
Subjects %>% left_join(Total_neoantigen ,by=c("Subject_ID"="Subject.No")) %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(x=1,y=Subject_ID,fill=value)) + geom_tile() + scale_fill_gradient(low="white",high="#4DA167",na.value="grey80", trans='log10') + scale_x_discrete(position = "top",labels=c("Neoantigen Count")) + theme(axis.text.x=element_text(angle=45, size=12,hjust=0)) 

#Tumor Plody
Subjects %>% left_join(Ploidy ,by=c("Subject_ID"="Subject.No")) %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% unique %>% ggplot(aes(x=1,y=Subject_ID,fill=value)) + geom_tile() + scale_fill_gradient(low="white",high="#4DA167",na.value="grey80",limits=c(2,5)) + scale_x_discrete(position = "top",labels=c("Neoantigen Count")) + theme(axis.text.x=element_text(angle=45, size=12,hjust=0)) 

#Viral count
Subjects %>% left_join(all_viral ,by=c("Subject_ID"="Subject.No")) %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(x=1,y=Subject_ID,fill=value)) + geom_tile() + scale_fill_gradient(low="white",high="#4DA167",na.value="grey80", trans='log10') + scale_x_discrete(position = "top",labels=c("Neoantigen Count")) + theme(axis.text.x=element_text(angle=45, size=12,hjust=0)) 

Subjects %>% left_join(with_data,by="Subject_ID") %>% mutate(Subject_ID = factor(Subject_ID,levels=rev(Subject_order))) %>% ggplot(aes(x=1,y=Subject_ID,fill=Combo)) + geom_tile() + scale_x_discrete(position = "top",labels=c("Neoantigen Count")) + theme(axis.text.x=element_text(angle=45, size=12,hjust=0)) 

with_data <- ihc_1_ratio %>% dfilter(Subject_ID %in% tissue_matched) %>% dselect(CD8_FOXP3_percent_log2_norm,CD8_FOXP3_density_log2_norm,CD8_FOXP3_log2_norm,Visit_type,Subject_ID,Combo,Anatomical_Location,Total_CD8_cells) %>% dfilter(Anatomical_Location=="ALL") %>% dselect(Subject_ID,Combo) %>% unique
```


```{r}
read.csv2(file="/Volumes/Picard/FLX/Clinical_data_internship/AD/GSE160501_Processed_gene_counts_matrix.txt" ,sep="\t")
```

