---
title: "Mosely lookup"
author: "Mengshu"
date: "7/16/2019"
output: html_document
---

In-house and Mosely microarray expression data for Mouse syngeneic models

mouse_dds is the in-house data of CT26, MC38 and Pan02 in triplicate
mouseData_tum is the Mosely data
samples.info_tum is the metadata for Mosely

```{r}
library(tidyverse)
library(dplyr)
library(magrittr)
library(tibble)
library(ggplot2)
library(data.table)
library(reshape2) #needed for the heatmap
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
source("D:/FLX/Files_from_Gene/R_functions.r")
```


Load data
```{r}
load("/Volumes/Enterprise/FLX/Silpa_data/mouseAnnot.rdata") # this is the mouse=human gene lookup table
load("/Volumes/Enterprise/FLX/Silpa_data/mouse_dds.rdata") # This is an in-house dataset, not the Mosely data
#This is the Mosely data
load("/Volumes/Enterprise/FLX/Silpa_data/mouseData_tum.rdata")
load("/Volumes/Enterprise/FLX/Silpa_data/samples.info_tum.rdata")

load("D:/FLX/Silpa_data/mouseData_tum.rdata")
load("D:/FLX/Silpa_data/samples.info_tum.rdata")

# Z_score
load("/Volumes/Enterprise/FLX/GCN2/ISR_data/RNA_seq/Mosely_data_zscore.rdata") # mouse_z
# ZZ_score
load("/Volumes/Enterprise/FLX/GCN2/ISR_data/RNA_seq/Mosely_data_zzscore.rdata") # mouse_zz
```
Examine data
Plot histogram of signal distribution
```{r}
samples.info_tum
rownames(mouseData_tum) # this is the gene expression data of 24818 genes in columns by 93 rows that are unnamed
mouseData_tum[1:10,1:10]
mouseData_tum %>% gather(Gene,Signal) %>% hist()
mouseData_tum %>% gather(Gene,Signal) %>% ggplot(aes(Signal, colour=Signal)) + geom_histogram(binwidth=0.2) + scale_x_continuous(name="Log2 Normalized Signal (Microarray)", breaks=seq(0,10,by=1)) +
ggsave("Mosely_Signal_Hist.jpg", width=5, height=5, dpi=600, plot= last_plot(), units = "in")

str(mouseAnnot) #this is the human mouse gene lookup table
mouseAnnot %>% dfilter(GENE.SYMBOL=="Asns")
```
Merge the Mosely data and metadata

```{r}
mouse <- data.frame(bind_cols(samples.info_tum,mouseData_tum))

mouse %>% dselect(cell.line, genes) %>% gather(gene,RNA_expression, -cell_line) %>% ggplot(aes(x=cell_line, y=RNA_expression, group_by(cell_line), colour= cell_line)) + geom_point(show.legend=FALSE) + facet_wrap(~gene, ncol=1, scales="free_y") + theme(axis.title = element_text(size=14), axis.text.y = element_text(size=14), axis.text.x = element_text(angle=45,size=14,hjust=1),strip.text.x=element_text(size=16)) +
ggsave("GCN2_off_target_Mouse_Vert.jpg", width=6, height=10, dpi=600, plot= last_plot(), units = "in")
names(mouse)[7] <- "cell_line"
mouse %>% dselect(cell_line, genes)  %>% gather(gene,RNA_expression, -cell_line) %>% ggplot(aes(x=cell_line, y=RNA_expression, group_by(cell_line), colour= cell_line)) + geom_point(show.legend=FALSE) + theme(axis.title = element_text(size=14), axis.text.y = element_text(size=14), axis.text.x = element_text(angle=45,size=14,hjust=1)) + stat_summary(inherit.aes = TRUE,quantiles = 0.5, colour="black")
```

Off-target Kinases plotting for Mike Z. 
```{r fig.height=6, fig.width=16}
ordered_cell_lines <- mouse %>% dselect(cell_line, genes) %>% gather(gene,RNA_expression, -cell_line) %>% dfilter(!is.na(RNA_expression)) %>% group_by(cell_line) %>% 
  summarise(value=median(RNA_expression)) %>% arrange(value) %>% pull(cell_line)


mouse %>% dselect(cell_line, genes) %>% gather(gene,RNA_expression, -cell_line) %>% dfilter(!is.na(RNA_expression)) %>% 
  mutate(cell_line = factor(cell_line, levels=ordered_cell_lines)) %>% mutate(gene = factor(gene, levels=c("Cdk6","Cdc7","Mapk8","Mapk9","Gsk3a","Gsk3b","Eif2ak4","Ly6g","gMDSC_score","Trib2"))) %>%
    ggplot(aes(x=cell_line, y=RNA_expression, group=cell_line, colour= cell_line)) + 
  stat_summary(fun.y=median, fun.ymax=median, fun.ymin=median, geom="crossbar", size=0.5, colour="black") + 
  geom_point(show.legend=FALSE) + 
  facet_wrap(~gene, ncol=10, scales= "free_x") + expand_limits(y = c(5,7)) +
  theme(axis.title = element_text(size=14), axis.text.y = element_text(size=12), axis.text.x = element_text(size=10,hjust=1),strip.text.x=element_text(size=14)) +  
  coord_flip() +

  ggsave("GCN2_off_target_Mouse_Horz.jpg", width=16, height=6, dpi=600, plot= last_plot(), units = "in")
```


CrownBio Syngeneic mouse line data

```{r}
library(limma)
crown <- read.table("C:/Users/mengs/Downloads/RNAseq_expr_syngeneicmodels.csv",sep=",",header=TRUE)

crown %<>% separate(Gene, sep="\\(", into=c("Gene","ENSG")) %>% dselect(-ENSG)
crown %>% gather(Cell_line,Signal,-Gene) %>% ggplot(aes(Cell_line, Signal)) + geom_boxplot(outlier.size=0.1, outlier.alpha = 0.1)
crown_n <- normalizeBetweenArrays(crown[,2:33])
crown_z <- crown %>% mutate_if(is.numeric,mx_zscore)

crown_n %>% data.frame %>% gather(Cell_line,Signal) %>% ggplot(aes(Cell_line, Signal)) + geom_boxplot(outlier.size=0.1, outlier.alpha = 0.1)

crown_z <- crown_n %>% data.frame %>% mutate_if(is.numeric,mx_zscore)

crown_z %>% data.frame %>% gather(Cell_line,Signal) %>% ggplot(aes(Cell_line, Signal)) + geom_boxplot(outlier.size=0.1, outlier.alpha = 0.1)

crown_z %<>% mutate(Gene= crown[,1]) 
crown_z %<>% dselect(Gene, everything())

crown_z <- crown_z[!duplicated(crown_z$Gene),]
```



Calculate the MDSC scores of each mouse cell line based on our MDSC signatures
Put the score back into the mouse table
```{r}
mouse[1:15,1:15]
mouse_gMDSC_genes <- mouseAnnot %>% dfilter(HumanName %in% gMDSC_sig) %>% dselect(GENE.SYMBOL)  # translate the gMDSC signature into mouse gene names
mouse_gMDSC_genes <- mouse_gMDSC_genes[,1] #this makes the data.frame into a simpler list of values
mouse_zscore <- mouse %>% dselect()
  
mx_zscore <- function(x) {result <- (x-mean(x,na.rm=TRUE)) / sd(x,na.rm=TRUE) 
                          return(result)}

mouse_gMDSC_sig <- mouse %>% tibble:: column_to_rownames("SampleID") %>% dplyr::select_if(is.double) %>% t %>% data.frame %>% tibble::rownames_to_column("Gene") %>% mutate_if(is.numeric, mx_zscore) %>% dfilter(Gene %in% mouse_gMDSC_genes) %>% summarize_if(is.numeric, sum)

#GCN2
mouse %>% tibble:: column_to_rownames("SampleID") %>% dplyr::select_if(is.double) %>% t %>% data.frame %>% tibble::rownames_to_column("Gene") %>% mutate_if(is.numeric, mx_zscore) %>% dfilter(Gene =="Eif2ak4") %>% ggplot(aes)

mouse %>% tibble:: column_to_rownames("SampleID") %>% dplyr::select_if(is.double) %>% t %>% data.frame %>% tibble::rownames_to_column("Gene") %>% mutate_if(is.numeric, mx_zscore) %>% gather(Gene,Signal) %>% ggplot(aes(Signal, colour=Signal)) + geom_histogram(binwidth=0.2) + scale_x_continuous(name="Log2 Normalized Signal (Microarray)") +
  ggsave("Mosely_Zscore_Signal_Hist.jpg", width=5, height=5, dpi=600, plot= last_plot(), units = "in")


mouse_gMDSC_sig %>% t %>% data.frame %>% tibble::rownames_to_column("SampleID")
rownames(mouse_gMDSC_sig) <- "gMDSC_score"
#mouse %<>% dselect(-gMDSC_score)
mouse <- left_join(mouse, mouse_gMDSC_sig %>% t %>% data.frame %>% tibble::rownames_to_column("SampleID"), by = "SampleID")
```

ISR response kinase target signature for George
1. Take Z-score of all genes for independent replicates of samples. 
2. Select markers for each Kinase to make a heatmap: examine it to see how specific each marker is for its intended kinase
3. Calculate a weighted score for GCN2-specific ISR marker score

```{r}
mouse_z <- mouse %>% tibble::column_to_rownames("SampleID") %>% dplyr::select_if(is.double) %>% t %>% data.frame %>% tibble::rownames_to_column("Gene") %>% mutate_if(is.numeric, mx_zscore)
save(mouse_z, file="Mosely_data_zscore.rdata")
load("D:/FLX/Syngeneic_mouse_models/Mosely_data_zscore.rdata") #mouse_z
load("/Volumes/Enterprise/FLX/Syngeneic_mouse_models/Mosely_data_zscore.rdata")
mouse_zz <- mouse_z %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("experimental.group") %>% mutate_if(is.numeric,mx_zscore2) %>% tibble::column_to_rownames("experimental.group") %>% t %>% as.data.frame %>% tibble::rownames_to_column("Gene")
save(mouse_zz, file="Mosely_data_zzscore.rdata")
```

ISR signatures
Missing: Cebpb, Ifit3
Lokup mouse gene names
Trfr2 == Tfr2
Stch ==Hspa13
```{r}
Common <- c("Ddit3","Trib3","Ppp1r15a","Asns")
#GCN2 <- c("Hspa5","Slc3a2","Hspa13")
#PRK <- c("Ifit1","Isg15","Usp18")
#PERK <- c("P4hb")
HRI <- c("Adm2","Slc7a11","Grb10","Nupr1","Soat2","Lrrc1","Glipr2","Arl14ep","Pycr1")
HRI_sub <-c("Slc7a11","Soat2","Arl14ep","Pycr1") 
aas_sig_m <- c("Clic4","Yars","Sars","Phgdh","Gars","Asns","Sat1","Slc3a2","Ddit3","Wars","Chac1","Herpud1","Pycr1","Trib3","Cebpb","Ppp1r15a","Atf4") # "C6orf48" is missing and removed
Arg_sig_u
ISR_markers <- unique(c(Common,PRK,PERK,HRI_sub,gMDSC_sig_m$GENE.SYMBOL, mMDSC_sig_m,"Eif2ak4",Arg_sig_m$GENE.SYMBOL,"Arg1"))

gMDSC_sig <- c("DYSF", "C5AR1", "TREM1", "CSF3R", "DEFA1", "CXCR2", "PLBD1", "CMTM2", "CXCR1", "TNFRSF10C", "LTF", "F13A1", "PPBP", "VNN3", "PADI4", "GLT1D1", "CLEC4D", "LCN2", "BPI", "CAMP", "CD24", "PGLYRP1", "CEACAM1", "S100P", "CYP4F3", "CLC", "S100A12", "MCEMP1", "BST1", "ARG1", "CDA", "ADGRG3", "CSF2RB", "IL1R2", "IL1RAP", "KCNJ15", "LIMK2", "DOCK5", "STX3", "FFAR2", "MEFV", "SIRPB1")
gMDSC_sig_m <- c("Dysf", "C5ar1", "Trem1", "Csf3r", "Defa1", "Cxcr2", "Plbd1", "Cmtm2", "Cxcr1", "Tnfrsf10c", "Ltf", "F13a1", "Ppbp", "Vnn3", "Padi4", "Glt1d1", "Clec4d", "Lcn2", "Bpi", "Camp", "Cd24", "Pglyrp1", "Ceacam1", "S100p", "Cyp4f3", "Clc", "S100a12", "Mcemp1", "Bst1", "Arg1", "Cda", "Adgrg3", "Csf2rb", "Il1r2", "Il1rap", "Kcnj15", "Limk2", "Dock5", "Stx3", "Ffar2", "Mefv", "Sirpb1")
mMDSC_sig_m <- c("Csf3r", "Slc6a6", "Trem1", "Clec4e", "Plbd1")
load("D:/FLX/Silpa_data/mouse_human_gene.rdata")
load("/Volumes/Enterprise/FLX/Silpa_data/mouse_human_gene.rdata")
mouseAnnot %>% dfilter(GENE.SYMBOL=="Hspa13")

hypoxia <- read.table("D:/FLX/GCN2/ISR_data/RNA_seq/GSEA_hypoxia.txt",sep="\n", header=TRUE) %>% pull(HALLMARK_HYPOXIA)
hypoxia <- read.table("/Volumes/Enterprise/FLX/GCN2/ISR_data/RNA_seq/GSEA_hypoxia.txt",sep="\n", header=TRUE) %>% pull(HALLMARK_HYPOXIA) %>% .[2:201]
hypoxia_m <- mouseAnnot %>% dfilter(HumanName %in% hypoxia) %>% dselect(GENE.SYMBOL) %>% pull(GENE.SYMBOL)

```

New ISR signatures Dec 2019
```{r fig.height=5, fig.width=5}
load("/Volumes/Enterprise/FLX/GCN2/ISR_data/RNA_seq/Dec24_2019_ISR_zz.rdata") #ISR_Dec

PERK_comb$Gene
GCN2_comb$Gene
HRI_marker$Gene
Common <- c("Ddit3","Trib3","Ppp1r15a","Asns")
Common_RAPT <- c("Adm2","Arl14ep","Chac1","Cth","Ddit3","Nupr1","Shmt2","Slc1a4","Soat2","Trib3")
#GPH <- c("Adm2","Arl14ep","Chac1","Cth", "Ddit3","Nupr1","Shmt2","Slc1a4","Soat2","Trib3")
PERK <- c("Mrgprf","Lzts1","Wnt1" ,"Ptpn5","Gem","Slc2a6", "Snx32","Gm11331", "Prob1" , "Ccl7","Dbhos" ,"Mir99ahg","Trpm6", "N4bp2os","Glis1" ,"Mmp19","Lrrc73","Cd74","Pkdcc",   "2700069I18Rik","B230217C12Rik","Osr2")
HRI <- c("Atf5","Lrrc1","Glipr2","Grb10")
GCN2 <-c( "Cep85l","Txnrd1","Angptl4","Antxr2","Tcim", "Ptgs1", "Bdkrb2", "Prune2","Rbpj","Wdr45","Znrf2","Cnnm4","Grem1","Tnfaip2","Gpr137b-ps", "Car8", "Wipi2","Tmem171","Ndst3","Foxl1")
aas_sig_m <- c("Clic4","Yars","Sars","Phgdh","Gars","Asns","Sat1","Slc3a2","Ddit3","Wars","Chac1","Herpud1","Pycr1","Trib3","Cebpb","Ppp1r15a","Atf4")
ISR_markers <- unique(c(Common,GCN2,PERK,HRI,gMDSC_sig_m,mMDSC_sig_m,"Eif2ak4"))
```

Pull out the signatures from the z_score data in order to do correlation analysis: this data is not input for double z score: zzscore generation
```{r}
ISR_RAPT <- mouse_z %>% dfilter(Gene %in% Common_RAPT) %>% summarise_if(is.numeric,sum)
ISR_common <- mouse_z %>% dfilter(Gene %in% Common) %>% summarise_if(is.numeric,sum)
GCN2_s <- mouse_z %>% dfilter(Gene %in% GCN2) %>% summarise_if(is.numeric,sum)
PERK_s <- mouse_z %>% dfilter(Gene %in% PERK) %>% summarise_if(is.numeric,sum)
HRI_s <- mouse_z %>% dfilter(Gene %in% HRI) %>% summarise_if(is.numeric,sum)
AAS_s <- mouse_z %>% dfilter(Gene %in% aas_sig_m) %>% summarise_if(is.numeric,sum)
gMDSC_s <- mouse_z %>% dfilter(Gene %in% gMDSC_sig_m) %>% summarise_if(is.numeric,function(x){sum(x,na.rm=TRUE)})
mMDSC_s <- mouse_z %>% dfilter(Gene %in% mMDSC_sig_m) %>% summarise_if(is.numeric,function(x){sum(x,na.rm=TRUE)})

#cor(ISR_RAPT,GCN2_s)

ISR_s <- bind_rows(ISR_RAPT,ISR_common,GCN2_s,PERK_s,HRI_s,AAS_s,gMDSC_s,mMDSC_s)
rownames(ISR_s) <- c("ISR_RAPT","Common","GCN2","PERK","HRI","AAS","gMDSC","mMDSC")
#ISR_s[,4:96]

ISR_s_t <- ISR_s %>% t %>% data.frame 
cor(ISR_s_t$Common,ISR_s_t$GCN2) #0.09
cor(ISR_s_t$AAS,ISR_s_t$GCN2) #0.09
plot(ISR_s_t$ISR_RAPT,ISR_s_t$GCN2) #-0.08

cor(ISR_s_t$Common,ISR_s_t$AAS) #0.88
plot(ISR_s_t$Common,ISR_s_t$AAS) 
ggplot(ISR_s_t,aes(x=ISR_s_t$Common,y=ISR_s_t$AAS)) + geom_point(size=1) + xlab("Common ISR score") + ylab("AAS score") + ggsave(file="Common_vs_AAS_score_Mosely.jpg", units ="in", width =3, height=3,dpi=300)
ggplot(ISR_s_t,aes(y=ISR_s_t$AAS,x=ISR_s_t$GCN2)) + geom_point(size=1) + xlab("GCN2 score") + ylab("AAS score") + ggsave(file="AAS_vs_GCN2_score_Mosely.jpg", units ="in", width =3, height=3,dpi=300)
ggplot(ISR_s_t,aes(x=ISR_s_t$ISR_RAPT,y=ISR_s_t$GCN2)) + geom_point(size=1) + xlab("Common ISR score") + ylab("GCN2 score") + ggsave(file="Common_vs_GCN2_score_Mosely.jpg", units ="in", width =3, height=3,dpi=300)

cor(ISR_s_t$gMDSC,ISR_s_t$mMDSC) #0.90
plot(ISR_s_t$gMDSC,ISR_s_t$mMDSC) #0.90

cor(ISR_s_t$ISR_RAPT,ISR_s_t$AAS) #0.88
cor(ISR_s_t$ISR_RAPT,(ISR_s_t$PERK)) #-0.06
cor(ISR_s_t$AAS,(ISR_s_t$PERK+ISR_s_t$GCN2+ISR_s_t$HRI)) #0.029
cor(ISR_s_t$ISR_RAPT,ISR_s_t$HRI) #0.13
```

New ISR

ZZ then Calculate signature

```{r}
isr_z <- mouse_z %>% dfilter(Gene %in% ISR_markers)
ISR_markers <- unique(c(GPH,GCN2,PERK,HRI,gMDSC_sig_m,mMDSC_sig_m))
isr_z_z <- isr_z %>% arrange(match(Gene,ISR_markers)) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("experimental.group") %>% mutate_if(is.numeric,mx_zscore2) %>% tibble::column_to_rownames("experimental.group") %>% t %>% as.data.frame %>% tibble::rownames_to_column("Gene")

Common_s <- mouse_zz %>% dfilter(Gene %in% Common) %>% summarise_if(is.numeric,sum)
GCN2_s <- mouse_zz %>% dfilter(Gene %in% GCN2) %>% summarise_if(is.numeric,sum)
PERK_s <- mouse_zz %>% dfilter(Gene %in% PERK) %>% summarise_if(is.numeric,sum)
HRI_s <- mouse_zz %>% dfilter(Gene %in% HRI) %>% summarise_if(is.numeric,sum)
AAS_s <- mouse_zz %>% dfilter(Gene %in% aas_sig_m) %>% summarise_if(is.numeric,sum)
gMDSC_s <- mouse_zz %>% dfilter(Gene %in% gMDSC_sig_m) %>% summarise_if(is.numeric,sum)
mMDSC_s <- mouse_zz %>% dfilter(Gene %in% mMDSC_sig_m) %>% summarise_if(is.numeric,sum)
Hypoxia <- mouse_zz %>% dfilter(Gene %in% hypoxia_m) %>% summarise_if(is.numeric,sum)

ISR_s <- bind_rows(Common_s,AAS_s,GCN2_s,PERK_s,HRI_s,gMDSC_s,mMDSC_s,Hypoxia)
rownames(ISR_s) <- c("ISR_Common","AAS","GCN2","PERK","HRI","gMDSC","mMDSC","Hypoxia")
ISR_s_t <- ISR_s %>% t %>% data.frame 
ISR_s_t <- bind_cols(samples.info_tum,ISR_s_t)

  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }

  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }

isr_genes_z <- isr_z %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame()

isr_genes_zz <- isr_z %>% arrange(match(Gene,ISR_markers)) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("experimental.group") %>% mutate_if(is.numeric,mx_zscore) %>% tibble::column_to_rownames("experimental.group")  

med_sig_cor <- round(cor(ISR_s_t),1)
med_sig_cor <- get_upper_tri(med_sig_cor)
melt_med_sig_cor <- melt(med_sig_cor)
melt_med_sig_cor %>% ggplot(aes(x=Var1, y=Var2, fill=value, label=value)) + geom_tile(color = "white")+
    geom_text(aes(Var1, Var2, label = value), color = "grey30", size = 2) +
 scale_fill_gradient2(low = "steelblue", high = "red", na.value="white", limit = c(-0.5,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
    scale_x_discrete(position = "top") +
 theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 0))+
 coord_fixed() +
ggsave("Sig_corr_median_heatmap_Dec27.jpg", width=6, height=6, dpi=600, plot= last_plot(), units = "in") 



ISR_s_t %>%  group_by(`experimental group`) %>% dplyr::summarise_if(is.double,mean) %>% dplyr::ungroup() %>% dfilter(grepl("_T$",`experimental group`)) %>% arrange(desc(GCN2)) %>% mutate(Group=`experimental group`) %>% dselect(`experimental group`,Group,ISR_Common,AAS,GCN2,PERK,HRI,gMDSC,mMDSC) %>% write.table(file="ISR_ZZscore_then_sig_Feb25_sortGCN.cdt", sep="\t",row.names = FALSE, quote=FALSE)

ISR_zz <- ISR_s_t %>%  group_by(`experimental group`) %>% dplyr::summarise_if(is.double,mean) %>% dplyr::ungroup() %>% dfilter(`experimental group` %like% "_T$") %>% arrange(desc(GCN2)) %>% mutate(Group=`experimental group`) %>% separate(Group,into=c("Model",NA), sep="_") %>% dselect(Model,ISR_Common,GCN2,PERK,HRI,gMDSC,mMDSC) %>% write.table(file="ISR_zz_Feb25.txt", sep="\t",quote=FALSE,row.names = FALSE)

```

Convert the signature score to Quantiles
```{r}
ISR_zz_WB <- read.table("Mouse_model_zz_WB_quant.txt", sep="\t", header=TRUE)
semi_quant_mx <- function(x) {round((x/max(x,na.rm=TRUE))*5,digits=0)}
quant_mx <- function(x) {ntile(x,5)-1} # quantiles will range from 0 to 4 
library(scales)
isr_semiquant <- ISR_zz_WB %>% mutate_if(is.numeric,semi_quant_mx) 
isr_quantiles %>% mutate(Summary=ISR_Common+GCN2+gMDSC+mMDSC) %>% arrange(desc(Summary)) 
#add in the WB data
model_order <- isr_quantiles %>% mutate(Summary=ISR_Common+GCN2+gMDSC+mMDSC) %>% arrange(desc(Summary)) %>% pull(Model)
signature_order <- colnames(isr_quantiles)[2:11]
isr_semiquant %>% 
  dselect(-GCN2_WB) %>%
  gather(Signature,SemiQuantitative_score,-Model) %>% 
  mutate(Model= factor(Model, levels=rev(model_order)), Signature= factor(Signature,levels=signature_order)) %>%
   ggplot(aes(y=Model, x=Signature, fill=SemiQuantitative_score)) + 
      geom_tile() +
      theme(legend.position="none") +
      geom_text(aes(y=Model, x=Signature, label = SemiQuantitative_score), color = "grey30", size = 3) +
      scale_fill_gradient2(low = "dodgerblue4", high = "red", na.value="grey80", limit = c(0,5),oob=squish) 


isr_semiquant %>% dselect(Model,contains("_WB")) %>% write.table(file="WB_Quant_zz_Feb25.txt", sep="\t",quote=FALSE,row.names = FALSE)
```

Correlation 

The grey area is the se= local standard error 
```{r fig.height=3, fig.width=3}
ISR_zz_WB %>% ggplot(aes(x=ATF4_WB, y=GCN2)) + geom_point() + geom_smooth(method="lm") + annotate("text", x=0.04,y=-10, label = "paste(italic(R) ^ 2, \" = 0.87\")", parse = TRUE) + ggsave("GCN2_ATF4_seq_WB_cor_plot.jpg", width=3, height=3, dpi=150, plot= last_plot(), units = "in") 

lm <- lm(ATF4_WB ~ GCN2, data=ISR_zz_WB)
lm <- lm(pGCN2_WB ~ GCN2, data=ISR_zz_WB)
lm <- lm(pGCN2_WB ~ ATF4_WB, data=ISR_zz_WB)
summary(lm)
ISR_zz_WB %>% ggplot(aes(x=pGCN2_WB, y=GCN2)) + geom_point() + geom_smooth(method="lm") + annotate("text", x=0.04,y=-8, label = "paste(italic(R) ^ 2, \" = 0.19\")", parse = TRUE)+ ggsave("GCN2_pGCN2_seq_WB_cor_plot.jpg", width=3, height=3, dpi=150, plot= last_plot(), units = "in") 

ISR_zz_WB %>% ggplot(aes(x=pGCN2_WB, y=ATF4_WB)) + geom_point() + geom_smooth(method="lm") + annotate("text", x=0.04,y=0, label = "paste(italic(R) ^ 2, \" = 0.16\")", parse = TRUE) + ggsave("ATF4_pGCN2_seq_WB_cor_plot.jpg", width=3, height=3, dpi=150, plot= last_plot(), units = "in") 

```


Mouse z_score histogram
```{r}

mouse_z %>% gather(Cell_line,Signal,-Gene) %>% ggplot(aes(Cell_line, Signal)) + geom_boxplot(outlier.size=0.1, outlier.alpha = 0.1)
```
ISR
gMDSC
Select marker genes out of dataset and merge back with cell line name
```{r fig.height=5, fig.width=3}
isr_z <- mouse_z %>% dfilter(Gene %in% ISR_markers)
colnames(isr_z) <- c("Gene",paste0(mouse$experimental.group,'_',1:93)) 
gMDSC_z_score <- isr_z %>% dfilter(Gene %in% gMDSC_sig_m$GENE.SYMBOL) %>% summarise_if(is.numeric,sum) %>% mutate(Gene="gMDSC_score") %>% dselect(Gene, everything())
colnames(gMDSC_z_score) <- c("Gene",paste0(mouse$experimental.group,'_',1:93))
mMDSC_z_score <- isr_z %>% dfilter(Gene %in% mMDSC_sig_m) %>% summarise_if(is.numeric,sum)
colnames(mMDSC_z_score) <- c(paste0(mouse$experimental.group,'_',1:93))
Arg_z_score <- isr_z_z %>% dfilter(Gene %in% Arg_sig_m$GENE.SYMBOL) %>% summarise_if(is.numeric,sum)
colnames(Arg_z_score) <- c(paste0(mouse$experimental.group,'_',1:93))
Arg_z_score <- Arg_z_score/5

 
isr_z %>% mutate(Gene2=Gene) %>% dselect(Gene,Gene2,everything()) %>% arrange(match(Gene,ISR_markers)) %>% write.table(file="ISR_zscore.cdt",row.names = FALSE,sep="\t")

isr_mdsc_z <- rbind(gMDSC_z_score,isr_z)

isr_z %>% arrange(match(Gene,ISR_markers)) %>% write.table(file="ISR_zscore.txt",row.names = FALSE,sep="\t")

isr_z %>% gather(Cell,zscore,-Gene) %>% mutate(Gene=factor(Gene,levels=c(rev(ISR_markers)))) %>% ggplot(aes(x=Gene,y=zscore, fill=Gene)) + geom_violin(show.legend=FALSE) + geom_point(size=0.5,show.legend=FALSE) + coord_flip() +  theme(axis.title = element_text(size=8), axis.text.y = element_text(size=16), axis.text.x = element_text(size=16,hjust=1))

#crown_z  %>% dfilter(Gene %in% ISR_markers) %>% gather(Cell,zscore,-Gene) %>% mutate(Gene=factor(Gene,levels=c(rev(ISR_markers)))) %>% ggplot(aes(x=Gene,y=zscore, fill=Gene)) + geom_violin(show.legend=FALSE) + geom_point(size=0.5,show.legend=FALSE) + coord_flip() +  theme(axis.title = element_text(size=8), axis.text.y = element_text(size=16), axis.text.x = element_text(size=16,hjust=1))
```

Take z-scores of genes across samples: I.e. take per sample cross_gene z-score, then take per gene cross_sample z-score
```{r fig.height=5, fig.width=3}

isr_z_z <- isr_z %>% arrange(match(Gene,ISR_markers)) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("experimental.group") %>% mutate_if(is.numeric,mx_zscore) %>% tibble::column_to_rownames("experimental.group") %>% t %>% as.data.frame %>% tibble::rownames_to_column("Gene")

#crown_z_z <- crown_z  %>% dfilter(Gene %in% ISR_markers) %>% arrange(match(Gene,ISR_markers)) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("experimental.group") %>% mutate_if(is.numeric,mx_zscore) %>% tibble::column_to_rownames("experimental.group") %>% t %>% as.data.frame %>% tibble::rownames_to_column("Gene")

isr_z_z %>% gather(Cell,zscore,-Gene) %>% mutate(Gene=factor(Gene,levels=rev(ISR_markers))) %>% ggplot(aes(x=Gene,y=zscore, fill=Gene)) + geom_violin(show.legend=FALSE) + geom_point(size=0.5,show.legend=FALSE) + coord_flip() +  theme(axis.title = element_text(size=8), axis.text.y = element_text(size=16), axis.text.x = element_text(size=16,hjust=1))

#crown_z_z %>% gather(Cell,zscore,-Gene) %>% mutate(Gene=factor(Gene,levels=rev(ISR_markers))) %>% ggplot(aes(x=Gene,y=zscore, fill=Gene)) + geom_violin(show.legend=FALSE) + geom_point(size=0.5,show.legend=FALSE) + coord_flip() +  theme(axis.title = element_text(size=8), axis.text.y = element_text(size=16), axis.text.x = element_text(size=16,hjust=1))

isr_z %>% arrange(match(Gene,ISR_markers)) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("experimental.group") %>% mutate_if(is.numeric,mx_zscore) %>% tibble::column_to_rownames("experimental.group") %>% t %>% as.data.frame %>% tibble::rownames_to_column("Gene") %>% mutate(Gene2=Gene) %>% dselect(Gene,Gene2,everything()) %>% write.table(file="ISR_zscore_twice.cdt",row.names = FALSE,sep="\t")

isr_z %>% arrange(match(Gene,ISR_markers)) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("experimental.group") %>% mutate_if(is.numeric,mx_zscore) %>% tibble::column_to_rownames("experimental.group") %>% t %>% as.data.frame %>% tibble::rownames_to_column("Gene") %>% dselect(Gene,everything()) %>% write.table(file="ISR_zscore_twice.txt",row.names = FALSE,sep="\t")
```

Correlation of markers
```{r fig.height=6, fig.width=6}
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }

  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }

isr_genes_z <- isr_z %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame()

isr_genes_zz <- isr_z %>% arrange(match(Gene,ISR_markers)) %>% tibble::column_to_rownames("Gene") %>% t %>% data.frame %>% tibble::rownames_to_column("experimental.group") %>% mutate_if(is.numeric,mx_zscore) %>% tibble::column_to_rownames("experimental.group")  

med_sig_cor <- round(cor(isr_genes_zz),1)
med_sig_cor <- get_upper_tri(med_sig_cor)
melt_med_sig_cor <- melt(med_sig_cor)
melt_med_sig_cor %>% ggplot(aes(x=Var1, y=Var2, fill=value, label=value)) + geom_tile(color = "white")+
    geom_text(aes(Var1, Var2, label = value), color = "grey30", size = 2) +
 scale_fill_gradient2(low = "steelblue", high = "red", na.value="white", limit = c(-0.5,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
    scale_x_discrete(position = "top") +
 theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 0))+
 coord_fixed() +
ggsave("Sig_corr_median_heatmap_Sep1.jpg", width=6, height=6, dpi=600, plot= last_plot(), units = "in")  

HRI_sub <-c("Slc7a11","Soat2","Arl14ep","Pycr1") 

``` 

##############
Old code: Before in-house GCN2 signature was known
############

Calculate GCN2 signature from other ISR signatures
```{r}
Common <- c("Ddit3","Trib3","Ppp1r15a","Asns")
GCN2 <- c("Hspa5","Slc3a2","Hspa13")
PRK <- c("Ifit1","Isg15","Usp18")
PERK <- c("P4hb")
#HRI <- c("Adm2","Slc7a11","Grb10","Nupr1","Soat2","Lrrc1","Glipr2","Arl14ep","Pycr1")
HRI_sub <-c("Slc7a11","Soat2","Arl14ep","Pycr1") 
aas_sig_m <- c("Clic4","Yars","Sars","Phgdh","Gars","Asns","Sat1","Slc3a2","Ddit3","Wars","Chac1","Herpud1","Pycr1","Trib3","Cebpb","Ppp1r15a","Atf4") # "C6orf48" is missing and removed
gMDSC_sig_m <- c("Dysf", "C5ar1", "Trem1", "Csf3r", "Defa1", "Cxcr2", "Plbd1", "Cmtm2", "Cxcr1", "Tnfrsf10c", "Ltf", "F13a1", "Ppbp", "Vnn3", "Padi4", "Glt1d1", "Clec4d", "Lcn2", "Bpi", "Camp", "Cd24", "Pglyrp1", "Ceacam1", "S100p", "Cyp4f3", "Clc", "S100a12", "Mcemp1", "Bst1", "Arg1", "Cda", "Adgrg3", "Csf2rb", "Il1r2", "Il1rap", "Kcnj15", "Limk2", "Dock5", "Stx3", "Ffar2", "Mefv", "Sirpb1")
mMDSC_sig_m <- c("Csf3r", "Slc6a6", "Trem1", "Clec4e", "Plbd1")

Common_s <- isr_z_z %>% dfilter(Gene %in% Common) %>% summarise_if(is.numeric,sum)
PKR_s <- isr_z_z %>% dfilter(Gene %in% PRK) %>% summarise_if(is.numeric,sum)
PERK_s <- isr_z_z %>% dfilter(Gene %in% PERK) %>% summarise_if(is.numeric,sum)
HRI_s <- isr_z_z %>% dfilter(Gene %in% HRI_sub) %>% summarise_if(is.numeric,sum)
#GCN2_s <- isr_z_z %>% dfilter(Gene %in% GCN2) %>% summarise_if(is.numeric,sum)
AAS_s <- isr_z_z %>% dfilter(Gene %in% aas_sig_m) %>% summarise_if(is.numeric,sum)
gMDSC_s <- gMDSC_z_score %>% dselect(-Gene)
mMDSC_s <- mMDSC_z_score
Arg_z_score

#Common_s <- isr_z %>% dfilter(Gene %in% Common) %>% summarise_if(is.numeric,sum)
#PRK_s <- isr_z %>% dfilter(Gene %in% PRK) %>% summarise_if(is.numeric,sum)
#PERK_s <- isr_z %>% dfilter(Gene %in% PERK) %>% summarise_if(is.numeric,sum)
#HRI_s <- isr_z %>% dfilter(Gene %in% HRI_sub) %>% summarise_if(is.numeric,sum)
#GCN2_s <- isr_z %>% dfilter(Gene %in% GCN2) %>% summarise_if(is.numeric,sum)

#CN2_s <- (Common_s - PRK_s - PERK_s - HRI_s)

ISR_s <- bind_rows(Common_s,PKR_s,PERK_s,HRI_s,gMDSC_s,mMDSC_s,Arg_z_score,AAS_s)
rownames(ISR_s) <- c("Common","PKR","PERK","HRI","gMDSC","mMDSC","ARG_AAS","Pan_AAS")
ISR_s_t <- ISR_s %>% t %>% data.frame 
ISR_s_t <- bind_cols(samples.info_tum,ISR_s_t)

ISR_s_t %>% mutate(PKRi= ifelse(PKR<0,0,PKR), PERKi=ifelse(PERK<0,0,PERK), HRIi=ifelse(HRI<0,0,HRI)) %>% mutate(GCN2_specific = Common - PKRi - PERKi - HRIi) %>% group_by(`experimental group`) %>% dplyr::summarise_if(is.double,mean) %>% dplyr::ungroup() %>% dfilter(`experimental group` %like% "_T$") %>% arrange(desc(ARG_AAS)) %>% mutate(Group=`experimental group`) %>% dselect(-PKRi,-PERKi, -HRIi,Group,Common,PKR,PERK,HRI,GCN2_specific,everything()) %>% write.table(file="ISR_scores_grouped_ZZscore_Sep26_2.txt", sep="\t",row.names = FALSE, quote=FALSE)
# %>% dselect(`experimental group`,Group,Common,PRK,PERK,HRI,GCN2_specific,AAS,gMDSC,mMDSC)


ISR_s_t %>% mutate(PRKi= ifelse(PRK<0,0,PRK), PERKi=ifelse(PERK<0,0,PERK), HRIi=ifelse(HRI<0,0,HRI)) %>% 
  mutate(GCN2_specific = Common - PRKi - PERKi - HRIi) %>% group_by(`experimental group`) %>% dplyr::summarise_if(is.double,mean) %>% dplyr::ungroup() %>% dfilter(`experimental group` %like% "_T$") %>% arrange(desc(GCN2_specific)) %>% mutate(Group=`experimental group`) %>% dselect(-PRKi,-PERKi, -HRIi) %>% dselect(`experimental group`,Group,Common,PRK,PERK,HRI,GCN2_specific,AAS,gMDSC,mMDSC) %>% melt %>% ggplot(aes(x=Group, y=variable, fill=value, label=value)) + geom_tile(color = "white") +
scale_fill_gradient2(low = "dodgerblue4", high = "red", na.value="white", limit = c(-6,6), name="Score") +
theme_minimal()+ 
scale_x_discrete(position = "top") +
theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 0, size=14),axis.text.y = element_text(size=14)) +
coord_fixed() 
ggsave("Tumor_signature_heatmap.jpg", width=12, height=6, dpi=600, plot= last_plot(), units = "in")  

```

Convert ISR_zz_score table to quantiles for semi-quantitative transformation

VV_low -2
v_low -1
Low 0
medium 1
high 2

```{r}
isr_zz <- read.table("ISR_scores_grouped_ZZscore_Sep26_2.txt", sep="\t", header=TRUE)

isr_zz %>% dselect(-PKRi,-PERKi,-HRIi,-ARG_AAS,-GCN2_specific) %>% dselect(experimental.group, Group, everything()) %>% write.table(file="ISR_zzscore_Dec2019.cdt",sep="\t", row.names=FALSE)
quant_mx <- function(x) {ntile(x,5)-3}

isr_quantiles <- isr_zz %>% mutate_if(is.numeric,quant_mx) 
isr_quantiles %>% dselect(-Group,-ARG_AAS,-GCN2_specific) %>% mutate(Summary=Common-median(PKRi,PERKi,HRIi)+gMDSC+mMDSC+Pan_AAS) %>%dselect(-PKRi,-PERKi,-HRIi) %>% arrange(desc(Summary)) %>% write.table(file="Scores_zzscore_Dec2019.txt",sep="\t", row.names=FALSE, quote=FALSE) 
other <- c("PKR","PERK","HRI")
isr_quantiles %>% dselect(-Group,-ARG_AAS,-GCN2_specific) %>% mutate(PERKi=ifelse(PERK<0,0,PERK),HRIi=ifelse(HRI<0,0,HRI),PKRi=ifelse(PKR<0,0,PKR)) %>% mutate(Other_ISR=(HRIi+PERKi+PKRi), MDSC=(mMDSC+gMDSC)/2) %>% dselect(-PKR,-PERK,-HRI,-mMDSC,-gMDSC,-PKRi,-PERKi,-HRIi) %>% mutate(Summary=Common-Other_ISR+MDSC+Pan_AAS) %>% arrange(desc(Summary)) %>% dselect(experimental.group,Common,Other_ISR,MDSC,Pan_AAS,Summary) %>% write.table(file="Simplifed_scores_zzscore_Dec2019.txt",sep="\t", row.names=FALSE, quote=FALSE)

median(2,0,2)
```



Calculate AAS scores
```{r}

aas_sig <- c("CLIC4","YARS","SARS","PHGDH","C6orf48","GARS","ASNS","SAT1","SLC3A2","DDIT3","WARS","CHAC1","HERPUD1","PYCR1","TRIB3","CEBPB","PPP1R15A","ATF4")

aas_sig_m <- c("Clic4","Yars","Sars","Phgdh","Gars","Asns","Sat1","Slc3a2","Ddit3","Wars","Chac1","Herpud1","Pycr1","Trib3","Cebpb","Ppp1r15a","Atf4") # "C6orf48" is missing and removed

mouseAnnot %>% dfilter(GENE.SYMBOL %in% aas_sig_m)
```
Signal Scores from Complete CrownBio data
```{r}
Common_s <- crown_z_z %>% dfilter(Gene %in% Common) %>% summarise_if(is.numeric,sum)
PRK_s <- crown_z_z %>% dfilter(Gene %in% PRK) %>% summarise_if(is.numeric,sum)
PERK_s <- crown_z_z %>% dfilter(Gene %in% PERK) %>% summarise_if(is.numeric,sum)
HRI_s <- crown_z_z %>% dfilter(Gene %in% HRI_sub) %>% summarise_if(is.numeric,sum)
GCN2_s <- crown_z_z %>% dfilter(Gene %in% GCN2) %>% summarise_if(is.numeric,sum)
AAS_s <- crown_z_z %>% dfilter(Gene %in% aas_sig_m) %>% summarise_if(is.numeric,sum)
gMDSC_s <- crown_z_z %>% dfilter(Gene %in% gMDSC_sig_m) %>% summarise_if(is.numeric,sum)
mMDSC_s <- crown_z_z %>% dfilter(Gene %in% mMDSC_sig_m) %>% summarise_if(is.numeric,sum)
crown_z_z %>% dfilter(Gene =="Csf3r")
ISR_s <- bind_rows(Common_s,PRK_s,PERK_s,HRI_s,GCN2_s,AAS_s,gMDSC_s,mMDSC_s)
rownames(ISR_s) <- c("Common","PRK","PERK","HRI","GCN2s","AAS","gMDSC","mMDSC")
ISR_s_t <- ISR_s %>% t %>% data.frame 
#ISR_s_t <- bind_cols(samples.info_tum,ISR_s_t)
lines <- rownames(ISR_s_t)
ISR_s_t %>% mutate(PRKi= ifelse(PRK<0,0,PRK), PERKi=ifelse(PERK<0,0,PERK), HRIi=ifelse(HRI<0,0,HRI)) %>% mutate(GCN2_specific = Common - PRKi - PERKi - HRIi)  %>% arrange(desc(GCN2s)) %>% mutate(Cell_line=lines, Cellline=lines) %>% dselect(-PRKi,-PERKi, -HRIi) %>% dselect(Cell_line,Cellline,GCN2s,Common,PRK,PERK,HRI,GCN2_specific,AAS,gMDSC,mMDSC) %>% write.table(file="ISR_scores_grouped_ZZscore_CrownBio_GCN2_s.cdt", sep="\t",row.names = FALSE, quote=FALSE)
```



Manually Downloaded CrownBio RNA signal
```{r}
crownbio <- read.table("D:/FLX/General\ working\ code/Mouse_syngeneic_RNA_CrownBio.txt",sep="\t",header=TRUE)
CrownBio_RNA <- crownbio %>% dselect(-CANCER.TYPE,-ID) %>% mutate() %>% spread(GENE,LOG2.FPKM.) %>% mutate(GENE=factor(GENE,levels=rev(ISR_markers)))
gene_order <- c("Clic4","Yars","Sars","Phgdh","Gars","Sat1","Slc3a2","Wars","Chac1","Herpud1","Pycr1","Cebpb","Atf4","Asns","Ddit3","Trib3","Ppp1r15a","Hspa5","Slc3a2","Hspa13","Ifit1","Isg15","Usp18","P4hb","Slc7a11","Soat2","Arl14ep")
CrownBio_RNA_Z <- CrownBio_RNA %>% mutate_if(is.numeric,mx_zscore)

CrownBio_RNA %>% mutate_if(is.numeric,mx_zscore) %>% mutate(Sample=SAMPLE.NAME) %>% select(SAMPLE.NAME, Sample, everything()) %>% arrange(desc(ASNS)) %>%  write.table(file="CrownBio_RNA_Z.txt", quote=FALSE, sep="\t", row.names = FALSE)


Common <- c("Ddit3","Trib3","Ppp1r15a","Asns")
GCN2 <- c("Hspa5","Slc3a2","Hspa13")
PRK <- c("Ifit1","Isg15","Usp18")
PERK <- c("P4hb")
HRI <- c("Adm2","Slc7a11","Grb10","Nupr1","Soat2","Lrrc1","Glipr2","Arl14ep","Pycr1")
HRI_sub <-c("Slc7a11","Soat2","Arl14ep","Pycr1") 
aas_sig_m <- c("Clic4","Yars","Sars","Phgdh","Gars","Asns","Sat1","Slc3a2","Ddit3","Wars","Chac1","Herpud1","Pycr1","Trib3","Cebpb","Ppp1r15a","Atf4")
```

