---
title: "Nanostring"
author: "Mengshu"
date: "6/11/2019"
output: html_document
---

Analyze Aparna's Mouse AD Nanostring dataset exported from nSolver software
Experimental setup: with no biological replicates
	Treatment	Challenge
1	Vehicle	  ETOH
2	Vehicle	  FITC
3	DEX	      ETOH
4	DEX	      FITC
5	FLX193	  ETOH
6	FLX193	  FITC

The challenge (ETOH mock, or FITC) was performed on the left and right ears of each mouse
Sample input was normalized by total ng of RNA across samples: due to sample limitations, 50ng was pipetted in as 1uL for each sample qc'd to 50ng/uL (this reduced precision of input amount)

```{r}
library(tidyverse)
library(data.table)
library(magrittr)
library(dplyr)
library(ggplot2)
library(limma) #quantileNormalization
library(tibble)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
```
AD samples: QC and Normalize
```{r}
raw <- read.table("Aparna_raw_June.txt", sep="\t", header=TRUE)
colnames(raw) <- c("Gene","Class",
                   "Vehicle_ETOH_A","Vehicle_FITC_A","DEX_ETOH_A","DEX_FITC_A","FLX193_ETOH_A","FLX193_FITC_A",
                   "Vehicle_ETOH_B","Vehicle_FITC_B","DEX_ETOH_B","DEX_FITC_B","FLX193_ETOH_B","FLX193_FITC_B",
                   "Vehicle_ETOH_C","Vehicle_FITC_C","DEX_ETOH_C","DEX_FITC_C","FLX193_ETOH_C","FLX193_FITC_C")
raw %<>% dselect(-Class)

boxplot(raw[,2:19], log="y")
raw %<>% tibble::column_to_rownames(var="Gene")
raw["BI_mid",] <- raw["BD",] * 100
raw["BI_high",] <- raw["BD",] * 1000
raw["BI_superhigh",] <- raw["BD",] * 10000
raw["Total_countx0.01",] <- colSums(raw)/100
QN <- normalizeQuantiles(raw, ties=TRUE)
rownames(QN) <- rownames(raw)
boxplot(QN,log="y")
#Plot Normalized Binding Density
col_sums <- colSums(QN)

barplot(col_sums)

barplot(as.matrix(QN["Total_countx0.01",]))
barplot(as.matrix(QN["BD",]))
barplot(as.matrix(QN["BI_mid",]))
barplot(as.matrix(QN["BI_high",]))
barplot(as.matrix(QN["BI_superhigh",]))
```

Gather for plotting

# Both they Key and Value are newly created for the gathered data. Specity the column that doesn't change in the 3rd term with -
```{r}
plot_AD <- QN  %>% rownames_to_column %>% gather(key= Sample, value= "Count", -rowname) # Both they Key and Value are newly created for the gathered data. Specity the column that doesn't change in the 3rd term with -
colnames(plot_AD) <- c("Gene","Sample","Count")
plot_AD$Sample <- factor(plot_AD$Sample, levels= c(
                   "Vehicle_ETOH_A","Vehicle_FITC_A","DEX_ETOH_A","DEX_FITC_A","FLX193_ETOH_A","FLX193_FITC_A",
                   "Vehicle_ETOH_B","Vehicle_FITC_B","DEX_ETOH_B","DEX_FITC_B","FLX193_ETOH_B","FLX193_FITC_B",
                   "Vehicle_ETOH_C","Vehicle_FITC_C","DEX_ETOH_C","DEX_FITC_C","FLX193_ETOH_C","FLX193_FITC_C"))

plot_raw <- raw  %>% rownames_to_column %>% gather(key= Sample, value= "Count", -rowname) 
colnames(plot_raw) <- c("Gene","Sample","Count")
plot_raw$Sample <- factor(plot_raw$Sample, levels= c(
                   "Vehicle_ETOH_A","Vehicle_FITC_A","DEX_ETOH_A","DEX_FITC_A","FLX193_ETOH_A","FLX193_FITC_A",
                   "Vehicle_ETOH_B","Vehicle_FITC_B","DEX_ETOH_B","DEX_FITC_B","FLX193_ETOH_B","FLX193_FITC_B",
                   "Vehicle_ETOH_C","Vehicle_FITC_C","DEX_ETOH_C","DEX_FITC_C","FLX193_ETOH_C","FLX193_FITC_C"))
```
Alternative method: Use Bioconductor package
```{r}
#source('http://bioconductor.org/biocLite.R')
#BiocManager::install('preprocessCore')
library(preprocessCore)

raw_QNBio <- normalize.quantiles(as.matrix(raw[,2:7]))
boxplot(log(raw[,2:7]))
#boxplot(log(raw_QN))
boxplot(log(raw_QNBio))
```

Plot Aparna Housekeeping genes
```{r fig.height=6, fig.width=9}

housekeeping= c("Cltc","Gapdh","Gusb","Hprt","Pgk1","Tubb5")

ggplot(plot_AD %>% dfilter(Gene == "Gapdh"| Gene == "Cltc" | Gene == "Gusb" | Gene == "Hprt" | Gene == "Pgk1" | Gene == "Tubb5"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y",ncol=2) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_HK_QN.jpg", width= 9, height=6, dpi=600, plot= last_plot(), units = "in")

ggplot( plot_raw %>% dfilter(Gene == "Gapdh"| Gene == "Cltc" | Gene == "Gusb" | Gene == "Hprt" | Gene == "Pgk1" | Gene == "Tubb5"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y", ncol=2) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_AD_HK_raw.jpg", width= 9, height=6, dpi=600, plot= last_plot(), units = "in")

#NEW HOUSEKEEPING GENES

ggplot(plot_AD %>% dfilter(Gene == "Mapkapk2"| Gene == "Mapk14" | Gene == "Map2k4" | Gene == "Map3k7" | Gene == "Atf2" | Gene == "Cfl1"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y", ncol=2) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_AD_QN_New_HK.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")

```
Find Housekeeping gene for Aparna's samples:

```{r}
t_AD <- data.frame(t(raw))
colnames(t_AD) <- rownames(raw)

gene_list <- rownames(raw)

rsq_t_AD <- function (x) cor(t_AD[,x],t_AD[,"Total_countx0.01"]) ^ 2  
rsquared_list <- t(data.frame(lapply(gene_list,rsq_t_AD)))
rownames(rsquared_list) <- rownames(raw)
sort(rsquared_list[,1], decreasing=TRUE)
rsquared_list[,1]
```
Plot Correlations

```{r fig.height=30, fig.width=30}
gene_gather <- t_AD %>% gather(key= Genes, value= "Count", -Total_countx0.01)

#ggplot(gene_gather, aes(x=Binding_index, y= Count)) + 
#  geom_point() +
#  facet_wrap(~Genes,scales="free_y", ncol=12)

ggplot(gene_gather, aes(x=Total_countx0.01, y= Count)) + 
  geom_point() +
  facet_wrap(~Genes,scales="free_y", ncol=12) +
  theme(axis.text.x = element_text(angle = 55, hjust = 1)) +
  ggsave("Nanostring_AD_CountsvsTotalCounts.jpg", width= 30, height=30, dpi=600, plot= last_plot(), units = "in")
```
Calculate R^2 values
```{r}
rsq <- function (x, y) cor(x, y) ^ 2  
rsq(t_AD[,"Gapdh"],t_AD[,"Total_countx0.01"]) #0.75
rsq(t_AD[,"Cltc"],t_AD[,"Total_countx0.01"]) #0.92
rsq(t_AD[,"Gusb"],t_AD[,"Total_countx0.01"]) #0.71
rsq(t_AD[,"Hprt"],t_AD[,"Total_countx0.01"]) #0.92
rsq(t_AD[,"Pgk1"],t_AD[,"Total_countx0.01"]) #0.73
rsq(t_AD[,"Tubb5"],t_AD[,"Total_countx0.01"]) #0.33

rsq(t_AD[,"Atf2"],t_AD[,"Total_countx0.01"]) #0.93
rsq(t_AD[,"Cfl1"],t_AD[,"Total_countx0.01"]) #0.93
rsq(t_AD[,"Map3k7"],t_AD[,"Total_countx0.01"]) #0.93
rsq(t_AD[,"Map2k4"],t_AD[,"Total_countx0.01"]) #0.94
rsq(t_AD[,"Mapkapk2"],t_AD[,"Total_countx0.01"]) #0.97
rsq(t_AD[,"Mapk14"],t_AD[,"Total_countx0.01"]) #0.95

```
Plot Markers

Plot genes of interest:
CCL22
FOXP3
GATA3
IFNB1
IFNG
IL13
IL17a
IL17b
IL17F
IL22
IL33
IL4
IL5
Pdcd1
Tgfb1
Tgfb2
Tgfb3
Tslp

"CCL22","IFNB1","IFNG","IL13","IL17a","IL22","IL4","IL5","Tgfb1","Tgfb2","Tgfb3","Tslp"
NOT in the dataset: "FOXP3" "GATA3" "IL17b" "IL17F" "IL33" "Pdcd1"
DEX_ETOH     DEX_FITC  FLX193_ETOH  FLX193_FITC Vehicle_ETOH Vehicle_FITC

Too low signal
Gene == "Ifnb1" 
Gene == "Ifng" 
Gene == "Il13" 
Gene == "Il17a" 
Gene == "Il22" 
Gene == "Il4" 
Gene == "Il4" 
Gene == "Il5" 
```{r fig.height=6, fig.width=12}
markers_AD <- c("Ccl22","Ifnb1","Ifng","Il13","Il17a","Il22","Il4","Il4","Il5","Tgfb1","Tgfb2","Tgfb3","Tslp")

ggplot(plot_AD %>% dfilter(Gene %in% markers_AD), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y") +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")) #+
ggsave("Nanostring_AD_QN_Markers.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")
```


Buvana samples
Quantile Normalize
```{r}
all <- read.table("/Volumes/Picard/FLX/Nanostring/Buvana_raw_June.txt", sep="\t", header=TRUE)
all <- all %>% select(-2)
boxplot(all[,3:20],log="y")
methods(plot)
#Find the total counts per sample, compare to binding density
col_sums <-colSums(all[,3:20])

length(colnames(all)) #20
BI <- as.matrix(all[1,3:20])
barplot(col_sums)
barplot(BI)
head(all)

boxplot(all[,3:20],log="y")
all_QN <- normalizeQuantiles(all[,3:20], ties=TRUE)
boxplot(all_QN,log="y")
BI_QN <- as.matrix(all_QN[1,])
barplot(BI_QN)
col_sums_QN <-colSums(all_QN[,1:18])
#Add back the gene names
rownames(all_QN) <- all[1:269,1]
write.table(all_QN, file="Nanostring_Buvana_QuantileNorm.txt", sep="\t", quote= FALSE, row.names=TRUE)

plot_QN <- all_QN %>% rownames_to_column %>% gather(key= Sample, value= "QN_Count", -rowname)
plot_raw <- all %>% select(-X..Samples.above.Threshold) %>% gather(key= Sample, value= "Count", -Probe.Name)
colnames(plot_raw) <- c("Gene","Sample","Count")
colnames(plot_QN) <- c("Gene","Sample","QN_Count")
plot_raw$Sample <- factor(plot_raw$Sample, levels=c("CT26Vehicle1A","CT26Vehicle2A","CT26_282_1A",  "CT26_282_2A",  "CT26_659_1A", "CT26_659_2A",  "CT26.TumorCtrl" ,"CT26.OVA.TumorOE","CT26..LMP1.TumorOE" ,"CT26..LMP1.TumorOE.1","X3.3Ctrl", "X6.1Treatment","CT26Vehicle1C","CT26Vehicle2C","CT26_282_1C",  "CT26_282_2C",  "CT26_659_1C",  "CT26_659_2C"))

plot_QN$Sample <- factor(plot_QN$Sample, levels=c("CT26Vehicle1A","CT26Vehicle2A","CT26_282_1A",  "CT26_282_2A",  "CT26_659_1A", "CT26_659_2A",  "CT26.TumorCtrl" ,"CT26.OVA.TumorOE","CT26..LMP1.TumorOE" ,"CT26..LMP1.TumorOE.1","X3.3Ctrl", "X6.1Treatment","CT26Vehicle1C","CT26Vehicle2C","CT26_282_1C",  "CT26_282_2C",  "CT26_659_1C",  "CT26_659_2C"))
levels(plot_raw$Sample) == colnames(all)[3:20]

# Parse New HK-normalized data table
#is below 

```

Plot Housekeeping genes
```{r fig.height=6, fig.width=8}
#Example data gather
#Buvi <- all %>% select(-Class) %>% gather(key= Sample, value= "Count", -Gene) # Both they Key and Value are newly created for the gathered data. Specity the column that doesn't change in the 3rd term with '-'

housekeeping= c("Cltc","Gapdh","Gusb","Hprt","Pgk1","Tubb5")

ggplot(plot_raw %>% dfilter(Gene == "Gapdh"| Gene == "Cltc" | Gene == "Gusb" | Gene == "Hprt" | Gene == "Pgk1" | Gene == "Tubb5"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y",ncol=2) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_CT26_HKraw.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")

ggplot(plot_QN %>% dfilter(Gene == "Gapdh"| Gene == "Cltc" | Gene == "Gusb" | Gene == "Hprt" | Gene == "Pgk1" | Gene == "Tubb5"), aes(x=Sample, y= QN_Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y", ncol=2) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_CT26_QN_HK.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")

#NEW HOUSEKEEPING GENES

ggplot(plot_QN %>% dfilter(Gene == "Daxx"| Gene == "Nfe2l2" | Gene == "Hmgb2" | Gene == "Mapk1" | Gene == "Mapkapk2" | Gene == "Rela"), aes(x=Sample, y= QN_Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y", ncol=2) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_CT26_QN_New_HK.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")


#NEW housekeeping genes after Normalizing to these genes

#ggplot(plot_HKcorr %>% dfilter(Gene == "Daxx"| Gene == "Nfe2l2" | Gene == "Hmgb2" | Gene == "Mapk1" | Gene == "Mapkapk2" | Gene == "Rela"), aes(x=Sample, y= Count,fill=Gene)) +
#  geom_bar(stat = "identity") +
#  facet_wrap(~Gene, scales="free_y", ncol=2) +
#  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
#ggsave("Nanostring_CT26_NormNewHK_HK.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")
```

Normalize to NEW Housekeeping genes
Take the geometric mean of the 6 New Housekeeping genes. Find the Mean of the means, and generate a correction factor for each sample and multiply across columns
To multiply a vector across a matric by rows or by columns, use sweep: sweep(matrix, MARGIN=2, vector, `*` )  MARGIN refers to 1= ROWs and 2 = COLUMNS, last term `*` says "perform multiplication"
```{r}
dim(all)
geomMean<-function(values){
  prod(values)^(1/length(values))
}
#test
values<-c(2,4,6,8)
geomMean(values)

HKs <- all %>% dfilter(Probe.Name == "Daxx"| Probe.Name == "Nfe2l2" | Probe.Name == "Hmgb2" | Probe.Name == "Mapk1" | Probe.Name == "Mapkapk2" | Probe.Name == "Rela") %>% dselect(-1,-2) %>% summarise_all(geomMean) 
rowMeans(HKs) # 538.4613
#Get the correction factors for each sample as a vector of 18 elements
correction_factor <- as.vector(t(538/HKs)[,1])
str(correction_factor)
#Apply correction factor across all
all_HKcorr <- sweep(as.matrix(all[,3:20]),MARGIN=2, correction_factor,`*`)
rownames(all_HKcorr) <- all$Probe.Name
all_HKcorr <- data.frame(all_HKcorr)
all_HKcorr %<>% tibble::rownames_to_column()
colnames(all_HKcorr)[1] <- "Gene"


plot_HKcorr <- all_HKcorr %>% gather(key= Sample, value= "Count", -Gene)
plot_HKcorr$Sample <- factor(plot_HKcorr$Sample, levels=c("CT26Vehicle1A","CT26Vehicle2A","CT26_282_1A",  "CT26_282_2A",  "CT26_659_1A", "CT26_659_2A",  "CT26.TumorCtrl" ,"CT26.OVA.TumorOE","CT26..LMP1.TumorOE" ,"CT26..LMP1.TumorOE.1","X3.3Ctrl", "X6.1Treatment","CT26Vehicle1C","CT26Vehicle2C","CT26_282_1C",  "CT26_282_2C",  "CT26_659_1C",  "CT26_659_2C"))
```

Markers Buvana

```{r fig.height=10, fig.width=20}

#genes_of_interest = c("Ccl22","Tgfb1","Tgfb2","Tgfb3","Il1b","Ifng","Ddit3")  
ggplot(plot_QN %>% dfilter(Gene == "Csf1" | Gene == "Il1b"| Gene == "Ccl22"|  Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Il1b" | Gene == "Ifng" | Gene == "Ddit3"| Gene == "Arg1" | Gene == "Cd163"| Gene == "Cd4"| Gene == "Daxx"| Gene == "Il10" | Gene == "Il2"| Gene == "Nos2"| Gene == "Ccl8" | Gene == "Il6" | Gene == "Mrc1"), aes(x=Sample, y= QN_Count,fill=Gene)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y", ncol=6) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_Markers_QN.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")

plot_QN %>% dfilter (Gene == "Ccl8")

ggplot(plot_raw %>% dfilter(Gene == "Csf1" | Gene == "Il1b"| Gene == "Ccl22"|  Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Il1b" | Gene == "Ifng" | Gene == "Ddit3"| Gene == "Arg1" | Gene == "Cd163"| Gene == "Cd4"| Gene == "Daxx"| Gene == "Il10" | Gene == "Il2"| Gene == "Nos2"| Gene == "Ccl8" | Gene == "Il6" | Gene == "Mrc1"), aes(x=Sample, y= Count,fill=Gene)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y", ncol=6) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_Markers_raw.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")



ggplot(plot_HKcorr %>% dfilter(Gene == "Csf1" | Gene == "Il1b"| Gene == "Ccl22"|  Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Il1b" | Gene == "Ifng" | Gene == "Ddit3"| Gene == "Arg1" | Gene == "Cd163"| Gene == "Cd4"| Gene == "Daxx"| Gene == "Il10" | Gene == "Il2"| Gene == "Nos2"| Gene == "Ccl8" | Gene == "Il6" | Gene == "Mrc1"), aes(x=Sample, y= Count,fill=Gene)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y", ncol=6) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_Markers_HKcorr.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")
```

```{r fig.height=10, fig.width=20}
plot_batch3 <- plot_QN %>% dfilter(Sample == "CT26Vehicle1C"| Sample == "CT26Vehicle2C" | Sample == "CT26_282_1C" | Sample == "CT26_282_2C" | Sample == "CT26_659_1C" | Sample == "CT26_659_2C")
plot_batch3$Sample <- factor(plot_batch3$Sample, levels = c("CT26Vehicle1C","CT26Vehicle2C","CT26_282_1C",  "CT26_282_2C",  "CT26_659_1C",  "CT26_659_2C"))

plot_batch3_HKcorr <- plot_HKcorr %>% dfilter(Sample == "CT26Vehicle1C"| Sample == "CT26Vehicle2C" | Sample == "CT26_282_1C" | Sample == "CT26_282_2C" | Sample == "CT26_659_1C" | Sample == "CT26_659_2C")
plot_batch3$Sample <- factor(plot_batch3$Sample, levels = c("CT26Vehicle1C","CT26Vehicle2C","CT26_282_1C",  "CT26_282_2C",  "CT26_659_1C",  "CT26_659_2C"))

ggplot(plot_batch3 %>% dfilter(Gene == "Csf1" | Gene == "Il1b"| Gene == "Ccl22"|  Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Il1b" | Gene == "Ifng" | Gene == "Ddit3"| Gene == "Arg1" | Gene == "Cd163"| Gene == "Tnf"| Gene == "Daxx"| Gene == "Il10" | Gene == "Il2"| Gene == "Nos2"| Gene == "Ccl8" | Gene == "Il6" | Gene == "Mrc1"), aes(x=Sample, y= QN_Count,fill=Gene)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y", ncol=6) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_Markers_QN_batch3.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")

ggplot(plot_batch3_HKcorr %>% dfilter(Gene == "Csf1" | Gene == "Il1b"| Gene == "Ccl22"|  Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Il1b" | Gene == "Ifng" | Gene == "Ddit3"| Gene == "Arg1" | Gene == "Cd163"| Gene == "Tnf"| Gene == "Daxx"| Gene == "Il10" | Gene == "Il2"| Gene == "Nos2"| Gene == "Ccl8" | Gene == "Il6" | Gene == "Mrc1"), aes(x=Sample, y= Count,fill=Gene)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y", ncol=6) +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_Markers_HKcorr_batch3.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")


```
Housekeeping gene and Total RNA correlations

("Cltc","Gapdh","Gusb","Hprt","Pgk1","Tubb5")
#```{r}
#plot(all_QN["Gapdh",],all_QN["Binding_index",])
#ggplot(all_QN, aes(x="Gapdh", y="Binding_Index")) +
 # geom_point()

t_all <- data.frame(t(all[,3:20]))
colnames(t_all) <- all[1:269,1]
t_all %<>% tibble::rownames_to_column()
rownames_save <- rownames(t_all)
t_all$rowname <- factor(t_all$rowname, levels=c("CT26Vehicle1A","CT26Vehicle2A","CT26_282_1A",  "CT26_282_2A",  "CT26_659_1A", "CT26_659_2A",  "CT26.TumorCtrl" ,"CT26.OVA.TumorOE","CT26..LMP1.TumorOE" ,"CT26..LMP1.TumorOE.1","X3.3Ctrl", "X6.1Treatment","CT26Vehicle1C","CT26Vehicle2C","CT26_282_1C",  "CT26_282_2C",  "CT26_659_1C",  "CT26_659_2C"))
ggplot(t_all, aes(x=Gapdh, y= Binding_index)) + 
  geom_point(aes(color=rowname)) +
  geom_abline(slope = 1.1/38000, intercept=0.1)


rsq <- function (x, y) cor(x, y) ^ 2  
rsq(t_all[,"Gapdh"],t_all[,"Binding_index"]) #0.95
rsq(t_all[,"Cltc"],t_all[,"Binding_index"]) #0.94
rsq(t_all[,"Gusb"],t_all[,"Binding_index"]) #0.94
rsq(t_all[,"Hprt"],t_all[,"Binding_index"]) #0.96
rsq(t_all[,"Pgk1"],t_all[,"Binding_index"]) #0.91
rsq(t_all[,"Tubb5"],t_all[,"Binding_index"]) #0.64

rsq(t_all[,"Mafk"],t_all[,"Binding_index"]) #0.92
rsq(t_all[,"Cebpb"],t_all[,"Binding_index"]) #0.92
rsq(t_all[,"Mapkapk2"],t_all[,"Binding_index"]) #0.97
rsq(t_all[,"Daxx"],t_all[,"Binding_index"]) #0.95
rsq(t_all[,"Mapk1"],t_all[,"Binding_index"]) #0.97
rsq(t_all[,"Hmgb2"],t_all[,"Binding_index"]) #0.96
rsq(t_all[,"Rela"],t_all[,"Binding_index"]) #0.95


rsq(t_all[,"Gapdh"],t_all[,"total_count"]) #0.97
rsq(t_all[,"Cltc"],t_all[,"total_count"]) #0.95
rsq(t_all[,"Gusb"],t_all[,"total_count"]) #0.96
rsq(t_all[,"Hprt"],t_all[,"total_count"]) #0.97
rsq(t_all[,"Pgk1"],t_all[,"total_count"]) #0.92
rsq(t_all[,"Tubb5"],t_all[,"total_count"]) #0.64

rsq(t_all[,"Mafk"],t_all[,"total_count"]) #0.96
rsq(t_all[,"Cebpb"],t_all[,"total_count"]) #0.97
rsq(t_all[,"Mapkapk2"],t_all[,"total_count"]) #0.98
rsq(t_all[,"Daxx"],t_all[,"total_count"]) #0.99
rsq(t_all[,"Mapk1"],t_all[,"total_count"]) #0.98
rsq(t_all[,"Hmgb2"],t_all[,"total_count"]) #0.98
rsq(t_all[,"Rela"],t_all[,"total_count"]) #0.98

gene_list <- list(colnames(t_all)[2:255])

rsq_t_all <- function (x) cor(t_all[,x],t_all[,"total_count"]) ^ 2  
rsquared_list <- lapply(gene_list,rsq_t_all)
rsquared_list
#```

Make a scatter plot of Binding_density versus all gene signals. I want to see whether it's true that at low Counts, Binding_density does not correlate well with Count:
i.e. quantile normalization is not possible at low Count. 

It's impossible to tell because a non-linear relationship can be due to:
1) sampling noise or 
2) actual biology

An actual good use for this plot is for finding appropriate HOUSEKEEPING GENES

#```{r fig.height=30, fig.width=30}
BI <- t_all[,"Binding_index"]
#t_all %<>% mutate(total_count= rowSums(.)) %>% mutate(Sample=rownames_save)

gene_gather <- t_all %>% gather(key= Genes, value= "Count", -Binding_index, -total_count, -Sample)

ggplot(gene_gather, aes(x=Binding_index, y= Count)) + 
  geom_point() +
  facet_wrap(~Genes,scales="free_y", ncol=12)

ggplot(gene_gather, aes(x=total_count, y= Count)) + 
  geom_point() +
  facet_wrap(~Genes,scales="free_y", ncol=12) +
  theme(axis.text.x = element_text(angle = 55, hjust = 1)) +
  ggsave("Nanostring_CountsvsTotalCounts.jpg", width= 30, height=30, dpi=600, plot= last_plot(), units = "in")
#```

Plot genes of interest from raw data
CCL22
FOXP3
GATA3
IFNB1
IFNG
IL13
IL17a
IL17b
IL17F
IL22
IL33
IL4
IL5
Pdcd1
Tgfb1
Tgfb2
Tgfb3
Tslp

"CCL22","IFNB1","IFNG","IL13","IL17a","IL22","IL4","IL5","Tgfb1","Tgfb2","Tgfb3","Tslp"
NOT in the dataset: "FOXP3" "GATA3" "IL17b" "IL17F" "IL33" "Pdcd1"
DEX_ETOH     DEX_FITC  FLX193_ETOH  FLX193_FITC Vehicle_ETOH Vehicle_FITC

Too low signal
Gene == "Ifnb1" 
Gene == "Ifng" 
Gene == "Il13" 
Gene == "Il17a" 
Gene == "Il22" 
Gene == "Il4" 
Gene == "Il4" 
Gene == "Il5" 
#```{r fig.height=8, fig.width=12}
#table(AD$Sample)

ggplot(AD_QN %>% dfilter(Gene == "Ccl22"|  Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Tslp"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y") +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,5), "cm")) 
ggsave("Nanostring_AD_ALLHK_All_robust.jpg", width= 11, height=6, dpi=600, plot= last_plot(), units = "in")

FITC_QN <- AD_QN %>% dfilter(Sample =="DEX_FITC" | Sample=="FLX193_FITC" | Sample=="Vehicle_FITC")

FITC_QN$Sample <- factor(FITC_QN$Sample, levels=c("Vehicle_FITC","DEX_FITC","FLX193_FITC"))
ggplot(FITC_QN %>% dfilter(Gene == "Ccl22"| Gene == "Ifnb1" | Gene == "Ifng" | Gene == "Il13" | Gene == "Il17a" | Gene == "Il22"| Gene == "Il4" | Gene == "Il4" | Gene == "Il5" | Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Tslp"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y") +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,5), "cm")) +
ggsave("Nanostring_AD_QN.jpg", width= 11, height=6, dpi=600, plot= last_plot(), units = "in")

FITC_QN$Sample <- factor(FITC_QN$Sample, levels=c("Vehicle_FITC","DEX_FITC","FLX193_FITC"))
ggplot(FITC_QN %>% dfilter(Gene == "Ccl22"|  Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Tslp"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y") +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,5), "cm")) +
ggsave("Nanostring_AD_QN_robust.jpg", width= 11, height=6, dpi=600, plot= last_plot(), units = "in")

FITC <- AD %>% dfilter(Sample =="DEX_FITC" | Sample=="FLX193_FITC" | Sample=="Vehicle_FITC")
FITC$Sample <- factor(FITC$Sample, levels=c("Vehicle_FITC","DEX_FITC","FLX193_FITC"))
ggplot(FITC %>% dfilter(Gene == "Ccl22"| Gene == "Ifnb1" | Gene == "Ifng" | Gene == "Il13" | Gene == "Il17a" | Gene == "Il22"| Gene == "Il4" | Gene == "Il4" | Gene == "Il5" | Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Tslp"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y") +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,5), "cm")) +
ggsave("Nanostring_AD_ALLHK_All_markers.jpg", width= 11, height=6, dpi=600, plot= last_plot(), units = "in")

ggplot(FITC %>% dfilter(Gene == "Ccl22"|  Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Tslp"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y") +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_AD_CltcGapdhGusb_robust_FITC.jpg", width= 6, height=6, dpi=600, plot= last_plot(), units = "in")
AD$Sample <- factor(AD$Sample, levels=c("Vehicle_ETOH","Vehicle_FITC","DEX_ETOH","DEX_FITC","FLX193_ETOH","FLX193_FITC"))
ggplot(AD %>% dfilter(Gene == "Ccl22"|  Gene == "Tgfb1" | Gene == "Tgfb2" | Gene == "Tgfb3" | Gene == "Tslp"), aes(x=Sample, y= Count,fill=Gene)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Gene, scales="free_y") +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.2), axis.text.x = element_text(angle = 55, hjust = 1),legend.position = "none",plot.margin = unit(c(1,1,1,1), "cm")) +
ggsave("Nanostring_AD_CltcGapdhGusb_robust_ALL.jpg", width= 6, height=6, dpi=600, plot= last_plot(), units = "in")

#```

