---
title: "Parse_human_Kallisto"
author: "MX"
date: "2/2/2021"
output: html_document
---

```{r}
library(limma)
library(tidyverse)
library(dplyr)
library(magrittr)
library(tibble)
library(ggplot2)
library(DESeq2)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
#source("D:/FLX/Files_from_Gene/R_functions.r")  # code for PC
sum_mx <- function(x) {sum(x,na.rm=TRUE)}
mean_mx <- function(x) {mean(x,na.rm=TRUE)}
```


Read in the data
```{r}
file_names <- list.files("/Volumes/Picard/FLX/HPK1/Bellshaped_curve_experiments/Kallisto/", pattern=".tsv") %>% lapply(function(i) {strsplit(i,"_abundance")[[1]][1]})
files <- base::list.files("/Volumes/Picard/FLX/HPK1/Bellshaped_curve_experiments/Kallisto/", pattern=".tsv", full.names=TRUE)

data_files <- lapply(files, function(i){read.table(i,header=TRUE)})
offtargets <- cbind.data.frame(lapply(data_files, function(j) {j %>% dselect(tpm)}))
rownames(offtargets) <- data_files[[1]]$target_id
colnames(offtargets) <- file_names

```


Normalize samples to each other
```{r}
library(limma)
boxplot(offtargets %>% as.matrix)
offtargets_norm <- normalizeBetweenArrays(offtargets %>% as.matrix)
boxplot(offtargets_norm)
```

Get the translation table for ENSEMBL to HUGO names
```{r}
require("biomaRt")
listMarts() # find the name for the ENSEMBL database
mart <- useMart("ENSEMBL_MART_ENSEMBL")
listDatasets(mart) #find the species
mart <- useDataset("hsapiens_gene_ensembl", mart)
listAttributes(mart) # find the data slot that you want to retrieve

annotLookup <- getBM(
  mart = mart,
  attributes = c(
    "external_gene_name","affy_hta_2_0"),values = c('TC01003833.hg', '2824546_st', 'TC01000020.hg'))
colnames(annotLookup) <- c("ID_REF","Gene") # change the column names
save(annotLookup,file="Human_ENSEMBL_transcript_2_HUGO.rdata")
load("Human_ENSEMBL_transcript_2_HUGO.rdata") # code to load it again later

annotLookup[10000:10010,]
```

Translate ENSEML transcript names to HUGO gene names, then take the rows that are the same gene and sum them together

The gene names are in this format:
ENST00000473358.1|ENSG00000243485.5|OTTHUMG00000000959.2|OTTHUMT00000002840.1|MIR1302-2HG-202|MIR1302-2HG|712|lincRNA| 
Have to parse out the first gene name and get rid of the rest of it using separate()
```{r}

offtargets_gene <- offtargets_norm %>% data.frame %>% tibble::rownames_to_column("Gencode") %>% separate(Gencode, into=c(Gencode,NA),sep="|", extra="drop") %>% left_join(annotLookup, by=c("Gencode"="ID_REF")) %>% dselect(Gene,everything()) %>% dselect(-Gencode) %>% dfilter(!is.na(Gene)) %>% group_by(Gene) %>% mutate_if(is.numeric,sum_mx) %>% unique %>% ungroup 

```

