---
title: "Limma regDC DE"
author: "Mengshu"
date: "February 21, 2019"
output: html_document
---
Differential timecourse analysis of microarray data using Limma
```{r}
#BiocManager::install("Biobase")
#BiocManager::install("GEOquery")
library(GEOquery)
library(Biobase)
library(limma)
library(dplyr)
```

Download datasets
```{r}


gse <- getGEO("GSE5151", destdir=".",GSEMatrix=TRUE)
gse <- getGEO("GSE59294", destdir=".",GSEMatrix=TRUE) #Guttman 2014 AD data
#garvan <- getGEO("GSE3982", destdir=".",GSEMatrix=TRUE)
show(gse)
eset <- gse[[1]]
e <- exprs(eset)
boxplot(log(e))
#see the experiments
show(pData(phenoData(eset)))
rownames(pData(phenoData(eset)))
```

```{r}
e_Df <- data.frame(e)
AD_2014 <- e_Df %>% dplyr::select("GSM1432944","GSM1432949", "GSM1432951","GSM1432954","GSM1432958", "GSM1432960" , "GSM1432961", "GSM1432962","GSM1432967", "GSM1432969", "GSM1432970" ,"GSM1432974","GSM1432978","GSM1432979", "GSM1432981","GSM1432945", "GSM1432948", "GSM1432950","GSM1432953","GSM1432955", "GSM1432959","GSM1432963", "GSM1432966", "GSM1432971","GSM1432975","GSM1432980","GSM1432982")

(e["221333_at",])
(AD_2014["221165_s_at",])
(AD_2014["221333_at",])

```



Set up design and the groups to compare
```{r}
design <- cbind(GSM116323=c(1,0,0,0),
                GSM116324=c(0,1,0,0),
                GSM116325=c(0,0,1,0),
                GSM116326=c(0,0,0,1)
              )
lev <- c("GSM116323", "GSM116324", "GSM116325", "GSM116326")
f <- factor(lev, levels=lev)
design2 <- model.matrix(~0+ f)
              )
design
design2
#what to compare
c <- makeContrasts(GSM116324-GSM116323,GSM116325-GSM116323,GSM116326-GSM116323, levels=design)
c
```
Fit Linear Model: Can't do this without replicates, so skip the lmFit step
```{r}
 fit <- lmFit(eset,design) 
 fit2<- contrasts.fit(fit,c) 
 lm.Fit(c,)
 fit2<- eBayes(fit2)
 toptable(fit2)
```

Forget trying to analyze this, just export and cluster the data. 
Limma does not take kindly to no replicates
```{r}
e <- tibble::rownames_to_column(data.frame(e))
e %<>% mutate(gene=rowname) %>% dselect(rowname, gene, everything())
gene_code <- e %>% dselect(gene)
write.table(e, file="DC_TGFB_Ju.txt", sep="\t", quote=FALSE, row.names=FALSE)
```

Look up the gene names
Array is HG_U95Av2
Problem is the tool returns
length(HUGO[,1]) 22283
length(OUT)    22300
length(unique(OUT$SYMBOL)) 
```{r}
source("http://bioconductor.org/biocLite.R")
biocLite("hgu95av2.db")
BiocManager::install("hgu95av2.db")
library("AnnotationDbi")
library("hgu95av2.db")    ##for Human
library("hgu133a2.db") 
columns(hgu95av2.db)
PROBES<- e[,1]
OUT <- select(hgu95av2.db,keys= PROBES, columns=c("SYMBOL","GENENAME"),multiVals="first")
OUT[1:100,]
HUGO[1:100,]
HUGO <- OUT[!duplicated(OUT[,1]),]
length(HUGO[,1])
e %<>% mutate(gene_name=HUGO[,2])
e %<>% dselect(-gene,gene_name,rowname,everything())
e %>% dfilter(gene_name=="MET")
select
#Back lookup the array name
mapIds(hgu95av2.db,keys="FOXP3", column="PROBEID", keytype='SYMBOL')
mapIds(hgu133a2.db,keys="IL22", column="PROBEID", keytype='SYMBOL')
select(hgu133a2.db,keys= c("221333_at"), columns=c("SYMBOL","GENENAME","ENSEMBL","ENSEMBLPROT"))
```
Normalize values

```{r}
fold_change_log <- e %>% mutate("4hour"=log2(GSM116324/GSM116323), "16hour" =log2(GSM116325/GSM116323), "36hour"=log2(GSM116326/GSM116323)) %>% dselect(-GSM116323,-GSM116324,-GSM116325,-GSM116326)
fold_change <- e %>% mutate("4hour"=(GSM116324/GSM116323), "16hour" =(GSM116325/GSM116323), "36hour"=(GSM116326/GSM116323)) %>% dselect(-GSM116323,-GSM116324,-GSM116325,-GSM116326)
write.table(fold_change, file="DC_TGFB_Ju_FC.txt", sep="\t", quote=FALSE, row.names=FALSE)

```

