---
title: "Limma regDC DE"
author: "Mengshu"
date: "February 21, 2019"
output: html_document
---
Differential timecourse analysis of microarray data using Limma
```{r}
BiocManager::install("Biobase")
BiocManager::install("GEOquery")
library(GEOquery)
library(Biobase)
library(limma)
library(dplyr)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
library(magrittr)
```

Download datasets
```{r}

gse <- getGEO("GSE5151", destdir=".",GSEMatrix=TRUE)
gse <- getGEO("GSE9241", destdir=".",GSEMatrix=TRUE)
gse <- getGEO("GSE18921", destdir=".",GSEMatrix=TRUE) # competitive Hybe
genes <- read.table("GSE18921_genes.txt", sep="\t", header=TRUE, na.strings="NA")

show(gse)
eset <- gse[[1]]
e <- exprs(eset)
dim(e)
boxplot(log(e))
#see the experiments
show(pData(phenoData(eset)))
```
Skip these steps to set up design and Linear model
Set up design and the groups to compare
```{r}
design <- cbind(GSM116323=c(1,0,0,0),
                GSM116324=c(0,1,0,0),
                GSM116325=c(0,0,1,0),
                GSM116326=c(0,0,0,1)
              )
lev <- c("GSM116323", "GSM116324", "GSM116325", "GSM116326")
f <- factor(lev, levels=lev)
design2 <- model.matrix(~0+ f)
              )
design
design2
#what to compare
c <- makeContrasts(GSM116324-GSM116323,GSM116325-GSM116323,GSM116326-GSM116323, levels=design)
c
```
Fit Linear Model: Can't do this without replicates, so skip the lmFit step
```{r}
 fit <- lmFit(eset,design) 
 fit2<- contrasts.fit(fit,c) 
 lm.Fit(c,)
 fit2<- eBayes(fit2)
 toptable(fit2)
```

Forget trying to analyze this, just export and cluster the data. 
Limma does not take kindly to no replicates
```{r}
e <- data.frame(e)
e %<>% mutate(rowname=c(1:19968))
gene<- data.frame(genes[,c(1,8)])
gene %<>% mutate(rowname=as.numeric(ID))
exp <- left_join(gene,e, by="rowname")
exp %<>% select(-ID,-rowname) %>% arrange(GSM468774)
exp %<>% arrange(-GSM468774)
#e %<>% mutate(gene=rowname) %>% dselect(rowname, gene, everything())
#gene_code <- e$rowname
write.table(exp, file="DC_Torres.txt", sep="\t", quote=FALSE, row.names=FALSE)
```

Look up the gene names
Array is HG_U95Av2
Problem is the tool returns
length(HUGO[,1]) 22283
length(OUT)    22300
length(unique(OUT$SYMBOL)) 
```{r}
source("http://bioconductor.org/biocLite.R")
biocLite("hgu95av2.db")
BiocManager::install("hgu95av2.db")
library("AnnotationDbi")
library("hgu95av2.db")    ##for Human
library("hgu133a2.db") 
columns(hgu95av2.db)
showMethods("keys")
PROBES<- e[,1]
#select(hgu95av2.db,keys= PROBES,columns=c("SYMBOL","GENENAME","ENSEMBL","ENSEMBLPROT"))
OUT <- AnnotationDbi::select(hgu95av2.db, keys=PROBES, columns=c("SYMBOL"),multiVals="first", keytype = "PROBEID")
??AnnotationDbi::select
OUT[1:100,]
HUGO[1:100,]
HUGO <- OUT[!duplicated(OUT[,1]),]
length(HUGO[,1])
e <- e %>% mutate(gene_name=HUGO[,2])
e <- e %>% dplyr::select(gene_name,rowname,everything())
#e %>% dfilter(gene_name=="MET")

#Back lookup the array name
mapIds(hgu95av2.db,keys="CD4", column="PROBEID", keytype='SYMBOL')
select(hgu95av2.db,keys= c("1608_at","1609_g_at","1812_s_at","35684_at"), columns=c("SYMBOL","GENENAME","ENSEMBL","ENSEMBLPROT"))
```
Normalize values

```{r}
#fold_change_log <- e %>% mutate("4hour"=log2(GSM116324/GSM116323), "16hour" =log2(GSM116325/GSM116323), "36hour"=log2(GSM116326/GSM116323)) %>% dselect(-GSM116323,-GSM116324,-GSM116325,-GSM116326)
colnames(e)
fold_change <- e %>% mutate("1hour" =(GSM234710/GSM234706), "3hour"=(GSM234712/GSM234706),"6hour"=(GSM234713/GSM234706),"12hour"=(GSM234716/GSM234706),"18hour"=(GSM234717/GSM234706),"36hour"=(GSM234718/GSM234706),"36h_Ctrl"=(GSM234708/GSM234706),"B1hour"=(GSM234719/GSM234706),"B6hour"=(GSM234720/GSM234706),"B12hour"=(GSM234722/GSM234706),"B12hour"=(GSM234723/GSM234706),"B18hour"=(GSM234757/GSM234706),"B36hour"=(GSM234758/GSM234706),"AB6hour"=(GSM234760/GSM234706),"AB12hour"=(GSM234761/GSM234706),"AB36hour"=(GSM234762/GSM234706)) %>% dplyr::select(-GSM234706,-GSM234708,-GSM234710,-GSM234712,-GSM234713,-GSM234716,-GSM234717,-GSM234718,-GSM234719,-GSM234720,-GSM234722,-GSM234723,-GSM234757, -GSM234758,-GSM234760 ,-GSM234761,-GSM234762)
write.table(fold_change, file="regDC_Jiang_FC_All.txt", sep="\t", quote=FALSE, row.names=FALSE)

```

Transpose table, retain the probe names because some of them are redundantly assigned to genes, and Seurat does not allow redundant cellids
```{r}
#fold_change[1:5,1:5]
colnames(fold_change)
genes <- fold_change[,1]
probes <- fold_change[,2]
seurat_in <- fold_change %>% dselect(-rowname,-gene_name)
seurat_in <- t(seurat_in)
#seurat_in[1:5,1:5]
colnames(seurat_in) <- probes
```

UMAP analysis on the data:
Pretend that the genes are "cells", and the timepoints are "genes", and plot UMAP based on the expression pattern across timepoints


Create Seurat object
For UMAP, I don't need to scale the data, but UMAP runs of the @scale.data slot, so I put raw.data into that slot
```{r}
jiang <- CreateSeuratObject(raw.data=seurat_in)
jiang <- FindClusters(object=jiang, genes.use = c(rownames(seurat_in)), resolution=1)
jiang@scale.data <- jiang@raw.data 
jiang <- RunUMAP(jiang, 
  genes.use = c(rownames(seurat_in)), max.dim = 2L,
  reduction.name = "umap", reduction.key = "UMAP", n_neighbors = 10L,
  min_dist = 0.2, metric = "manhattan", seed.use = 42)

DimPlot(object = jiang, do.label = TRUE, reduction.use = 'umap', pt.size = 0.8, label.size = 4) 
ggsave("Azizi_PBMCs_UMAP.jpg", width= 12, height=8, dpi=600, plot= last_plot(), units = "in")
```

Seurat is doing something wonky with the raw.data again
```{r fig.height=30, fig.width=30}
jiang@imputed <- data.frame(seurat_in)
FeaturePlot(object=jiang, features.plot = c(rownames(seurat_in)),max.cutoff = 3, cols.use=c("grey","blue"), no.legend=FALSE,reduction.use="umap") 
VlnPlot(jiang, features.plot = c(rownames(seurat_in)), use.raw=TRUE)
```
Export Clusters

```{r}
write.table(jiang@ident, file="Seurat_Clusters_Jiang.txt", sep="\t", quote=FALSE)
```


UMAP directly

UMAP takes objects in rows, features in columns. 
```{r}
install.packages("umap")
library(umap)
??umap
dc.data <- fold_change[,3:17]
dc.label <- fold_change[,2]
dc.gene <- fold_change[,1]
jiang_umap <- umap(dc.data)
jiang_umap
#this contains the coordinates
head(jiang_umap$layout)
ggplot(data.frame(jiang_umap$layout), aes()) +
  geom_point()
```

Insert a model expression gene, and correlate the rest with it
The genes have to be the columns
```{r}
library(Hmisc)
install.packages("corrplot")
library(corrplot)
head(fold_change)
tail(fold_change)

ModelGene <- c(2,2.2,2,3,4,5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5)
cd276_corr <- rcorr(DC_counts,type="pearson")
Model_Gene_corr <- rcorr(fold_change[],type=c("pearson"))

new_col <- as.matrix(c(2,2.2,2,3,4,5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5))
colnames(new_col) <- "ModelGene"
seurat_in[,1:5]
input_1 <- as.matrix(seurat_in)
ModelGene_corr <- rcorr(input_1,new_col,type=c("pearson"))
ModelGene_corr_Spear <- rcorr(input_1,new_col,type=c("spearman"))

write.table(sort(ModelGene_corr$r[,"ModelGene"],decreasing=TRUE) , file="Jiang_Pearson_correlations_ModelGene.txt", sep="\t", quote=FALSE)
write.table(sort(ModelGene_corr_Spear$r[,"ModelGene"],decreasing=TRUE), file="Jiang_Spearman_correlations_ModelGene.txt", sep="\t", quote=FALSE)
sort(ModelGene_corr$r[,"ModelGene"],decreasing=TRUE)
names(sort(ModelGene_corr$r[,"ModelGene"],decreasing=TRUE))
```

Reorder fold_change tables
```{r}
pearson_order <- names(sort(ModelGene_corr$r[,"ModelGene"],decreasing=TRUE))[2:length(ModelGene_corr$r)]
spearman_order <- names(sort(ModelGene_corr_Spear$r[,"ModelGene"],decreasing=TRUE))[2:length(ModelGene_corr_Spear$r)]

fold_change[1:5,1:5]
fold_change <- rbind()
pearson_matrix <- fold_change %>% arrange(factor(rowname, levels=pearson_order))
spearman_matrix <- fold_change %>% arrange(factor(rowname, levels=spearman_order))
write.table(pearson_matrix , file="Jiang_ModelGene_Pearson_matrix.txt", sep="\t", quote=FALSE, row.names=FALSE)
write.table(spearman_matrix, file="Jiang_ModelGene_Spearman_matrix.txt", sep="\t", quote=FALSE, row.names=FALSE)
```

Replace names
```{r}
biocLite("hgu95av2.db")
BiocManager::install("hgu95av2.db")
library("AnnotationDbi")
library("hgu95av2.db")    ##for Human

hi <-select(hgu95av2.db,keys= "34959_at", columns=c("SYMBOL","GENENAME"),multiVals="first")
OUT <- select(hgu95av2.db,keys= PROBES, columns=c("SYMBOL","GENENAME"),multiVals="first")
OUT[1:100,]
HUGO[1:100,]
HUGO <- OUT[!duplicated(OUT[,1]),]

corr_pear <- read.csv("Jiang_Pearson_correlations_ModelGene.txt", sep="\t")
corr_spear <- read.csv("Jiang_Spearman_correlations_ModelGene.txt", sep="\t")
length(rownames(corr_pear))
corr_pear <- tibble::rownames_to_column(data.frame(corr_pear))
PROBES <- corr_pear[,1]
length(PROBES)
PROBES <- PROBES[2:12626]
substring(2,lenght(PROBES))
sub(PROBES,"X","")
```