---
title: "Seurat with Fit-SNE perplexity annealing"
author: "Mengshu"
date: "11/2/2018"
output: html_document
---
Main steps of snRNA analysis:
0. Library normalization (if combining datasets)
1. Feature selection for variable genes (try Gene's method)
2. Non-linear transformation
3. PCA analysis and selection
4. Clustering using _______ method
5. t-SNE clustering for visualization

Here we combine 4. and 5. and use t-SNE for both clustering and visualization
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(Seurat)
library(ggplot2)
library(magrittr)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")
source("bin/fast_tsne.R") # wrapper code to adjust "perplexity" during t-SNE clustering
```

## This is a trial run using a Seurat pipeline, implementing the t-SNE clustering step with FIt-SNE from Berens, which uses the "perplexity" variable to optimize global and close

Load data, filter:
```{r}
pbmc_8k <- Read10X_h5("pbmc8k_raw_gene_bc_matrices_h5.h5", ensg.names= FALSE)
pbmc_FitSNE <- CreateSeuratObject(pbmc_8k, project="Seurat_pbmc", min.cells=3, min.genes=200, is.expr = 0)
#gathers the values for all MT rows
mito.genes <- grep(pattern="^MT-", x= rownames(x=pbmc_FitSNE@data), value=TRUE)
 #calculate percent of genes that are MITO for each cell. Summing total expression for each cell, sumMITO/sumALL
percent.mito <- Matrix::colSums(pbmc_FitSNE@raw.data[mito.genes,])/Matrix::colSums(pbmc_FitSNE@raw.data)
#add this information into pbmc under metadata
pbmc_FitSNE <- AddMetaData(object=pbmc_FitSNE, metadata = percent.mito, col.name= "percent.mito")

pbmc_FitSNE <- FilterCells(object=pbmc_FitSNE, subset.names=c("nGene", "percent.mito"), low.thresholds =c(500,-Inf), high.thresholds=c(Inf,0.1))
#LogNormalize: normalize each cell by total expression, multiply by scale factor 10000, and log2 transform
pbmc_FitSNE <- NormalizeData(object=pbmc_FitSNE, normalization.method="LogNormalize", scale.factor=10000)
```

Find variable Genes: Use Seurat default

```{r}
pbmc_FitSNE <- FindVariableGenes(pbmc_FitSNE, mean.function = ExpMean,
  dispersion.function = LogVMR, do.plot = TRUE, set.var.genes = TRUE,
  x.low.cutoff = 0.0125, x.high.cutoff = 3.5, y.cutoff = 0.5,
  y.high.cutoff = Inf, num.bin = 20, binning.method = "equal_width",
  selection.method = "mean.var.plot", top.genes = 1000, do.recalc = TRUE,
  sort.results = TRUE, do.cpp = TRUE, display.progress = TRUE)

length(x=pbmc_FitSNE@var.genes) #1554
```
Remove uninteresting features by regression
```{r}
pbmc_FitSNE <- ScaleData(object=pbmc_FitSNE, vars.to.regress=c("nUMI","percent.mito"), display.progress = TRUE, assay.type = "RNA", check.for.norm = TRUE)
```

Run PCA: Run with 29 PC groups

```{r}
pbmc_FitSNE <- RunPCA(object = pbmc_FitSNE, pc.genes = pbmc_FitSNE@var.genes, do.print = TRUE, pcs.print = 1:10, 
    genes.print = 5, pcs.compute = 29)

# this saves in pbmc_FitSNE@dr$pca
#however, there's no viewable data in View(pbmc_FitSNE@dr$pca)
```
Find Clusters with 29 PC groups, but reduce resolution from 1.5 to 1.0 
Resolution 1.0 only gives 16 groups, Resolution 2.0 gives 25 groups and looks reasonable. 

```{r}
# save 29 PC, resolution 1.5 cluster labels
pbmc_FitSNE <- StashIdent(object = pbmc_FitSNE, save.name = "ClusterNames_29PCresolution1.5")

pbmc_FitSNE <- FindClusters(pbmc_FitSNE,reduction.type = "pca",
  dims.use = 1:29, k.param = 30, plot.SNN = TRUE, prune.SNN = 1/15,
  print.output = FALSE, distance.matrix = NULL, save.SNN = TRUE,
  reuse.SNN = FALSE, force.recalc = FALSE, nn.eps = 0,
  modularity.fxn = 1, resolution = 2.0, algorithm = 1, n.start = 100,
  n.iter = 10, random.seed = 0, temp.file.location = NULL,
  edge.file.name = NULL)
```


Save results
```{r}
saveRDS(pbmc_FitSNE, file="10X_8K_PBMC_Seurat_FItSNE.rds")
pbmc_FitSNE <- readRDS("10X_8K_PBMC_Seurat_FItSNE.rds")

```


Here, instead of clustering using Seurat's FindCLusters tool, use FIt-SNE to do the "perplexity annealing" protocol
## Can't use FIt-SNE: the Seurat matrix format is not compatible with the fftRtsne function, and the function's required input format is beyond my understanding
##Can try running the entire analysis using FIt-SNE, but the example input format is not something that I'm familiar with, and it's only in python 
## Running on the raw data using various perplexities, without using feature selected genes of PCs, gives comically bad results
```{r}
#problem running this Seurat object because it's not a standard matrix, because it is a combination of many matrices. 
#pbmc_FitSNE@data is its own matrix with 6 slots, Barcodes vs. genes, with normalized UMI values
#explicitly convert that table into a matrix with as.matrix()
#pbmc_matrix <- as.matrix(pbmc_FitSNE@data) #This data has not been scaled and log2 transformed and should not be used for T-SNE clustering input. It's known to give bad results. 
pbmc_matrix <- as.matrix(c(pbmc_FitSNE@data,pbmc_FitSNE@scale.data,pbmc_FitSNE@var.genes,pbmc_FitSNE@dr$pca))
##### I'm pretty sure the extra data is not being transferred over correctly##########
FItSNE5 <- fftRtsne(pbmc_matrix,dims=2, perplexity=5, theta=0.5, initialization=PCAinit)
FItSNE50 <- fftRtsne(pbmc_matrix,dims=2, perplexity=50, theta=0.5)
FItSNE500 <- fftRtsne(pbmc_matrix,dims=2, perplexity=500, theta=0.5) #running. Takes minutes. 
plot(FItSNE500)
plot()
```

The Strategy to get better T-SNE clustering:
1. Cluster with perplexity=500, or some high number, to get the global clusters positioned correctly
  Using the result from above as seed to cluster with perplexity = 50, to resolve closer relationships: use the add.iter function to use the existing T-SNE clusters to rerun
2. Cluster using the first 2 PCAs as the seeding structure, which also should preserve global structures, 
3. FIt-SNE is already built-in to Seurat!!! RunTSNE sith tsne.method=FIt-SNE, but the instructions don't say how FIt-SNE is implemented. I think FIt-SNE here does not refer to the "perplexity" stratery that I'm trying to implement, but the underlying T-SNE algorithm
A) First iteration of "perplexity annealing" with t-SNE, I ran Perplexity 1000 -> 50 + 5000 iterations. Clusters are not very well separated start 10:16
B) Second iteration, run Perplexity 500 -> 50 + 5000 iterations. The result is almost identical

```{r fig.height=10, fig.width=12}
pbmc_FitSNE <- RunTSNE(object= pbmc_FitSNE, reduction.use = "pca", cells.use = NULL, dims.use = 1:29,
  genes.use = NULL, seed.use = 1, add.iter = 0,
  dim.embed = 2, distance.matrix = NULL, reduction.name = "tsne",
  reduction.key = "tSNE_", perplexity= 500)
TSNEPlot(object = pbmc_FitSNE, do.label= TRUE) 
ggsave("T-SNE_Perplexity=500.jpg", width= 12, height=10, dpi=600, plot= last_plot(), units = "in")
```

Get a rough idea of cluster identities

```{r fig.height=10, fig.width=25}
FeaturePlot(object=pbmc_FitSNE, features.plot = c("CD3D","CD8A","CD79A","GNLY","LYZ","CD14","KLF4","PPBP", "CD4","KLRB1","CD69","CCR7","GZMB","PRF1","MT-CYB"), cols.use=c("grey","blue"), nCol=5, no.legend=FALSE) +
ggsave("T-SNE_Perplexity=500_markers.jpg", width= 24, height=12, dpi=600, plot= last_plot(), units = "in")
```
# Rerun on the previous tight cluster with Perplexity = 50, with 5000 iterations, which is pretty large. (10X CellRanger uses 1000 by default) Started at 11:13 a.m. Iteration 1000 at 2:30 Iteration 2000 at end of work day
From the error at iterations, 2000 iterations is enough, 1000 is sufficient
```{r}
pbmc_FitSNE <- RunTSNE(object= pbmc_FitSNE, reduction.use = "pca", cells.use = NULL, dims.use = 1:29,
  genes.use = NULL, seed.use = 1, add.iter = 1000,
  dim.embed = 2, distance.matrix = NULL, reduction.name = "tsne",
  reduction.key = "tSNE_", perplexity= 100)

TSNEPlot(object = pbmc_FitSNE, do.label= TRUE) +
  ggsave("T-SNE_Perplexity=500+100.jpg", width= 10, height=6, dpi=600, plot= last_plot(), units = "in")
```

```{r fig.height=10, fig.width=20}
FeaturePlot(object=pbmc_FitSNE, features.plot = c("CD3D","CD8A","CD79A","GNLY","LYZ","CD14","KLF4","PPBP", "CD4","KLRB1","CD69","CCR7","GZMB","PRF1","MT-CYB"), cols.use=c("grey","blue"), nCol=5, no.legend=FALSE) +
ggsave("T-SNE_Perplexity=500+100_markers.jpg", width= 24, height=12, dpi=600, plot= last_plot(), units = "in")
```

