---
title: "scRNA Seurat Oct"
author: "Mengshu"
date: "10/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(Seurat)
library(ggplot2)
library(magrittr)
source("/Volumes/Enterprise/FLX/Files_from_Gene/R_functions.r")

```

#Seurat does not have a User Manual. All the help functions are in R. Pasted below for quick reference
??Seurat
monocle::importCDS		Import a seurat or scatter/scran CellDataSet object and convert it to a monocle cds.
RBGL::isKuratowskiSubgraph		isKuratowskiSubgraph
Seurat::AddSamples		Add samples into existing Seurat object.
Seurat::Convert		Convert Seurat objects to other classes and vice versa
Seurat::CreateSeuratObject		Initialize and setup the Seurat object
Seurat::FilterCells		Return a subset of the Seurat object
Seurat::MergeSeurat		Merge Seurat Objects
Seurat::Seurat-deprecated		Deprecated function(s) in the Seurat package
Seurat::SubsetByPredicate		Return a subset of the Seurat object.
Seurat::SubsetData		Return a subset of the Seurat object
Seurat::UpdateSeuratObject		Update old Seurat object to accomodate new features
Seurat::seurat		The Seurat Class

??Read10X
data_dir <- 'path/to/data/directory'
list.files(data_dir) # Should show barcodes.tsv, genes.tsv, and matrix.mtx
expression_matrix <- Read10X(data.dir = data_dir)
seurat_object = CreateSeuratObject(raw.data = expression_matrix)

??Read10X_h5
Read10X_h5(path/to/filename, ensg.names = FALSE) #labels the genes with ENSG names rather than the HUGO names

??CreateSeuratObject 
CreateSeuratObject(raw.data, project = "SeuratProject", min.cells = 0,
  min.genes = 0, is.expr = 0, normalization.method = NULL,
  scale.factor = 10000, do.scale = FALSE, do.center = FALSE,
  names.field = 1, names.delim = "_", meta.data = NULL,
  display.progress = TRUE, ...)
  
??VlnPlot
  VlnPlot(object, features.plot, ident.include = NULL, nCol = NULL,
  do.sort = FALSE, y.max = NULL, same.y.lims = FALSE, size.x.use = 16,
  size.y.use = 16, size.title.use = 20, adjust.use = 1,
  point.size.use = 1, cols.use = NULL, group.by = NULL, y.log = FALSE,
  x.lab.rot = FALSE, y.lab.rot = FALSE, legend.position = "right",
  single.legend = TRUE, remove.legend = FALSE, do.return = FALSE,
  return.plotlist = FALSE, ...)
  
??GenePlot(object, gene1, gene2, cell.ids = NULL, col.use = NULL,
  pch.use = 16, cex.use = 1.5, use.imputed = FALSE, use.scaled = FALSE,
  use.raw = FALSE, do.hover = FALSE, data.hover = "ident",
  do.identify = FALSE, dark.theme = FALSE, do.spline = FALSE,
  spline.span = 0.75, ...)
  
??FilterCells(object, subset.names, low.thresholds, high.thresholds,
  cells.use = NULL)
  
??FindVariableGenes(object, mean.function = ExpMean,
  dispersion.function = LogVMR, do.plot = TRUE, set.var.genes = TRUE,
  x.low.cutoff = 0.1, x.high.cutoff = 8, y.cutoff = 1,
  y.high.cutoff = Inf, num.bin = 20, binning.method = "equal_width",
  selection.method = "mean.var.plot", top.genes = 1000, do.recalc = TRUE,
  sort.results = TRUE, do.cpp = TRUE, display.progress = TRUE, ...)
  
??ScaleData(object, genes.use = NULL, data.use = NULL, vars.to.regress,
  model.use = "linear", use.umi = FALSE, do.scale = TRUE,
  do.center = TRUE, scale.max = 10, block.size = 1000,
  min.cells.to.block = 3000, display.progress = TRUE, assay.type = "RNA",
  do.cpp = TRUE, check.for.norm = TRUE, do.par = FALSE, num.cores = 1)
  
??RunPCA(object, pc.genes = NULL, pcs.compute = 20, use.imputed = FALSE,
  rev.pca = FALSE, weight.by.var = TRUE, do.print = TRUE,
  pcs.print = 1:5, genes.print = 30, reduction.name = "pca",
  reduction.key = "PC", assay.type = "RNA", seed.use = 42, ...)
Returns Seurat object with the PCA calculation stored in object@dr$pca.

??VizPCA(object, pcs.use = 1:5, num.genes = 30, use.full = FALSE,
  font.size = 0.5, nCol = NULL, do.balanced = FALSE)

PCA visualization
PrintPCA, VizPCA, PCAPlot PCHeatmap
  
FindClusters(object, genes.use = NULL, reduction.type = "pca",
  dims.use = NULL, k.param = 30, plot.SNN = FALSE, prune.SNN = 1/15,
  print.output = TRUE, distance.matrix = NULL, save.SNN = FALSE,
  reuse.SNN = FALSE, force.recalc = FALSE, nn.eps = 0,
  modularity.fxn = 1, resolution = 0.8, algorithm = 1, n.start = 100,
  n.iter = 10, random.seed = 0, temp.file.location = NULL,
  edge.file.name = NULL)
  
RunTSNE(object, reduction.use = "pca", cells.use = NULL, dims.use = 1:5,
  genes.use = NULL, seed.use = 1, tsne.method = "Rtsne", add.iter = 0,
  dim.embed = 2, distance.matrix = NULL, reduction.name = "tsne",
  reduction.key = "tSNE_", ...)

FindAllMarkers(object, genes.use = NULL, logfc.threshold = 0.25,
  test.use = "wilcox", min.pct = 0.1, min.diff.pct = -Inf,
  print.bar = TRUE, only.pos = FALSE, max.cells.per.ident = Inf,
  return.thresh = 0.01, do.print = FALSE, random.seed = 1,
  min.cells.gene = 3, min.cells.group = 3, latent.vars = NULL,
  assay.type = "RNA", ...)


#Import data. Save as a sparse data matrix, where missing values are not recorded, and the positins of data that do exist is specified by a second data object.
10X data is in the h5.h5 format
```{r}
pbmc_8k <- Read10X_h5("pbmc8k_raw_gene_bc_matrices_h5.h5", ensg.names= FALSE)
```

Create a seurat object from the loaded raw data: Filter out noise expression, and GEMs without cells 
Usually normalization is not performed at this point
```{r}
pbmc <- CreateSeuratObject(pbmc_8k, project="Seurat_pbmc", min.cells=3, min.genes=200, is.expr = 0)
```

Standard Preprocessing workflow: These used to be blackbox standardized, but make explicit in the newest version for transparency
This involves data exploration by graphing and setting thresholds based on how the data looks

Mitochondrial genes: Calculate the percent of mitochondrial genes expressed in each cell, as an indicator of apoptosis
It looks like > 0.1 percent.mito should be filtered out
```{r fig.height=6, fig.width=16}
#gathers the values for all MT rows
mito.genes <- grep(pattern="^MT-", x= rownames(x=pbmc@data), value=TRUE)
 #calculate percent of genes that are MITO for each cell. Summing total expression for each cell, sumMITO/sumALL
percent.mito <- Matrix::colSums(pbmc@raw.data[mito.genes,])/Matrix::colSums(pbmc@raw.data)
#add this information into pbmc under metadata
pbmc <- AddMetaData(object=pbmc, metadata = percent.mito, col.name= "percent.mito")
VlnPlot(object=pbmc, features.plot=c("nGene", "nUMI", "percent.mito"), nCol=3,  point.size.use = 0.1)

```

Plot percent.mito vs. nUMI, and nGene vs. nUMI
Again, percent.mito > 0.1 looks like outliers, and are associated with very low nUMI, probably apoptosed cell debris
```{r}
#GenePlot is usually used for correlation between two genes, but will do a 2D scatter plot of anythng
par(mfrow=c(1,2)) # I think this is a way to do facet plotting of two graphs together
GenePlot(object=pbmc, gene1= "nUMI", gene2 = "percent.mito")
GenePlot(object=pbmc, gene1= "nUMI", gene2= "nGene")

```
Filter dataset
```{r}
# pbmc <- FilterCells(object = pbmc, subset.names = c("nGene", "percent.mito"), low.thresholds = c(200, -Inf), high.thresholds = c(2500, 0.05))
#I'm not going to apply an upper game for nGenes, because I can't see a clustering that indices doublets. 
pbmc <- FilterCells(object=pbmc, subset.names=c("nGene", "percent.mito"), low.thresholds =c(500,-Inf), high.thresholds=c(Inf,0.1))
```

Normalize data
LogNormalize: normalize each cell by total expression, multiply by scale factor 10000, and log2 transform
```{r}
pbmc <- NormalizeData(object=pbmc, normalization.method="LogNormalize", scale.factor=10000)
```
Variable genes are useful for analysis. This step identifies them for downstream analysis
Cutoff for high average expression, which should exclude the RPL genes, which are not very informative)
```{r fig.height=15, fig.width=20}
pbmc <- FindVariableGenes(pbmc, mean.function = ExpMean,
  dispersion.function = LogVMR, do.plot = TRUE, set.var.genes = TRUE,
  x.low.cutoff = 0.0125, x.high.cutoff = 3.5, y.cutoff = 0.5,
  y.high.cutoff = Inf, num.bin = 20, binning.method = "equal_width",
  selection.method = "mean.var.plot", top.genes = 1000, do.recalc = TRUE,
  sort.results = TRUE, do.cpp = TRUE, display.progress = TRUE)
```
```{r}
length(x=pbmc@var.genes)
```
Remove uninteresting variable by regression: Here since the PBMCs are not cycling cells, regress by UMI and percent mito gene content. This regression is part of ScaleData
ScaleData also scales and centers the data. 

```{r}
pbmc <- ScaleData(object=pbmc, vars.to.regress=c("nUMI","percent.mito"), display.progress = TRUE, assay.type = "RNA", check.for.norm = TRUE)
```

Perform linear dimensional reduction
```{r}
pbmc <- RunPCA(object = pbmc, pc.genes = pbmc@var.genes, pcs.compute = 29, do.print = TRUE, pcs.print = 1:29, 
    genes.print = 5)
```
Examine PCA results visually
```{r}
#Just print PCA results
PrintPCA(object=pbmc, pcs.print=1:10, genes.print = 10, use.full = FALSE)
```
# What do the values mean? 
# The values are eigenvalues, which represent their contribution to the PC, the higher the number to 1, the more effect
# What does a positive or a negative value mean?
# Negative values mean the feature has a negative contribution to the PC: lower expression of the gene more strongly defines the PC 
```{r fig.height=20, fig.width=20}
VizPCA(object=pbmc, pcs.use=20:29,font.size = 1)

```

Only the first 12 PCs really change the shape of sample clusters. Starting with 13 on, the samples mostly fall into one large smooth ovoid cluster

```{r}
PCAPlot(object=pbmc, dim.1 = 15, dim.2 = 16)
```
PC1 doesn't separate cells from each other, it's continuous: IL7, CD3G are specific to T cells
PC2 clearly separates cells into 2 groups: CD79A, CD79B are B cell markers
PC3:All negative eigenvalues: NKG7 is a natural killer cell granule protein
PC4: All negative eigenvalues: CD160 is a NK and T cell marker, and , PTCRA marks immature T cells
PC5: All negative eigenvalues: IL3RA is CD123 Dendritic, granulocyte marker
PC6: CD14, granulocyte, monocyte marker
Divisive: 2,5,6,10,14*      Continuous: 1,3,7,8,9,11,12,13,15,16        No resolution: 4


```{r fig.height=20, fig.width=20}
PCHeatmap(object=pbmc, pc.use=1:18, cells.use = 500, do.balanced= TRUE, label.columns= FALSE)
# All clusters look pretty distinctive
```
Determine statistically significant components
A resampling test inspired by the JackStraw procedure. Randomly permutate a subset of data (1% default) and rerun PCA, looking for PCs that are enriched for genes that change PC grouping when changed. 
This step takes a long time
```{r fig.height=20, fig.width=20}
#this process can take a long time
pbmc <- JackStraw(object = pbmc, num.replicate = 100, display.progress = FALSE)
JackStrawPlot(object = pbmc, PCs = 1:20) 
ggsave("JackstrawPlot of PCs.jpg", width= 14, height= 5, dpi=500, plot= last_plot(), units = "in")

```
The PCE elbow plot is a more ad hoc wat of determining a cut-off. Here it looks like the cutoff is around 14, but from the JackStraw and heatmaps, all PCs are significant. Therefore, I will try keeping all PCs.

```{r}
PCElbowPlot(object = pbmc)
#Elbow looks like it's at PC 14
```
#############################
#############################

          CLUSTERING

#############################
#############################
Cluster cells into groups. Using graph-cased clustering. The method embeds cells in a graph structure, like a K-nearest neighbour graph, and then refines the edges. The cells are grouped iteratively. 
Using the first 20 PCs is enough. The PCs after 13 don't really separate cells. Clustering using 29 didn't give meaningful sub-divisions
```{r}
pbmc <- FindClusters(pbmc,reduction.type = "pca",
  dims.use = 1:20, k.param = 30, plot.SNN = TRUE, prune.SNN = 1/15,
  print.output = FALSE, distance.matrix = NULL, save.SNN = TRUE,
  reuse.SNN = FALSE, force.recalc = FALSE, nn.eps = 0,
  modularity.fxn = 1, resolution = 1.5, algorithm = 1, n.start = 100,
  n.iter = 10, random.seed = 0, temp.file.location = NULL,
  edge.file.name = NULL)

PrintFindClustersParams(object=pbmc)
```
Run t-SNE to cluster cells for visualization
I removed the do.fast= TRUE term to try to improve results. 24
```{r}
pbmc <- RunTSNE(object = pbmc, dims.use = 1:20)

```
Plot tSNE
```{r fig.height=6, fig.width=12}
TSNEPlot(object = pbmc, do.label = TRUE, pt.size = 0.8, label.size = 4) 
ggsave("tSNE_PBMC_8K_seurat_Nov8.jpg", width=12, height=8, dpi=600, plot= last_plot(), units="in")

write.table(table(pbmc@ident), "PBMC_Cluster_counts.txt")

```

Save results
```{r}
saveRDS(pbmc, file="10X_8K_PBMC_Seurat.rds")
saveRDS(pbmc_label, file="10X_8K_PBMC_Seurat.rds")
pbmc <- readRDS("10X_8K_PBMC_Seurat.rds")
rm(pbmc)
ls()
```
#################################################
#################################################

               DIFFERENTIAL ANALYSIS

#################################################
#################################################
Cluster Biomarkers: Find differentially expressed genes
Seurat can perform differential expression analysis. It compares each group against all other groups, or only certain groups to each other
First, identify cell types by running group against all
```{r}
all.markers <- FindAllMarkers(object=pbmc, min.pct=0.5, logfc.threshold = 0.5)
all.markers2 <- FindAllMarkers(object=pbmc, min.pct=0.5, logfc.threshold = 0.3)
#min.pct is the minimum proportion of cells expression the marker in either group. 
all.markers %>% group_by(cluster) %>% top_n(20, avg_logFC)
all.markers %>% dfilter(gene=="GNLY") %>% .$cluster %>% table
all.markers %>% dfilter(cluster==18)
```
Plots

VlnPlot()
FeaturePlot() # darkers the gene on the dotplot
RidgePlot() # 
CellPlot()
DotPlot()
DoHeatmap() Heatmap of clusters and gene subset


```{r fig.height=10, fig.width=20}
plot_pbmc <- FeaturePlot(object=pbmc, features.plot = c("CD3D","CD8A","CD4","CD79A","GNLY","GZMB","PRF1","LYZ","CD14","KLF4","PPBP","KLRB1","CD69","CCR7","nUMI","percent.mito"), cols.use=c("grey","blue"), nCol=4, no.legend=FALSE) +
ggsave("ClusterGenePlot_PBMC.jpg", width= 25, height=22, dpi=600, plot= last_plot(), units = "in")

FeaturePlot(object=pbmc, features.plot = c("MAP4K1"), cols.use=c("grey","blue"), no.legend=FALSE) +
ggsave("ClusterGenePlot_HPK1.jpg", width= 8, height=6, dpi=600, plot= last_plot(), units = "in")

pbmc@data %>% dfilter(var.genes=="MAP4K1")
```
T cells heatmaps

```{r fig.height=15, fig.width=16}
FeaturePlot(object=pbmc, features.plot = c("CD3D","CD8A","CD4","CCR7","GZMB","KLRB1","SELL","PRF1","GNLY","NKG7","CD14","ITGAE","CD69","LEF1"), cols.use=c("grey","blue"), no.legend=FALSE) +
ggsave("Tcell_ClusterGenePlot.jpg", width= 30, height=18, dpi=600, plot= last_plot(), units = "in")
VlnPlot(object=pbmc, features.plot = c("CD3D","CD8A","TRGC2","IFITM3","ZNF683","NCR3","XCL2"))
```
```{r fig.height=6, fig.width=16}
FeaturePlot(object=pbmc, features.plot = c("CCR7","KLRB1","IFNG","IL2RB"), cols.use=c("grey","blue"), no.legend=FALSE) +
ggsave("Naive&Active_ClusterGenePlot.jpg", width= 13, height=8, dpi=600, plot= last_plot(), units = "in")

VlnPlot(object=pbmc, features.plot = c("nGene", "nUMI")) +
ggsave("Cell_metrics_violin.jpg", width= 13, height=8, dpi=600, plot= last_plot(), units = "in")
FeaturePlot(object=pbmc, features.plot = c("nGene", "nUMI"), cols.use=c("grey","blue"), no.legend=FALSE) +
ggsave("Cell_metrics.jpg", width= 13, height=8, dpi=600, plot= last_plot(), units = "in")
```

What is the difference between cluter 0 and the rest of the CD8 T cells? 
CCR7 expression, which differentiates Naive and effector T cells
```{r}
Tcells_vs_0 <- FindMarkers(object=pbmc, ident.1=0, ident.2 =c(14,7,16), min.pct=0.5, logfc.threshold = 0.5)
Tcells_vs_0 %<>% mutate(gene=rownames(Tcells_vs_0)) %>% arrange(avg_logFC)
tail(Tcells_vs_0)
write.csv(Tcells_vs_0, file="Cluster0vsTcells.txt", sep="\t")
```
T cells: 
cluster 16 from 14 and 7
Cluster 15 from 14, 7, 16
```{r}
Tcells_vs_16 <- FindMarkers(object=pbmc, ident.1=16, ident.2 =c(7), min.pct=0.5, logfc.threshold = 0.3)
Tcells_vs_16 %<>% mutate(gene=rownames(Tcells_vs_16)) %>% arrange(avg_logFC)

Tcells_vs_15 <- FindMarkers(object=pbmc, ident.1=15, ident.2 =c(14,7,16), min.pct=0.5, logfc.threshold = 0.3)
Tcells_vs_15 %<>% mutate(gene=rownames(Tcells_vs_15)) %>% arrange(avg_logFC)
```
NK cell: cluster 20 is SELL high
20 vs 9 and 10
```{r}
NKcells_vs_20 <- FindMarkers(object=pbmc, ident.1=20, ident.2 =c(9,10), min.pct=0.5, logfc.threshold = 0.3)
NKcells_vs_20 %<>% mutate(gene=rownames(Bcells_vs_20)) %>% arrange(avg_logFC)
```

CD4 T cells
Cluster 1 vs 3 and 6 Cluster 6 vs 1 and 3
```{r}
CD4cells_vs_1 <- FindMarkers(object=pbmc, ident.1=1, ident.2 =c(3,6), min.pct=0.5, logfc.threshold = 0.3)
CD4cells_vs_1 %<>% mutate(gene=rownames(CD4cells_vs_1)) %>% arrange(avg_logFC)

CD4cells_vs_6 <- FindMarkers(object=pbmc, ident.1=6, ident.2 =3, min.pct=0.2, logfc.threshold = 0.3)
CD4cells_vs_6 %<>% mutate(gene=rownames(CD4cells_vs_6)) %>% arrange(avg_logFC)
```
Monocyte heatmaps
```{r fig.height=20, fig.width=20}
FeaturePlot(object=pbmc, features.plot = c("LYZ","CD14","ITGAX", "HLA-DPB1","FCGR3A","CD1C","CXCL8","IL3RA","CLEC4C","CD33"), cols.use=c("grey","blue"),no.legend=FALSE) 
ggsave("Myeloid_ClusterGenePlot.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")
VlnPlot(object=pbmc, features.plot = c("LYZ","CD14","ITGAX", "HLA-DPB1","FCGR3A","CD1C","CXCL8","IL3RA","CLEC4C","CD33"), use.raw = TRUE, y.log = TRUE, x.lab.rot = TRUE)
ggsave("Myeloid_VlnGenePlot.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")
```
Monocyte
```{r}
Monocyte_vs_13 <- FindMarkers(object=pbmc, ident.1=13, ident.2 =c(2,5,11,12), min.pct=0.5, logfc.threshold = 0.3)
Monocyte_vs_13 %<>% mutate(gene=rownames(Monocyte_vs_13)) %>% arrange(avg_logFC)

Monocyte_vs_12 <- FindMarkers(object=pbmc, ident.1=12, ident.2 =c(2,5,11,13), min.pct=0.5, logfc.threshold = 0.3)
Monocyte_vs_12 %<>% mutate(gene=rownames(Monocyte_vs_12)) %>% arrange(avg_logFC)

Monocyte_vs_11 <- FindMarkers(object=pbmc, ident.1=11, ident.2 =c(2,5), min.pct=0.5, logfc.threshold = 0.3)
Monocyte_vs_11 %<>% mutate(gene=rownames(Monocyte_vs_11)) %>% arrange(avg_logFC)
```
B cell
```{r fig.height=5, fig.width=13}
Monocyte_vs_17 <- FindMarkers(object=pbmc, ident.1=17, ident.2 =c(4,8), min.pct=0.5, logfc.threshold = 0.3)
Monocyte_vs_17 %<>% mutate(gene=rownames(Monocyte_vs_17)) %>% arrange(avg_logFC)

Monocyte_4_vs_8 <- FindMarkers(object=pbmc, ident.1=4, ident.2 =c(8), min.pct=0.5, logfc.threshold = 0.3)
Monocyte_4_vs_8 %<>% mutate(gene=rownames(Monocyte_4_vs_8)) %>% arrange(avg_logFC)

Monocyte_4_vs_8 <- FindMarkers(object=pbmc, ident.1=4, ident.2 =c(8), min.pct=0.5, logfc.threshold = 0.3)
Monocyte_4_vs_8 %<>% mutate(gene=rownames(Monocyte_4_vs_8)) %>% arrange(avg_logFC)

Monocyte_2_vs_5 <- FindMarkers(object=pbmc, ident.1="2. CD14+ Monocytes", ident.2 ="5. CD14+ Monocytes", min.pct=0.5, logfc.threshold = 0.3)
Monocyte_2_vs_5 %<>% mutate(gene=rownames(Monocyte_2_vs_5)) %>% arrange(avg_logFC)
```
#################################################
#################################################

               Relabel Cell Groups

#################################################
#################################################


```{r}
current.cluster.ids <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)
new.cluster.ids <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)
current.cluster.ids <- c("0. Naive CD8 T","1. Active CD4 T", "2. Active Monocyte", "3. Naive CD4 T", "4. B cells","5. Monocyte","6. Naive CD4 T","7. T EM","8. B cells","9. NK cell","10. NKT cell","11. Immature Monocyte","12. FCGR3A+ DC4", "13. DC2", "14. T effector", "15. Apoptotic T","16. T RM","17. CD3+ B","18. DC1","19. Megakaryocyte","20. NK precursor" )
new.cluster.ids <- c("0. Naive CD8 T","1. Active CD4 T", "2. Active Monocyte", "3. Naive CD4 T", "4. B cell precursor","5. Monocyte","6. Naive CD4 T","7. T EM","8. B cells","9. NK cell","10. NKT cell","11. Immature Monocyte","12. FCGR3A+ DC4", "13. DC2", "14. T effector", "15. Apoptotic T","16. T RM","17. CD3+ B","18. DC1","19. Megakaryocyte","20. NK precursor" )
pbmc@ident <- plyr::mapvalues(x = pbmc@ident, from = current.cluster.ids, to = new.cluster.ids)
TSNEPlot(object = pbmc, do.label = TRUE, pt.size = 1, label.size = 3) 
ggsave("Labelled_ClusterGenePlot.jpg", width= 14, height=7, dpi=600, plot= last_plot(), units = "in")


```

Violin Plots for markers
```{r fig.height=7, fig.width=20}
VlnPlot(object = pbmc, features.plot = c("CD3D","CD3E","CD3G","CD79A"), use.raw = TRUE, y.log = TRUE, x.lab.rot = TRUE) +
ggsave("CD3+ B cell.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")
```
T RM heatmaps
```{r fig.height=15, fig.width=20}
FeaturePlot(object=pbmc, features.plot = c("CD69","CD8A","ITGAE","KLRB1","PRF1","GZMB","HAVCR2","CTLA4","PDCD1","SELL","KLRG1"), cols.use=c("grey","blue"),no.legend=FALSE) 
#ggsave("Trm_ClusterGenePlot.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")
VlnPlot(object=pbmc, features.plot = c("CD69","CD8A","ITGAE","KLRB1","PRF1","GZMB","HAVCR2","CTLA4","PDCD1","SELL","CCR7","KLRG1"), use.raw = TRUE, y.log = TRUE, x.lab.rot = TRUE)
VlnPlot(object=pbmc, features.plot = c("nGene","percent.mito"), use.raw = TRUE, y.log = TRUE, x.lab.rot = TRUE) +
ggsave("VlnGenePlot_mito_gene.jpg", width= 20, height=10, dpi=600, plot= last_plot(), units = "in")
```

HPK1
```{r}
library(plotly)
FeaturePlot(object=pbmc, features.plot = c("MAP4K1"), cols.use=c("grey","blue"), no.legend=FALSE, do.hover=TRUE, data.hover = c("ident", "nGene")) 

FeaturePlot(object=pbmc, features.plot = c("MAP4K1","GZMB"), cols.use=c("grey","blue","red","green"), overlay=TRUE, no.legend=FALSE, do.hover=TRUE, data.hover = c("ident", "nGene")) 

VlnHPK1 <- VlnPlot(object=pbmc, features.plot = c("MAP4K1"), x.lab.rot = TRUE) 
ggsave("Vln_Feature_HPK1.jpg", width= 9, height=5, dpi=600, plot= VlnHPK1, units = "in")

FeaturePlot(object=pbmc_30, features.plot = c("IL4","IL6","IL10","IL13","nGene","nUMI","percent.mito"), cols.use=c("grey","blue"), no.legend=FALSE) 

ggsave("Vln_Feature_HPK1.jpg", width= 9, height=7, dpi=600, plot= last_plot(), units = "in")
```

##########
###########
Re-CLUSTER

################
#####################



```{r fig.height=6, fig.width=12}
#Save original labels for clusters before re-clustering

pbmc <- StashIdent(object = pbmc, save.name = "ClusterNames_1.5")


#Re-run with higher resolution number: 2.0, to subdivide clusters further
pbmc <- FindClusters(pbmc,reduction.type = "pca",
  dims.use = 1:20, k.param = 30, plot.SNN = TRUE, prune.SNN = 1/15,
  print.output = FALSE, distance.matrix = NULL, save.SNN = TRUE,
  reuse.SNN = FALSE, force.recalc = FALSE, nn.eps = 0,
  modularity.fxn = 1, resolution = 3.0, algorithm = 1, n.start = 100,
  n.iter = 10, random.seed = 0, temp.file.location = NULL,
  edge.file.name = NULL)

plot1 <- TSNEPlot(object = pbmc, do.return = TRUE, no.legend = TRUE, do.label = TRUE)
plot2 <- TSNEPlot(object = pbmc, do.return = TRUE, group.by = "ClusterNames_1.5", 
    no.legend = TRUE, do.label = TRUE)
plot_grid(plot1, plot2)
```
```{r}
naiveCD8.markers <- FindMarkers(object = pbmc, ident.1 = 10, ident.2 = 4)
naiveCD8.markers %<>% mutate(gene=rownames(naiveCD8.markers)) %>% arrange(avg_logFC)

```


Re-do PCA with 30 PCs
```{r}
pbmc_30 <- pbmc
pbmc_30 <- RunPCA(object = pbmc, pc.genes = pbmc@var.genes, do.print = TRUE, pcs.print = 1:5, 
    genes.print = 5, pcs.compute = 30)
```

```{r}
PCHeatmap(object = pbmc_30, pc.use = 20:30, cells.use = 500, do.balanced = TRUE, 
    label.columns = FALSE, use.full = FALSE)
```
Re-run jackstraw with 30 PCs, 29 of them look significant
```{r fig.height=24, fig.width=12}
pbmc_30 <- JackStraw(object = pbmc_30, num.replicate = 100, display.progress = FALSE, num.pc=30)
JackStrawPlot(object = pbmc_30, PCs = 1:30)

```
```{r}
pbmc_30 <- FindClusters(object = pbmc_30, reduction.type = "pca", dims.use = 1:29, 
    resolution = 1.5, print.output = 0, save.SNN = TRUE)
pbmc_30 <- RunTSNE(object = pbmc, dims.use = 1:29, do.fast = TRUE)
TSNEPlot(object = pbmc_30, do.label=TRUE)
ggsave("Labelled_ClusterGenePlot30.jpg", width= 9, height=5, dpi=600, plot= last_plot(), units = "in")
```

##### Gene plots

```{r fig.height=12, fig.width=16}
FeaturePlot(object=pbmc_30, features.plot = c("CD3D","CD8A","CD79A","GNLY","LYZ","CD14","CD1C","KLF4","PPBP", "CD4","KLRB1","CD69","MAP4K1","CCR7","GZMB","PRF1","IL7R","MT-CYB","nGene", "nUMI", "percent.mito"), cols.use=c("grey","blue"), no.legend=FALSE) 
FeaturePlot(object=pbmc_30, features.plot = c("MAP4K1"), cols.use=c("grey","blue"), no.legend=FALSE) 
VlnPlot(object=pbmc_30, features.plot = c("MAP4K1")) 

FeaturePlot(object=pbmc_30, features.plot = c("IL4","IL6","IL10","IL13","nGene","nUMI","percent.mito"), cols.use=c("grey","blue"), no.legend=FALSE) 

ggsave("Feature_HPK1.jpg", width= 9, height=7, dpi=600, plot= last_plot(), units = "in")
```
DC markers

```{r fig.height=12, fig.width=16}
FeaturePlot(object=pbmc_30, features.plot = c("CD3D","CD8A","CD79A","HLA-DRB1","LYZ","CD14","CD1C","FCER1A","CLEC9A","S100A9","CD1D","FCGR3A","ITGAX","nGene","nUMI","percent.mito"), cols.use=c("grey","blue"), no.legend=FALSE) 
```
Differential Analysis of T cell clusters 19 and 20
```{r}
Tvs19.markers <- FindMarkers(object = pbmc_30, ident.1 = 19, ident.2 = 20)
Tvs19.markers %<>% mutate(gene=rownames(Tvs19.markers)) %>% arrange(avg_logFC)

Tvs7.markers <- FindMarkers(object = pbmc_30, ident.1 = 7, ident.2 = c(19))
Tvs7.markers %<>% mutate(gene=rownames(Tvs7.markers)) %>% arrange(avg_logFC)

T1vs11.markers <- FindMarkers(object = pbmc_30, ident.1 = 1, ident.2 = c(11))
T1vs11.markers %<>% mutate(gene=rownames(T1vs11.markers)) %>% arrange(avg_logFC)

T1vs22.markers <- FindMarkers(object = pbmc_30, ident.1 = 22, ident.2 = c(3))
T1vs22.markers %<>% mutate(gene=rownames(T1vs22.markers)) %>% arrange(avg_logFC)
```

#################################################
#################################################

    CALCULATE AVERAGE EXPRESSION PER CLUSTER

#################################################
#################################################

```{r}
cluster.averages <- AverageExpression(object = pbmc, return.seurat = TRUE)
```
```{r}
head(x = cluster.averages[, 1:5])
```




```{r}
# How can I plot the average expression of NK cells vs. CD8 T cells?  Pass
# do.hover = T for an interactive plot to identify gene outliers
library(plotly)
CellPlot(object = cluster.averages, cell1 = "7. T EM", cell2 = "16. T RM", do.hover = TRUE, hoverinfo="text")
CellPlot(object = cluster.averages, cell1 = "4. B cells", cell2 = "8. B cells", do.hover = TRUE, hoverinfo="text")
CellPlot(object = cluster.averages, cell1 = "9. NK cell", cell2 = "10. NKT cell", do.hover = TRUE, hoverinfo="text")
CellPlot(object = cluster.averages, cell1 = "9. NK cell", cell2 = "20. NK precursor", do.hover = TRUE, hoverinfo="text")
??do.hover
```
PC3: NK
PC4: Megakaryocyte
PC5: DC1
PC6: DC4
PC8: DC2
PC9: T em
PC10: B cell precursor vs mature
PC11: not sure, Monocyte
PC12: T eff
PC18: B cell precursor
PC20: T rm
```{r}
hi <-DoHeatmap(object = cluster.averages, genes.use = PCTopGenes(object = pbmc, pc.use = 20, do.balanced = TRUE), group.label.rot = TRUE, group.cex = 0)
ggsave("Heatmap_PC20_Trm.jpg", width= 9, height=7, dpi=600, plot= hi, units = "in")
```
Differential expression table

```{r}
View(cluster.averages)
```

