---
title: "Talmadge.Rmd"
author: "MX"
date: "11/12/2020"
output: html_document
---

In this Rmd, I'm starting a fresh code page because the old one is getting too cluttered with the difference batches of data and things I've tried.
This data will have batch 1 to batch 4 datasets (batch 5 all failed)

```{r}
library(Seurat)
library(sctransform)
library(data.table)
library(magrittr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(stringr)
library(tibble)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
#source("D:/FLX/Files_from_Gene/R_functions.r")
#install.packages("hdf5r")
#devtools::install_github("hhoeflin/hdf5r")
```

#######################
Save and Load data
#######################


```{r}
#batch3
load("/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/Cellranger3.1.0_batch3/batch3c.Robj")
#batch4
load("/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/Cellranger3.1.0_batch4/batch4_SCT.Robj")
rownames(batch4$CITE)
save(batch4,file="/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/Cellranger3.1.0_batch4/batch4_SCT.Robj")
save(batch4_2,file="/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/Cellranger3.1.0_batch4/batch4_SCT_2.Robj")
```

Batch 4  
```{r}
cancer_aml3 <- Read10X_h5("/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/Cellranger3.1.0_batch4/CPB04MAY20_CITE_filtered_feature_bc_matrix.h5") 
```

#######################
Get rid of the mouse genes
#######################

Seurat suggests keeping the top 100, but since we didn't spike in any mouse genes, there's no point, filter them all out
This automatically gets rid of the GRCh38_ prefix in front of all of the gene names as well. 

batch4
```{r}
cancer_aml3$'Gene Expression' <- CollapseSpeciesExpressionMatrix(cancer_aml3$'Gene Expression',prefix ="GRCh38_", controls = "mm10___", ncontrols = 0)
```

Insert CITE data

batch4
```{r}
aml3 <- CreateSeuratObject(counts= cancer_aml3$'Gene Expression', project= "aml3", min.cells=1, min.features=100, assay="RNA")

aml3[["CITE"]] <- CreateAssayObject(counts= cancer_aml3$'Antibody Capture'[,colnames(x=aml3)])

aml3$CITE@data@Dimnames[[1]]
new_CITE_names <- rownames(aml3$CITE) %>% str_replace("-human","-CITE")
aml3$CITE@data@Dimnames[[1]] <- new_CITE_names
```

#######################
Merge Batch 4 (AML3) into batch3c
#######################
```{r}
batch4 <- merge(x=batch3c,y=aml3, add.cell.ids=c("","aml3"), merge.data=FALSE,project="MDSC_batch4")
batch4$CITE
```

######################
Normalize CITE seq data
#######################
 aml1  aml3 pbmc1 pbmc2 pbmc3 pdac1 pdac2 pdac3 
 6436  2418  2460  3464  4699  5110  4108  2611 
```{r}
batch4 <- NormalizeData(batch4,assay="CITE",normalization.method = "CLR")
batch4 <- ScaleData(batch4, assay="CITE")
tail(colnames(batch4))
table(batch4$orig.ident)
```

######################
Mitochondrial gene expression
Check Apoptosis
nCount_RNA is # of unique UMIs I think
Cut off percent.mt at > 25
Clusters 12, 8, 9 are very low in RNA content, but aren't neutrophils
#######################
```{r fig.height=8, fig.width=8}
batch4[["percent.mt"]] <- PercentageFeatureSet(batch4, pattern = "^MT")
batch4$percent.mt
batch4[["percent.HLA.D"]] <- PercentageFeatureSet(batch4, pattern = "^HLA-D")
batch4[["percent.HLA.DR"]] <- PercentageFeatureSet(batch4, pattern = "^HLA-DR")
batch4[["RNA.Per.Feature"]] <- batch4[["nCount_RNA"]] / batch4[["nFeature_RNA"]]
batch4[["percent.ribo"]] <- PercentageFeatureSet(batch4, pattern = c("^RPL","^RPS"))

VlnPlot(batch4, features=c("nFeature_RNA","nCount_RNA", "nFeature_CITE","nCount_CITE"),ncol =2,pt.size = 0.1, slot="data", sort=TRUE, log=TRUE) # ,"percent.mt","percent.ribo"

VlnPlot(batch4, features=c("nFeature_RNA","nCount_RNA","percent.mt","percent.ribo" ),ncol =2,pt.size = 0.1, slot="data", sort=T, log=TRUE) # ,"percent.mt","percent.ribo" 
VlnPlot(batch3b, features=c("nCount_RNA"),pt.size = 0.1, slot="counts", y.max=20000)
VlnPlot(batch3c, features=c("nCount_RNA","nFeature_RNA","percent.mt"),ncol =3,pt.size = 0.1)

VlnPlot(batch4, features=c("BCL2","BAX","MLKL","HMGB1"),ncol=4,pt.size = 0.1, slot="data") # DR5
VlnPlot(batch4, features=c("CD63","FCGR3B","LTF","S100A9","percent.ribo"),ncol =5,pt.size = 0.1, log=TRUE)
## Cell cycle
VlnPlot(batch4, features=c("MKI67","CCNA2","CCNB1","GAPDH"),ncol =4,pt.size = 0.1, slot="data")

# Is there an anti-correlation between MT-content and nFeature_RNA?
plot1 <- FeatureScatter(batch4, feature1 = "nCount_RNA", feature2 = "percent.mt",slot="data")
plot2 <- FeatureScatter(batch4, feature1 = "BAK1", feature2 = "percent.mt",slot="data")
plot3 <- FeatureScatter(batch4, feature1 = "BCL2", feature2 = "percent.mt",slot="data")
plot4 <- FeatureScatter(batch4, feature1 = "BAX", feature2 = "percent.mt",slot="data")
plot5 <- FeatureScatter(batch4, feature1 = "BAX", feature2 = "nCount_RNA",slot="data")
plot6 <- FeatureScatter(batch4, feature1 = "percent.ribo", feature2 = "percent.mt",slot="data")
CombinePlots(plots = list(plot1, plot2,plot3,plot4))

# It looks like cluster 21 is high in mt genes

FeatureScatter(batch4, feature1 = "nFeature_RNA", feature2 = "percent.mt",slot="data", cells = dying_cells)
low_rna_cells <- FetchData(batch4,vars=c("ident")) %>% data.frame %>% tibble::rownames_to_column("Cell") %>% dfilter(ident %in% c(8,9,12)) %>% pull(Cell)
dying_cells <- FetchData(batch4,vars=c("ident")) %>% data.frame %>% tibble::rownames_to_column("Cell") %>% dfilter(ident %in% c(21,19,20)) %>% pull(Cell)
```
There are dying and actively dividing cells in the gMDSC group
```{r fig.height=4, fig.width=5}
gMDSC <- FetchData(batch4,vars=c("ident")) %>% data.frame %>% tibble::rownames_to_column("Cell") %>% dfilter(ident %in% c(20)) %>% pull(Cell)
FeatureScatter(batch4, feature1 = "GAPDH", feature2 = "percent.mt",slot="counts", cells = gMDSC)
```


######################
Normalize data: @data slot calculation
#######################
By default, we employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result. Normalized values are stored in pbmc[["RNA"]]@data
```{r}
batch4 <- NormalizeData(batch4,normalization.method = "LogNormalize",scale.factor=10000) # this goes into the @data slot
```

######################
Feature Selection
#######################
Top variable genes: these are mostly Myeloid markers: PPBP and DEFA3 are both gMDSC markers
 [1] "CAMP","PPBP","LTF"      "DEFA3"    "PF4"      "PTGDS"    "FCER1A","HLA-DQA1" "DEFA4"    "JCHAIN"  
```{r fig.height=5, fig.width=12}
batch4 <- FindVariableFeatures(batch4, selection.method = "vst", nfeatures = 2000)
top30 <- head(VariableFeatures(batch4),30)

plot1 <- VariableFeaturePlot(batch4)
plot2 <- LabelPoints(plot = plot1, points = top30, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))
```

######################
Data scaling to median = 0
Z-score
Stored in @scale.data
#######################
```{r}
all.genes <- rownames(batch4)
batch4 <- ScaleData(batch4,features = all.genes, assay = "RNA")
```

######################
scTransform

This step supersedes NormalizeData, ScaleData and FindVariableFeatures, in that it becomes the default data (it doesn't over write data)
Regularized negative binomial regression to normalize UMI counts: saved in SCT, and this becomes marked default
Optional step to regress out confounding conditions: sctransform
https://satijalab.org/seurat/v3.0/sctransform_vignette.html
I should regress out the mt percentage, because they vary between samples, and shouldn't be biologically relevant
Also regress out RPL/RPS ribosomal protein %, which removed the RPL and RPS genes from the PCA feature selection
#######################

```{r}
batch4 <- SCTransform(batch4,vars.to.regress=c("orig.ident","percent.mt","percent.ribo"),assay="RNA",verbose=FALSE) 
batch4@assays$SCT
batch4[["SCT"]]@scale.data[1:5,1:5] 

batch4_orig <- SCTransform(batch4,vars.to.regress=c("orig.ident"),assay="RNA",verbose=FALSE) # try regression with orig.ident only
```

####################################
PCA analysis
I will choose variable RNA features
####################################

```{r}
batch4 <- RunPCA(batch4, features =c(VariableFeatures(object=batch4)))

batch4_mt <- RunPCA(batch4_mt, features =c(VariableFeatures(object=batch4_mt)))
```

#################################
Visualize PCA
################################

```{r fig.height=5, fig.width=10}
VizDimLoadings(batch4, dims=1:2, reduction ="pca")
DimPlot(batch4, reduction="pca", dims= 1:2)
```


Cut off should be at 20 from this plot
```{r fig.height=14, fig.width=9}
DimHeatmap(batch4, dims=1:24, cells=500, balanced = TRUE)
#Looks like cut-off at 20
```

####################################
JackStraw permutation test for significance of PCs

Batch4: choosing 25 PC as cutoff
####################################
```{r}
ElbowPlot(batch4,ndims=30)
```

###################################
###################################
Graph-based clustering : embed cells in a graph structure, like a KNN graph
##################################
###################################
```{r}
batch4 <- FindNeighbors(batch4, dims=1:30)
batch4 <- FindClusters(batch4, resolution=0.6,random.seed=42) # 0.8, finds 41 clusters, 0.6 35 clusters, 0.55 31 clusters
batch4@active.ident %>% table
batch4$orig.ident %>% table

```

UMAP and tSNE

Make an experimenting Seurat object called Batch_Exp. I don't want to change the UMAP version to a new one 
But then accidentally ran the analysis in batch1, so now batch_exp is the permanent one and batch1 is the test
```{r fig.height=6, fig.width=9}
batch4 <- RunUMAP(batch4, dims=1:30, min.dist = 0.75,random.seed=42)

#min.dist = 0.75 increases inter-cell distance, but setting spread=2 returns the cells to equivalent to min.dist=0.5
UMAPPlot(batch4_2, label=TRUE) 
DimPlot(batch4_2, reduction="umap",label=FALSE, group.by="orig.ident", cols=c("red","firebrick2","green4","green3","blue","magenta3","orange1","violet"))
```

```{r}
DimPlot(batch4_2, reduction="umap",label=F, group.by="orig.ident")
```

Plot per Donor
```{r fig.height=8, fig.width=10}
library(grid)
library(gridExtra)
  p1 <- DimPlot(batch4_2, cells.highlight = WhichCells(batch4,expression= orig.ident =="pdac1"),cols.highlight="purple",sizes.highlight=0.2) + NoLegend()
  p2 <- DimPlot(batch4_2, cells.highlight = WhichCells(batch4,expression= orig.ident =="pdac2"),cols.highlight="purple",sizes.highlight=0.2) + NoLegend()
  p3 <- DimPlot(batch4_2, cells.highlight = WhichCells(batch4,expression= orig.ident =="pdac3"),cols.highlight="purple",sizes.highlight=0.2) + NoLegend()
  n1 <- DimPlot(batch4_2, cells.highlight = WhichCells(batch4,expression= orig.ident =="pbmc1"),cols.highlight="blue",sizes.highlight=0.2) + NoLegend()
  n2 <- DimPlot(batch4_2, cells.highlight = WhichCells(batch4,expression= orig.ident =="pbmc2"),cols.highlight="blue",sizes.highlight=0.2) + NoLegend()
  n3 <- DimPlot(batch4_2, cells.highlight = WhichCells(batch4,expression= orig.ident =="pbmc3"),cols.highlight="blue",sizes.highlight=0.2) + NoLegend()
  a1 <- DimPlot(batch4_2, cells.highlight = WhichCells(batch4,expression= orig.ident =="aml1"),cols.highlight="red",sizes.highlight=0.05) + NoLegend()
  a3 <- DimPlot(batch4_2, cells.highlight = WhichCells(batch4,expression= orig.ident =="aml3"),cols.highlight="red",sizes.highlight=0.05) + NoLegend()
grid.arrange(p1,p2,p3,n1,n2,n3,a1,a3, ncol=3)
```

Plot lineage markers
```{r fig.height=12, fig.width=25}
#lymphocytes
FetchData(batch4_2, vars=orig.ident=="aml1")
FeaturePlot(batch4_2,slot = "data",features = c("PTPRC","CD3G","CD8A","CD4-CITE","GZMB","NCAM1","XCL1","GZMA","nFeature_RNA","percent.mt","XCL1","GAPDH"), ncol=4, cols=c("grey80","red")) 
FeaturePlot(batch4_2,slot = "data",features = c("PTPRC","CD3G","CD8A","CD4-CITE","FOXP3"), ncol=4, cols=c("grey80","red")) 
#myeloid
FeaturePlot(batch4_2,slot = "data",features = c("HLA-DRA","CD11c-CITE","CD33-CITE","CD15-CITE","CD14","CXCR2","CD24","FCGR3A","FCGR3B","nFeature_RNA","percent.mt","percent.ribo","CD163","CLEC4C","CD1C","CD68","mMDSC_s"), ncol=5, cols=c("grey80","red")) 
#CD4 cells that express neutrophil markers
FeaturePlot(batch4_2,slot = "data",features = c("CXCR2","NEAT1","NAMPT","SOD2","IFITM2","FCGR3B",order=T), ncol=3, cols=c("grey80","red")) 
#this instantly makes a Rshiny style page
IFeaturePlot(batch4,slot = "scale.data",feature = c("CD24","CXCR2")) 
```

```{r fig.height=5, fig.width=6}
library(scales)
library(viridis)
FeaturePlot(batch4,slot = "counts",features = c("percent.mt")) + scale_color_viridis(option="inferno", limits=c(0,60), oob=squish)
FeaturePlot(batch4,slot = "counts",features = c("percent.ribo")) + scale_color_viridis(option="inferno", limits=c(0,30), oob=squish)
FeaturePlot(batch4_2,slot = "counts",features = c("nFeature_RNA")) + scale_color_viridis(option="inferno", limits=c(0,3000), oob=squish)
```
Filter dataset for percent mito
Filter away clusters 7, 8, 11, 20
Clusters 25 and 34 are neutrophil-like but with weak markers, and very low RNA content, remove
```{r}
batch4_2 <- subset(batch4, idents = c(7,8,11,20),invert=TRUE )
batch4_2 <- subset(batch4_2, subset= percent.mt <= 25 )
batch4_2 <- subset(batch4_2, idents = c(25,34),invert=TRUE )
batch4_2@active.ident %>% unique
VlnPlot(batch4_2, features=c("percent.mt"), pt.size=0.1) + NoLegend()
save(batch4_2, file="/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/Cellranger3.1.0_batch4/batch4_SCT_2.Robj")
```

Find markers

```{r fig.height=11, fig.width=10}
batch4.markers <- FindAllMarkers(batch4_2, only.pos = TRUE, min.pct = 0.5, logfc.threshold = 0.75)
top.markers <- batch4.markers %>% group_by(cluster) %>% top_n(n=8, wt= avg_logFC) 
DoHeatmap(batch4_2, features= top.markers$gene, size=3,group.bar=T,slot="scale.data") 
```

```{r fig.height=11, fig.width=8}
small_clusters <- subset(batch4_2, idents=c(23,24,25,26,27,28,29,30,31,32,33,34))
top.markers_small <- batch4.markers %>% group_by(cluster) %>% dfilter(cluster %in% c(23,24,25,26,27,28,29,30,31,32,33,34)) %>% top_n(n=8, wt= avg_logFC) 
DoHeatmap(small_clusters, features= top.markers$gene, size=3,group.bar=T,slot="scale.data") 
```

Get cell type and donor data and tabulate & plot
Seurat tidyr
```{r}
#install.packages("tidyseurat")
library(tidyseurat)
library(purrr)
batch4_tidy <- tidy(batch4_2)
batch4_tidy@meta.data
```

Assign cell types to the clusters
```{r}

batch4_tidy %>% dselect(orig.ident,seurat_clusters,nFeature_RNA,percent.mt ) %>% dfilter(seurat_clusters %nin% c(7,8,11,20,25,34)) %>%  mutate(Cell_type = case_when(
 seurat_clusters %in% c(5,6,10,16,26,27,31) ~ "CD4_T",
 seurat_clusters %in% c(4,17,30) ~ "CD8_T",
 seurat_clusters %in% c(9,24,28) ~ "NK",
 seurat_clusters %in% c(28) ~ "NK_XCL1",
 seurat_clusters %in% c(19,21) ~ "B_cell",
 seurat_clusters %in% c(33) ~ "pDC",
 seurat_clusters %in% c(29) ~ "DC2",
 seurat_clusters %in% c(32) ~ "RBC",
 seurat_clusters %in% c(23) ~ "gMDSC",
 seurat_clusters %in% c(12) ~ "Basophil",
 seurat_clusters %in% c(0,2,3,14,18) ~ "Neutrophil",
 seurat_clusters %in% c(1,15,22) ~ "Monocyte_classic",
 seurat_clusters %in% c(13) ~ "Monocyte_nonclassic",
)) 
  
  group_by(orig.ident) %>% mu
```

