---
title: "Talmadge_AML.Rmd"
author: "MX"
date: "9/28/2020"
output: html_document
---

The complete Talmadge dataset is finalized, with 4 batches of data, 9 samples. Batch 5 with 2 AMLs and 1 apheresis did not make it, no cDNA detected.

This Rmd includes analysis for AML1 and AML3 individually
```{r}
library(Seurat)
library(data.table)
library(magrittr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(stringr)
source("/Volumes/Picard/FLX/Files_from_Gene/R_functions.r")
```

AML data split off from batch4 grouped clustering
This will contain all the meta data that is relevant to the global analysis, which is useful for comparing back
```{r}
load("/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/Cellranger3.1.0_batch4/batch4.Robj")
aml1 <- subset(batch4, subset= orig.ident %in% c("aml1"))
save(aml1,file="Batch4_AML1.rdata")
aml3 <- subset(batch4, subset= orig.ident %in% c("aml3"))
save(aml3,file="Batch4_AML3.rdata")
```


Start analysis fresh with just the AML1/AML3 data
```{r}
AML1_alone <- subset(batch4, subset= orig.ident %in% c("aml1"))
AML3_alone <- subset(batch4, subset= orig.ident %in% c("aml3"))
```

######################
Feature Selection
#######################
Top variable genes: these are mostly Myeloid markers: PPBP and DEFA3 are both gMDSC markers
 [1] "CAMP","PPBP","LTF"      "DEFA3"    "PF4"      "PTGDS"    "FCER1A","HLA-DQA1" "DEFA4"    "JCHAIN"  
```{r fig.height=5, fig.width=12}
AML3_alone <- FindVariableFeatures(AML3_alone, selection.method = "vst", nfeatures = 2000)
top30 <- head(VariableFeatures(AML3_alone),30)

plot1 <- VariableFeaturePlot(AML3_alone)
plot2 <- LabelPoints(plot = plot1, points = top30, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))

AML1_alone <- FindVariableFeatures(AML1_alone, selection.method = "vst", nfeatures = 2000)
top30 <- head(VariableFeatures(AML1_alone),30)

plot1 <- VariableFeaturePlot(AML1_alone)
plot2 <- LabelPoints(plot = plot1, points = top30, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))

```

######################
Data scaling to median = 0
Z-score
Stored in @scale.data
#######################
```{r}
all.genes <- rownames(AML3_alone)
AML3_alone <- ScaleData(AML3_alone,features = all.genes, assay = "RNA")

all.genes <- rownames(AML1_alone)
AML1_alone <- ScaleData(AML1_alone,features = all.genes, assay = "RNA")

```

######################
scTransform

This step supersedes NormalizeData, ScaleData and FindVariableFeatures
Regularized negative binomial regression to normalize UMI counts: saved in SCT, and this becomes marked default
Optional step to regress out confounding conditions: sctransform
https://satijalab.org/seurat/v3.0/sctransform_vignette.html
I should regress out the mt percentage, because they vary between samples, and shouldn't be biologically relevant
Also regress out RPL/RPS ribosomal protein %, which removed the RPL and RPS genes from the PCA feature selection
#######################

```{r}
AML1_alone <- SCTransform(AML1_alone, vars.to.regress = c("percent.mt","percent.ribo"), verbose = FALSE, assay="RNA")
#this did not finish, also still not sure that I need to run this
AML3_alone <- SCTransform(AML3_alone, vars.to.regress = c("percent.mt","percent.ribo"), verbose = FALSE, assay="RNA")
```
####################################
PCA analysis
I will choose variable RNA features, and all CITE features
Noticed some RPL and RPS genes in there, go back to scTransform and regress out percent.ribo
####################################

```{r}
AML1_alone <- RunPCA(AML1_alone, features =c(VariableFeatures(object=AML1_alone)))
AML3_alone <- RunPCA(AML3_alone, features =c(VariableFeatures(object=AML3_alone)))
```


#################################
Visualize PCA
################################

```{r fig.height=5, fig.width=10}
DimPlot(AML1_alone, reduction="pca", dims= 1:2)
DimPlot(AML1_alone, reduction="pca", dims= 2:3)
DimPlot(AML1_alone, reduction="pca", dims= 3:4)

DimPlot(AML3_alone, reduction="pca", dims= 1:2)
```


Cut off should be at 20 from this plot
```{r fig.height=14, fig.width=9}
DimHeatmap(AML1_alone, dims=1:24, cells=500, balanced = TRUE)
DimHeatmap(AML3_alone, dims=1:20, cells=500, balanced = TRUE)
#Looks like cut-off at 20
```


Graph-based clustering : embed cells in a graph structure, like a KNN graph

Take many groups, and then merge them later if necessary
PCs: 24
Resolution: 1.2
21 Groups

Batch3 resolution 1.0: 28 clusters
0.9 -> 28 clusters
0.8 -> 25

Batch3b 20 PCs
resolution 
0.8 -> 23 clusters
##################################
###################################
```{r}
AML1_alone <- FindNeighbors(AML1_alone, dims=1:20)
AML1_alone <- FindClusters(AML1_alone, resolution=0.8,random.seed=42) #finds 17 communities
FindCluster
table(Idents(AML1_alone))

AML3_alone <- FindNeighbors(AML3_alone, dims=1:20)
AML3_alone <- FindClusters(AML3_alone, resolution=0.8,random.seed=42) #finds 12 communities
table(Idents(AML3_alone))

```


UMAP and tSNE

Make an experimenting Seurat object called Batch_Exp. I don't want to change the UMAP version to a new one 
But then accidentally ran the analysis in batch1, so now batch_exp is the permanent one and batch1 is the test
```{r fig.height=4, fig.width=6}
AML1_alone <- RunUMAP(AML1_alone, dims=1:20, min.dist = 0.5,random.seed=42)
#AML1_alone <- RunUMAP(AML1_alone, features=VariableFeatures(AML1_alone), min.dist = 0.5, random.seed=42) #this doesn't separate cluters within major clusters, it moves superclusters apart

AML3_alone <- RunUMAP(AML3_alone, dims=1:20, min.dist = 0.5,random.seed=42)

AML1_alone <- RunTSNE(AML1_alone, dims=1:20, min.dist = 0.5, random.seed=42)
AML3_alone <- RunTSNE(AML3_alone, dims=1:20, min.dist = 0.5, random.seed=42)

batch4 <- RunUMAP(batch4, dims=1:24, min.dist = 0.5,random.seed=42)

UMAPPlot(AML1_alone, label=TRUE) #+ ggsave("UMAP_batch3b.jpg", width=10, height=6, dpi=300, plot= last_plot(), units = "in")
UMAPPlot(AML3_alone, label=TRUE) 
TSNEPlot(AML1_alone, label=TRUE) #+ ggsave("TSNE_batch2_mindist.jpg", width=10, height=6, dpi=300, plot= last_plot(), units = "in")
TSNEPlot(AML3_alone, label=TRUE)

```
Try the GF-ICF method
https://github.com/dibbelab/gficf
Can only be installed from source, but it's not happening despite R and Rstudio upgrades
Need to upgrade R. My Current version is 3.5.1. Current version is 4.0.0
```{r}
install.packages(pkgs = "gficf",repos = c("https://dibbelab.github.io/Rrepo/","https://cloud.r-project.org")) #doesn't work
install.packages("gficf") #doesn't work, says not available
devtools::install_github("dibbelab/gficf")
```


#######
Identify Clusters by Top markers
#######

```{r}
pbmc.markers <- FindAllMarkers(AML1_alone, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.5)
#save(pbmc.markers, file="MDSC_all_vs_all_DE_batch2.rdata")
save(pbmc.markers, file="/Volumes/Picard/FLX/scRNA_data/MDSC_Talmadge/All_vs_all_DE_AML1.rdata")

write.table(pbmc.markers, file="Talmadge_scRNA_cluster_markers_batch4.txt", sep="\t", quote=FALSE,row.names = FALSE)

pbmc.markers  %>% top_n(n = 100, wt = avg_logFC) %>% arrange(cluster)
pbmc.markers %>% group_by(cluster) %>% top_n(n = 30, wt = avg_logFC)

FindMarkers(AML1_alone, ident.1=10, ident.2=14,only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)


#mRNA_ident <- c("0 gMDSC","1 T cell CD4+ TCF7+","2 NK","3 T cell CD4 CCR7+","4 T cell CD4 CCR7+ TCF7+","5 gMDSC SELL+","6 T cell CD8+","7 mMDSC CD15+","8 Granulocyte","9 mMDSC CD14+/- CD34 med","10 Platelet","11 B cell","12 T cell CD4 RPhigh","13 mMDSC CD15+ CSF1R+","14 Granulocyte CD16+","15 Granulocyte CD16+","16 gMDSC DEFA3+","17 Apoptotic","18 Unknown GAPDH++","19 DC","20 NK XCL+")
```

Plot CITE markers
```{r fig.height=15, fig.width=20}
FeaturePlot(AML1_alone,slot = "data",features = c("CD11b-CITE","CD11c-CITE","CD14-CITE","CD33-CITE","BPI","CD15-CITE","CSF-1R-CITE","CD34-CITE","FCGR3B","CD45RA-CITE","CD45RO-CITE","KIT","MKI67","RPL37"))

DimPlot(AML1_alone) + NoLegend()
```

